### 2022-08-17 manuscript.Rmd
init manuscript ch2

---
title: "Diversification of Australian blind snakes"
date: "`r Sys.Date()`"
author: 
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [C:/Users/ST/Documents/repo/AusARG_blindsnake/custom-reference-doc.docx]
    number_sections: true
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
# biblio-style: "apalike"
csl: evolution.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
# knitr::opts_knit$set(root.dir = normalizePath("..")) #use grandparent folder working directory
# for including graphics we need full path with knitr
WD <- getwd()
```

```{r packages}
library(dplyr); library(readr); library(janitor); 
library(ape); library(phytools);
library(kableExtra); library(RPANDA)
```

**Abstract**

**Keywords**
scolecophidia, Diversification, paleoclimate, RPANDA, BioGeoBEARS

\newpage

# Introduction
<!-- COMMUNITY words - Reported, accepted, widely  -->
<!-- TRANSITION / FLOW words - Nonetheless, however, although, unless, because, and, but/ -->
<!-- 1. Why phylogeny is important to readers of the target journals? -->

#  Materials and methods

# Results

\newpage

# Discussion

# Conclusion

\newpage

**References**
<!-- # References -->

### 2022-08-18 code/utility/00_diversification_tree_data_prep.R
init

@@ -1,0 +1,162 @@
+# 00_diversification_tree_data_prep.R
+# Putter Tiatragul
+# Started on 27 May 2022
+
+# The purpose of this script is to prepare the data required for diversification analyses.
+# --- Tree data -- SqCL exon-capture tree (two versions depends on prior)
+# --- Environment data -- global paleo climate, aridity, and precipitation (Alex Skeels)
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse); library(RPANDA)
+library(phytools); library(ape)
+library(gridExtra); library(ggplot2)
+
+# phylogeny ---------------------------------------------------------------
+# Time calibrated tree from SqCL probe and mcmctree analysis with fossils
+
+# Full tree with fossils
+fos_tree_full <- ape::read.nexus("data/tree/20220420mcmctree.tre")
+phy_full <- ape::read.nexus("data/tree/anilios_v3_st.tre")
+phy_full_b <- ape::read.nexus("data/tree/anilios_v3_b.tre")
+
+# Rename cfwaitii
+phy_full$tip.label[which(phy_full$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_SAMA_R51244"
+phy_full_b$tip.label[which(phy_full_b$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_SAMA_R51244"
+
+# Rename polygrammicus to Sundatyphlops
+phy_full$tip.label[which(phy_full$tip.label == "Anilios_polygrammicus_R98715")] <- "Sundatyphlops_polygrammicus_R98715"
+
+
+# Subset species tree to Acutotyphlops & Anilios ------------------------------------------
+
+# Additional species to drop
+# A. splendidus, we decided to synonymise this species with A. pinguis
+# A. centralis R51244 is redundant
+# only need one unguirostris at the moment. Will only keep 180002
+to_drop <- c("Anilios_splendidus_R119900", "Anilios_centralis_R51244",
+             "Anilios_unguirostris_R115861", "Anilios_unguirostris_R115861", "Anilios_unguirostris_R21669", "Anilios_unguirostris_R54430")
+
+plot(phy_full);nodelabels()
+
+phy_small <- phy_full %>% 
+  ape::extract.clade(phy = ., 159) %>% 
+  ape::drop.tip(phy = ., tip = to_drop)
+
+phy_small_b <- phy_full_b %>% 
+  ape::extract.clade(phy = ., 159) %>% 
[…]

### 2022-08-19 code/01_diversification_analyses.R
init, correct number of species

@@ -1,0 +1,296 @@
+# 02_diversification_analyses.R
+# Putter Tiatragul
+# June 2022
+
+# The purpose of this script is to fit diversification models with RPANDA
+#### Pre-reqs
+# source("code/utility/00_diversification_env_data_prep.R")
+# source("code/utility/00_diversification_tree_data_prep.R")
+####
+
+## IMPORTANT ## 
+# These analyses can be slow, especially when running multiple models. 
+# From initial testing, using R on WSL2 is faster than R on Windows even when using the same source files on Windows. 
+# So when running analyses, I'm using Rstudio server launched from WSL2. 
+
+# Library -----------------------------------------------------------------
+
+library(RPANDA); library(tidyverse)
+library(phytools); library(gridExtra)
+library(parallel) # can do this on WSL2 (if using Windows)
+
+
+# Load --------------------------------------------------------------------
+
+load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
+trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
+
+is.ultrametric(trees[[1]])
+fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+plotTree(fos_tree)
+is.ultrametric(fos_tree)
+
+# Fit models ---------------------------------------------------------------
+
+# Crown age
+tot_time <- max(node.age(fos_tree)$ages)
+cond = "crown"
+
+## Fraction of diversity represented. In our tree we have 
+## 48 Described species of Anilios + 2 more grypus undescribed + 1 more ligatus - A. splendidus
+## So our "true" species at the moment is 50 and we have 41 tips in our tree
+
+length(fos_tree$tip.label)
+frac = 38/50
+
+# Explanation of the models we'll fit:
+# BCST
+#   + standard Yule model, speciation rate is constant and no extinction
+#	BCST_DCST
+#		+ standard birth-death model, with constant speciation and extinction rates
+#	BEnvVar_EXPO
[…]

### 2022-08-19 code/02_diversification_plots.R

@@ -1,0 +1,33 @@
+# 02_diversification_plots.R
+# Putter
+# Plotting diversification models
+
+
+# libraries ---------------------------------------------------------------
+
+library(RPANDA); library(dplyr)
+library(phytools); library(gridExtra)
+
+# Load results from 01_diversification_analyses ---------------------------
+# These models take a lot of time to run so I ran it on WSL2 and saved the image file. 
+
+load("output/20220818_model_output_env.RData")
+
+
+# Sort result table -------------------------------------------------------
+
+result_table <- rbind(tab_func(tree_BCSTDenvVar_EXPO_res, .mu = TRUE, gen_name = "BCSTDenvVar_expo"),
+      tab_func(tree_BEnvVarDCST_EXPO_res, .mu = TRUE, gen_name = "BEnvVarDCST_expo"),
+      tab_func(tree_BEnvVarDEnvVar_EXPO_res, .mu = TRUE, gen_name = "BEnvVarDEnvVar_expo"),
+      tab_func(tree_BEnvVar_EXPO_res, .mu = FALSE, gen_name = "BEnvVar_expo")) %>% 
+  arrange(aicc)
+
+head(result_table)
+
+result_table %>% 
+  dplyr::mutate_if(is.numeric, round, digits =2)
+
+
+tree_BEnvVar_EXPO_res$arid_sco
+
+RPANDA::plot_fit_env(tree_BEnvVar_EXPO_res$arid_str, env_data = env_data_list$arid_str, tot_time = tot_time)

### 2022-08-19 code/utility/00_diversification_env_data_prep.R

@@ -1,0 +1,93 @@
+# 00_diversification_env_data_prep.R
+# Putter Tiatragul
+# Started on 27 May 2022
+
+# The purpose of this script is to prepare the data required for diversification analyses.
+# --- Environment data -- global paleo climate, aridity, and precipitation (Alex Skeels)
+# Temperature variation based on combination of paleo digital elevation models (DEMs) and paleo climate bands. 
+# Contemporary mean values for Australian Koppen bands [@scoteseAtlas2021] projected back in time by Alex Skeels (quoted).
+# Reconstructions based different global DEMs by (1) Scotese @scotesePALEOMAP2018, and (2) 
+# Straume @straumeGlobal2020 is reconstructed by Alex Skeels using `gen3sis` @hagenGen3sis2020.
+
+
+# libraries ---------------------------------------------------------------
+library(tidyverse); library(RPANDA); library(stringr)
+
+# Climate data ------------------------------------------------------------
+
+# Global temp from RPANDA (isotopes)
+data("InfTemp")
+
+# Australia temp and aridity from Alex Skeels
+# --- aridity index (coded as p = precipitation because mislabeled from original data set
+# --- higher values means more arid)
+climate_data_sco <- read.csv("data/paleo_env/Australia_climate_data_45Mya_Scotese.csv")
+climate_data_str <- read.csv("data/paleo_env/Australia_climate_data_45Mya_Straume.csv")
+climate_data_val <- read.csv("data/paleo_env/Australia_climate_data_45Mya_Valdes_160K.csv") %>% dplyr::select(-prec_mean, -prec_median, -prec_sd)
+
+# Multiply p_mean by 100
+climate_data_sco$p_mean <- climate_data_sco$p_mean * 100
+climate_data_str$p_mean <- climate_data_str$p_mean * 100
+climate_data_val$p_mean <- climate_data_val$p_mean * 100
+
+# Temporal resolution in RPANDA::InfTemp is greater than reconstructions Alex Skeels' models 
+# so need to make same length by sub sampling RPANDA temp to similar timestep
+
+# choose one random Age within a % range of Age from Scotese temperature data. 
+climate_data_sco$time_mya[2]; climate_data_sco$time_mya[3]
+
+#  get timesteps from RPANDA's dataset that already match Alex's
+panda_true <- InfTemp %>% 
+  dplyr::filter(Age < 46) %>% 
[…]

### 2022-08-19 code/utility/00_diversification_tree_data_prep.R

@@ -4,28 +4,26 @@
 
 # The purpose of this script is to prepare the data required for diversification analyses.
 # --- Tree data -- SqCL exon-capture tree (two versions depends on prior)
-# --- Environment data -- global paleo climate, aridity, and precipitation (Alex Skeels)
+# --- Only Anilios taxa are considered.
 
 # libraries ---------------------------------------------------------------
 
-library(tidyverse); library(RPANDA)
-library(phytools); library(ape)
-library(gridExtra); library(ggplot2)
+library(dplyr); library(phytools); library(ape); library(stringr)
 
 # phylogeny ---------------------------------------------------------------
 # Time calibrated tree from SqCL probe and mcmctree analysis with fossils
 
 # Full tree with fossils
-fos_tree_full <- ape::read.nexus("data/tree/20220420mcmctree.tre")
 phy_full <- ape::read.nexus("data/tree/anilios_v3_st.tre")
 phy_full_b <- ape::read.nexus("data/tree/anilios_v3_b.tre")
 
 # Rename cfwaitii
-phy_full$tip.label[which(phy_full$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_SAMA_R51244"
-phy_full_b$tip.label[which(phy_full_b$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_SAMA_R51244"
+phy_full$tip.label[which(phy_full$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_R51244"
+phy_full_b$tip.label[which(phy_full_b$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_R51244"
 
 # Rename polygrammicus to Sundatyphlops
 phy_full$tip.label[which(phy_full$tip.label == "Anilios_polygrammicus_R98715")] <- "Sundatyphlops_polygrammicus_R98715"
+phy_full_b$tip.label[which(phy_full_b$tip.label == "Anilios_polygrammicus_R98715")] <- "Sundatyphlops_polygrammicus_R98715"
 
 
 # Subset species tree to Acutotyphlops & Anilios ------------------------------------------
@@ -35,7 +33,8 @@ phy_full$tip.label[which(phy_full$tip.label == "Anilios_polygrammicus_R98715")] 
 # A. centralis R51244 is redundant
 # only need one unguirostris at the moment. Will only keep 180002
[…]

### 2022-08-19 code/utility/color_by_CF_Ian.R
belong in another repo

@@ -1,149 +1,0 @@
-# The function requires a few things:
-## phy: your tree file in phylo format
-## cf.file: the path to your concordance factor file (e.g. "~/Desktop/cf.stat")
-## terminal.color: terminal branches don't have CFs, so how do you want them colored? Default is the color set for CF=100
-## col.palette: the color palette from RColorBrewer that you'd like. Default is "RdYlBu"
-## legend: TRUE/FALSE, do you want a legend of the gCF colors?
-## value.check: check to see if the applied values match those from the .cf.stat file
-## label.size: how big should the tip label size be?
-
-#  Add ability to subset and return two plots (full and subset)
-
-library(dplyr); library(phytools)
-
-color.by.CF <- function(phy, cf.file, terminal.color=NULL, col.palette="RdYlBu", 
-                        legend=TRUE, value.check=FALSE, label.size=0.5,
-                        subset = FALSE, subset_tips,
-                        CF=c("sCF", "gCF")){
-  require(RColorBrewer); require(dplyr); require(phytools)
-  
-  # set the plot layout
-  if(legend==T){layout(matrix(1:2,ncol=2), width = c(4,1),height = c(1,1))}
-  
-  # read in the gCF file
-  cf.data = read.delim(cf.file, header=T, comment.char="#")
-  
-  # rename the first column from 'ID' to 'child', 
-  # the naming convention for branches in IQTREE is not the same for edges in APE
-  # but it is the same as the number of the child node below an edge in APE, so we can use that
-  names(cf.data)[1] = "child"
-  # add root node, and apply an arbitrary value, this may just be funny behavior because of 
-  cf.data[1,"gCF"] <- 100; # arbitrary CF value for 2nd node
-  cf.data[1,"sCF"] <- 100;
-  cf.data <- rbind(cf.data, c(min(cf.data$child)-1, 100, NA,NA,NA,NA,NA)) # arbitrary CF value for root node
-  
-  # create a data frame of all the edges in the tree, listed by (parent, child) nodes
-  all.edges <- data.frame(parent=phy$edge[,1], child=phy$edge[,2])
-  
-  # extract just the branches that lead to tips
[…]

### 2022-08-19 code/utility/color_by_CF_Ian_breakdown.R

@@ -1,108 +1,0 @@
-
-support = read.delim("data/tree/concord.cf.stat", header=T, comment.char="#") # astral tree
-astral.tree <- read.tree("data/tree/concord.cf.tree")
-
-phy = astral.tree
-cf.file="data/tree/concord.cf.stat"
-value.check = T
-legend=T
-CF="gCF"
-col.palette="RdYlBu"
-
-library(RColorBrewer); 
-
-col.pallette
-
-  # set the plot layout
-  if(legend==T){layout(matrix(1:2,ncol=2), width = c(4,1),height = c(1,1))}
-  
-  # read in the gCF file
-  cf.data = read.delim(cf.file, header=T, comment.char="#")
-  
-  # rename the first column from 'ID' to 'child', 
-  # the naming convention for branches in IQTREE is not the same for edges in APE
-  # but it is the same as the number of the child node below an edge in APE, so we can use that
-  names(cf.data)[1] = "child"
-  # add root node, and apply an arbitrary value, this may just be funny behavior because of 
-  cf.data[1,"gCF"] <- 100; # arbitrary CF value for 2nd node
-  cf.data[1,"sCF"] <- 100;
-  cf.data <- rbind(cf.data, c(min(cf.data$child)-1, 100, NA,NA,NA,NA,NA)) # arbitrary CF value for root node
-  
-  # View(rbind(cf.data, c(min(cf.data$child)-1, 100, NA,NA,NA,NA,NA)))
-  
-  # create a data frame of all the edges in the tree, listed by (parent, child) nodes
-  all.edges <- data.frame(parent=phy$edge[,1], child=phy$edge[,2])
-  
-  # extract just the branches that lead to tips
-  tip.edges <- subset(all.edges, all.edges$child <= Ntip(phy))
-  tip.edges$gCF <- 100 # give these branches arbitrary values of 100
-  tip.edges$sCF <- 100 # give these branches arbitrary values of 100
-  
-  
-  # extract just the internal branches
-  int.edges <- subset(all.edges, all.edges$child > Ntip(phy))
-  
-  # get the branch-appropriate concordance factors
-  int.edges <- left_join(int.edges, cf.data)
-  int.edges <- int.edges[,c("parent", "child", "gCF", "sCF")] # subset to just the data we care about
-  
[…]

### 2022-08-19 code/utility/concordance_factor_plot.R

@@ -1,39 +1,0 @@
-library(dplyr)
-library(patchwork)
-
-source('code/utility/color_by_CF_Ian.R')
-
-# read in the CF file and appropriate tree (ASTRAL)
-support = read.delim("data/tree/concord.cf.stat", header=T, comment.char="#") # astral tree
-astral.tree <- read.tree("data/tree/concord.cf.tree")
-# astral.tree <- read.tree("data/tree/con")
-
-# Visualize the concordance factors as branch colors
-dev.off()
-
-# s1 <- astral.tree$tip.label[stringr::str_detect(astral.tree$tip.label, pattern = "Anilios")]
-# s2 <- astral.tree$tip.label[stringr::str_detect(astral.tree$tip.label, pattern = "Acutotyphlops")]
-# s3 <- astral.tree$tip.label[stringr::str_detect(astral.tree$tip.label, pattern = "Ramphotyphlops")]
-# s4 <- astral.tree$tip.label[stringr::str_detect(astral.tree$tip.label, pattern = "Indotyphlops")]
-# subset_things <- c(s1, s2, s3, s4)
-
-
-# CF.obj <- color.by.CF(astral.tree, cf.file="data/tree/concord.cf.stat", value.check = T, legend=T, CF="gCF", subset = TRUE, subset_tips = subset_things)
-CF.obj <- color.by.CF(astral.tree, cf.file="data/tree/concord.cf.stat", value.check = T, legend=T, CF="gCF", subset = FALSE)
-
-
-CF.scf <- color.by.CF(astral.tree, cf.file="data/tree/concord.cf.stat", value.check = T, legend=T, CF="sCF")
-# CF.obj <- color.by.CF(mcmctree_anilios, cf.file="data/tree/concord_ahe.cf.stat", value.check = T, legend=T, CF="gCF")
-
-
-dev.off()
-
-length(astral.tree$tip.label)
-
-# Plot gene concordance factors as a result of branch lengths
-CF_comparison(CF.obj, type="gCFxBL")
-CF_comparison(CF.obj, type="sCFxBL")
-CF_comparison(CF.obj, type="sCFxgCF")
-
-
-

### 2022-09-25 code/04_BioGeoBEARS_results_plots.R
initial script for presenting result

@@ -1,0 +1,184 @@
+# 04_BioGeoBEARS_results_plots.R
+# August 2022
+
+# Test best model fit from BioGeoBEARS analyses
+
+# load results from analysis script
+# source("code/03_BioGeoBEARS_analyses_parallel.R")
+
+
+# Libraries ---------------------------------------------------------------
+
+library(GenSA);library(FD)      
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+# library(parallel); library(snow)
+
+
+# Tree --------------------------------------------------------------------
+
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
+
+
+# Saved results -----------------------------------------------------------
+## These saved results are from source('code/03_BioGeoBEARS_analyses_parallel.R)
+
+bio_geo_bears_files_list <- list.files(path = "./data/intermediate_data/bears/",
+                                       pattern = "*.Rdata", full.names = T)
+
+# Load individual saved files
+lapply(bio_geo_bears_files_list, load, .GlobalEnv)
+
+
+# Results table -----------------------------------------------------------
+# Extract parameters and make results table to compare AICc scores
+
+restable = NULL
+teststable = NULL
+
+fin.dec = extract_params_from_BioGeoBEARS_results_object(results_object=resDEC, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decj = extract_params_from_BioGeoBEARS_results_object(results_object=resDECj, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decx = extract_params_from_BioGeoBEARS_results_object(results_object=resDECx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decjx = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.diva = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKE, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
[…]

### 2022-09-26 manuscript.Rmd
update results biogeobears

@@
 
+The aridification on continental Australia over the past 20 million years resulted in a diversity of habitats and spurred multiple radiations of flora and fauna across the continent, many of which are endemic [@byrneBirth2008; @melvilleEvolutionary2011].
+The arid zone makes up the majority of the Australian landscape and provides habitat to ~43% of all Australia’s terrestrial reptiles.
+Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which have undergone rapid diversification in the last 10 – 6 myr [@brennanMass2017; @oliverPhylogenetic2017].
+For example, some Australian squamate clades like dragons [@melvilleIntegrating2019], sphenomorphine skinks [@raboskyAnalysis2014], geckos [@ashmanDiversification2018; @brennanMass2017] and elapid snakes [@leeDiversification2016; @grundlerTrophic2014] show increased rates of diversification coincidentally with the onset of aridification.
+However, these studies did not account for heterogeneity of the arid zone [@pepperLife2021].
+
+Roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises 48 described species [@uetzQuarter2021].
+All _Anilios_ species are found in Australia except for _A.
+eryncinus_, which is only found in Papua New Guinea.
+While morphologically conserved and difficult to distinguished, extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013; @pyronSystematics2014; @ellisTyphlopid2019].
+The diversification is reported to have originated from a single colonisation event from Asia between 19-39 mya [@vidalBlindsnake2010; @marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
+However, the study did not account for alternative biogeographic areas (e.g.
+WWF Ecoregions) and models, or uncertainty in phylogenetic relationships and divergence times.
+Despite representing a substantial portion of snake biodiversity there has been few investigations into the macroevolutionary dynamics of blind snakes.
+Recent phylogenomic tree and monophyly of the genus makes the Australian _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in almost every biome [citefirstchapter?].
+
+<!-- Environmental correlates of diversification -->
+<!-- agamid lizards [@harmonTempo2003], passerine birds [@ricklefsGlobal2006], amphibians [@rollandContribution2019], rosid plants [@sunRecent2020]).  -->
+A variety of ecological correlates have been associated with diversification processes of a clade.
+Among others, studies on the influence of climate (e.g.
+temperature, precipitation, and latitude) [@ricklefsGlobal2006; @mayhewBiodiversity2012; @condamineMacroevolutionary2013; @cantalapiedraDietary2014; @condamineClimate2019; @rollandContribution2019; @sunRecent2020]], elevation [@esquerreHow2019; @boschmanMountain2021], biological interactions [@ramirezAsynchronous2011; @druryEstimating2016; @brennanPhylogenomicsMonitorLizards2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007; @pepperGenetic2011; @deepakAridification2018; @catulloAridification2014] have been documented.
+In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010; @pepperGenetic2011; @pepperPalaeoclimate2011; @oliverIndependent2014; @oliverContrasting2014], frogs [@catulloAridification2014; @vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017; @toussaintUnveiling2015], and other plants [reviewed in @crispHow2013].
+
+There are several links between temperature and diversification rates of reptiles.
+First, warm temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
+Second, spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011; @lawsonLatitudinal2014].
+Third, change in climate change (e.g.
+glaciation) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
+Fourth, ectotherms are subjected to higher metabolic rates in warm climates (i.e.
+Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006; @makarievaMean2008].
+Finally, aridification of Australia may offer new ecological niche for species to exploit, which have been shown to result in deep divergences in many arid-adapted [@byrneBirth2008; @pepperGenetic2011; @marinTracing2013].
+
+The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013; @kergoatOpposite2018; @condamineClimate2019; @rollandContribution2019], these temperature data are reconstructed (interpolated) based on conversion of concentration of oxygen isotopes [@condamineMacroevolutionary2013].
+However, the relationship between diversification rates and aridity are attributed to some proxy such as the the timing of aridification or global temperature rather than quantifying the extent of aridity itself [@dornStrange2014; @deepakAridification2018; @catulloAridification2014; @fujitaDiversification2010; @pepperGenetic2011; @pepperPalaeoclimate2011; @oliverIndependent2014; @oliverContrasting2014; @owenHow2017; @crispHow2013; @toussaintUnveiling2015].
+As such, no study has been able to quantitatively assess aridity as a continuous variable in environment-dependent models and compare it to other ecological processes.
+Recent development in geographic-based methods such as `gen3sis` modeling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
+In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia (instead of relying on global paleoclimate data) in the time span of interest (between the 45 mya and the present).
+These datasets will allow us to empirically test the influence of localised ecological processes on diversification rates over time.
+
+In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates in arid and non-arid zones by testing hypotheses about the processes that promote speciation.
+Applying the cutting edge localised geographic-based paleoenvironment reconstruction, we evaluated (1) the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, (2) the model fit between different ecological processes on diversification rates, and (3) What is the ancestral range of Australian blindsnakes?
+(4) Are there differences in the rates of speciation between arid and non-arid zones?
+
 #  Materials and methods
 
+## Reconstruction of Australian paleo climate 
+<!-- Consult Alex Skeels -->
+Two paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from Scotese (2021) [@scoteseAtlas2021] and Straume et al (2020) [@straumeGlobal2020].
+We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in Hagen et al (2021) @hagenGen3sis2021.
+From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day (\@ref(fig:fig-aus-region).
+
+## Phylogenetic hypothesis
+<!-- Based on our phylogenetic tree -->
+We employ the recent time-calibrated phylogenomic tree of Australian blindsnake with outgroups.
+This tree includes __ taxa representing __/48 described species of _Anilios_.
+However, because we our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to x/52 species.
+
+## Biogeography
+
+To evaluate the ancestral biome of blindsnakes in Australia, we used the R package `BioGeoBEARS` 1.1.1 [@nicholas_j_matzke_2018_1478250].
+BioGeoBEARS provides a flexible likelihood-based framework that models how the branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
+These shifts in biomes can occur within a single branch (anagenetic events including processes like dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant and founder-event speciation).
+The fundamental inputs for the software include a time-calibrated phylogeny (described above), information about the biome/range each species are found, and a historical biogeographic model.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) in which the based on occurrence records from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Supplementary information S2 for how data was compiled).
+Biome designations were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017; @pavon-vazquezComprehensive2021].
+
+The phylogeny and distribution information were then applied to widely used historical biogeographical methods including Dispersal-Vicariance Analysis (DIVA; [@ronquistDispersalVicariance1997]), and likelihood implementations of Dispersal-Extinction-Cladogenesis (DEC; [@reeMaximum2008]) or DECLIKE, and BayArea [@landisBayesian2013] or BAYAREALIKE [@matzkeBioGeoBEARS2013].
+Additionally, we tested these models with distance matrix as a free parameter (+X).
+We then compared these models and selected the best models based on AICc (corrected Akaike Information Criterion) scores.
+The best-fitting model was then used to calculate the probability of the ancestral states by using the BioGeoBEARS stochastic mapping function (`N` runs).
+
+<!-- Include environments on range evolution ?? -->
+
+The best-fitting model was then used to calculate the probability of the ancestral states by using the BioGeoBEARS stochastic mapping function (`N` runs).
+
+We tested several biogeographic models including  and their modifiers (See supplementary).
+
+<!-- SUPPLEMENT TABLE WITH OCCURENCE RECORDS AND BIOME MAP -->
+
+## Diversification analyses
+
+To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+These models will be a combination of pure-birth and birth-death models with different speciation/extinction rates specified as constant, linearly, or exponentially correlated with paleoenvironment.
+We fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 global paleo temperature, 3 Australian paleo temperature, 3 Australian paleo aridity, and 3 Australian paleo proportion arid area.
+Missing taxa is accounted for by fractioning function in RPANDA.
+We assessed model fit with corrected Akaike Information Criteria (AICc) scores.
+
+- Initial lamba parameters were changed for mean_sco
+
+## Biogeographic-dependent 
+To test whether diversification rates differs between arid and non-arid zones test for biogeographic-depdendent diversification using geographic state and hidden state speciation-extinction model (GeoHiSSE) analysis with different parameters [@goldbergPhylogenetic2011].
+
+MiSSE https://www.biorxiv.org/content/10.1101/2021.11.02.466977v1
+
 # Results
+
+## Paleoclimate conditions
+
+Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
+Both aridity index shows increase in aridity post 10 mya, consistent with aridification of the continent [@byrneBirth2008].
+Temperature for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
+Reconstructed temperature values from Australia were higher than global temperature and stayed relatively constant compared to global temperature that shows slight decrease.
+
+## Biogeographic history
+
+We found that the best fitting model is DEC plus jump dispersal and distance matrix (DEC+j+x). 
+Likelihood ratio test (LRT) indicates that given the number of extra free parameters for the complex model, it is a statistically significant improvement over the simpler DEC, DEC+j, and DEC+x models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+
+- Topography dyanmics [@brazModelling2021]
+- @marinTracing2013 concluded that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+- Range expansion was important for diversification ...
+
+Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
+
+## Diversification
+The best fitting model among constant, time-dependent, temperature-dependent, and aridity dependent models is 
+This model indicate that X associate with speciation (alpha = ) and .....
+
+Fit with Australia specific mean temperature receives better support than model with global temperature and exponential dependency of speciation rate with time.
+
+- Solomon Island 40 MYA [@hallCenozoic2002]
 
@@
 
+## Localised climate model outperforms global models
+
+## Ancestral biome reconstruction
+- Consistent with Marin et al.
+- Timing is similar to other squamates arrival
+
+## Drivers of diversifcation 
+-
+
 # Conclusion
+
+# Tables
+
+Table BioGeoBEARS. BioGeoBEARS model estimates and model AICc scores. Lower AICc scores indicate model is preferred. + Log likelihood
+Table RPANDA. Diversification model fit estimates and AICc score of model fit
+
+# Figure legends
+
+Fig. Ancestral biome: Stochastic map of Ancestral biome from Australo-Papuan blind snakes
+Fig.
+Paleoclimate: Reconstructed paleoclimate values from different data sets.
+Mean temperature data.
+Aridity index based on precipitation threshold.
+Abbreviations Pli = Paleocene and Ple = Pleistocene.
+Fig. Speciation rates + phylogeny over time + lineages through time
 

### 2022-09-30 manuscript.Rmd
update methods and biogeobears results

@@
 In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates in arid and non-arid zones by testing hypotheses about the processes that promote speciation.
-Applying the cutting edge localised geographic-based paleoenvironment reconstruction, we evaluated (1) the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, (2) the model fit between different ecological processes on diversification rates, and (3) What is the ancestral range of Australian blindsnakes?
+Applying the cutting edge localised geographic-based paleoenvironment reconstruction,
+
+We (1) evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, (2) compared the model fit between different ecological processes on diversification rates, and (3) reconstructed the ancestral biome of Australian blindsnakes?
+
 (4) Are there differences in the rates of speciation between arid and non-arid zones?
@@
 
-## Phylogenetic hypothesis
+## Phylogenetic hypothesis and taxon sampling
 <!-- Based on our phylogenetic tree -->
@@
 However, because we our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to x/52 species.
+The species were pruned to single operational taxonomic unit (OTU).
 
@@
 We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) in which the based on occurrence records from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Supplementary information S2 for how data was compiled).
-Biome designations were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017; @pavon-vazquezComprehensive2021].
+Biome designations were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
 
 The phylogeny and distribution information were then applied to widely used historical biogeographical methods including Dispersal-Vicariance Analysis (DIVA; [@ronquistDispersalVicariance1997]), and likelihood implementations of Dispersal-Extinction-Cladogenesis (DEC; [@reeMaximum2008]) or DECLIKE, and BayArea [@landisBayesian2013] or BAYAREALIKE [@matzkeBioGeoBEARS2013].
-Additionally, we tested these models with distance matrix as a free parameter (+X).
-We then compared these models and selected the best models based on AICc (corrected Akaike Information Criterion) scores.
-The best-fitting model was then used to calculate the probability of the ancestral states by using the BioGeoBEARS stochastic mapping function (`N` runs).
-
-<!-- Include environments on range evolution ?? -->
-
-The best-fitting model was then used to calculate the probability of the ancestral states by using the BioGeoBEARS stochastic mapping function (`N` runs).
-
-We tested several biogeographic models including  and their modifiers (See supplementary).
+, we included two additional free parameters.
+The first is the *j* parameter that assigns per-event weight to founder-events [@matzkeModel2014].
+The second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Supplementary information).
+We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on AICc (corrected Akaike Information Criterion) scores.
 
@@
 Missing taxa is accounted for by fractioning function in RPANDA.
-We assessed model fit with corrected Akaike Information Criteria (AICc) scores.
+We assessed model fit with corrected AICc scores.
 
@@
 Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
-Both aridity index shows increase in aridity post 10 mya, consistent with aridification of the continent [@byrneBirth2008].
+Both aridity index shows increase in aridity post 10 mya, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
 Temperature for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
@@
 ## Biogeographic history
+We reconstructed the tropical grassland as the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears). The preferred biogeographic model is the constrained-dispersal matrix, DEC+j+x. 
+Likelihood ratio test (LRT) indicates that founder event (parameter j) and distance matrix (free parameter _x_) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+ *j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+Our result also supports the hypothesis that ancestor of the Australian clade invaded the continent, but exactly where they lived prior to this invasion is equivocal between the Solomon Islands and New Guinea.
+After arriving to Australia in the late Oligocene – early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years. 
+- Our analysis is consistent with @marinTracing2013 conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
 
-We found that the best fitting model is DEC plus jump dispersal and distance matrix (DEC+j+x). 
-Likelihood ratio test (LRT) indicates that given the number of extra free parameters for the complex model, it is a statistically significant improvement over the simpler DEC, DEC+j, and DEC+x models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+## Biome transitions and current distributions
+- The earliest biome transition was from tropical grassland into XXX by XX MYA (95% HPD: XX – XX). 
+- Subsequent splits show transitions the arid zone coinciding with with the onset of aridification between 20 – 10 MYA. 
+- Most species are still found in tropical grasslands, but also in arid zone. 
+- Greatest number of species can be found in ...
+
+- Range expansion was important for diversification ...
 
 - Topography dyanmics [@brazModelling2021]
-- @marinTracing2013 concluded that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
-- Range expansion was important for diversification ...
 
@@
 ## Ancestral biome reconstruction
-- Consistent with Marin et al.
-- Timing is similar to other squamates arrival
+- Our result — that the ancestor of Anilios occupied tropical grasslands – is consistent with the result presented in @marinTracing2013. 
+- We found transitions from XXX to XXX to be common as have been shown / as opposed to other Australian squamates (CITATIONS)
+- Arid zone was not arid until ~20 Million Years ago. Ancestors in deeper time probably occupied temperate or tropical areas that subsequently became the arid zone we know see today. 
+- The Australo-Papuan clade seems to have stayed in Australia for the most part but we were not able to include New Guinean endemic _Anilios erycinus_, the only recognised species of _Anilios_ not found in Australia.
+Therefore, we cannot make any conclusions about transitions between Australia and New Guinea even though at least three species (_A.
+torresianus_, _A.
+leucoproctus_, ...) are found in Northern Australia Torres Strait Islands, and southern parts of New Guinea.
 
 ## Drivers of diversifcation 
--
+- We expected the biogeographic models with jump dispersal to be the best fitting model because _Anilios_ are believed to have non-Australian origins [@marinTracing2013;@vidalBlindsnake2010;@oliverPhylogenetic2017].
+- Blind snakes are thought to not move very much and so formation of geological barriers during aridification may accelerate speciation between 20 – 10 MYA. 
+- The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
+Events like floods are likely to have faciliated dispersals but in short distances.
+Discuss species that are far apart...
+- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
+- Speciation could even be sympatric but no evidence to support this.
 

### 2022-09-30 code/01_diversification_analyses_individual.R
individual models updated

@@ -1,23 +1,15 @@
 # Individual_env_fit.R
 
-library(picante)
-library(pspline)
-library(RPANDA)
-library(dplyr)
-library(phytools); library(gridExtra)
-library(parallel) # can do this on WSL2 (if using Windows)
-
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
 trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
 
-fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
+is.ultrametric(trees[[1]])
+fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
 plot(fos_tree); axisPhylo()
 is.ultrametric(fos_tree)
-Anilios <- fos_tree
 
 # Fit models ---------------------------------------------------------------
 
@@ -36,10 +28,8 @@ env_data_list <- lapply(env_data_list, subset_env_age)
 ## 48 Described species of Anilios + 2 more grypus undescribed + 1 more ligatus - A. splendidus
 ## So our "true" species at the moment is 50 and we have 41 tips in our tree
 
-phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-tot_time<-max(node.age(phylo)$ages)
-f<-Ntip(phylo)/50 # As I found in Reptile Database
-cond="crown"
+length(fos_tree$tip.label)
+frac = 38/50
 
 env_data <- env_data_list$mean_sco
 
@@ -105,8 +95,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
[…]

### 2022-09-30 code/01_diversification_analyses_parallell.R

@@ -42,8 +42,11 @@ library(parallel) # can do this on WSL2 (if using Windows)
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
 trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
 
-is.ultrametric(trees[[1]])
-fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+names(env_data_list)
+sub_env_data_list <- env_data_list[c("min_sco", "arid_sco", "mean_str", "min_str", "arid_str", "min_val", "max_val", "arid_val")]
+
+
+fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
 plot(fos_tree); axisPhylo()
 is.ultrametric(fos_tree)
@@ -60,6 +63,7 @@ subset_env_age <- function(df) {
 }
 
 env_data_list <- lapply(env_data_list, subset_env_age)
+sub_env_data_list <- lapply(sub_env_data_list, subset_env_age)
 
 ## Fraction of diversity represented. In our tree we have 
 ## 48 Described species of Anilios + 2 more grypus undescribed + 1 more ligatus - A. splendidus
@@ -315,12 +319,9 @@ env_fitter <- function(env_data_list){
   return(final_Anilios, Anilios_res)
 }
 
-result_fit <- mclapply(X = env_data_list, FUN=function(i)env_fitter(i), mc.cores = 6)
+result_fit <- mclapply(X = sub_env_data_list, FUN=function(i)env_fitter(i), mc.cores = 6)
 
 # save.image(file = 'output/20220823_multiRPANDA_fit.Rdata')
-
-write.table(final_Anilios,file="/YOUR_PATH/Results_Anilios_EnvDep.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="/YOUR_PATH/final_AniliosTrees_EnvDep.RData")
-save(Anilios_res,file="/YOUR_PATH/Anilios_EnvDep.RData")
+# save.image(file = 'output/20220823_multiRPANDA_fit_b.Rdata')
 
 

### 2022-09-30 code/01_fit_arid_sco.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$arid_sco
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_sco.Rdata")
[…]

### 2022-09-30 code/01_fit_arid_str.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$arid_str
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_str.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_str.Rdata")
[…]

### 2022-09-30 code/01_fit_arid_val.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$arid_val
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_val.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_arid_val.Rdata")
[…]

### 2022-09-30 code/01_fit_global.R

@@ -9,7 +9,7 @@ library(parallel) # can do this on WSL2 (if using Windows)
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
 trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
 
-fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
 Anilios <- fos_tree
 
@@ -292,7 +292,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_global_temp.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp.Rdata")
+# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_global_temp.Rdata") 
+write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
+save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_global_temp_b.Rdata")
+save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_global_temp_b.Rdata")
 

### 2022-09-30 code/01_fit_max_sco.R

@@ -9,7 +9,7 @@ library(parallel) # can do this on WSL2 (if using Windows)
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
 trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
 
-fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
 Anilios <- fos_tree
 
@@ -290,7 +290,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_sco.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco.Rdata")
+# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_sco.Rdata")
+write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
+save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_sco_b.Rdata")
+save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_sco_b.Rdata")
 

### 2022-09-30 code/01_fit_max_str.R

@@ -9,8 +9,8 @@ library(parallel) # can do this on WSL2 (if using Windows)
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
 trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
 
-is.ultrametric(trees[[1]])
-fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+is.ultrametric(trees[[2]])
+fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
 plot(fos_tree); axisPhylo()
 is.ultrametric(fos_tree)
@@ -293,7 +293,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_str.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str.Rdata")
+# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_str.Rdata")
+write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
+save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_str_b.Rdata")
+save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_str_b.Rdata")
 

### 2022-09-30 code/01_fit_max_val.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$max_val
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_val.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_max_val.Rdata")
[…]

### 2022-09-30 code/01_fit_mean_str.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$mean_str
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_str.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_str.Rdata")
[…]

### 2022-09-30 code/01_fit_mean_val.R

@@ -1,12 +1,8 @@
-# Individual_env_fit.R
+# mean_val.R
 
-library(picante)
-library(pspline)
-library(RPANDA)
-library(dplyr)
+library(RPANDA); library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
-
 
 # Load --------------------------------------------------------------------
 
@@ -15,8 +11,6 @@ trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/
 
 fos_tree <- phytools::force.ultrametric(trees[[2]],"extend")
 fos_tree$edge.length <- fos_tree$edge.length * 100
-plot(fos_tree); axisPhylo()
-is.ultrametric(fos_tree)
 Anilios <- fos_tree
 
 # Fit models ---------------------------------------------------------------
@@ -41,8 +35,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
-
+env_data <- env_data_list$mean_val
 #################
 ## Fit the models ##
 #################
@@ -61,7 +54,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -105,8 +98,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
-  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2]) # difference 
-  lamb_par<-c(0.1,0)
+  lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2])
   mu_par<-c(0.01)
[…]

### 2022-09-30 code/01_fit_min_sco.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$min_sco
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_sco.Rdata")
[…]

### 2022-09-30 code/01_fit_min_str.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$min_str
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_str.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_str.Rdata")
[…]

### 2022-09-30 code/01_fit_min_val.R

@@ -7,7 +7,6 @@ library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
 
-
 # Load --------------------------------------------------------------------
 
 load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
@@ -41,7 +40,7 @@ tot_time<-max(node.age(phylo)$ages)
 f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
-env_data <- env_data_list$mean_sco
+env_data <- env_data_list$min_val
 
 #################
 ## Fit the models ##
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -303,10 +302,10 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   Anilios_res<-c(Anilios_res,list(resi))
 }
 
-# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.txt", quote=FALSE,sep="\t",row.names=FALSE)
-# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco.Rdata")
-# save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata")
-write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
-save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
-save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
+# write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_val.txt", quote=FALSE,sep="\t",row.names=FALSE)
+# save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_min_val.Rdata")
[…]

### 2022-09-30 code/00_anilios_BioGeoBEARS.R
update areas range and other parameters

@@ -6,23 +6,24 @@
 
 library(GenSA)    # GenSA is better than optimx (although somewhat slower)
 library(FD)       # for FD::maxent() (make sure this is up-to-date)
-# library(parallel)
+library(parallel)
 library(snow)
 library(rexpokit)
 library(cladoRcpp)
 library(BioGeoBEARS)
 
-tree.c <- ape::read.tree('data/tree/anilios_newick_st.tre')
-
+# Check tree
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
 plot(tree.c)
 
-trfn <- "data/tree/anilios_newick_st.tre"
+trfn <- "data/intermediate_data/bears/blindsnake_b.tre"
+dist.mat <- "data/bears_txt/biome_distance.txt"
 geogfn <- "data/bears_txt/geofile.txt"
 # adj <- "data/bears_txt/area_adj.txt"
 
 # Set the maximum number of areas any species may occupy; this cannot be larger 
 # than the number of areas you set up, but it can be smaller.
-max_range_size = 6
+max_range_size = 5
 
 # Look at your geographic range data:
 tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
@@ -33,8 +34,9 @@ max(rowSums(dfnums_to_numeric(tipranges@df)))
 
 coln <- c("A", "B", "C", "D", "E", "F", "G", "H", "I")
 
+# Manually setting up area-adjacency using Octavio's code
 #                     A   B   C   D   E   F   G   H   I
-adjacency <- matrix(c(1,	1,	1,	1,	1,	1,	1,	0,	1,  # A
+adjacency <- matrix(c(1,	1,	1,	1,	1,	1,	1,	1,	1,  # A
                       1,	1,	1,	1,	1,	1,	0,	0,	0,  # B
                       1,	1,	1,	1,	1,	1,	0,	0,	0,  # C
                       1,	1,	1,	1,	0,	1,	0,	0,	0,  # D
@@ -45,8 +47,7 @@ adjacency <- matrix(c(1,	1,	1,	1,	1,	1,	1,	0,	1,  # A
                       0,	0,	0,	0,	0,	0,	1,	1,	1), # I
                       ncol=9, nrow=9, byrow=T)
 
-
-maxAreas = 6
+maxAreas = 5
[…]

### 2022-09-30 code/01_diversification_analyses.R
text

@@ -11,7 +11,7 @@
 ## IMPORTANT ## 
 # These analyses can be slow, especially when running multiple models. 
 # From initial testing, using R on WSL2 is faster than R on Windows even when using the same source files on Windows. 
-# So when running analyses, I'm using Rstudio server launched from WSL2. 
+# So when running analyses, I'm using R (WSL2 console) or Rstudio server launched from WSL2 via browser. 
 
 # Library -----------------------------------------------------------------
 

### 2022-09-30 code/02_diversification_plots.R
plotting

@@ -13,21 +13,12 @@ library(phytools); library(gridExtra)
 
 load("output/20220818_model_output_env.RData")
 
+source("code/02_table_fit_env_results.R")
 
-# Sort result table -------------------------------------------------------
-
-result_table <- rbind(tab_func(tree_BCSTDenvVar_EXPO_res, .mu = TRUE, gen_name = "BCSTDenvVar_expo"),
-      tab_func(tree_BEnvVarDCST_EXPO_res, .mu = TRUE, gen_name = "BEnvVarDCST_expo"),
-      tab_func(tree_BEnvVarDEnvVar_EXPO_res, .mu = TRUE, gen_name = "BEnvVarDEnvVar_expo"),
-      tab_func(tree_BEnvVar_EXPO_res, .mu = FALSE, gen_name = "BEnvVar_expo")) %>% 
-  arrange(aicc)
-
-head(result_table)
-
-result_table %>% 
-  dplyr::mutate_if(is.numeric, round, digits =2)
+# Best fitting model
+result_fit$arid_sco[[1]]$BEnvVarDCST_EXPO
+result_fit$min_sco[[1]]$BCSTDEnvVar_EXPO
 
 
-tree_BEnvVar_EXPO_res$arid_sco
-
-RPANDA::plot_fit_env(tree_BEnvVar_EXPO_res$arid_str, env_data = env_data_list$arid_str, tot_time = tot_time)
+RPANDA::plot_fit_env(result_fit$arid_sco[[1]]$BEnvVar_EXPO, env_data = env_data_list$arid_sco, tot_time = tot_time)
+RPANDA::plot_fit_env(result_fit$min_sco[[1]]$BCSTDEnvVar_EXPO, env_data = env_data_list$min_sco, tot_time = tot_time)

### 2022-09-30 code/03_BioGeoBEARS_analyses_parallel.R
init RPANDA

@@ -1,0 +1,1020 @@
+# 03_anilios_BioGeoBEARS_parallel.R
+# August 2022
+
+# Script to fit different biogeography models with BioGeoBEARS and save results.
+# base code from BioGeoBEARS website, Octavio Jimenez Robles and Carlos Pavon Vazquez
+
+# Notes:
+# +J is for jump dispersal; + X is to account for distance between biome
+# Models: DEC, BAYAREALIKE, DIVALIKE
+# Fit models, models+j, models+x, models+j+x
+
+# Libraries ---------------------------------------------------------------
+
+library(GenSA)    # GenSA is better than optimx (although somewhat slower)
+library(FD)       # for FD::maxent() (make sure this is up-to-date)
+library(parallel); library(snow)
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+
+# Check tree
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
+plot(tree.c)
+
+
+# Load inputs -------------------------------------------------------------
+
+# Tree file
+trfn <- "data/intermediate_data/bears/blindsnake_b.tre"
+# Distance matrix file 
+dist.mat <- "data/bears_txt/biome_distance.txt"
+# Geo file
+geogfn <- "data/bears_txt/geofile.txt"
+# adj <- "data/bears_txt/area_adj.txt"
+
+
+
+# Specifications ----------------------------------------------------------
+
+# Set the maximum number of areas any species may occupy; this cannot be larger 
+# than the number of areas you set up, but it can be smaller.
+max_range_size = 5
+
+# Look at your geographic range data:
+tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
+tipranges
+
+# Maximum range size observed:
+max(rowSums(dfnums_to_numeric(tipranges@df)))
+
+coln <- c("A", "B", "C", "D", "E", "F", "G", "H", "I")
+
+# Manually setting up area-adjacency using Octavio's code
+#                     A   B   C   D   E   F   G   H   I
+adjacency <- matrix(c(1,	1,	1,	1,	1,	1,	1,	1,	1,  # A
+                      1,	1,	1,	1,	1,	1,	0,	0,	0,  # B
[…]

### 2022-09-30 code/03_BioGeoBEARS_analyses_parallel_st.R

@@ -1,0 +1,1021 @@
+# 03_anilios_BioGeoBEARS_parallel_st.R
+# August 2022
+# THIS one runs with tree using skewT and skew normal priors only
+
+# Script to fit different biogeography models with BioGeoBEARS and save results.
+# base code from BioGeoBEARS website, Octavio Jimenez Robles and Carlos Pavon Vazquez
+
+# Notes:
+# +J is for jump dispersal; + X is to account for distance between biome
+# Models: DEC, BAYAREALIKE, DIVALIKE
+# Fit models, models+j, models+x, models+j+x
+
+# Libraries ---------------------------------------------------------------
+
+library(GenSA)    # GenSA is better than optimx (although somewhat slower)
+library(FD)       # for FD::maxent() (make sure this is up-to-date)
+library(parallel); library(snow)
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+
+# Check tree
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_st.tre')
+plot(tree.c)
+
+# Load inputs -------------------------------------------------------------
+
+# Tree file
+trfn <- "data/intermediate_data/bears/blindsnake_st.tre"
+# Distance matrix file 
+dist.mat <- "data/bears_txt/biome_distance.txt"
+# Geo file
+geogfn <- "data/bears_txt/geofile.txt"
+# adj <- "data/bears_txt/area_adj.txt"
+
+
+
+# Specifications ----------------------------------------------------------
+
+# Set the maximum number of areas any species may occupy; this cannot be larger 
+# than the number of areas you set up, but it can be smaller.
+max_range_size = 5
+
+# Look at your geographic range data:
+tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
+tipranges
+
+# Maximum range size observed:
+max(rowSums(dfnums_to_numeric(tipranges@df)))
+
+coln <- c("A", "B", "C", "D", "E", "F", "G", "H", "I")
+
+# Manually setting up area-adjacency using Octavio's code
+#                     A   B   C   D   E   F   G   H   I
+adjacency <- matrix(c(1,	1,	1,	1,	1,	1,	1,	1,	1,  # A
[…]

### 2022-09-30 code/04_BioGeoBEARS_results_plots.R
plots only no code for tables

@@ -1,133 +1,369 @@
 # 04_BioGeoBEARS_results_plots.R
 # August 2022
+# Code adapted from Carlos Pavon Vazquez
 
 # Test best model fit from BioGeoBEARS analyses
 
 # load results from analysis script
-# source("code/03_BioGeoBEARS_analyses_parallel.R")
+# source("code/03_BioGeoBEARS_analyses_parallel.R") or load from saved intermediate data same thing.
 
 
 # Libraries ---------------------------------------------------------------
 
 library(GenSA);library(FD)      
 library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
-# library(parallel); library(snow)
+# library(dplyr)
+library(parallel); library(snow)
 
 
 # Tree --------------------------------------------------------------------
 
 tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
-
+tree.c$edge.length <- tree.c$edge.length * 100
 
 # Saved results -----------------------------------------------------------
-## These saved results are from source('code/03_BioGeoBEARS_analyses_parallel.R)
 
-bio_geo_bears_files_list <- list.files(path = "./data/intermediate_data/bears/",
-                                       pattern = "*.Rdata", full.names = T)
-
-# Load individual saved files
-lapply(bio_geo_bears_files_list, load, .GlobalEnv)
+load("data/intermediate_data/bears/bears_model_fit.Rdata")
 
 
-# Results table -----------------------------------------------------------
-# Extract parameters and make results table to compare AICc scores
+#######################################################
+# Plot ancestral states - DEC+j+x
+#######################################################
+resjx$inputs$states_list
 
-restable = NULL
-teststable = NULL
+colour_lists <- data.frame(biome = c("A", "B", "C", "D", "E", "F", "G", "H", "I"),
+                           colours = c("#dcc674", #0
+                                       "#1f4733", #1
+                                       "#e03f28", #2
+                                       "#4f937f", #3
+                                       "#179d48", #4 
+                                       "#5265ae", #5
+                                       "#d2e6ae", #6
+                                       "#92d1bd", #7
[…]

### 2022-09-30 code/04_BioGeoBEARS_results_tables.R
code to get tables AICc fit

@@ -1,0 +1,205 @@
+# 04_BioGeoBEARS_results_table.R
+# Sept 2022
+
+# Format best table and likelihood ratio tests
+# load results from analysis script
+# source("code/03_BioGeoBEARS_analyses_parallel.R") or load from saved intermediate data same thing.
+
+# Libraries ---------------------------------------------------------------
+
+library(GenSA);library(FD)      
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+library(dplyr)
+# library(parallel); library(snow)
+
+# Tree --------------------------------------------------------------------
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
+tree.c$edge.length <- tree.c$edge.length * 100
+
+# Saved results -----------------------------------------------------------
+load("data/intermediate_data/bears/bears_model_fit.Rdata")
+
+# Results table -----------------------------------------------------------
+# Extract parameters and make results table to compare AICc scores
+
+restable = NULL
+teststable = NULL
+
+fin.dec = extract_params_from_BioGeoBEARS_results_object(results_object=resDEC, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decj = extract_params_from_BioGeoBEARS_results_object(results_object=resDECj, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decx = extract_params_from_BioGeoBEARS_results_object(results_object=resDECx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.decjx = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.diva = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKE, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.divaj = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEj, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.divax = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
+fin.divajx = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjx, returnwhat="table", addl_params=c("j","x"), paramsstr_digits=4)
[…]

### 2022-09-30 code/utility/00_bears_tree_data_prep.R
prepare data for BioGeoBEARS analysis

@@ -1,0 +1,76 @@
+# Utility/tree_data_match.R
+# 2022-06-14
+
+# For BioGeoBEARS analysis, we need a species-level time-calibrated phylogeny. 
+# Will only include Australian species of Anilios as we don't have data on A. erycinus 
+
+### NOTE ####
+# DEC, DEC+J, and most methods in use in historical biogeography implicitly assume that the observed tree is the complete tree, 
+# and that lineage-level speciation and extinction are independent of the range evolution process*
+
+# The purpose of this script is to prepare the data required for BioGeoBEARS.
+# --- Tree data -- SqCL exon-capture tree (two versions depends on prior)
+# --- Anilios and close outgroups are considered
+
+# libraries ---------------------------------------------------------------
+
+library(dplyr); library(phytools); library(ape); library(stringr)
+
+# phylogeny ---------------------------------------------------------------
+# Time calibrated tree from SqCL probe and mcmctree analysis with fossils
+
+# Full tree with fossils
+phy_full <- ape::read.nexus("data/tree/anilios_v3_st.tre")
+phy_full_b <- ape::read.nexus("data/tree/anilios_v3_b.tre")
+
+# Rename cfwaitii
+phy_full$tip.label[which(phy_full$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_R51244"
+phy_full_b$tip.label[which(phy_full_b$tip.label == "Ramphotyphlops_cfwaitii_R51244")] <- "Anilios_centralis_R51244"
+
+# Rename polygrammicus to Sundatyphlops
+phy_full$tip.label[which(phy_full$tip.label == "Anilios_polygrammicus_R98715")] <- "Sundatyphlops_polygrammicus_R98715"
+phy_full_b$tip.label[which(phy_full_b$tip.label == "Anilios_polygrammicus_R98715")] <- "Sundatyphlops_polygrammicus_R98715"
+
+
+# Subset species tree to Acutotyphlops & Anilios ------------------------------------------
+
+# Additional species to drop
+# A. splendidus, we decided to synonymise this species with A. pinguis
+# A. centralis R51244 is redundant
+# only need one unguirostris at the moment. Will only keep 180002
[…]

### 2022-09-30 code/utility/access_server_R.sh
script to access NICK server

@@ -1,0 +1,4 @@
+#!/bin/bash
+sshpass -f "/home/st_wsl/kak.txt" ssh putter@nick.rsb.anu.edu.au 'cd blindsnakebiogeography; R --vanilla'
+
+# then run source'code/....' in the R console

### 2022-09-30 code/utility/designate_bioregions.R
update scripts

@@ -45,6 +45,7 @@ ggplot(data = top_3_ecoregions, aes(fill = ECO_NAME, y = prop, x = species)) +
 
 sp_biomes <- wwf_biome_records %>% 
   dplyr::group_by(species, BIOME) %>% 
+  dplyr::filter(species != "Anilios splendidus") %>% 
   dplyr::summarise(n = n()) %>% 
   dplyr::mutate(prop = n / sum(n)) %>% 
   dplyr::arrange(desc(prop)) %>% 
@@ -54,9 +55,11 @@ sp_biomes <- wwf_biome_records %>%
 
 # Pivot long for BioGeoBears format
 output_sp_biomes <- sp_biomes %>% 
-  select(-BIOME, -prop, -full_name) %>% 
+  dplyr::select(-BIOME, -prop, -full_name) %>% 
   tidyr::pivot_wider(names_from = short_name, values_from = n, values_fill = 0) %>% 
   janitor::clean_names() %>% 
   dplyr::mutate(species = stringr::str_replace(species, pattern = ' ', replacement = '_'))
+
+# Writing this out helps decide where to designate the species.
 
 write.csv(output_sp_biomes, file = 'data/intermediate_data/geo_file_precursor.csv', quote = FALSE, row.names = FALSE)

### 2022-09-30 code/utility/get_min_distrance.R

@@ -1,27 +1,50 @@
 library(raster)
 library(sp)
-library(rgeos)
+library(sf)
 library(maps)
 library(mapproj)
 
-#Each realm's shape was obtained by intersecting the realm shapes with the shape containing all varanid ranges.
-#That way realms are not artificially close.
+# Each .shp file was extracted using QGIS 
+# CRS == EPSG:4326 - WGS 84 - Geographic
+# Australia is from WWF Biomes
 
-afr <- shapefile("afrotropical.shx")
-aus <- shapefile("australian.shx")
-pap <- shapefile("papuan.shx")
-ori <- shapefile("oriental.shx")
-pal <- shapefile("palearctic.shx")
-sah <- shapefile("saharo_arabian.shx")
-sin <- shapefile("sino_japanese.shx")
-phi <- shapefile("philippines.shx")
-mel <- shapefile("melanesian.shx")
+# From Natural Earth -- Admin 0-Countries
+# PNG (ex Bismarck) is manually cut out in QGIS
+# Bismarck + Solomon Island is manually cut out in QGIS
+# Lesser Sunda manually cut out from Carlos
 
+reg.aus <- shapefile("data/bioregions_map/map/aus_biomes.shx")
+reg.sun <- shapefile("data/bioregions_map/map/LesserSundas.shx")
+reg.bis <- shapefile("data/bioregions_map/map/Bismarck.shx")
+reg.spg <- shapefile("data/bioregions_map/map/Papua.shx")
+
+shp.list <- list(subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[1]), # Tropical Grassland (Savannah)
+                 subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[9]), # Temperate Forest
+                 subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[2]), # Arid
+                 subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[5]), # Mediterranean
+                 subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[3]), # Tropical Forest
+                 subset(reg.aus,reg.aus$BIOME==unique(reg.aus$BIOME)[8]), # Temperate Grasslands
+                 reg.sun, # Lesser Sunda
[…]

### 2022-09-30 code/utility/tree_data_match.R

@@ -56,7 +56,9 @@ blindsnakes_samples <- c('Anilios_affinis_J93635','Anilios_ammodytes_R158097','A
                          'Anilios_nigrescens_R31022','Anilios_obtusifrons_R146400','Anilios_pilbarensis_R108813',
                          'Anilios_pinguis_R146995',
                          'Anilios_proximus_R132486', 
-                         'Anilios_silvia_J46128','Anilios_splendidus_R119900','Anilios_systenos_R114894', 
+                         'Anilios_silvia_J46128',
+                         # 'Anilios_splendidus_R119900',
+                         'Anilios_systenos_R114894', 
                          'Anilios_torresianus_J47640','Anilios_tovelli_R21063','Anilios_troglodytes_R146048', 
                          'Anilios_unguirostris_R21669','Anilios_waitii_R113302','Anilios_wiedii_J59020', 
                          # 'Anilios_ligatus_R19109', # eastern
@@ -86,16 +88,12 @@ fos_tree_full_b$tip.label <- stringr::str_replace(string = fos_tree_full_b$tip.l
 fos_tree_full$tip.label <- stringr::str_replace(string = fos_tree_full$tip.label, pattern = regex("([A-z][0-9]+)"), replacement = "")
 fos_tree_full_b$tip.label <- stringr::str_replace(string = fos_tree_full_b$tip.label, pattern = regex("_([A-z][0-9]+)"), replacement = "")
 
-
-
-compare.chronograms(fos_tree_full, fos_tree_full_b)
-
-
+# compare.chronograms(fos_tree, fos_tree_b)
 
 # write as newick for BioGeoBears -----------------------------------------
 
-# ape::write.tree(fos_tree, file = 'data/tree/anilios_newick_st.tre')
-# ape::write.tree(fos_tree_b, file = 'data/tree/anilios_newick_b.tre')
+ape::write.tree(fos_tree, file = 'data/intermediate_data/bears/anilios_newick_st.tre')
+ape::write.tree(fos_tree_b, file = 'data/intermediate_data/bears/anilios_newick_b.tre')
 
 # Script to plot the tree is in 'tree_compare_chronogram_plot.R'
 

### 2022-10-03 code/02_table_fit_env_results_b.R
init script to organise values from RPANDA model fit using bounded (cauchy) time-calibrated tree

@@ -1,0 +1,197 @@
+# 02_table_fit_env_results_b.R
+# September 2022
+
+# This script will organise the results from the parallel fit and the individual fit together in one table from fitting on B tree (with cauchy priors).
+
+# From 01_fit_*.R. I successfully fitted several environment-dependent birth-death models.
+# The following had to be fitted individually using initial lamb_par 0.1 and/or 0. 
+# min_sco ;arid_sco; mean_str; min_str;  arid_str;  min_val; max_val; arid_val. 
+# mean_val; # max_str; mean_sco; max_sco; global temp
+
+# libraries ---------------------------------------------------------------
+
+library(RPANDA); library(dplyr); library(phytools);
+library(stringr)
+
+# Load results ------------------------------------------------------------
+
+# Get full path to load data the txt files with the results
+file_n <- list.files(path = "data/intermediate_data/diversification_analyses/", pattern = ".txt", full.names = T)
+
+# get the names of the model
+file_n_m <- stringr::str_remove(list.files(path = "data/intermediate_data/diversification_analyses/", pattern = ".txt"), pattern = "Results_Anilios_")
+mod_names <- stringr::str_remove(file_n_m, pattern = ".txt")
+
+# List of models with their names
+filelist <- lapply(file_n, read.delim)
+names(filelist) <- mod_names
+
+# Environmental data in list form
+load('data/intermediate_data/diversification_analyses/env_data_list.RData')
+
+# Organise table
+
+# Put environment information in the table
+for(i in 1:length(filelist)){
+  filelist[[i]]$env <- names(filelist)[i]
+}
+
+
+
+
+# plot_fit_env the models to check 
+
[…]

### 2022-10-04 code/01_fit_global.R
changed initial lamb par to 0 or 0.1 or 0.01 to test if improve model fit. These models spit out very large lamb par in previous models

@@ -37,7 +37,6 @@ cond="crown"
 
 env_data <- env_data_list$global_temp
 
-
 #################
 ## Fit the models ##
 #################
@@ -100,7 +99,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
-  lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2])
+  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2]) # difference 
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01)
   cst.lamb=F; cst.mu=T; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -112,7 +112,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BCST DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
-  lamb_par<-c(treei_BCSTDCST$lamb_par[1])
+  # lamb_par<-c(treei_BCSTDCST$lamb_par[1])
+  lamb_par<-c(0.1)
   mu_par<-c(0.01,0)
   cst.lamb=T; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -124,7 +125,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
-  lamb_par<-c(abs(treei_BEnvVarDCST_EXPO$lamb_par[1]),treei_BEnvVarDCST_EXPO$lamb_par[2])
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01,0)
   cst.lamb=F; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
   
[…]

### 2022-10-04 code/01_fit_max_sco.R

@@ -36,6 +36,7 @@ f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
 env_data <- env_data_list$max_sco
+
 #################
 ## Fit the models ##
 #################
@@ -98,6 +99,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
+  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2]) # difference 
   lamb_par<-c(0.1,0)
   mu_par<-c(0.01)
   cst.lamb=F; cst.mu=T; expo.lamb=F; expo.mu=F; fix.mu=F
@@ -110,6 +112,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BCST DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
+  # lamb_par<-c(treei_BCSTDCST$lamb_par[1])
   lamb_par<-c(0.1)
   mu_par<-c(0.01,0)
   cst.lamb=T; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
@@ -122,7 +125,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
-  lamb_par<-c(abs(treei_BEnvVarDCST_EXPO$lamb_par[1]),treei_BEnvVarDCST_EXPO$lamb_par[2])
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01,0)
   cst.lamb=F; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -140,6 +143,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar LIN")
[…]

### 2022-10-04 code/01_fit_max_str.R

@@ -39,6 +39,7 @@ f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
 env_data <- env_data_list$max_str
+
 #################
 ## Fit the models ##
 #################
@@ -101,7 +102,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
-  lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2])
+  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2]) # difference 
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01)
   cst.lamb=F; cst.mu=T; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -113,7 +115,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BCST DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
-  lamb_par<-c(treei_BCSTDCST$lamb_par[1])
+  # lamb_par<-c(treei_BCSTDCST$lamb_par[1])
+  lamb_par<-c(0.1)
   mu_par<-c(0.01,0)
   cst.lamb=T; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -125,7 +128,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
-  lamb_par<-c(abs(treei_BEnvVarDCST_EXPO$lamb_par[1]),treei_BEnvVarDCST_EXPO$lamb_par[2])
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01,0)
[…]

### 2022-10-04 code/01_fit_mean_sco.R

@@ -1,12 +1,11 @@
 # Individual_env_fit.R
 
-library(picante)
-library(pspline)
-library(RPANDA)
-library(dplyr)
+## SAME Initial lamb par fixed at 0 or 0.1. Fitting other values cause extreme values.
+
+library(picante); library(pspline)
+library(RPANDA); library(dplyr)
 library(phytools); library(gridExtra)
 library(parallel) # can do this on WSL2 (if using Windows)
-
 
 # Load --------------------------------------------------------------------
 
@@ -61,7 +60,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 {
   print(i)
   phylo<-Anilios # If a posterior trees distribution is used, then it would be "posteriors[[i]]"
-
+  
   # BCST DCST (constant Birth-death)
   print("BCST DCST")
   f.lamb<-function(x,y){y[1]}
@@ -309,4 +308,3 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
 write.table(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.txt", quote=FALSE,sep="\t",row.names=FALSE)
 save(final_Anilios,file="data/intermediate_data/diversification_analyses/Results_Anilios_mean_sco_b.Rdata")
 save(Anilios_res,file="data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco_b.Rdata")
-

### 2022-10-04 code/01_fit_mean_val.R

@@ -36,6 +36,7 @@ f<-Ntip(phylo)/50 # As I found in Reptile Database
 cond="crown"
 
 env_data <- env_data_list$mean_val
+
 #################
 ## Fit the models ##
 #################
@@ -98,7 +99,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar DCST EXPO")
   f.lamb<-function(t,x,y){y[1]*exp(y[2]*x)}
   f.mu<-function(t,x,y){y[1]}
-  lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2])
+  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),treei_BEnvVar_EXPO$lamb_par[2]) # difference 
+  lamb_par<-c(0.1,0)
   mu_par<-c(0.01)
   cst.lamb=F; cst.mu=T; expo.lamb=F; expo.mu=F; fix.mu=F
   
@@ -110,6 +112,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BCST DEnvVar EXPO")
   f.lamb<-function(t,x,y){y[1]}
   f.mu<-function(t,x,y){y[1]*exp(y[2]*x)}
+  # lamb_par<-c(treei_BCSTDCST$lamb_par[1])
   lamb_par<-c(0.1)
   mu_par<-c(0.01,0)
   cst.lamb=T; cst.mu=F; expo.lamb=F; expo.mu=F; fix.mu=F
@@ -140,7 +143,8 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
   print("BEnvVar LIN")
   f.lamb<-function(t,x,y){y[1]+y[2]*x}
   f.mu<-function(t,x,y){0}
-  lamb_par<-c(0.1,0) # change initial lamb par
+  # lamb_par<-c(abs(treei_BEnvVar_EXPO$lamb_par[1]),0)
+  lamb_par<-c(0.1,0)
   mu_par<-c()
   cst.lamb=F; cst.mu=T; expo.lamb=F; expo.mu=F; fix.mu=T
   
@@ -152,6 +156,7 @@ for (i in 1: nb.trees) # This can be useful to run over a posterior trees distri
[…]

### 2022-10-04 code/01_env_data_plots.R
init script to plot env data

@@ -1,0 +1,58 @@
+# name: 01_env_data_plots.R
+
+# libraries ---------------------------------------------------------------
+
+library(deeptime) # for deeptime axis
+library(ggplot2); library(viridis)
+
+# load data
+load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
+
+# paleoclimate plots ------------------------------------------------------
+
+str(env_data_list)
+
+# Combine all data into one dataframe to plot using GGPLOT2. Need to make dataframe
+# Make new column for each data set to have name of data as a column
+
+for(i in 1:length(env_data_list)){
+  env_data_list[[i]]$source <- names(env_data_list)[i]
+}
+
+# New column names so we can rbind them AGE, VAL(UES), SOURCE
+env_data_list <- lapply(env_data_list, setNames, nm = c("age", "vals", "source"))
+
+# Merge as df
+env_df <- do.call(rbind, env_data_list)
+rownames(env_df) <- NULL
+
+env_df
+
+# Function to plot
+env_plottr <- function(.df){
+  .df %>% 
+  ggplot(data = ., aes(x = age, colour = source)) +
+    geom_line(aes(y = vals), size = 1) +
+    deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
+    # scale_colour_manual(values = viridis(4), name = "Data source", labels = c("RPANDA Global", "Scotese Australia", "Straume Australia", "Valdes Australia")) +
+    scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Mean temperature (°C)") +
+    scale_x_reverse("Age (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() 
+}
+
+
+# pdf(file = "output/paleo_env_data_plots.pdf", height = 10, width = 10)
+
+env_df[which(env_df$source %in% c("mean_sco", "max_sco", "min_sco", "global_temp")),] %>% env_plottr()
+
+env_df[which(env_df$source %in% c("mean_str", "max_str", "min_str")),] %>% env_plottr()
+
+env_df[which(env_df$source %in% c("mean_val", "max_val", "min_val")),] %>% env_plottr()
+
+env_df[which(env_df$source %in% c("arid_sco", "arid_str", "arid_val")),] %>% 
[…]

### 2022-10-04 manuscript.Rmd
update new subheading

@@
 <!-- Consult Alex Skeels -->
-Two paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from Scotese (2021) [@scoteseAtlas2021] and Straume et al (2020) [@straumeGlobal2020].
-We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in Hagen et al (2021) @hagenGen3sis2021.
+Two paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021 and @straumeGlobal2020.
+We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
 From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day (\@ref(fig:fig-aus-region).
@@
 - Greatest number of species can be found in ...
-
 - Range expansion was important for diversification ...
-
 - Topography dyanmics [@brazModelling2021]
@@
 - Solomon Island 40 MYA [@hallCenozoic2002]
+
+## Arid vs. non-arid speciation
 

### 2022-10-04 code/02_table_fit_env_results.R
code to compile fit env results together.

@@ -1,0 +1,169 @@
+#02_table_fit_env_results.R
+# September 2022
+
+# This script will organise the results from the parallel fit and the individual fit together in one table
+
+# From 01_diversification_analyses_parallel.R I successfully fitted several environment-dependent birth-death models to
+# min_sco ;arid_sco; mean_str; min_str;  arid_str;  min_val; max_val; arid_val. 
+# The following had to be fitted individually using initial lamb_par 0.1 and/or 0. 
+# mean_val; # max_str; mean_sco; max_sco; global temp
+
+
+# libraries ---------------------------------------------------------------
+
+library(RPANDA); library(dplyr); library(phytools)
+
+# Load results ------------------------------------------------------------
+# from parallel fit and individual fit
+load('output/20220823_multiRPANDA_fit.Rdata')
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_global_temp.Rdata'); global_temp_res <- Anilios_res
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_sco.Rdata'); max_sco_res <- Anilios_res
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_str.Rdata'); max_str_res <- Anilios_res
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_mean_sco.Rdata'); mean_sco_res <- Anilios_res
+
+# Data --------------------------------------------------------------------
+load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
+trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
+
[…]

### 2022-10-04 code/02_table_fit_env_results_b.R
combine results together. Waiting for rerun

@@ -16,7 +16,7 @@ library(stringr)
 # Load results ------------------------------------------------------------
 
 # Get full path to load data the txt files with the results
-file_n <- list.files(path = "data/intermediate_data/diversification_analyses/", pattern = ".txt", full.names = T)
+file_n <- list.files(path = "data/intermediate_data/diversification_analyses/", pattern = "_b.txt", full.names = T)
 
 # get the names of the model
 file_n_m <- stringr::str_remove(list.files(path = "data/intermediate_data/diversification_analyses/", pattern = ".txt"), pattern = "Results_Anilios_")
@@ -29,21 +29,31 @@ names(filelist) <- mod_names
 # Environmental data in list form
 load('data/intermediate_data/diversification_analyses/env_data_list.RData')
 
-# Organise table
-
 # Put environment information in the table
 for(i in 1:length(filelist)){
   filelist[[i]]$env <- names(filelist)[i]
 }
 
-
-
-
-# plot_fit_env the models to check 
-
+# Plot_fit_env the models to check 
 RPANDA::plot_fit_env(global_res[[1]]$BEnvVar_EXPO, env_data = env_data_list$global_temp, tot_time = tot_time)
 
-arid_sco_res <- 
+
+
+# Combine tables into one big table ----------------------------------------
+
+res_table <- do.call(rbind, filelist)
+
+options("scipen"=2, "digits"=3)
+res_table[3:4] <- lapply(res_table[3:4], round, 2)
+
+res_table %>% arrange(desc(Lambda))
+
+final_Anilios[, 2:4] <- lapply(final_Anilios[, 2:4], round, 2)
+final_Anilios$AICcWt <- round(final_Anilios$AICcWt, 2)
+final_Anilios[5:8] <- lapply(final_Anilios[5:8], formatC, format = "e", digits = 2)
+
+
+
 
 
 load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_max_str.Rdata'); max_str_res <- Anilios_res

### 2022-10-19 manuscript.Rmd
update after geohisse initial run

@@
 
-To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
 These models will be a combination of pure-birth and birth-death models with different speciation/extinction rates specified as constant, linearly, or exponentially correlated with paleoenvironment.
@@
 
-## Biogeographic-dependent 
-To test whether diversification rates differs between arid and non-arid zones test for biogeographic-depdendent diversification using geographic state and hidden state speciation-extinction model (GeoHiSSE) analysis with different parameters [@goldbergPhylogenetic2011].
+## Biogeographic-dependent diversification analyses
 
-MiSSE https://www.biorxiv.org/content/10.1101/2021.11.02.466977v1
+To test whether diversification rates differs between arid and non-arid zones we used the R package GeoHiSSE [@caetanoHidden2018].
+
+ for biogeographic-depdendent diversification using geographic state and hidden state speciation-extinction model (GeoHiSSE) analysis with different parameters [@goldbergPhylogenetic2011].
+
+## Statistical analyses
+All statistical analyses were run in R version 4.2 (CITE R).
 
@@
 We reconstructed the tropical grassland as the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears). The preferred biogeographic model is the constrained-dispersal matrix, DEC+j+x. 
-Likelihood ratio test (LRT) indicates that founder event (parameter j) and distance matrix (free parameter _x_) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+ *j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+Likelihood ratio test (LRT) indicates that founder event (parameter j) and distance matrix (free parameter _x_) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+*j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 Our result also supports the hypothesis that ancestor of the Australian clade invaded the continent, but exactly where they lived prior to this invasion is equivocal between the Solomon Islands and New Guinea.
@@
 ## Biome transitions and current distributions
-- The earliest biome transition was from tropical grassland into XXX by XX MYA (95% HPD: XX – XX). 
-- Subsequent splits show transitions the arid zone coinciding with with the onset of aridification between 20 – 10 MYA. 
-- Most species are still found in tropical grasslands, but also in arid zone. 
-- Greatest number of species can be found in ...
+Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occured very rapidly.
+Descendents of the earliest split remained in the Tropical Grassland biome but within a million years later another split occured.
+The earliest biome transition was from tropical grassland into Temperate Forest around 25 MYA (95% HPD: 28 – 22 MYA) and the Arid zone 24 MYA (95% HPD: 28 – 21 MYA). 
+Subsequent splits show transitions from the arid zone into the Mediterranean biome starting after 15 MYA. 
+Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the arid zone and Mediterranean biome. 
+Closely related species tend to share the same biome and occur around the same geographical areas.
+
 - Range expansion was important for diversification ...

### 2022-10-19 code/00_anilios_BioGeoBEARS.R

@@ -6,10 +6,8 @@
 
 library(GenSA)    # GenSA is better than optimx (although somewhat slower)
 library(FD)       # for FD::maxent() (make sure this is up-to-date)
-library(parallel)
-library(snow)
-library(rexpokit)
-library(cladoRcpp)
+library(parallel); library(snow)
+library(rexpokit); library(cladoRcpp)
 library(BioGeoBEARS)
 
 # Check tree

### 2022-10-19 code/01_env_data_plots.R

@@ -39,7 +39,6 @@ env_plottr <- function(.df){
     scale_x_reverse("Age (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() 
 }
 
-
 # pdf(file = "output/paleo_env_data_plots.pdf", height = 10, width = 10)
 
 env_df[which(env_df$source %in% c("mean_sco", "max_sco", "min_sco", "global_temp")),] %>% env_plottr()

### 2022-10-19 code/04_BioGeoBEARS_results_bsm.R

@@ -1,0 +1,381 @@
+# 04_BioGeoBEARS_results_bsm.R
+# August 2022
+# Code adapted from Carlos Pavon Vazquez
+
+# Stochastic mapping
+
+# load results from analysis script
+
+
+# Libraries ---------------------------------------------------------------
+
+library(GenSA);library(FD)      
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+# library(dplyr)
+library(parallel); library(snow)
+
+
+# Tree --------------------------------------------------------------------
+
+tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
+tree.c$edge.length <- tree.c$edge.length * 100
+
+# Saved results -----------------------------------------------------------
+
+load("data/intermediate_data/bears/bears_model_fit.Rdata")
+
+
+############################
+#####Stochastic mapping#####
+############################
+
+load("data/intermediate_data/bears/AniliosDEC_JX_full.Rdata")
+resjx
+
+# BSM = Biogeographic Stochastic Mapping for the best fitting model
+
+model_name = "DECjx"
+res = resjx
+
+pdffn = paste0("output/","Anilios_", model_name, "_v1.pdf")
+pdf(pdffn, width=6, height=6)
+
+analysis_titletxt = paste0(model_name, " on Anilios")
+
+# Setup
+results_object = res
+scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))
+
+# States
+res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j", "x"), plotwhat="text", 
+                                label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, 
+                                plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
+
+# Pie chart
+plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j", "x"), plotwhat="pie", label.offset=0.45, 
+                         tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, 
+                         include_null_range=TRUE, tr=tr, tipranges=tipranges)
+
+dev.off()  # Turn off PDF
+cmdstr = paste("open ", pdffn, sep="")
+system(cmdstr) # Plot it
+
+
+
+
+#######################################################
[…]

### 2022-10-19 code/04_BioGeoBEARS_results_plots.R

@@ -76,13 +76,18 @@ tr=tree.c
 
 dev.off()
 # States
-resstates = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45,
+
+# pdf('output/bgb_letter_ancestral.pdf', width = 11.5, height = 8.5)
+resstates = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j","x"), plotwhat="text", label.offset=0.45,
                                      tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, 
                                      include_null_range=TRUE, tr=tr,
                                      colors_list_for_states=colors_list_for_states,
                                      tipranges=tipranges)
+tiplabels(tr$tip.label, adj = c(0,0.5), bg = NULL, col = NULL, frame = "none")
 
 dev.off()
+
+
 
 # pdf('output/pie_ancestral.pdf')
 # Pie chart
@@ -96,325 +101,8 @@ tiplabels(tr$tip.label, adj = c(0,0.5), bg = NULL, col = NULL, frame = "none")
 dev.off()
 
 
-############################
-#####Stochastic mapping#####
-############################
+# Results write up --------------------------------------------------------
 
-load("data/intermediate_data/bears/AniliosDEC_JX_full.Rdata")
-resjx
+plot(tr)
+nodelabels()
 
-# BSM = Biogeographic Stochastic Mapping for the best fitting model
-
-model_name = "DECjx"
-res = resjx
-
-pdffn = paste0("output/","Anilios_", model_name, "_v1.pdf")
-pdf(pdffn, width=6, height=6)
-
-analysis_titletxt = paste0(model_name, " on Anilios")
-
-# Setup
-results_object = res
-scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))
-
-# States
-res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j", "x"), plotwhat="text", 
-                                label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, 
-                                plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)
-
-# Pie chart
[…]

### 2022-10-19 code/04_BioGeoBEARS_results_tables.R

@@ -15,7 +15,7 @@ library(dplyr)
 # Tree --------------------------------------------------------------------
 tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
 tree.c$edge.length <- tree.c$edge.length * 100
-
+tr <- tree.c
 # Saved results -----------------------------------------------------------
 load("data/intermediate_data/bears/bears_model_fit.Rdata")
 
@@ -86,48 +86,54 @@ write.csv(sorted_resttable, file="output/supp_biogeobears_table_weights.csv")
 # Top five
 
 # Is DEC+j+x significantly better than DECj
-lrt.fin_base <- BioGeoBEARS::lrttest(LnL_1= restable$LnL[which(rownames(restable)=="fin.decjx")],
-                                LnL_2= restable$LnL[which(rownames(restable)=="fin.dec")],
-                                numparams1= restable$numparams[which(rownames(restable)=="fin.decjx")],
-                                numparams2= restable$numparams[which(rownames(restable)=="fin.dec")],
+lrt.fin_base <- BioGeoBEARS::lrttest(LnL_1= restable$LnL[which(rownames(restable)=="DEC+J+X")],
+                                LnL_2= restable$LnL[which(rownames(restable)=="DEC")],
+                                numparams1=restable$numparams[which(rownames(restable)=="DEC+J+X")],
+                                numparams2=restable$numparams[which(rownames(restable)=="DEC")],
+                                returnwhat = "all")
+lrt.fin_base$alt <- "DEC+J+X"; lrt.fin_base$null <- "DEC"
+
+
+lrt.fin_j <- BioGeoBEARS::lrttest(LnL_1= restable$LnL[which(rownames(restable)=="DEC+J+X")],
+                                LnL_2= restable$LnL[which(rownames(restable)=="DEC+J")],
+                                numparams1=restable$numparams[which(rownames(restable)=="DEC+J+X")],
+                                numparams2=restable$numparams[which(rownames(restable)=="DEC")],
                                 returnwhat = "all")
 
-lrt.fin_j <- BioGeoBEARS::lrttest(LnL_1= restable$LnL[which(rownames(restable)=="fin.decjx")],
-                                LnL_2= restable$LnL[which(rownames(restable)=="fin.decj")],
-                                numparams1= restable$numparams[which(rownames(restable)=="fin.decjx")],
[…]

### 2022-10-19 code/05_GeoHiSSE_fit.R

@@ -1,0 +1,194 @@
+# 05_GeoHiSSE_fit.R
+# October 2022
+
+# Are rates equal between arid and non-arid zones? 
+# How does it differ? Is it better explained by hidden states? 
+
+# install -----------------------------------------------------------------
+
+# library( devtools )
+# install_github(repo = "thej022214/hisse", ref = "master")
+
+# libraries ---------------------------------------------------------------
+
+library(hisse); library(ape)
+library(diversitree)
+library(ape)
+
+
+# Data --------------------------------------------------------------------
+
+phy <- ape::read.tree("data/tree/anilios_newick_b.tre")
+
+# Plot to check dates
+plot(phy); axisPhylo()
+
+# Check to make sure tree is ultrametric
+is.ultrametric(phy)
+
+
+# Prune -------------------------------------------------------------------
+
+# List of species to drop from analysis. In this case we only want Anilios
+droplist <- c("Acutotyphlops_subocularis", "Sundatyphlops_polygrammicus",
+              "Sundatyphlops_polygrammicus", "Ramphotyphlops_multilineatus",
+              "Anilios_splendidus")
+
+pruned.tree <- ape::drop.tip(phy = phy, tip = droplist)
+
+plot(pruned.tree); ape::axisPhylo()
+
+# Write to file
+write.tree(pruned.tree, file = "data/tree/subset_anilios_newick_b.tre")
+
+# Re read in the subset tree
+phy <- read.tree("data/tree/subset_anilios_newick_b.tre")
+
+taxon_list <- data.frame(taxon = phy$tip.label)
+taxon_list <- sort(taxon_list)
+
+# write.csv(taxon_list, file = "data/intermediate_data/geohisse/taxon_list.csv", row.names = FALSE)
+
+# State data --------------------------------------------------------------
+# State data indicates whether the species is classified as (state 1) arid, (state 2) non-arid (1), (state 0) widespread between both regions.
+# Classification is based on distribution of 'species'. I plotted ALA distribution on biome map of Australia in QGIS. 
+
+# Test whether arid habitation influences rates of diversification 
+# and to reconstruct ancestral states while accounting for diversification rate heterogeneity.
+
[…]

### 2022-10-19 code/utility/00_diversification_env_data_prep.R

@@ -13,6 +13,21 @@
 # libraries ---------------------------------------------------------------
 library(tidyverse); library(RPANDA); library(stringr)
 
+
+# tree --------------------------------------------------------------------
+
+trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
+
+is.ultrametric(trees[[1]])
+fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+fos_tree_b <- phytools::force.ultrametric(trees[[2]],"extend")
+plotTree(fos_tree)
+
+tot_time <- max(node.age(fos_tree)$ages)
+tot_time <- max(node.age(fos_tree_b)$ages)
+
+
+
 # Climate data ------------------------------------------------------------
 
 # Global temp from RPANDA (isotopes)
@@ -38,7 +53,7 @@ climate_data_sco$time_mya[2]; climate_data_sco$time_mya[3]
 
 #  get timesteps from RPANDA's dataset that already match Alex's
 panda_true <- InfTemp %>% 
-  dplyr::filter(Age < 46) %>% 
+  dplyr::filter(Age < 26) %>% 
   dplyr::mutate(choose = ifelse(Age %in% c(climate_data_sco$time_mya), "TRUE", "FALSE")) %>% 
   dplyr::filter(choose == "TRUE") %>% dplyr::distinct(Age, .keep_all = TRUE)
 
@@ -89,5 +104,9 @@ env_data_list <- list(mean_sco = climate_data_sco[, c("time_mya", "t_mean")],
                       # arid_prop_val = climate_data_val[, c("time_mya", "prop_arid")],
                       global_temp = subset_InfTemp
 )
+
+dev.off()
+plot(climate_data_sco$t_max ~ climate_data_sco$time_mya)
+plot(subset_InfTemp$t_mean ~ subset_InfTemp$time_mya)
 
 save(env_data_list, file = "data/intermediate_data/diversification_analyses/env_data_list.RData")

### 2022-10-19 code/utility/batch_fit_diversification.R

@@ -1,0 +1,148 @@
+# batch_fit_diversification.R
+
+# Fit diversification by environmental data rather than full list
+
+# Library -----------------------------------------------------------------
+
+library(RPANDA); library(tidyverse)
+library(phytools); library(gridExtra)
+library(parallel) # can do this on WSL2 (if using Windows)
+
+# Load --------------------------------------------------------------------
+
+load(file = "data/intermediate_data/diversification_analyses/env_data_list.RData")
+trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
+
+is.ultrametric(trees[[1]])
+fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+plotTree(fos_tree)
+is.ultrametric(fos_tree)
+
+# Fit models ---------------------------------------------------------------
+
+# Crown age
+tot_time <- max(node.age(fos_tree)$ages)
+cond = "crown"
+
+## Fraction of diversity represented. In our tree we have 
+## 48 Described species of Anilios + 2 more grypus undescribed + 1 more ligatus - A. splendidus
+## So our "true" species at the moment is 50 and we have 41 tips in our tree
+
+length(fos_tree$tip.label)
+frac = 38/50
+
+# Explanation of the models we'll fit:
+# BCST
+#   + standard Yule model, speciation rate is constant and no extinction
+#	BCST_DCST
+#		+ standard birth-death model, with constant speciation and extinction rates
+#	BEnvVar_EXPO
+#		+ speciation rate is correlated exponentially to the rate of environmental variation, with no extinction (mu=0)
+#	BEnvVarDCST_EXPO
+#		+ speciation rate is correlated exponentially to the rate of environmental variation, with constant extinction
+#	BCSTDEnvVar_EXPO
+#		+ speciation rate is constant, and extinction rate is correlated exponentially to the rate of environmental variation
+#	BEnvVarDEnvVar
[…]

### 2022-10-19 code/utility/sse_practice_script.R

@@ -1,0 +1,156 @@
+# geoHiSSE
+
+# library( devtools )
+# install_github(repo = "thej022214/hisse", ref = "master")
+
+# library(hisse)
+# library(diversitree)
+
+phy <- read.tree("subset_2mill.tree")
+library(ape)
+
+# put the names of tips you want to drop in a vector
+
+droplist<-c(
+  "Chiropodomys_gliroides_Z25153_EC1",
+  "Chiropodomys_karlkoopmani_NHM19782959_EC4",
+  "Chiropodomys_muroides_BMNH290_EC4",
+  "Haeromys_minahassae_Z23000_EC1",
+  "Apomys_lubangensis_JAE169_EXOME",
+  "Archboldomys_luzonensis_EAR1826_EXOME",
+  "Chrotomys_mindorensis_JAE520_EXOME",
+  "Rhynchomys_isarogensis_JAE2195_EXOME")
+
+pruned.tree <- drop.tip(phy,phy$tip.label[match(droplist, phy$tip.label)])
+write.tree(pruned.tree, file="subset_2mill_just_sahul.tree")
+
+
+phy <- read.tree("subset_2mill_just_sahul.tree")
+phy$tip.label
+
+state <- read.csv("shelf_vs_island_binary.txt",header=F)
+state  
+
+state <- data.frame(taxon=state$V1, ranges=as.numeric(state$V2))
+state
+
+
+# Setting up the models
+# We will fit a total of four models. Two models with a range-indendent diversification process and 
+# two other models in which the range have an effect on the diversification rate of the lineages (each with either one or two rate classes).
+
+#turnover - turnover parameter tau00, tau11, tau01
+#f = estimated proportion of extant species in the different state (areas)
+#eps = extinct fraction parameters. 
+
+## Model 1 - Dispersal parameters vary only, no range-dependent diversification.
+#### Assume equal rates regardless of biogeographic region. ###
+turnover <- c(1,1,0)
+eps <- c(1,1)
+trans.rate <- TransMatMakerGeoHiSSE(hidden.traits=0)
+trans.rate.mod <- ParEqual(trans.rate, c(1,2))
+mod1 <- GeoHiSSE(phy = phy, data = sim.dat, f=c(1,1,1),
+                 turnover=turnover, eps=eps,
+                 hidden.states=FALSE, trans.rate=trans.rate.mod, 
+                 turnover.upper=100, trans.upper=10)
+
+## Model 2. Canonical GeoSSE model, range effect on diversification
[…]

### 2022-10-23 code/05_GeoHiSSE_fit.R
update code GeoHiSSE

@@ -13,8 +13,9 @@
 
 library(hisse); library(ape)
 library(diversitree)
-library(ape)
-
+library(ape); library(RColorBrewer)
+library(dplyr); library(tidyr)
+library(ggplot2)
 
 # Data --------------------------------------------------------------------
 
@@ -49,8 +50,9 @@ taxon_list <- sort(taxon_list)
 
 # write.csv(taxon_list, file = "data/intermediate_data/geohisse/taxon_list.csv", row.names = FALSE)
 
+
 # State data --------------------------------------------------------------
-# State data indicates whether the species is classified as (state 1) arid, (state 2) non-arid (1), (state 0) widespread between both regions.
+# State data indicates whether the species is classified as (state 1) arid, (state 2) non-arid, (state 0) widespread between both regions.
 # Classification is based on distribution of 'species'. I plotted ALA distribution on biome map of Australia in QGIS. 
 
 # Test whether arid habitation influences rates of diversification 
@@ -60,7 +62,34 @@ state <- read.csv("data/intermediate_data/geohisse/arid_nonarid_both_states.csv"
 state <- data.frame(taxon=state$taxon, ranges=as.numeric(state$state))
 head(state)
 
+state$new_tip_names <- paste(state$taxon, state$ranges, sep = "_")
 
+# Rename tips
+
+state_temp <- data.frame(tip.labels = phy$tip.label)
+
+new_tip_names <- state_temp %>% 
+  dplyr::left_join(state, by = c("tip.labels" = "taxon")) %>% 
+  select(new_tip_names)
+
+phy$tip.label <- new_tip_names$new_tip_names
+plot(phy)
+
+state <- data.frame(taxon = state$new_tip_names, ranges = state$ranges)
+head(state)
+
+## Proportion of species for input in `f` argument, estimated proportion of extant species in states
+prop_df <- read.csv("data/intermediate_data/geohisse/2022_species_list_arid_nonarid_widespread.csv", header = T)
+
+prop_df$state <- as.character(prop_df$state)
+
+proportion_f <- prop_df %>% 
+  dplyr::group_by(state, in_tree) %>% 
+  dplyr::summarise(n = n()) %>% 
+  dplyr::mutate(freq = n / sum(n)) %>% 
+  dplyr::filter(in_tree == "yes")
+  
+proportion_f
 
 # Model fitting -----------------------------------------------------------
 
[…]

### 2022-10-24 manuscript.Rmd
update script and write about geohisse

@@
 
+- The emergence aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
+- XX species today occupy arid zones exclusively and part of the range of YY species includes the arid zone as defined by WWF Bioregions. 
+To evaluate whether diversification rates are associated with geographic ranges, we used a series of geographic range-dependent speciation and extinction models including GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] implemented in the R package `hisse` v.
+2.1.9 [@beaulieuDetecting2016].
+- Both GeoSSE and GeoHiSSE allow only two geographic states, as such we categorised species that primarily occur in the arid zone as arid, and species that occur primarily outside of the arid zone as non-arid.
+
+Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
+Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
+Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
+Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
+Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.
+
 \newpage

### 2022-10-24 code/05_GeoHiSSE_fit.R

@@ -228,6 +228,7 @@ recon.models <- list(recon.mod1, recon.mod2, recon.mod3, recon.mod4)
 model.ave.rates <- hisse::GetModelAveRates(recon.models, type = "both")
 head( model.ave.rates )
 
+model.ave.rates$tips
 
 # add new column for range and change species name
 rates_df <- model.ave.rates$tips %>% 
@@ -269,23 +270,28 @@ ratecolours <- colorRampPalette(brewer.pal(6, 'RdYlBu'))(100)
 #                      # state.colors = statecolours,
 #                      fsize = 0.8)
 
-plot.geohisse.states(x = recon.models, rate.param = "net.div", type = "fan",
-                     show.tip.label = F, legend = FALSE,
+
+head( model.ave.rates$nodes )
+
+plot.geohisse.states(x = recon.models, rate.param = "net.div", type = "phylogram",
+                     show.tip.label = T, legend = FALSE,
                      # rate.colors = ratecolours,
                      state.colors = statecolours,
                      fsize = 0.8)
 
+plot(phy)
+nodelabels()
 
-# Net diversification based on range
-rates_df %>% 
-  ggplot(., aes(x = range, y = net.div)) +
-  geom_boxplot() +
-  theme_classic()
-
-rates_df %>% 
-  ggplot(., aes(x = range, y = speciation)) +
-  geom_boxplot() +
-  theme_classic()
+# # Net diversification based on range
+# rates_df %>% 
+#   ggplot(., aes(x = range, y = net.div)) +
+#   geom_boxplot() +
+#   theme_classic()
+# 
+# rates_df %>% 
+#   ggplot(., aes(x = range, y = speciation)) +
+#   geom_boxplot() +
+#   theme_classic()
 
 
 # AIC ---------------------------------------------------------------------

### 2022-10-31 manuscript.Rmd
update

@@
 # Introduction
-<!-- COMMUNITY words - Reported, accepted, widely  -->
-<!-- TRANSITION / FLOW words - Nonetheless, however, although, unless, because, and, but/ -->
-<!-- 1. Why phylogeny is important to readers of the target journals? -->
+
+<!-- Aridification  -->
 
@@
 The arid zone makes up the majority of the Australian landscape and provides habitat to ~43% of all Australia’s terrestrial reptiles.
-Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which have undergone rapid diversification in the last 10 – 6 myr [@brennanMass2017; @oliverPhylogenetic2017].
-For example, some Australian squamate clades like dragons [@melvilleIntegrating2019], sphenomorphine skinks [@raboskyAnalysis2014], geckos [@ashmanDiversification2018; @brennanMass2017] and elapid snakes [@leeDiversification2016; @grundlerTrophic2014] show increased rates of diversification coincidentally with the onset of aridification.
+Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which have undergone rapid diversification in the last 10 – 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
+For example, some Australian squamate clades like dragons [@melvilleIntegrating2019], sphenomorphine skinks [@raboskyAnalysis2014], geckos [@ashmanDiversification2018;@brennanMass2017] and elapid snakes [@leeDiversification2016;@grundlerTrophic2014] show increased rates of diversification coincidentally with the onset of aridification.
 However, these studies did not account for heterogeneity of the arid zone [@pepperLife2021].
 
-Roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises 48 described species [@uetzQuarter2021].
+Roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021].
 All _Anilios_ species are found in Australia except for _A.
 eryncinus_, which is only found in Papua New Guinea.
-While morphologically conserved and difficult to distinguished, extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013; @pyronSystematics2014; @ellisTyphlopid2019].
-The diversification is reported to have originated from a single colonisation event from Asia between 19-39 mya [@vidalBlindsnake2010; @marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
+While morphologically conserved and difficult to distinguished, extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@ellisTyphlopid2019].
+The diversification is reported to have originated from a single colonisation event by a non-arid adapted ancestor from Asia between 19-39 mya [@vidalBlindsnake2010;@marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
 However, the study did not account for alternative biogeographic areas (e.g.
 WWF Ecoregions) and models, or uncertainty in phylogenetic relationships and divergence times.
-Despite representing a substantial portion of snake biodiversity there has been few investigations into the macroevolutionary dynamics of blind snakes.
-Recent phylogenomic tree and monophyly of the genus makes the Australian _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in almost every biome [citefirstchapter?].
+Despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
+The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in almost every biome.
 
@@
 Among others, studies on the influence of climate (e.g.
-temperature, precipitation, and latitude) [@ricklefsGlobal2006; @mayhewBiodiversity2012; @condamineMacroevolutionary2013; @cantalapiedraDietary2014; @condamineClimate2019; @rollandContribution2019; @sunRecent2020]], elevation [@esquerreHow2019; @boschmanMountain2021], biological interactions [@ramirezAsynchronous2011; @druryEstimating2016; @brennanPhylogenomicsMonitorLizards2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007; @pepperGenetic2011; @deepakAridification2018; @catulloAridification2014] have been documented.
-In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010; @pepperGenetic2011; @pepperPalaeoclimate2011; @oliverIndependent2014; @oliverContrasting2014], frogs [@catulloAridification2014; @vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017; @toussaintUnveiling2015], and other plants [reviewed in @crispHow2013].
+temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@cantalapiedraDietary2014;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2021], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomicsMonitorLizards2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
+In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and other plants (reviewed in @crispHow2013).
 
@@
 First, warm temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
-Second, spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011; @lawsonLatitudinal2014].
-Third, change in climate change (e.g.
-glaciation) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
+Second, spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
+Third, change in climate (e.g.
+glaciation and aridification) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
 Fourth, ectotherms are subjected to higher metabolic rates in warm climates (i.e.
-Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006; @makarievaMean2008].
-Finally, aridification of Australia may offer new ecological niche for species to exploit, which have been shown to result in deep divergences in many arid-adapted [@byrneBirth2008; @pepperGenetic2011; @marinTracing2013].
+Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
+Finally, aridification of Australia may offer new ecological niche for species to exploit, which have been shown to result in deep divergences in many arid-adapted [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013].
 
-The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013; @kergoatOpposite2018; @condamineClimate2019; @rollandContribution2019], these temperature data are reconstructed (interpolated) based on conversion of concentration of oxygen isotopes [@condamineMacroevolutionary2013].
-However, the relationship between diversification rates and aridity are attributed to some proxy such as the the timing of aridification or global temperature rather than quantifying the extent of aridity itself [@dornStrange2014; @deepakAridification2018; @catulloAridification2014; @fujitaDiversification2010; @pepperGenetic2011; @pepperPalaeoclimate2011; @oliverIndependent2014; @oliverContrasting2014; @owenHow2017; @crispHow2013; @toussaintUnveiling2015].
+The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019], these temperature data are reconstructed (interpolated) based on conversion of concentration of oxygen isotopes [@condamineMacroevolutionary2013].
+However, the relationship between diversification rates and aridity are attributed to some proxy such as the the timing of aridification or global temperature rather than quantifying the extent of aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
 As such, no study has been able to quantitatively assess aridity as a continuous variable in environment-dependent models and compare it to other ecological processes.
 Recent development in geographic-based methods such as `gen3sis` modeling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
-In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia (instead of relying on global paleoclimate data) in the time span of interest (between the 45 mya and the present).
+In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia in the time span of interest, between the 45 MYA to present.
 These datasets will allow us to empirically test the influence of localised ecological processes on diversification rates over time.
 
-In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates in arid and non-arid zones by testing hypotheses about the processes that promote speciation.
-Applying the cutting edge localised geographic-based paleoenvironment reconstruction,
+In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
 
-We (1) evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, (2) compared the model fit between different ecological processes on diversification rates, and (3) reconstructed the ancestral biome of Australian blindsnakes?
+<!--Why people should care-->
+- Applying the cutting edge localised geographic-based paleoenvironment reconstruction,
 
+We 
+(1) evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
+(2) compared the model fit between different ecological processes on diversification rates, and 
+(3) reconstructed the ancestral biome of Australian blindsnakes? 
 (4) Are there differences in the rates of speciation between arid and non-arid zones?
@@
 
-To test whether diversification rates differs between arid and non-arid zones we used the R package GeoHiSSE [@caetanoHidden2018].
+- Aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
+- XX species today occupy arid zones exclusively and part of the range of YY species includes the arid zone as defined by WWF Bioregions. 
+- To evaluate whether broad patterns of diversification are associated with geographic range, we used a series of biogeographic-dependent speciation-extinction models 
+including GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] implemented in the R package `hisse` v. 2.1.9 [@beaulieuDetecting2016].
+
+- GeoHiSSE differs from GeoSSE by incorporating hidden states, which accounts for heteregenous diversification caused by unsampled traits.
+
+- These models only allow three states including two geographic states and one widespread state.
+As such we categorised species that primarily occur in the arid zone as arid, species that occur primarily outside of the arid zone as non-arid, and those that occupy both the arid and non-arid zone as widespread.
+We assumed uniform taxon sampling for 89% in arid areas, 71 in non-arid areas, and 83% in widespread areas.
+-
+
+Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
+Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
+Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
+Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
+Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.
 
@@
 ## Arid vs. non-arid speciation
-
-- The emergence aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
-- XX species today occupy arid zones exclusively and part of the range of YY species includes the arid zone as defined by WWF Bioregions. 
-To evaluate whether diversification rates are associated with geographic ranges, we used a series of geographic range-dependent speciation and extinction models including GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] implemented in the R package `hisse` v.
-2.1.9 [@beaulieuDetecting2016].
-- Both GeoSSE and GeoHiSSE allow only two geographic states, as such we categorised species that primarily occur in the arid zone as arid, and species that occur primarily outside of the arid zone as non-arid.
-
-Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
-Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
-Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
-Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
-Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.
 

### 2022-10-31 code/05_GeoHiSSE_fit.R
update script and add how to add arc clade labels but not working so well

@@ -105,7 +105,7 @@ proportion_f
 #eps = extinct fraction parameters. 
 
 #porportion of extant species probably not fully sampled
-fraction <- c(0.88, # arid
+fraction <- c(0.89, # arid
               0.71, # non-arid
               0.83) # widespread
 
@@ -119,7 +119,8 @@ trans.rate.mod <- ParEqual(trans.rate, c(1,2))
 mod1 <- GeoHiSSE(phy = phy, data = state, f=fraction,
                  turnover=turnover, eps=eps,
                  hidden.states=FALSE, trans.rate=trans.rate.mod, 
-                 turnover.upper=100, trans.upper=10)
+                 turnover.upper=100, trans.upper=10,
+                 assume.cladogenetic = TRUE)
 
 ## Model 2. Canonical GeoSSE model, range effect on diversification
 #### Range evolution affects diversification by adding turnover rate for widespread range. 
@@ -148,6 +149,7 @@ mod3 <- GeoHiSSE(phy = phy, data = state, f=fraction,
                  hidden.states=TRUE, trans.rate=trans.rate.mod,
                  turnover.upper=100, trans.upper=10)
 
+mod3$index.par
 
 ## Model 4. GeoHiSSE model with 1 hidden trait, range-dependent diversification.
 turnover <- c(1,2,3,4,5,6)
@@ -172,7 +174,7 @@ trans.rate.mod <- ParEqual(trans.rate.mod, c(2,3))
 mod5 <- GeoHiSSE(phy = phy, data = state, f=fraction,
                  turnover=turnover, eps=eps,
                  hidden.states=FALSE, trans.rate=trans.rate.mod,
-                 turnover.upper=100, trans.upper=11, # need to change trans.upper to 11 instead of 10 for model to run. 
+                 turnover.upper=100, trans.upper=10, # need to change trans.upper to 11 instead of 10 for model to run. 
                  sann=FALSE, 
                  assume.cladogenetic = FALSE)
 
@@ -186,6 +188,8 @@ mod5 <- GeoHiSSE(phy = phy, data = state, f=fraction,
 
 hisse::GetAICWeights(list(model1 = mod1, model2 = mod2, model3 = mod3,
                           model4 = mod4, model5 = mod5), criterion="AIC")
+
+
 
 ### Model averaging and plotting
 
[…]

### 2022-10-31 code/06_CLaDS_Anilios.R
new script based on E Roycroft's

@@ -1,0 +1,98 @@
+# 06_CLaDS_Anilios.R
+
+# Cladogenetic diversification in rate shifts 
+# Estimates lineage specific diversification rates on a phylogeny. 
+# Accounts for diverse sources of variation in diversification rates that occur
+# during the evolutionary history of clades. 
+
+library(ape)
+library(phytools)
+library(RPANDA)
+# library(diversitree)
+
+full.tree <-read.tree("finalthesis_2mill_run_FigTree_run4_newick.tre")
+full.tree
+
+keep <- read.table("keep_tips_mcmctree_names.txt", header = T, row.names = NULL)
+
+tree_sp <- full.tree$tip.label
+table_sp <- keep[,2]
+table_sp
+matched_tips <- na.omit(match(table_sp,tree_sp))
+
+drop_tips <- tree_sp[-matched_tips]
+
+subset.tree.noout <- drop.tip(full.tree,drop_tips)
+
+#CLADS
+
+clads.out <- fit_ClaDS0(subset.tree.noout, name = NULL, 
+                        pamhLocalName = "pamhLocal_nooutgroup",
+                        iteration = 1e+07, thin = 20000, 
+                        update = 1000,adaptation = 10, seed = 4821, nCPU = 1)
+clads.out
+
+
+rates <- getMAPS_ClaDS0(subset.tree.noout, clads.out, burn=0.5, thin=1)
+rates
+
+plot_ClaDS0_chains(clads.out, burn = 1/2, thin = 1, 
+                   param = c("sigma", "alpha", "l_0", "LP"))
+
+
+plot_ClaDS_phylo(subset.tree.noout, rates[-(1:3)], rates2 = NULL,same.scale = T, main = NULL, lwd = 2, log = T, show.tip.label = F)
+
+plot_ClaDS_phylo(subset.tree, rates[-(1:3)], rates2 = NULL,same.scale = T, main = NULL, lwd = 2, log = T, show.tip.label = F)
+
+
+#plot with magma palette
+source("plot_ClaDS_phylo_colours.R")
+
+plot_ClaDS_phylo_col(subset.tree.noout, rates[-(1:3)], rates2 = NULL,same.scale = T, main = NULL, lwd = 2, log = T, show.tip.label = F)
+axisPhylo() # plots timescale
+
+
+pdffn = "CLaDs_Sahul_magma1.pdf"
+
[…]

### 2022-10-31 code/05_GeoHiSSE_all_arid.R
fit GeoHiSSE with different data

@@ -1,0 +1,307 @@
+# 05_GeoHiSSE_binary.R
+# October 2022
+
+# Are rates equal between arid and non-arid zones? 
+# How does it differ? Is it better explained by hidden states? 
+
+# install -----------------------------------------------------------------
+
+# library( devtools )
+# install_github(repo = "thej022214/hisse", ref = "master")
+
+# libraries ---------------------------------------------------------------
+
+library(hisse); library(ape)
+library(diversitree)
+library(ape); library(RColorBrewer)
+library(dplyr); library(tidyr)
+library(ggplot2)
+
+# Data --------------------------------------------------------------------
+
+phy <- ape::read.tree("data/tree/anilios_newick_b.tre")
+
+# Plot to check dates
+plot(phy); axisPhylo()
+
+# Check to make sure tree is ultrametric
+is.ultrametric(phy)
+
+
+# Prune -------------------------------------------------------------------
+
+# List of species to drop from analysis. In this case we only want Anilios
+droplist <- c("Acutotyphlops_subocularis", "Sundatyphlops_polygrammicus",
+              "Sundatyphlops_polygrammicus", "Ramphotyphlops_multilineatus",
+              "Anilios_splendidus")
+
+pruned.tree <- ape::drop.tip(phy = phy, tip = droplist)
+
+plot(pruned.tree); ape::axisPhylo()
+
+# Write to file
+write.tree(pruned.tree, file = "data/tree/subset_anilios_newick_b.tre")
+
+# Re read in the subset tree
+phy <- read.tree("data/tree/subset_anilios_newick_b.tre")
+
+taxon_list <- data.frame(taxon = phy$tip.label)
+taxon_list <- sort(taxon_list)
+
+# write.csv(taxon_list, file = "data/intermediate_data/geohisse/taxon_list.csv", row.names = FALSE)
+
+
+# State data --------------------------------------------------------------
+
+state <- read.csv("data/intermediate_data/geohisse/arid_nonarid_binary_all_arid.csv", header = T)
+state <- data.frame(taxon=state$taxon, ranges=as.numeric(state$state))
+head(state)
+
+state$new_tip_names <- paste(state$taxon, state$ranges, sep = "_")
+
+# Rename tips
+
+state_temp <- data.frame(tip.labels = phy$tip.label)
+
+new_tip_names <- state_temp %>% 
+  dplyr::left_join(state, by = c("tip.labels" = "taxon")) %>% 
+  select(new_tip_names)
+
[…]

### 2022-10-31 code/05_GeoHiSSE_all_non_arid.R

@@ -1,0 +1,307 @@
+# 05_GeoHiSSE_binary_non-arid.R
+# October 2022
+
+# Are rates equal between arid and non-arid zones? 
+# How does it differ? Is it better explained by hidden states? 
+
+# install -----------------------------------------------------------------
+
+# library( devtools )
+# install_github(repo = "thej022214/hisse", ref = "master")
+
+# libraries ---------------------------------------------------------------
+
+library(hisse); library(ape)
+library(diversitree)
+library(ape); library(RColorBrewer)
+library(dplyr); library(tidyr)
+library(ggplot2)
+
+# Data --------------------------------------------------------------------
+
+phy <- ape::read.tree("data/tree/anilios_newick_b.tre")
+
+# Plot to check dates
+plot(phy); axisPhylo()
+
+# Check to make sure tree is ultrametric
+is.ultrametric(phy)
+
+
+# Prune -------------------------------------------------------------------
+
+# List of species to drop from analysis. In this case we only want Anilios
+droplist <- c("Acutotyphlops_subocularis", "Sundatyphlops_polygrammicus",
+              "Sundatyphlops_polygrammicus", "Ramphotyphlops_multilineatus",
+              "Anilios_splendidus")
+
+pruned.tree <- ape::drop.tip(phy = phy, tip = droplist)
+
+plot(pruned.tree); ape::axisPhylo()
+
+# Write to file
+write.tree(pruned.tree, file = "data/tree/subset_anilios_newick_b.tre")
+
+# Re read in the subset tree
+phy <- read.tree("data/tree/subset_anilios_newick_b.tre")
+
+taxon_list <- data.frame(taxon = phy$tip.label)
+taxon_list <- sort(taxon_list)
+
+# write.csv(taxon_list, file = "data/intermediate_data/geohisse/taxon_list.csv", row.names = FALSE)
+
+
+# State data --------------------------------------------------------------
+
+state <- read.csv("data/intermediate_data/geohisse/arid_nonarid_binary_all_non_arid.csv", header = T)
+state <- data.frame(taxon=state$taxon, ranges=as.numeric(state$state))
+head(state)
+
+state$new_tip_names <- paste(state$taxon, state$ranges, sep = "_")
+
+# Rename tips
+
+state_temp <- data.frame(tip.labels = phy$tip.label)
+
+new_tip_names <- state_temp %>% 
+  dplyr::left_join(state, by = c("tip.labels" = "taxon")) %>% 
+  select(new_tip_names)
+
[…]

### 2022-10-31 code/05_GeoHiSSE_binary.R

@@ -1,0 +1,319 @@
+# 05_GeoHiSSE_binary.R
+# October 2022
+
+# Are rates equal between arid and non-arid zones? 
+# How does it differ? Is it better explained by hidden states? 
+
+# install -----------------------------------------------------------------
+
+# library( devtools )
+# install_github(repo = "thej022214/hisse", ref = "master")
+
+# libraries ---------------------------------------------------------------
+
+library(hisse); library(ape)
+library(diversitree)
+library(ape); library(RColorBrewer)
+library(dplyr); library(tidyr)
+library(ggplot2)
+
+# Data --------------------------------------------------------------------
+
+phy <- ape::read.tree("data/tree/anilios_newick_b.tre")
+
+# Plot to check dates
+plot(phy); axisPhylo()
+
+# Check to make sure tree is ultrametric
+is.ultrametric(phy)
+
+
+# Prune -------------------------------------------------------------------
+
+# List of species to drop from analysis. In this case we only want Anilios
+droplist <- c("Acutotyphlops_subocularis", "Sundatyphlops_polygrammicus",
+              "Sundatyphlops_polygrammicus", "Ramphotyphlops_multilineatus",
+              "Anilios_splendidus")
+
+pruned.tree <- ape::drop.tip(phy = phy, tip = droplist)
+
+plot(pruned.tree); ape::axisPhylo()
+
+# Write to file
+write.tree(pruned.tree, file = "data/tree/subset_anilios_newick_b.tre")
+
+# Re read in the subset tree
+phy <- read.tree("data/tree/subset_anilios_newick_b.tre")
+
+taxon_list <- data.frame(taxon = phy$tip.label)
+taxon_list <- sort(taxon_list)
+
+# write.csv(taxon_list, file = "data/intermediate_data/geohisse/taxon_list.csv", row.names = FALSE)
+
+
+# State data --------------------------------------------------------------
+
+# In this version, we classify the species as either arid or non-arid, no widespread stages. 
+# Classification is based on distribution of 'species'. I plotted ALA distribution on biome map of Australia in QGIS. 
+# Species that were previously classifed as widespread includes the following:
+# Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
[…]

### 2022-11-03 manuscript.Rmd
prep for j of biogeography

@@
 # biblio-style: "apalike"
-csl: evolution.csl
+csl: jbiogeography.csl
 link-citations: true
@@
 
+**Taxon**: Blind snakes
+
+**Location**: Australasia
+
+**Methods**:
+
+**Results**:
+
+**Main conclusions**:
+
 **Keywords**
-scolecophidia, Diversification, paleoclimate, RPANDA, BioGeoBEARS
+scolecophidia, Diversification, paleoclimate, BioGeoBEARS
 
@@
 <!-- Aridification  -->
-
 The aridification on continental Australia over the past 20 million years resulted in a diversity of habitats and spurred multiple radiations of flora and fauna across the continent, many of which are endemic [@byrneBirth2008; @melvilleEvolutionary2011].
@@
 (3) reconstructed the ancestral biome of Australian blindsnakes? 
-(4) Are there differences in the rates of speciation between arid and non-arid zones?
+(4) reconstructed ancestral niche and model niche evolution of the radiation
+(5) Are there differences in the rates of speciation between arid and non-arid zones?
 
@@
 
+## Niche evolution
+
+We reconstructed ancestral niche from present-only data using NEMo package.
+
 Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
@@
 
-## Paleoclimate conditions
+## Paleoclimate estimates
 
@@
 - Topography dyanmics [@brazModelling2021]
+- Solomon Island 40 MYA [@hallCenozoic2002]
 
@@
 
-- Solomon Island 40 MYA [@hallCenozoic2002]
+Rates of diversification between arid and non-arid, and widespread.
 
-## Arid vs. non-arid speciation
+## Niche evolution
 
@@
 
+## Niche evolution
+
 # Conclusion

### 2022-11-03 code/01_env_data_plots.R
update script to plot points

@@ -33,6 +33,7 @@ env_plottr <- function(.df){
   .df %>% 
   ggplot(data = ., aes(x = age, colour = source)) +
     geom_line(aes(y = vals), size = 1) +
+    geom_point(aes(y = vals)) + 
     deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
     # scale_colour_manual(values = viridis(4), name = "Data source", labels = c("RPANDA Global", "Scotese Australia", "Straume Australia", "Valdes Australia")) +
     scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Mean temperature (°C)") +
@@ -55,3 +56,6 @@ env_df[which(env_df$source %in% c("arid_sco", "arid_str", "arid_val")),] %>%
   labs(y = "Aridity index")
 
 dev.off()  
+
+
+

### 2022-11-03 code/02_diversification_plots_b.R
update script to prepare RPANDA results

@@ -1,0 +1,145 @@
+ # 02_diversification_plots_b.R
+# Putter Tiatragul
+# Plotting diversification _b models
+
+# libraries ---------------------------------------------------------------
+
+library(RPANDA); library(dplyr)
+library(phytools); library(gridExtra)
+library(deeptime) # for deeptime axis
+library(ggplot2); library(viridis)
+library(patchwork)
+
+# source("code/01_env_data_plots.R")
+# Load results from 01_diversification_analyses ---------------------------
+source("code/02_table_fit_env_results_b.R") 
+source("code/utility/plot_fit_env_new.R") # load custom functions and some data
+
+b_table$AICc <- as.numeric(b_table$AICc)
+
+# Check best fitting model by checking lowest AICc values
+b_table %>% 
+  dplyr::arrange(AICc) %>% 
+  head()
+
+# Plot best fitting -------------------------------------------------------
+
+# Best fitting models based on lowest AICc
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_arid_sco_b.Rdata'); arid_sco <- Anilios_res
+load('data/intermediate_data/diversification_analyses/Anilios_EnvDep_arid_str_b.Rdata'); arid_str <- Anilios_res
+
+# arid_sco[[1]]$BEnvVarDCST_EXPO
+
+fit_env_arid <- arid_sco[[1]]$BEnvVarDCST_EXPO
+
+# RPANDA Base R plot
+plot_fit_env_new(fit_env_arid, 
+                 env_data = env_data_list$arid_sco, tot_time = tot_time,
+                 name_env = 'Aridity')
+
+
+# GGPLOT
+# fit_env_df_make is a custom function 
+arid_sco_df <- fit_env_df_make(env_data = env_data_list$arid_sco, 
+                               tot_time = tot_time,
+                              fit.env = fit_env_arid)
+
+# Custom theme
+theme_1 <- function (base_size = 11, base_family = "", base_line_size = base_size/22, base_rect_size = base_size/22) {
+  theme_grey(base_size = base_size,
+             base_family = base_family, base_line_size = base_line_size,
+             base_rect_size = base_rect_size
+             ) %+replace%
+    theme(panel.background = element_rect(fill = "white", colour = NA),      
[…]

### 2022-11-03 code/utility/plot_fit_env_new.R
update script to calculate values to plot from RPANDA

@@ -1,0 +1,129 @@
+# plot_fit_env_new.R
+# modified function of plot_fit_env from RPANDA.
+
+library(RPANDA); library(dplyr)
+library(pspline)
+
+plot_fit_env_new <- function (fit.env, env_data, tot_time, name_env) 
+{
+  if (!inherits(fit.env, "fit.env")) 
+    stop("object is not of class \"fit.env\"")
+  t <- seq(0, tot_time, length.out = 100)
+  # dev.new()
+  plot(-t, sapply(t, fit.env$f.lamb), type = "l", xlab = "time", 
+       ylab = "speciation rate", main = "Fitted speciation rate")
+  df <- smooth.spline(env_data[, 1], env_data[, 2])$df
+  spline_result <- sm.spline(env_data[, 1], env_data[, 2], 
+                             df = df)
+  env_func <- function(t) {
+    predict(spline_result, t)
+  }
+  # dev.new()
+  plot(env_func(t), sapply(t, fit.env$f.lamb), xlab = paste(name_env, "Environmental data", sep=" "), 
+       ylab = "speciation rate", main = "Fitted speciation rate")
+  if ("f.mu" %in% attributes(fit.env)$names) {
+    # dev.new()
+    plot(-t, sapply(t, fit.env$f.mu), type = "l", xlab = "time", 
+         ylab = "extinction rate", main = "Fitted extinction rate")
+    # dev.new()
+    plot(env_func(t), sapply(t, fit.env$f.mu), xlab = paste(name_env, "Environmental data", sep=" "), 
+         ylab = "extinction rate", main = "Fitted extinction rate")
+    r <- function(t) {
+      fit.env$f.lamb(t) - fit.env$f.mu(t)
+    }
+    # dev.new()
+    plot(-t, sapply(t, r), type = "l", xlab = "time", ylab = "net diversification rate", 
+         main = "Fitted net diversification rate")
+    # dev.new()
+    plot(env_func(t), sapply(t, r), xlab = paste(name_env, "Environmental data", sep=" "), 
+         ylab = "net diversification rate", main = "Fitted net diversification rate")
+  }
+  else {
+    # dev.new()
+    plot(-t, sapply(t, fit.env$f.lamb), type = "l", xlab = "time", 
+         ylab = "net diversification rate", main = "Fitted net diversification rate")
+    # dev.new()
+    plot(env_func(t), apply(t, fit.env$f.lamb), xlab = paste(name_env, "Environmental data", sep=" "), 
[…]

### 2022-11-03 code/utility/plot_ClaDS_phylo_colour.R
colour variety for plotting ClaDS

@@ -1,0 +1,420 @@
+# plot_ClaDS_phylo_colour.R
+
+library(fields)
+
+plot_ClaDS_phylo_colour <- function (phylo, rates, rates2 = NULL, same.scale = T, main = NULL,
+    lwd = 2, log = T, show.tip.label = F, 
+    colour = colorRampPalette(c("steelblue2", "red4"))(100), ...) 
+{
+    Colors = colour
+    if (is.null(rates2)) {
+        if (log) 
+            rates = log(rates)
+        if (isTRUE(all.equal(rep(as.numeric(rates[1]), length(rates)), 
+            as.numeric(rates)))) {
+            col = rep(1, length(rates))
+            plot(phylo, edge.color = Colors[col], show.tip.label = show.tip.label, 
+                main = main, edge.width = lwd, ...)
+            if (log) {
+                image.plot(z = c(exp(rates[1]), 2 * exp(rates[1])), 
+                  col = Colors, horizontal = T, legend.only = T)
+            }
+            else {
+                image.plot(z = c(rates[1], 2 * rates[1]), col = Colors, 
+                  horizontal = T, legend.only = T)
+            }
+        }
+        else {
+            col = round((rates - min(rates))/diff(range(rates)) * 
+                99) + 1
+            plot(phylo, edge.color = Colors[col], show.tip.label = show.tip.label, 
+                main = main, edge.width = lwd, ...)
+            if (log) {
+                min = min(rates)
+                max = max(rates)
+                m10 = floor(min/log(10))
+                M10 = ceiling(max/log(10))
+                if ((M10 - m10) < 4) {
+                  ticks = c(1, 2, 5)
+                }
+                else {
+                  ticks = 1
+                }
+                ticks = as.vector(sapply(m10:M10, function(k) {
+                  return(ticks * 10^k)
+                }))
+                lt = length(ticks[ticks > exp(min) & ticks < 
+                  exp(max)])
+                if (lt < 4) {
+                  ticks = c(round(exp(min), max(0, -1 * m10 + 
+                    (lt < 2))), ticks, round(exp(max), max(0, 
+                    -1 * M10 + 1 + (lt < 2))))
+                }
+                image.plot(z = c(min, max), col = Colors, horizontal = T, 
+                  legend.only = T, axis.args = list(at = log(ticks), 
+                    labels = ticks))
+            }
+            else {
+                image.plot(z = as.matrix(rates), col = Colors, 
+                  horizontal = T, legend.only = T)
+            }
[…]

### 2022-11-03 code/06_CLaDS_plot_tips.R
script to plot tips from clads

@@ -1,0 +1,114 @@
+# 06_CLaDS_plot_tips.R
+
+# Cladogenetic diversification in rate shifts 
+# Estimates lineage specific diversification rates on a phylogeny. 
+# Accounts for diverse sources of variation in diversification rates that occur
+# during the evolutionary history of clades. 
+
+# Plot the MCMC chains obtained with fit_ClaDS.
+
+library(ape); library(phytools); library(RPANDA)
+
+# Load data ---------------------------------------------------------------
+s_tree <- ape::read.tree("data/tree/subset_anilios_newick_b.tre")
+
+# load data from 06_ClaDS_Anilios.R
+
+
+# ClaDS0 ------------------------------------------------------------------
+
+plot_ClaDS0_chains(clads.out, burn = 1/2, thin = 1, 
+                   param = c("sigma", "alpha", "l_0", "LP"))
+
+clads0_rates <- getMAPS_ClaDS0(s_tree, clads.out, burn=0.5, thin=1)
+clads0_rates
+
+# 
+pdf(file = 'output/ClaDS0_plot.pdf', width = 30, height = 40)
+plot_ClaDS_phylo(s_tree, clads0_rates[-(1:3)], rates2 = NULL,
+                 same.scale = T, main = NULL, 
+                 lwd = 2, log = T, show.tip.label = F)
+dev.off()
+
+
+# ClaDS1 ------------------------------------------------------------------
+
+
+
+plot_ClaDS_chains(clads1.out)
+# Extract the maxima A posteriori for each parameter
+
+maps = getMAPS_ClaDS(clads1.out, thin = 1)
+print(paste0("sigma = ", maps[1], " ; alpha = ", 
+             maps[2], " ; epsilon = ", maps[3], " ; l_0 = ", maps[4] ))
+
+# Plot 
+plot_ClaDS_phylo(phylo = s_tree, rates = maps[-(1:4)], 
+                 main = "ClaDS1",
+                 same.scale = T, lwd = 2, log = T, show.tip.label = F)
+axisPhylo() # plots timescale
+
+source('code/utility/plot_ClaDS_phylo_colour.R')
+plot_ClaDS_phylo_colour(phylo = s_tree, rates = maps[-(1:4)], 
+                        main = "ClaDS1",
[…]

### 2022-11-03 code/02_table_fit_env_results_b.R
clean up script

@@ -1,7 +1,8 @@
 # 02_table_fit_env_results_b.R
 # September 2022
 
-# This script will organise the results from the parallel fit and the individual fit together in one table from fitting on B tree (with cauchy priors).
+# This script will organise the results from the parallel fit and the individual
+# fit together in one table from fitting on B tree (with cauchy priors).
 
 # From 01_fit_*.R. I successfully fitted several environment-dependent birth-death models.
 # The following had to be fitted individually using initial lamb_par 0.1 and/or 0. 
@@ -19,189 +20,38 @@ library(stringr)
 file_n <- list.files(path = "data/intermediate_data/diversification_analyses/", pattern = "_b.txt", full.names = T)
 
 # get the names of the model
-file_n_m <- stringr::str_remove(list.files(path = "data/intermediate_data/diversification_analyses/", pattern = ".txt"), pattern = "Results_Anilios_")
+file_n_m <- stringr::str_remove(list.files(path = "data/intermediate_data/diversification_analyses/",
+                                           pattern = "_b.txt"), pattern = "Results_Anilios_")
 mod_names <- stringr::str_remove(file_n_m, pattern = ".txt")
 
 # List of models with their names
 filelist <- lapply(file_n, read.delim)
-names(filelist) <- mod_names
 
 # Environmental data in list form
 load('data/intermediate_data/diversification_analyses/env_data_list.RData')
 
-# Put environment information in the table
+# Put environment information in the table as a column
 for(i in 1:length(filelist)){
-  filelist[[i]]$env <- names(filelist)[i]
+  filelist[[i]]$env <- mod_names[i]
 }
-
-# Plot_fit_env the models to check 
[…]

### 2022-11-04 manuscript.Rmd
update methods

@@
 Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which have undergone rapid diversification in the last 10 – 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
+- Aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
 For example, some Australian squamate clades like dragons [@melvilleIntegrating2019], sphenomorphine skinks [@raboskyAnalysis2014], geckos [@ashmanDiversification2018;@brennanMass2017] and elapid snakes [@leeDiversification2016;@grundlerTrophic2014] show increased rates of diversification coincidentally with the onset of aridification.
@@
 We 
-(1) evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
-(2) compared the model fit between different ecological processes on diversification rates, and 
-(3) reconstructed the ancestral biome of Australian blindsnakes? 
-(4) reconstructed ancestral niche and model niche evolution of the radiation
-(5) Are there differences in the rates of speciation between arid and non-arid zones?
+- evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
+- compared the model fit between different ecological processes on diversification rates, and 
+- reconstructed ancestral niche and model niche evolution of the radiation
+- reconstructed the ancestral biome of Australian blindsnakes? 
+- Are there differences in the rates of speciation between arid and non-arid zones?
 
 #  Materials and methods
-
-## Reconstruction of Australian paleo climate 
-<!-- Consult Alex Skeels -->
-Two paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021 and @straumeGlobal2020.
-We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
-From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day (\@ref(fig:fig-aus-region).
 
@@
 We employ the recent time-calibrated phylogenomic tree of Australian blindsnake with outgroups.
-This tree includes __ taxa representing __/48 described species of _Anilios_.
-However, because we our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to x/52 species.
-The species were pruned to single operational taxonomic unit (OTU).
+This tree includes 38 operational taxonomic units (OTUs) representing  __taxa representing __ described species of _Anilios_.
+However, because our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to XX/XX species.
+The genetically distinct OTUs that are not fully recognised species include 3 'Anilios grypus' and 2 'Anilios ligatus'.
+
+## Reconstruction of Australian paleo climate and diversification analyses
+<!-- Consult Alex Skeels -->
+Three paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021, @straumeGlobal2020, [VALDES].
+We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
+From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day.
+
+To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 global paleo temperature, 3 Australian paleo temperature, 3 Australian paleo aridity, and 3 Australian paleo proportion arid area.
+Missing taxa is accounted for by fractioning function in RPANDA.
+We then selected the model that best fit the data based on corrected Akaike Information Criterion (AICc) scores.
+
+## Biogeographic-dependent diversification analyses
+
+We further investigate whether broad patterns of diversification are associated with geographic range.
+Using the R package `hisse` v.
+2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction models from the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the tree and geographic range data.
+Since state speciation-extinction (SSE) models only allow three states including two geographic states and one widespread state.
+Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
+We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas.
+The five models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved hidden states.
+Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
+Model 3 and 4 are derivation of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
+The last model is a complementary model that removes cladogenetic effect entirely.
+We compared the AIC scores to select the best supported models, then averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights following @caetanoHidden2018.
+
+## Niche evolution and spatial data
+We reconstructed ancestral niche from present-only data using NEMo package. 
+from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).
 
@@
 
-To evaluate the ancestral biome of blindsnakes in Australia, we used the R package `BioGeoBEARS` 1.1.1 [@nicholas_j_matzke_2018_1478250].
+We used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013] to estimate the ancestral biome of blindsnakes in Australia.
 BioGeoBEARS provides a flexible likelihood-based framework that models how the branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
 These shifts in biomes can occur within a single branch (anagenetic events including processes like dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant and founder-event speciation).
-The fundamental inputs for the software include a time-calibrated phylogeny (described above), information about the biome/range each species are found, and a historical biogeographic model.
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) in which the based on occurrence records from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Supplementary information S2 for how data was compiled).
-Biome designations were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
-
+The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range each species are found, and a historical biogeographic model.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) based on their occurence records as described above.
+Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
 The phylogeny and distribution information were then applied to widely used historical biogeographical methods including Dispersal-Vicariance Analysis (DIVA; [@ronquistDispersalVicariance1997]), and likelihood implementations of Dispersal-Extinction-Cladogenesis (DEC; [@reeMaximum2008]) or DECLIKE, and BayArea [@landisBayesian2013] or BAYAREALIKE [@matzkeBioGeoBEARS2013].
-, we included two additional free parameters.
-The first is the *j* parameter that assigns per-event weight to founder-events [@matzkeModel2014].
-The second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Supplementary information).
-We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on AICc (corrected Akaike Information Criterion) scores.
-
-<!-- SUPPLEMENT TABLE WITH OCCURENCE RECORDS AND BIOME MAP -->
-
-## Diversification analyses
-
-To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
-These models will be a combination of pure-birth and birth-death models with different speciation/extinction rates specified as constant, linearly, or exponentially correlated with paleoenvironment.
-We fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 global paleo temperature, 3 Australian paleo temperature, 3 Australian paleo aridity, and 3 Australian paleo proportion arid area.
-Missing taxa is accounted for by fractioning function in RPANDA.
-We assessed model fit with corrected AICc scores.
-
-- Initial lamba parameters were changed for mean_sco
-
-## Biogeographic-dependent diversification analyses
-
-- Aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
-- XX species today occupy arid zones exclusively and part of the range of YY species includes the arid zone as defined by WWF Bioregions. 
-- To evaluate whether broad patterns of diversification are associated with geographic range, we used a series of biogeographic-dependent speciation-extinction models 
-including GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] implemented in the R package `hisse` v. 2.1.9 [@beaulieuDetecting2016].
-
-- GeoHiSSE differs from GeoSSE by incorporating hidden states, which accounts for heteregenous diversification caused by unsampled traits.
-
-- These models only allow three states including two geographic states and one widespread state.
-As such we categorised species that primarily occur in the arid zone as arid, species that occur primarily outside of the arid zone as non-arid, and those that occupy both the arid and non-arid zone as widespread.
-We assumed uniform taxon sampling for 89% in arid areas, 71 in non-arid areas, and 83% in widespread areas.
--
-
-## Niche evolution
-
-We reconstructed ancestral niche from present-only data using NEMo package.
-
-Anilios_bituberculatus - distributed in the arid zone but also many many occurence in coastal areas in southern Australia. Classified as non-arid.
-Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
-Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
-Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
-Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.
-
- for biogeographic-depdendent diversification using geographic state and hidden state speciation-extinction model (GeoHiSSE) analysis with different parameters [@goldbergPhylogenetic2011].
+We included two additional free parameters.
+The first parameter *j* assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Appendix-method-biogeobears).
+We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on AICc scores.
 
 ## Statistical analyses
-All statistical analyses were run in R version 4.2 (CITE R).
+All statistical analyses were run in R version 4.2 (CITE R). We visualised diversification data using `ggplot2`, made map with `QGIS`.
 

### 2022-11-09 code/utility/00_geohisse_tree_subset.R
script to make subset tree

@@ -1,0 +1,36 @@
+# 05_subset_st_tree.R
+# November 2022
+
+# libraries ---------------------------------------------------------------
+
+library(ape)
+library(dplyr); library(tidyr)
+
+# Data --------------------------------------------------------------------
+
+trees <- ape::read.tree(file = "data/intermediate_data/diversification_analyses/blindsnake.trees", tree.names = c("st", "b"))
+
+fos_tree <- phytools::force.ultrametric(trees[[1]],"extend")
+fos_tree$edge.length <- fos_tree$edge.length * 100
+phy <- fos_tree
+
+# Plot to check dates
+plot(phy); axisPhylo()
+
+# Check to make sure tree is ultrametric
+is.ultrametric(phy)
+
+# Prune -------------------------------------------------------------------
+
+# List of species to drop from analysis. In this case we only want Anilios
+droplist <- c("Acutotyphlops_subocularis", "Sundatyphlops_polygrammicus",
+              "Sundatyphlops_polygrammicus", "Ramphotyphlops_multilineatus",
+              "Anilios_splendidus")
+
+pruned.tree <- ape::drop.tip(phy = phy, tip = droplist)
+
+plot(pruned.tree); ape::axisPhylo()
+
+# Write to file
+write.tree(pruned.tree, file = "data/tree/subset_anilios_newick_st.tre")
+

### 2022-11-09 code/utility/squish_trans.R
script for squishing function ggplot

@@ -1,0 +1,40 @@
+# squish_trans.R
+
+# used for squishing axis
+
+library(scales)
+squish_trans <- function(from, to, factor) {
+  
+  trans <- function(x) {
+    
+    if (any(is.na(x))) return(x)
+    
+    # get indices for the relevant regions
+    isq <- x > from & x < to
+    ito <- x >= to
+    
+    # apply transformation
+    x[isq] <- from + (x[isq] - from)/factor
+    x[ito] <- from + (to - from)/factor + (x[ito] - to)
+    
+    return(x)
+  }
+  
+  inv <- function(x) {
+    
+    if (any(is.na(x))) return(x)
+    
+    # get indices for the relevant regions
+    isq <- x > from & x < from + (to - from)/factor
+    ito <- x >= from + (to - from)/factor
+    
+    # apply transformation
+    x[isq] <- from + (x[isq] - from) * factor
+    x[ito] <- to + (x[ito] - (from + (to - from)/factor))
+    
+    return(x)
+  }
+  
+  # return the transformation
+  return(trans_new("squished", trans, inv))
+}

### 2022-11-10 manuscript.Rmd
update writing about aims and clads

@@
 We 
-- evaluated the model fit performance of localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
-- compared the model fit between different ecological processes on diversification rates, and 
-- reconstructed ancestral niche and model niche evolution of the radiation
-- reconstructed the ancestral biome of Australian blindsnakes? 
+- Estimate the ancestral biome of Australian blindsnakes based on extant taxa.
+- Compare fit_env models between localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
+- Assess model fit between paleo-aridity and paleo-climate on diversification rates, and 
+- Estimate ancestral niche and model niche evolution of the blind snake radiation
 - Are there differences in the rates of speciation between arid and non-arid zones?
@@
 However, because our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to XX/XX species.
-The genetically distinct OTUs that are not fully recognised species include 3 'Anilios grypus' and 2 'Anilios ligatus'.
-
-## Reconstruction of Australian paleo climate and diversification analyses
-<!-- Consult Alex Skeels -->
-Three paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021, @straumeGlobal2020, [VALDES].
-We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
-From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day.
-
-To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time-variable likelihood diversification models using the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
-Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 global paleo temperature, 3 Australian paleo temperature, 3 Australian paleo aridity, and 3 Australian paleo proportion arid area.
-Missing taxa is accounted for by fractioning function in RPANDA.
-We then selected the model that best fit the data based on corrected Akaike Information Criterion (AICc) scores.
-
-## Biogeographic-dependent diversification analyses
-
-We further investigate whether broad patterns of diversification are associated with geographic range.
-Using the R package `hisse` v.
-2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction models from the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the tree and geographic range data.
-Since state speciation-extinction (SSE) models only allow three states including two geographic states and one widespread state.
-Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
-We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas.
-The five models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved hidden states.
-Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
-Model 3 and 4 are derivation of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
-The last model is a complementary model that removes cladogenetic effect entirely.
-We compared the AIC scores to select the best supported models, then averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights following @caetanoHidden2018.
-
-## Niche evolution and spatial data
-We reconstructed ancestral niche from present-only data using NEMo package. 
-from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).
+The genetically distinct OTUs that are not fully recognised species include 3 '_Anilios grypus_' and 2 '_Anilios ligatus_'.
 
@@
 The first parameter *j* assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Appendix-method-biogeobears).
-We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on AICc scores.
+We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on corrected Akaike Information Criterion (AICc) scores.
+
+## Reconstruction of Australian paleo climate and environment-dependent diversification analyses
+<!-- Consult Alex Skeels -->
+Three paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021, @straumeGlobal2020, [VALDES].
+We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
+From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day.
+Because reptiles are sensitive to extremely low and high temperature, we decided to to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature.
+
+To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` function in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 mean global paleo temperature, 3 Australian paleo temperature, and 3 Australian paleo aridity.
+We accounted for missing taxa using the fractioning function (f = XX) in RPANDA.
+We then assessed which diversification model best fit the data based on AICc scores.
+
+To visualise the dynamics of the radiation, we use ClaDS [@malietFast2022] implemented in `julia` to estimate speciation rates at the origin of each branch (hencforth branch-specific rates).
+ClaDS is designed to capture frequent shifts with small effects.
+At each speciation event, daughter lineages are assigned its own speciation rate sampled from a distribution parameterised on the parental rate [@malietFast2022].
+
+## Biogeographic-dependent diversification analyses
+
+We further investigate whether broad patterns of diversification are associated with geographic range.
+Using the R package `hisse` v.
+2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction models from the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the tree and geographic range data.
+Since state speciation-extinction (SSE) models only allow three states, we restricted our geographic states to arid, non-arid, and widespread.
+Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
+We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas based on number of species in the phylogeny and all currently described species of _Anilios_.
+The five models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved hidden states.
+Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
+Model 3 and 4 are derivation of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
+The last model is a complementary model that removes cladogenetic effect entirely.
+We compared the AIC scores to select the best supported models, then averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights following @caetanoHidden2018.
+
+## Niche evolution and spatial data
+We reconstructed ancestral niche from present-only data using NEMo package. 
+from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).
 
@@
 
-## Paleoclimate estimates
-
-Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
-Both aridity index shows increase in aridity post 10 mya, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
-Temperature for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
-Reconstructed temperature values from Australia were higher than global temperature and stayed relatively constant compared to global temperature that shows slight decrease.
-
 ## Biogeographic history
@@
 Likelihood ratio test (LRT) indicates that founder event (parameter j) and distance matrix (free parameter _x_) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+*j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-Our result also supports the hypothesis that ancestor of the Australian clade invaded the continent, but exactly where they lived prior to this invasion is equivocal between the Solomon Islands and New Guinea.
-After arriving to Australia in the late Oligocene – early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years. 
+Our result also supports the hypothesis that the ancestor of the Australian clade invaded the continent, but exactly where they lived prior to this invasion is equivocal between the Solomon Islands and New Guinea.
+After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years. 
 - Our analysis is consistent with @marinTracing2013 conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
@@
 Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
+
+## Paleoclimate estimates
+
+Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
+Both aridity index shows increase in aridity post 10 mya, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
+Temperature for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
+Reconstructed temperature values from Australia were higher than global temperature and stayed relatively constant compared to global temperature that shows slight decrease.
 

### 2022-11-10 code/07_NEMo.R
init

@@ -1,0 +1,2 @@
+# 07_NEMo.R
+

### 2022-11-15 manuscript.Rmd
updated results

@@
 However, because our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to XX/XX species.
-The genetically distinct OTUs that are not fully recognised species include 3 '_Anilios grypus_' and 2 '_Anilios ligatus_'.
+The genetically distinct OTUs that are not fully recognised species include 3 '_Anilios grypus_' (2 the west and 1 in the east) and 2 '_Anilios ligatus_' (one from the east and one from the north).
+For ancestral range estimation analyses we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
 
@@
 
-We used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013] to estimate the ancestral biome of blindsnakes in Australia.
+To estimate the ancestral biome of blind snakes in Australia, we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible likelihood-based framework that models how the branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
@@
 The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range each species are found, and a historical biogeographic model.
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) based on their occurence records as described above.
-Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of previous Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
-The phylogeny and distribution information were then applied to widely used historical biogeographical methods including Dispersal-Vicariance Analysis (DIVA; [@ronquistDispersalVicariance1997]), and likelihood implementations of Dispersal-Extinction-Cladogenesis (DEC; [@reeMaximum2008]) or DECLIKE, and BayArea [@landisBayesian2013] or BAYAREALIKE [@matzkeBioGeoBEARS2013].
-We included two additional free parameters.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) based on their occurrence records as described above.
+Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
+The phylogeny and distribution information were then applied to widely used historical biogeographical methods including likelihood implementations of Dispersal-Vicariance Analysis [@ronquistDispersalVicariance1997] (DIVALIKE), Dispersal-Extinction-Cladogenesis [@reeMaximum2008] (DECLIKE), and BayArea [@landisBayesian2013] (BAYAREALIKE) in `BioGeoBEARS` [@matzkeBioGeoBEARS2013].
+We included two additional free parameters that we thought is relevant.
 The first parameter *j* assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Appendix-method-biogeobears).
-We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the best overall model based on corrected Akaike Information Criterion (AICc) scores.
+We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the model that fitted the data best based on corrected Akaike Information Criterion (AICc) scores.
 
-## Reconstruction of Australian paleo climate and environment-dependent diversification analyses
+## Reconstruction of Australian paleo climate
 <!-- Consult Alex Skeels -->
-Three paleographic digital elevation models (DEMs) providing paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution  were obtained from @scoteseAtlas2021, @straumeGlobal2020, [VALDES].
+To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES].
+Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution.
 We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
@@
 
+--More information about summarising values and determining aridity index ---
+
+## Time ane environment-dependent diversification analyses
+
 To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` function in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
-Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with the following environment variables: 1 mean global paleo temperature, 3 Australian paleo temperature, and 3 Australian paleo aridity.
+Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleo environment as described above.
+We also fit the model to widely use mean global paleo temperature data set (CITE).
 We accounted for missing taxa using the fractioning function (f = XX) in RPANDA.
@@
 
-To visualise the dynamics of the radiation, we use ClaDS [@malietFast2022] implemented in `julia` to estimate speciation rates at the origin of each branch (hencforth branch-specific rates).
-ClaDS is designed to capture frequent shifts with small effects.
+To visualise the dynamics of the radiation, we use ClaDS [@malietFast2022] implemented in `julia` to estimate speciation rates at the origin of each branch (hencforth branch-specific rates) and at the tip.
+ClaDS is designed to capture frequent shifts in speciation rates with small effects.
 At each speciation event, daughter lineages are assigned its own speciation rate sampled from a distribution parameterised on the parental rate [@malietFast2022].
@@
 
-We further investigate whether broad patterns of diversification are associated with geographic range.
+We further investigated whether broad patterns of diversification are associated with geographic range.
 Using the R package `hisse` v.
-2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction models from the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the tree and geographic range data.
-Since state speciation-extinction (SSE) models only allow three states, we restricted our geographic states to arid, non-arid, and widespread.
+2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction (SSE) models under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the phylogeny and geographic range data.
+Since SSE models only allow three states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
 Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
 We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas based on number of species in the phylogeny and all currently described species of _Anilios_.
-The five models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved hidden states.
+The five SSE models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved 'hidden states'.
 Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
@@
 The last model is a complementary model that removes cladogenetic effect entirely.
-We compared the AIC scores to select the best supported models, then averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights following @caetanoHidden2018.
+We compared the AIC scores to select the best models.
+Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
 
-## Niche evolution and spatial data
-We reconstructed ancestral niche from present-only data using NEMo package. 
-from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).
+<!-- ## Niche evolution and spatial data -->
+<!-- We reconstructed ancestral niche from present-only data using NEMo package.  -->
+<!-- from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).  -->
 
 ## Statistical analyses
-All statistical analyses were run in R version 4.2 (CITE R). We visualised diversification data using `ggplot2`, made map with `QGIS`.
+Statistical analyses were run in R version 4.2 (CITE R) and Julia (CITE).
+We visualised dynamics using baseR and `ggplot2` in R.
+The Australian bioregion map visualised in `QGIS` v (CITE Australian Map).
 
@@
 ## Biogeographic history
-We reconstructed the tropical grassland as the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears). The preferred biogeographic model is the constrained-dispersal matrix, DEC+j+x. 
-Likelihood ratio test (LRT) indicates that founder event (parameter j) and distance matrix (free parameter _x_) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+*j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-Our result also supports the hypothesis that the ancestor of the Australian clade invaded the continent, but exactly where they lived prior to this invasion is equivocal between the Solomon Islands and New Guinea.
+
+Based on our best fitting biogeographic model, the constrained-dispersal matrix DEC+_j_+_x_ model, we inferred tropical grassland biome to be ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears).
+Likelihood ratio test (LRT) indicates that founder event (parameter *j*) and distance matrix (free parameter *x*) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+*j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+Our result also supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the continent [@marinTracing2013], but exactly where they lived prior to this invasion is equivocal between the Bismarck Archipelago and New Guinea.
 After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
-- Our analysis is consistent with @marinTracing2013 conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
 
 ## Biome transitions and current distributions
-Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occured very rapidly.
-Descendents of the earliest split remained in the Tropical Grassland biome but within a million years later another split occured.
+
+Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly.
+Descendents of the earliest split remained in the Tropical grassland biome but within a million years later another split occured.
 The earliest biome transition was from tropical grassland into Temperate Forest around 25 MYA (95% HPD: 28 – 22 MYA) and the Arid zone 24 MYA (95% HPD: 28 – 21 MYA).
-Subsequent splits show transitions from the arid zone into the Mediterranean biome starting after 15 MYA. 
-Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the arid zone and Mediterranean biome. 
-Closely related species tend to share the same biome and occur around the same geographical areas.
-
-- Range expansion was important for diversification ...
-- Topography dyanmics [@brazModelling2021]
-- Solomon Island 40 MYA [@hallCenozoic2002]
-
-Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
+Subsequent splits show transitions from the Arid zone into the Mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears).
+Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome.
+We observed that closely related species tend to share the same biome and occur around the same geographical areas.
 
@@
 Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
-Both aridity index shows increase in aridity post 10 mya, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
-Temperature for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
-Reconstructed temperature values from Australia were higher than global temperature and stayed relatively constant compared to global temperature that shows slight decrease.
+Both aridity index shows increase in aridity post 10 MYA, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
+Paleoclimate estimates for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
+However, all three paleo temperature models indicates that Australia was hotter than the average global temperature.
+Paleoclimate in Australia were also relatively constant compared to global temperature that shows greater fluctuation then slight decrease towards the present.
 
-## Diversification
-The best fitting model among constant, time-dependent, temperature-dependent, and aridity dependent models is 
-This model indicate that X associate with speciation (alpha = ) and .....
+## Environment-dependent diversification
 
-Fit with Australia specific mean temperature receives better support than model with global temperature and exponential dependency of speciation rate with time.
+We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) with aridity index derived from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA).
+This model indicates that increase aridity is exponentially associate with decrease speciation rate (lambda = 1.32, alpha = -0.14) and with a constant background extinction (mu = 0.23).
+The second and third most competitive models (delta AICc < 2) also found increase aridity is exponentially associated with decrease speciation rate but without estimating extinction (Table-RPANDA-TOP10).
+Among exponential environment-dependent models fitted to our blind snake phylogeny, we found that models that uses Australia-specific paleoclimate data received better AICc support compared to fit with global paleoclimate (see supp-TABLE_FULL_RPANDA_model_fit).
+
+## Rates of diversification
 
 Rates of diversification between arid and non-arid, and widespread.
-
-## Niche evolution
 
@@
 torresianus_, _A.
-leucoproctus_, ...) are found in Northern Australia Torres Strait Islands, and southern parts of New Guinea.
+leucoproctus_, ...) are found in Northern
+Australia Torres Strait Islands, and southern parts of New Guinea.
+- Range expansion was important for diversification ...
+- Topography dyanmics [@brazModelling2021]
+- Solomon Island 40 MYA [@hallCenozoic2002]
+
+- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
 

### 2022-11-15 manuscript.Rmd
update manuscript intro

--- 2022-11-10 manuscript.Rmd
@@
 
-<!-- Aridification  -->
-The aridification on continental Australia over the past 20 million years resulted in a diversity of habitats and spurred multiple radiations of flora and fauna across the continent, many of which are endemic [@byrneBirth2008; @melvilleEvolutionary2011].
-The arid zone makes up the majority of the Australian landscape and provides habitat to ~43% of all Australia’s terrestrial reptiles.
-Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which have undergone rapid diversification in the last 10 – 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
-- Aridification has been hypothesised as a driver of diversification of many groups of reptiles and amphibians (CITE). 
-For example, some Australian squamate clades like dragons [@melvilleIntegrating2019], sphenomorphine skinks [@raboskyAnalysis2014], geckos [@ashmanDiversification2018;@brennanMass2017] and elapid snakes [@leeDiversification2016;@grundlerTrophic2014] show increased rates of diversification coincidentally with the onset of aridification.
-However, these studies did not account for heterogeneity of the arid zone [@pepperLife2021].
+<!-- General background  -->
+
+Present-day distrubtion of extant species offers clues about their biogeographic history and how the lineages may have diversified.
+
+The aridification on continental Australia over the past 20 million years resulted in a diversity of habitats and spurred multiple radiations of flora and fauna across the continent, many of which are endemic [@byrneBirth2008;@melvilleEvolutionary2011].
+The arid zone makes up the majority of the Australian landscape and provides habitat to ~43% of all Australia’s terrestrial squamate reptiles (lizards and snakes).
+Reptiles represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified throughout the Miocene (starting ~25 MYA) to the present [@oliverPhylogenetic2017].
+For example, agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013] are reptiles that colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene.
 
@@
 The diversification is reported to have originated from a single colonisation event by a non-arid adapted ancestor from Asia between 19-39 mya [@vidalBlindsnake2010;@marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
+
 However, the study did not account for alternative biogeographic areas (e.g.
@@
 - Assess model fit between paleo-aridity and paleo-climate on diversification rates, and 
-- Estimate ancestral niche and model niche evolution of the blind snake radiation
 - Are there differences in the rates of speciation between arid and non-arid zones?
@@
 
-## Niche evolution and spatial data
-We reconstructed ancestral niche from present-only data using NEMo package. 
-from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).
-
 ## Statistical analyses
@@
 
+We found no statistical difference between diversification rates between arid and non-arid adapted lineages. 
 Rates of diversification between arid and non-arid, and widespread.
-
-## Niche evolution
 
@@
 
+Account for heterogeneity of the arid zone [@pepperLife2021].
+
 # Tables

### 2022-11-15 06_ClaDS0_1_plots.R
delete this file in root dir

@@ -1,48 +1,0 @@
-# 06_ClaDS0_1_plots.R
-
-# Cladogenetic diversification in rate shifts 
-# Estimates lineage specific diversification rates on a phylogeny. 
-# Accounts for diverse sources of variation in diversification rates that occur
-# during the evolutionary history of clades. 
-
-library(ape); library(phytools); library(RPANDA)
-library(phyloch); data(strat2012)
-
-# R fit ClaDS0 and ClaDS1
-load('data/intermediate_data/ClaDS/clads_models.Rdata')
-
-# ClaDS0 ------------------------------------------------------------------
-
-plot_ClaDS0_chains(clads.out, burn = 1/2, thin = 1, 
-                   param = c("sigma", "alpha", "l_0", "LP"))
-
-clads0_rates <- getMAPS_ClaDS0(s_tree, clads.out, burn=0.5, thin=1)
-clads0_rates
-
-# 
-pdf(file = 'output/ClaDS0_plot.pdf', width = 30, height = 40)
-plot_ClaDS_phylo(s_tree, clads0_rates[-(1:3)], rates2 = NULL,
-                 same.scale = T, main = NULL, 
-                 lwd = 2, log = T, show.tip.label = F)
-dev.off()
-
-
-# ClaDS1 ------------------------------------------------------------------
-
-plot_ClaDS_chains(clads1.out)
-# Extract the maxima A posteriori for each parameter
-
-maps = getMAPS_ClaDS(clads1.out, thin = 1)
-print(paste0("sigma = ", maps[1], " ; alpha = ", 
-             maps[2], " ; epsilon = ", maps[3], " ; l_0 = ", maps[4] ))
-
-# Plot 
-plot_ClaDS_phylo(phylo = s_tree, rates = maps[-(1:4)], 
-                 main = "ClaDS1",
-                 same.scale = T, lwd = 2, log = T, show.tip.label = F)
-axisPhylo() # plots timescale
-
-source('code/utility/plot_ClaDS_phylo_colour.R')
-plot_ClaDS_phylo_colour(phylo = s_tree, rates = maps[-(1:4)], 
-                        main = "ClaDS1",
-                        same.scale = T, lwd = 2, log = T, show.tip.label = F)

### 2022-11-15 code/06_ClaDS0_1_plots.R
tip rates for ClaDS0

@@ -1,0 +1,48 @@
+# 06_ClaDS0_1_plots.R
+
+# Cladogenetic diversification in rate shifts 
+# Estimates lineage specific diversification rates on a phylogeny. 
+# Accounts for diverse sources of variation in diversification rates that occur
+# during the evolutionary history of clades. 
+
+library(ape); library(phytools); library(RPANDA)
+library(phyloch); data(strat2012)
+
+# R fit ClaDS0 and ClaDS1
+load('data/intermediate_data/ClaDS/clads_models.Rdata')
+
+# ClaDS0 ------------------------------------------------------------------
+
+plot_ClaDS0_chains(clads.out, burn = 1/2, thin = 1, 
+                   param = c("sigma", "alpha", "l_0", "LP"))
+
+clads0_rates <- getMAPS_ClaDS0(s_tree, clads.out, burn=0.5, thin=1)
+clads0_rates
+
+# 
+pdf(file = 'output/ClaDS0_plot.pdf', width = 30, height = 40)
+plot_ClaDS_phylo(s_tree, clads0_rates[-(1:3)], rates2 = NULL,
+                 same.scale = T, main = NULL, 
+                 lwd = 2, log = T, show.tip.label = F)
+dev.off()
+
+
+# ClaDS1 ------------------------------------------------------------------
+
+plot_ClaDS_chains(clads1.out)
+# Extract the maxima A posteriori for each parameter
+
+maps = getMAPS_ClaDS(clads1.out, thin = 1)
+print(paste0("sigma = ", maps[1], " ; alpha = ", 
+             maps[2], " ; epsilon = ", maps[3], " ; l_0 = ", maps[4] ))
+
+# Plot 
+plot_ClaDS_phylo(phylo = s_tree, rates = maps[-(1:4)], 
+                 main = "ClaDS1",
+                 same.scale = T, lwd = 2, log = T, show.tip.label = F)
+axisPhylo() # plots timescale
+
+source('code/utility/plot_ClaDS_phylo_colour.R')
+plot_ClaDS_phylo_colour(phylo = s_tree, rates = maps[-(1:4)], 
+                        main = "ClaDS1",
+                        same.scale = T, lwd = 2, log = T, show.tip.label = F)

### 2022-11-15 code/06_CLaDS_plot_tips.R
plot tip rates for ClaDS

@@ -5,8 +5,6 @@
 # Accounts for diverse sources of variation in diversification rates that occur
 # during the evolutionary history of clades. 
 
-
-
 # libraries ---------------------------------------------------------------
 library(ape); library(phytools); library(RPANDA)
 library(phyloch); data(strat2012); library(treeio)
@@ -31,7 +29,7 @@ source('code/utility/squish_trans.R') # squish continuous axis
 dev.off()
 pdf(file = 'output/supp_ClaDS_magma.pdf', width = 8.5, height = 11.33)
 plot_ClaDS_phylo_colour(s_tree, rates = CladsOutput$lambdai_map, 
-                 show.tip.label = T, log = T)
+                 show.tip.label = T, log = T, direction = "leftwards")
 axisGeo(GTS = strat2012, unit = "epoch"); axisPhylo()
 dev.off()
 
@@ -47,6 +45,11 @@ plot_ClaDS_phylo_colour(s_tree, rates = CladsOutput$lambdatip_map,
                                         show.tip.label = T, log = T)
 axisGeo(GTS = strat2012, unit = "epoch"); axisPhylo()
 dev.off()
+
+plot_ClaDS_phylo_colour(s_tree, rates = CladsOutput$RTT_map, 
+                        show.tip.label = T, log = T)
+
+lapply(CladsOutput$rtt_chains[[1]], max)
 
 # Get tip rates -----------------------------------------------------------
 

### 2022-11-16 code/04_BioGeoBEARS_results_bsm.R
biogeobears stochastic mapping 50. Extracting count values for anagenetic_dispersal and plot with ggplot. Needs to be cleaned

@@ -6,7 +6,6 @@
 
 # load results from analysis script
 
-
 # Libraries ---------------------------------------------------------------
 
 library(GenSA);library(FD)      
@@ -31,7 +30,7 @@ load("data/intermediate_data/bears/bears_model_fit.Rdata")
 
 load("data/intermediate_data/bears/AniliosDEC_JX_full.Rdata")
 resjx
-
+tr <- tree.c
 # BSM = Biogeographic Stochastic Mapping for the best fitting model
 
 model_name = "DECjx"
@@ -80,7 +79,7 @@ lnum = 0
 # Update: 2015-09-23 -- now, if you used multicore functionality for the ML analysis,
 # the same settings will be used for get_inputs_for_stochastic_mapping().
 #######################################################
-BSM_inputs_fn = "BSM_inputs_file.Rdata"
+BSM_inputs_fn = "data/intermediate_data/bears/BSM_inputs_file.Rdata"
 runInputsSlow = FALSE
 if (runInputsSlow)
 {
@@ -243,7 +242,8 @@ clado_events_tables = BSMs_w_sourceAreas$clado_events_tables
 ana_events_tables = BSMs_w_sourceAreas$ana_events_tables
 
 # Count all anagenetic and cladogenetic events
-counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables, areanames, actual_names)
+counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables,
+                                     areanames, actual_names)
 
 summary_counts_BSMs = counts_list$summary_counts_BSMs
 print(conditional_format_table(summary_counts_BSMs))
@@ -252,12 +252,77 @@ print(conditional_format_table(summary_counts_BSMs))
 hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
 
 
+# Count individual stuff --------------------------------------------------
+
+### Sum up dispersal event to different locations ###
+
+# Name each column by letter of each biome LETTERS[1:length(number_of_areas))]
+colnames(counts_list$anagenetic_dispersals_counts_cube) <- LETTERS[1:9]
+
+# Find sum of each column (to get gross number of dispersal TO each biome per stochastic map)
[…]

### 2022-11-16 manuscript.Rmd
rewrote some parts of intro, included more info on Biogeobears and GeoHiSSE methods.

@@
 
-Present-day distrubtion of extant species offers clues about their biogeographic history and how the lineages may have diversified.
+Present-day distrubtion of extant species offers clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
 
-The aridification on continental Australia over the past 20 million years resulted in a diversity of habitats and spurred multiple radiations of flora and fauna across the continent, many of which are endemic [@byrneBirth2008;@melvilleEvolutionary2011].
-The arid zone makes up the majority of the Australian landscape and provides habitat to ~43% of all Australia’s terrestrial squamate reptiles (lizards and snakes).
-Reptiles represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified throughout the Miocene (starting ~25 MYA) to the present [@oliverPhylogenetic2017].
+The aridification on continental Australia began around 20 million years ago with pockets of arid zone developing and subsequent extreme warming and cooling periods in last ~5 million years creating what is now known as the 'outback' [@byrneBirth2008].
+From there arid-adapted vegetation expanded and the mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
+This period of aridification is believed to involve rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+
+Today, the arid zone covers the majority of area in the Australian landscape (70%) and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
+Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified in the last 20 - 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
+
+Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [[@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches .
+
+Other clades like geckos and blind snakes remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@marinHidden2013;@moritzCryptic2018;@zozayaAre2019].
+
+The species richness and distribution of squamate reptiles across different biomes makes them a good group to test hypotheses about the tempo and mode of diversification in Australia.
+
 For example, agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013] are reptiles that colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene.
@@
 Roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021].
-All _Anilios_ species are found in Australia except for _A.
-eryncinus_, which is only found in Papua New Guinea.
-While morphologically conserved and difficult to distinguished, extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@ellisTyphlopid2019].
-The diversification is reported to have originated from a single colonisation event by a non-arid adapted ancestor from Asia between 19-39 mya [@vidalBlindsnake2010;@marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
+All currently described species of _Anilios_ are found in Australia except for _A.
+eryncinus_, which is endemic to New Guinea.
+Blind snakes are morphologically conserved and often difficult to distinguis.
+Extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
 
-However, the study did not account for alternative biogeographic areas (e.g.
-WWF Ecoregions) and models, or uncertainty in phylogenetic relationships and divergence times.
+The diversification of _Anilios_ is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19-39 mya [@vidalBlindsnake2010;@marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
+However, how and
+
+However, the study did not account for alternative biogeographic areas (e.g. WWF Ecoregions) and models, or uncertainty in phylogenetic relationships and divergence times.
+
 Despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
@@
 
-## Biogeography
+## Biogeography and biome transitions
 
-To estimate the ancestral biome of blind snakes in Australia, we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
+To investigate how the distribution of biome evolved in blind snakes we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible likelihood-based framework that models how the branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
@@
 The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range each species are found, and a historical biogeographic model.
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to specific biome(s) based on their occurrence records as described above.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records.
 Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
@@
 We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the model that fitted the data best based on corrected Akaike Information Criterion (AICc) scores.
+Based on the best model we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and trasition between biomes, and to estimate the frequency and directionality of dispersal events across biomes [@mat.
 
@@
 
-## Biogeographic-dependent diversification analyses
+## Biogeographic-dependent speciation
 
-We further investigated whether broad patterns of diversification are associated with geographic range.
+We further investigated whether diversification rates and biome transitions differ between geographic areas.
 Using the R package `hisse` v.
 2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction (SSE) models under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the phylogeny and geographic range data.
-Since SSE models only allow three states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
+Since SSE models only allow two states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
 Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
@@
 
-## Statistical analyses
-Statistical analyses were run in R version 4.2 (CITE R) and Julia (CITE).
-We visualised dynamics using baseR and `ggplot2` in R.
+## Data visualisation
+Base data visualisations were performed in run in R version 4.2 (CITE R) using baseR and `ggplot2` and post-processed in Adobe Illustrator.
 The Australian bioregion map visualised in `QGIS` v (CITE Australian Map).
@@
 
+Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
+
+Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa.
+Two pulses of speciation early Miocene and then another early Pleistocene.
+Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have a hard time crossing the lines [@bryantLines2016]
+
 ## Drivers of diversifcation 

### 2022-11-17 code/04_BioGeoBEARS_bsm_plots.R
init plot

@@ -1,0 +1,248 @@
+# 04_BioGeoBEARS_bsm_plots.R
+# Nov 2022
+# Plotting figures from 50 BioGeoBEARS biostochastic maps
+
+
+# Libraries ---------------------------------------------------------------
+library(dplyr)
+library(stringr)
+library(patchwork)
+
+# Load data ---------------------------------------------------------------
+load('data/intermediate_data/bears/BSB_50.Rdata')
+
+
+
+# EXTRACT BSM OUTPUT ------------------------------------------------------
+
+# Range-switching dispersal (all observed 'a' dispersals): fixed to 0 in DEC models
+# Range-expansion dispersal (all observed 'd' dispersals): controlled by the parameter 'd', which is freely estimated in DEC and related models.
+                                                         # If the d parameter is 0, these counts are zero, but otherwise there should be a few events.
+
+# Anagenetic dispersal (mean of all observed anagenetic 'a' or 'd' dispersals): since a = 0, it's just d
+# Cladogenetic dispersal (mean of all observed jump 'j' dispersals): founder event 'j '
+
+# ALL dispersal (mean of all observed anagenetic 'a', 'd' dispersals, PLUS cladogenetic founder/jump dispersal):
+
+summary_counts_BSMs = counts_list$summary_counts_BSMs
+
+# Summary of BSM
+conditional_format_table(summary_counts_BSMs)
+
+# Histogram of event counts
+hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
+
+# Count individual stuff --------------------------------------------------
+
+### Sum up dispersal event to different locations ###
+
+# Change row and column names to be more informative (from and to)
+dimnames(counts_list$anagenetic_dispersals_counts_cube) <- list(paste0('from', LETTERS[1:9]), paste0('to', LETTERS[1:9]), 1:50)
+
+# Find sum of each column (to get gross number of dispersal TO each biome per stochastic map)
+sum_disp_area <- rowSums(counts_list$anagenetic_dispersals_counts_cube, dims = 2)
+
+counts_list$anagenetic_dispersals_counts_cube
+
[…]

### 2022-11-17 code/04_BioGeoBEARS_results_bsm.R
move plotting section to new script

@@ -13,7 +13,6 @@ library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
 # library(dplyr)
 library(parallel); library(snow)
 
-
 # Tree --------------------------------------------------------------------
 
 tree.c <- ape::read.tree('data/intermediate_data/bears/blindsnake_b.tre')
@@ -22,7 +21,6 @@ tree.c$edge.length <- tree.c$edge.length * 100
 # Saved results -----------------------------------------------------------
 
 load("data/intermediate_data/bears/bears_model_fit.Rdata")
-
 
 ############################
 #####Stochastic mapping#####
@@ -58,9 +56,6 @@ plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"
 dev.off()  # Turn off PDF
 cmdstr = paste("open ", pdffn, sep="")
 system(cmdstr) # Plot it
-
-
-
 
 #######################################################
 # Stochastic mapping on DEC+J+x
@@ -232,7 +227,7 @@ actual_names
 dmat_times = get_dmat_times_from_res(res=res, numstates=NULL)
 dmat_times
 
-# Extract BSM output
+
 clado_events_tables = BSM_output$RES_clado_events_tables
 ana_events_tables = BSM_output$RES_ana_events_tables
 
@@ -244,169 +239,6 @@ ana_events_tables = BSMs_w_sourceAreas$ana_events_tables
 # Count all anagenetic and cladogenetic events
 counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables,
                                      areanames, actual_names)
-
-summary_counts_BSMs = counts_list$summary_counts_BSMs
-print(conditional_format_table(summary_counts_BSMs))
-
-# Histogram of event counts
-hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
-
-
-# Count individual stuff --------------------------------------------------
-
-### Sum up dispersal event to different locations ###
-
-# Name each column by letter of each biome LETTERS[1:length(number_of_areas))]
-colnames(counts_list$anagenetic_dispersals_counts_cube) <- LETTERS[1:9]
-
[…]

### 2022-11-17 code/04_BioGeoBEARS_bsm_plots.R
notes about clado and ana genesis

@@ -15,7 +15,10 @@ load('data/intermediate_data/bears/BSB_50.Rdata')
 
 # EXTRACT BSM OUTPUT ------------------------------------------------------
 
-# Range-switching dispersal (all observed 'a' dispersals): fixed to 0 in DEC models
+# For DEC+j+x 
+## CLADOGENETIC Events -- sympatry (subset) + vicariance (range splitting)
+## ANAGETIC Events -- Founder (jump j parameter) + extinction
+
 # Range-expansion dispersal (all observed 'd' dispersals): controlled by the parameter 'd', which is freely estimated in DEC and related models.
                                                          # If the d parameter is 0, these counts are zero, but otherwise there should be a few events.
 
@@ -23,7 +26,6 @@ load('data/intermediate_data/bears/BSB_50.Rdata')
 # Cladogenetic dispersal (mean of all observed jump 'j' dispersals): founder event 'j '
 
 # ALL dispersal (mean of all observed anagenetic 'a', 'd' dispersals, PLUS cladogenetic founder/jump dispersal):
-
 summary_counts_BSMs = counts_list$summary_counts_BSMs
 
 # Summary of BSM
@@ -145,7 +147,7 @@ inwards_dispersal_grp <- ggplot(data = inward_disp_area, aes(fill= subset_from, 
 ### stitch together with patchwork package
 
 # Save PDF to be included as inset in BioGeoBEARS plot
-pdf(file = 'output/supp_biogeobears_dispersal.pdf', width = 11.33, height = 8.5)
+pdf(file = 'output/biogeobears_dispersal_freq.pdf', width = 11.5, height = 5.7)
 outward_dispersal_grp +inwards_dispersal_grp
 dev.off()
 

### 2022-11-17 manuscript.Rmd
modified parts of biogeobear results

@@
 
-Based on our best fitting biogeographic model, the constrained-dispersal matrix DEC+_j_+_x_ model, we inferred tropical grassland biome to be ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears).
-Likelihood ratio test (LRT) indicates that founder event (parameter *j*) and distance matrix (free parameter *x*) in the more complex model provides statistically significant improvement over the simpler DEC, DEC+*j*, and DEC+*x* models (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-Our result also supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the continent [@marinTracing2013], but exactly where they lived prior to this invasion is equivocal between the Bismarck Archipelago and New Guinea.
-After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
-Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 
+Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+We inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears), but the origin of crown *Anilios* is ambiguous.
 
@@
 - Our result — that the ancestor of Anilios occupied tropical grasslands – is consistent with the result presented in @marinTracing2013.
+
+- Our result also supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the continent [@marinTracing2013], but exactly where they lived prior to this invasion is equivocal between the Bismarck Archipelago and New Guinea.
+
+After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
+Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+
 - We found transitions from XXX to XXX to be common as have been shown / as opposed to other Australian squamates (CITATIONS)

### 2022-11-18 manuscript.Rmd
update talk about biome transitions

@@
 
-## Biogeographic history
+## Biogeographic history and biome transitions
 
@@
 Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-We inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears), but the origin of crown *Anilios* is ambiguous.
-
-## Biome transitions and current distributions
+Our best model inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
 
 Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly.
-Descendents of the earliest split remained in the Tropical grassland biome but within a million years later another split occured.
-The earliest biome transition was from tropical grassland into Temperate Forest around 25 MYA (95% HPD: 28 – 22 MYA) and the Arid zone 24 MYA (95% HPD: 28 – 21 MYA).
-Subsequent splits show transitions from the Arid zone into the Mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears).
+Descendents of the earliest split remained in the tropical grassland biome but within a million years later another split occured.
+The earliest biome transition was from tropical grassland biome into the temperate forest biome around 25 MYA (95% HPD: 28 – 22 MYA) for branch leading _A.
+siliva_, and another branch that included current arid and mediterranean species starting 24 MYA (95% HPD: 28 – 21 MYA).
+Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (642 dispersals, mean = 71.3 per BSM) almost double the number of dispersals from the arid biome (395 dispersals, mean = 43.9 per BSM) (\@ref(fig:fig-biogeobears)).
+
+The top recipients of dispersals from tropical grassland biome is the temperate grassland (5.06 ± 0.79), follotropical forest (4.12 ± 1.29), ,
+
+Subsequent splits show transitions from the arid biome into the mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears).
 Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome.
@@
 
+- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands (e.g. _A. longissimus_)
+
 - Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].

### 2022-11-18 manuscript.Rmd
biome transitions

@@
 siliva_, and another branch that included current arid and mediterranean species starting 24 MYA (95% HPD: 28 – 21 MYA).
-Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (642 dispersals, mean = 71.3 per BSM) almost double the number of dispersals from the arid biome (395 dispersals, mean = 43.9 per BSM) (\@ref(fig:fig-biogeobears)).
+Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (mean = XX) almost double the number of dispersals from the arid biome (mean = XX) (\@ref(fig:fig-biogeobears)).
+On the other hand, the largest sink for all inward dispersal is the mediterranean biome (
 

### 2022-11-18 code/04_BioGeoBEARS_bsm_plots.R
plot means instead of raw numberes

@@ -7,11 +7,10 @@
 library(dplyr)
 library(stringr)
 library(patchwork)
+library(corrplot)
 
 # Load data ---------------------------------------------------------------
 load('data/intermediate_data/bears/BSB_50.Rdata')
-
-
 
 # EXTRACT BSM OUTPUT ------------------------------------------------------
 
@@ -70,6 +69,56 @@ biome_colours <- c("#dcc674", #0
                    "#92d1bd", #7
                    "#eaf400")
 
+# Mean transitions
+mean_in_disp <- counts_list$all_dispersals_counts_fromto_means 
+
+# change row names
+mean_in_disp$to <- paste0("to", rownames(mean_in_disp))
+
+
+mean_in_disp_area <- mean_in_disp %>% 
+  tidyr::pivot_longer(!to, names_to = "from", values_to = "count") %>% 
+  dplyr::mutate(state_from = ifelse(from == "C", "arid", "non_arid")) %>% 
+  dplyr::mutate(state_to = ifelse(to == "toC", "arid", "non_arid")) %>% 
+  dplyr::mutate(subset_from = ifelse(from %in% c("G", "H", "I"), "island", from)) %>% 
+  dplyr::mutate(subset_to = ifelse(to %in% c("toG", "toH", "toI"), "tozisland", to))
+
+# outwards dispersals 
+outwards_disp <- ggplot(data = inward_disp_area, aes(fill= from, y = count, x = to)) + 
+  geom_bar(position="stack", stat="identity") +
+  labs(x = "Inward dispersal", y = "frequency") + 
+  scale_x_discrete(labels = biome_labs) + 
+  scale_fill_manual(labels = biome_labs, values = biome_colours) +
+  guides(fill = guide_legend(title = "From")) +
+  theme(legend.position = "none")
+
+# outwards dispersals geo state
+outwards_disp_geo <- ggplot(data = inward_disp_area, aes(fill= state_from, y = count, x = state_to)) + 
+  geom_bar(position="stack", stat="identity") +
+  labs(x = "Inward dispersal", y = "frequency") + 
+  scale_fill_manual(labels = c('arid', 'non-arid'), values = c(biome_colours[3], "grey")) +
+  guides(fill = guide_legend(title = "From")) +
+  theme(legend.position = "bottom")
+
+# outwards dispersals grp
[…]

### 2022-11-21 manuscript.Rmd
values from BSM into results section

--- 2022-11-18 manuscript.Rmd
@@
 Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-Our best model inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
-
-Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly.
+Based on 50 biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
+We estimated an average of 71.2 ± 1.3 total events (anagenesis and cladogenesis) occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2±1.3 events) and founder events (6.1±1.5 events) (see supp-summary_counts_bsm_table).
+We found that the largest source of outward dispersal is the tropical grassland biome (16.2±4.9 events mean±SD; 43.6%) making up almost double the number of dispersals from the arid biome (9.7±3.2 events; 26.0%) (\@ref(fig:fig-biogeobears)).
+On the other hand, the largest sink for all inward dispersal are the temperate grassland biome (7.9±4.2 events; 21.2%) and the mediterranean biome (7.0±2.9 events; 18.8%).
+The largest directional transition is from the arid biome to the mediterranean biome (5.3±1.5 events) followed by transitions from the tropical grassland into temperate forest biome (5.1±1.3 events).
+Following the hypothesized colonisation event from a single common ancestor, the diversification of _Anilios_ occurred very rapidly.
 Descendents of the earliest split remained in the tropical grassland biome but within a million years later another split occured.
@@
 siliva_, and another branch that included current arid and mediterranean species starting 24 MYA (95% HPD: 28 – 21 MYA).
-Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (642 dispersals, mean = 71.3 per BSM) almost double the number of dispersals from the arid biome (395 dispersals, mean = 43.9 per BSM) (\@ref(fig:fig-biogeobears)).
-
-The top recipients of dispersals from tropical grassland biome is the temperate grassland (5.06 ± 0.79), follotropical forest (4.12 ± 1.29), ,
-
 Subsequent splits show transitions from the arid biome into the mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears).
-Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome.
+Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the arid and mediterranean biomes.
 We observed that closely related species tend to share the same biome and occur around the same geographical areas.

### 2022-11-21 code/04_BioGeoBEARS_bsm_plots.R
code to inform BSM values in manuscript results

@@ -8,6 +8,9 @@ library(dplyr)
 library(stringr)
 library(patchwork)
 library(corrplot)
+library(GenSA);library(FD)      
+library(rexpokit);library(cladoRcpp); library(BioGeoBEARS)
+library(ggplot2)
 
 # Load data ---------------------------------------------------------------
 load('data/intermediate_data/bears/BSB_50.Rdata')
@@ -31,30 +34,75 @@ summary_counts_BSMs = counts_list$summary_counts_BSMs
 conditional_format_table(summary_counts_BSMs)
 
 # Histogram of event counts
-hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
+# hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
 
-# Count individual stuff --------------------------------------------------
+counts_list$all_dispersals_counts_fromto_means
+rowSums(counts_list$all_dispersals_counts_fromto_means)
+rowSums(counts_list$all_dispersals_counts_fromto_sds)
 
-### Sum up dispersal event to different locations ###
 
-# Change row and column names to be more informative (from and to)
-dimnames(counts_list$anagenetic_dispersals_counts_cube) <- list(paste0('from', LETTERS[1:9]), paste0('to', LETTERS[1:9]), 1:50)
+df %>% 
+  mutate(tot = sum(Freq)) %>% 
+  group_by(Var2) %>% # For each Gr in Var2 
+  summarise(Freq = sum(Freq)) %>% 
+  mutate(Prop = 100*Freq/sum(Freq))
 
-# Find sum of each column (to get gross number of dispersal TO each biome per stochastic map)
-sum_disp_area <- rowSums(counts_list$anagenetic_dispersals_counts_cube, dims = 2)
+ggplot(data = df,
+       aes(x = Var2, y = Freq)) +
+  geom_bar(stat = "identity",
+           aes(fill = Var1)) +
+  geom_text(data = df2,
+            aes(y = Freq + 1,
+                label = sprintf('%.2f%%', Prop)))
 
-counts_list$anagenetic_dispersals_counts_cube
 
-rownames(sum_disp_area) <- LETTERS[1:9]
-sum_disp_area_df <- data.frame(sum_disp_area)
-sum_disp_area_df$from <- rownames(sum_disp_area_df)
 
[…]

### 2022-11-21 manuscript.Rmd
annotate where data came from

@@
 
+<!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
 We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 

### 2022-11-21 code/04_BioGeoBEARS_bsm_plots.R
clean up code for plotting

@@ -37,24 +37,17 @@ conditional_format_table(summary_counts_BSMs)
 # hist_event_counts(counts_list, pdffn=paste0("output/", model_name, "_histograms_of_event_counts.pdf"))
 
 counts_list$all_dispersals_counts_fromto_means
+counts_list$all_dispersals_counts_fromto_sds
+
+
+## Largest outwards dispersal
 rowSums(counts_list$all_dispersals_counts_fromto_means)
 rowSums(counts_list$all_dispersals_counts_fromto_sds)
 
+## Largest inwards dispersals
 
-df %>% 
-  mutate(tot = sum(Freq)) %>% 
-  group_by(Var2) %>% # For each Gr in Var2 
-  summarise(Freq = sum(Freq)) %>% 
-  mutate(Prop = 100*Freq/sum(Freq))
-
-ggplot(data = df,
-       aes(x = Var2, y = Freq)) +
-  geom_bar(stat = "identity",
-           aes(fill = Var1)) +
-  geom_text(data = df2,
-            aes(y = Freq + 1,
-                label = sprintf('%.2f%%', Prop)))
-
+rowSums(t(counts_list$all_dispersals_counts_fromto_means))
+rowSums(t(counts_list$all_dispersals_counts_fromto_sds))
 
 
 # Mean transitions --------------------------------------------------------
@@ -106,7 +99,7 @@ mean_in_disp_prop <- mean_in_disp_area %>%
 
 # x-labels
 biome_labs <- c("Trop Grass", "Temp Forest", "Arid", "Mediterranean", "Trop Forest", "Temp Grass", "Sunda", "N Guinea", "Bismarck Arc")
-biome_values <- unique(outward_disp_df$to)
+biome_values <- unique(mean_in_disp_prop$to)
 biome_colours <- c("#dcc674", #0
                    "#1f4733", #1
                    "#e03f28", #2
@@ -125,8 +118,8 @@ outwards_disp <- ggplot(data = mean_out_disp_area, aes(fill= to, y = count, x = 
   labs(x = "Outwards dispersal", y = "Mean outwards transition frequency") + 
   scale_x_discrete(labels = biome_labs) + 
   scale_fill_manual(labels = biome_labs, values = biome_colours) +
-  guides(fill = guide_legend(title = "From")) +
-  theme(legend.position = "none")
+  guides(fill = guide_legend(title = "From")) 
+  # theme(legend.position = "none")
 
 # outwards dispersals geo state
[…]

### 2022-11-21 code/04_BioGeoBEARS_results_bsm.R
try with time slicing

@@ -238,7 +238,15 @@ ana_events_tables = BSMs_w_sourceAreas$ana_events_tables
 
 # Count all anagenetic and cladogenetic events
 counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables,
-                                     areanames, actual_names)
+                                     areanames, actual_names, timeperiod = c(0,0.1))
+
+count_ana_clado_events(clado_events_tables, ana_events_tables,
+                       areanames, actual_names)
+
+counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables,
+                                     areanames, actual_names, timeperiod = c(5,10))
+
+
 
 #######################################################
 # Print counts to files

### 2022-11-24 code/05_GeoHiSSE_fit.R
only include four models

@@ -191,8 +191,12 @@ mod5 <- GeoHiSSE(phy = phy, data = state, f=fraction,
 # Models with higher weight show better fit to the data and, as a result, 
 # have more weight when performing model averaging.
 
+# hisse::GetAICWeights(list(model1 = mod1, model2 = mod2, model3 = mod3, model4 = mod4, model5 = mod5), criterion="AICc")
+
+hisse::GetAICWeights(list(model1 = mod1, model2 = mod2, model3 = mod3, model4 = mod4), criterion="AIC")
+
 hisse::GetAICWeights(list(model1 = mod1, model2 = mod2, model3 = mod3,
-                          model4 = mod4, model5 = mod5), criterion="AICc")
+                          model4 = mod4), criterion="AIC")
 
 
 ### Model averaging and plotting

### 2022-11-24 manuscript.Rmd
add results from GeoHiSSE

--- 2022-11-18 manuscript.Rmd
@@
 We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the model that fitted the data best based on corrected Akaike Information Criterion (AICc) scores.
-Based on the best model we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and trasition between biomes, and to estimate the frequency and directionality of dispersal events across biomes [@mat.
+Based on the best model we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and trasition between biomes, and to estimate the frequency and directionality of dispersal events across biomes [@matzkeBioGeoBEARS2013].
+
+## Biogeographic-dependent speciation
+
+We further investigated whether diversification rates and biome transitions differ between arid and non-arid biomes.
+Using the R package `hisse` v.
+2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction (SSE) models under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the phylogeny and geographic range data.
+Since SSE models only allow two states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
+Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
+We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas based on number of species in the phylogeny and all currently described species of _Anilios_.
+The five SSE models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved 'hidden states.' Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
+Model 3 and 4 are derivations of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
+The last model is a complementary model that removes cladogenetic effect entirely.
+We compared the AIC scores to select the best models.
+Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
 
@@
 
-## Biogeographic-dependent speciation
-
-We further investigated whether diversification rates and biome transitions differ between geographic areas.
-Using the R package `hisse` v.
-2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction (SSE) models under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the phylogeny and geographic range data.
-Since SSE models only allow two states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
-Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
-We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas based on number of species in the phylogeny and all currently described species of _Anilios_.
-The five SSE models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved 'hidden states'.
-Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
-Model 3 and 4 are derivation of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
-The last model is a complementary model that removes cladogenetic effect entirely.
-We compared the AIC scores to select the best models.
-Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
-
 <!-- ## Niche evolution and spatial data -->
@@
 
+<!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
+We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
 We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 
 Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-Our best model inferred tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
 
-Following the hypothesized colonisation event from a single common ancestor, the diversifcation of _Anilios_ occurred very rapidly.
-Descendents of the earliest split remained in the tropical grassland biome but within a million years later another split occured.
-The earliest biome transition was from tropical grassland biome into the temperate forest biome around 25 MYA (95% HPD: 28 – 22 MYA) for branch leading _A.
-siliva_, and another branch that included current arid and mediterranean species starting 24 MYA (95% HPD: 28 – 21 MYA).
-Based on 50 biogeographic stochastic mapping, we found that the largest source of total outward dispersal (anagenetic and cladogenetic) is the tropical grassland biome (mean = XX) almost double the number of dispersals from the arid biome (mean = XX) (\@ref(fig:fig-biogeobears)).
-On the other hand, the largest sink for all inward dispersal is the mediterranean biome (
+Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
+We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
+We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
+Together the tropical grassland biome and the arid biome make up the majority (69.6%) of outward transitions.
 
-The top recipients of dispersals from tropical grassland biome is the temperate grassland (5.06 ± 0.79), follotropical forest (4.12 ± 1.29), ,
+On the other hand, the largest sink for all inward dispersal are the temperate grassland biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
+The largest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
 
-Subsequent splits show transitions from the arid biome into the mediterranean biome starting after 15 MYA (\@ref(fig:fig-biogeobears).
-Current distributions of _Anilios_ are mostly concentrated in the tropical grassland biome closely followed by the Arid zone and Mediterranean biome.
-We observed that closely related species tend to share the same biome and occur around the same geographical areas.
+We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
+We found that the model where where geographic range was not linked to variation in diversification rates but rather to a hidden unobserved state fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
+The best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see suupp-table-geohiss).
+The weighted average parameter estimates indicate both diversification rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE).
+
+Results from our ClaDS analysis is consistent with that of GeoHiSSE, showing branch-specific diversification rates slow down towards the present.
+
+The analysis shows slow down in diversification rates over time.
 
@@
 Both aridity index shows increase in aridity post 10 MYA, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
-Paleoclimate estimates for Australia differed between the three methods \@ref(fig:fig-paleoclimate-compare)b.
+Paleoclimate estimates for Australia differed between the three methods (\@ref(fig:fig-paleoclimate-compare))b.
 However, all three paleo temperature models indicates that Australia was hotter than the average global temperature.
@@
 ## Rates of diversification
-
-Rates of diversification between arid and non-arid, and widespread.
 

### 2022-11-24 code/04_BioGeoBEARS_bsm_plots.R
code to show percentage from each dispersal out/in

@@ -130,18 +130,18 @@ outwards_disp_geo <- ggplot(data = mean_in_disp_area, aes(fill= state_to, y = co
   theme(legend.position = "bottom")
 
 # outwards dispersals grp
-outwards_disp_grp <- ggplot(data = mean_in_disp_area, aes(y = count, x = subset_from)) + 
+outwards_disp_grp <- ggplot(data = mean_in_disp_area, aes(y = count, fill= subset_to, x = subset_from)) + 
   geom_bar(position="stack", stat="identity") +
   labs(x = "Outwards dispersal", y = "Mean outwards transition frequency") + 
   scale_x_discrete(labels = c(biome_labs[1:6], 'islands')) +
   scale_fill_manual(values = c(biome_colours[1:6], 'grey')) +
   guides(fill = guide_legend(title = "From")) +
   theme(legend.position = "none") +
-  geom_text(data = mean_out_disp_prop,
-            aes(y = Freq + 1, x = subset_from,
-                label = sprintf('%.2f%%', Prop)))
+  # geom_text(data = mean_out_disp_prop,
+  #           aes(y = Freq + 1, x = subset_from,
+  #               label = sprintf('%.2f%%', Prop)))
 
-  # geom_text(aes(label = count), position = position_stack(vjust = 0.5))
+  geom_text(aes(label = count), position = position_stack(vjust = 0.5))
 
 ### Inward transitions
 

### 2022-11-24 code/04_BioGeoBEARS_results_bsm.R
bsm change width height and pie

@@ -191,7 +191,7 @@ stratified = stratified
 
 # Loop through the maps and plot to PDF
 pdffn = paste0("output/", model_name, "_", length(clado_events_tables), "BSMs_v1.pdf")
-pdf(file=pdffn, width=6, height=6)
+pdf(file=pdffn, width=11.33, height=7.5)
 
 nummaps_goal = 50
 for (i in 1:nummaps_goal)
@@ -199,7 +199,7 @@ for (i in 1:nummaps_goal)
   clado_events_table = clado_events_tables[[i]]
   analysis_titletxt = paste0(model_name, " - Stochastic Map #", i, "/", nummaps_goal)
   plot_BSM(results_object=res, clado_events_table=clado_events_table, stratified=stratified, analysis_titletxt=analysis_titletxt, addl_params=list("j", "x"),
-           label.offset=0.5, plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, show.tip.label=TRUE, 
+           label.offset=0.5, plotwhat="pie", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, show.tip.label=TRUE, 
            include_null_range=include_null_range)
 } # END for (i in 1:nummaps_goal)
 

### 2022-11-29 manuscript.Rmd
small update

@@
 We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
-We found that the model where where geographic range was not linked to variation in diversification rates but rather to a hidden unobserved state fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
-The best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see suupp-table-geohiss).
-The weighted average parameter estimates indicate both diversification rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE).
+We found that the model where variation in diversification rates were not linked to geographic range fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
+The best-fitting geographic range independent model included a hidden unobserved state, indicating diversification rates is better explained by a hidden unobserved state.
+Compared to the three other models ,the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
+The weighted average parameter estimates from all models indicate both diversification rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE).
+However, net diversification rates for the widespread lineages were slightly higher (see figure), but this may be due to low amount of lineages.
 
-Results from our ClaDS analysis is consistent with that of GeoHiSSE, showing branch-specific diversification rates slow down towards the present.
-
-The analysis shows slow down in diversification rates over time.
+Results from our ClaDS analysis show that variation in branch-specific diversification rates generally slowed down towards the present (fig-ClaDS), and did not differ between arid, non-arid, and widespread lineages (fig-ClaDS).
 
@@
 The second and third most competitive models (delta AICc < 2) also found increase aridity is exponentially associated with decrease speciation rate but without estimating extinction (Table-RPANDA-TOP10).
-Among exponential environment-dependent models fitted to our blind snake phylogeny, we found that models that uses Australia-specific paleoclimate data received better AICc support compared to fit with global paleoclimate (see supp-TABLE_FULL_RPANDA_model_fit).
+Among exponential environment-dependent models, we found Australia-specific paleoclimate data provides a better fit to our phylogeny than global paleoclimate (see supp-TABLE_FULL_RPANDA_model_fit).
 
@@
 Discuss species that are far apart...
-- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
 - Speciation could even be sympatric but no evidence to support this.
 
-## Niche evolution
+- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation.
+
+## Environment-dependent diversification
+- Diversification rate decrease with aridification. 
+- Our RPANDA model did not model environment-dependent extinction. 
+- Extinction is a big cavaet especially because we suspect there are more species
+-
 

### 2022-11-29 code/05_GeoHiSSE_fit.R
remove 5th model

@@ -229,7 +229,7 @@ recon.mod5 <- MarginReconGeoSSE(phy = mod5$phy, data = mod5$data, f = mod5$f,
 
 # Marginal reconstruction computed above is used to compute the model average.
 
-recon.models <- list(recon.mod1, recon.mod2, recon.mod3, recon.mod4, recon.mod5)
+recon.models <- list(recon.mod1, recon.mod2, recon.mod3, recon.mod4)
 
 # save.image(file = 'data/intermediate_data/geohisse/geohisse_three_states.Rdata')
 load('data/intermediate_data/geohisse/geohisse_three_states.Rdata')
@@ -347,10 +347,15 @@ plot.geohisse.states(x = recon.models, rate.param = "turnover", type = "fan",
 
 # AIC ---------------------------------------------------------------------
 
-recon.mod1$AIC; recon.mod2$AIC; recon.mod3$AIC; recon.mod4$AIC; recon.mod5$AIC
-mod1$AIC; mod2$AIC; mod3$AIC; mod4$AIC; mod5$AIC
+# recon.mod1$AIC; recon.mod2$AIC; recon.mod3$AIC; recon.mod4$AIC; recon.mod5$AIC
+recon.mod1$AIC; recon.mod2$AIC; recon.mod3$AIC; recon.mod4$AIC
+# mod1$AIC; mod2$AIC; mod3$AIC; mod4$AIC; mod5$AIC
+mod1$AIC; mod2$AIC; mod3$AIC; mod4$AIC
 
-mod1$AICc; mod2$AICc; mod3$AICc; mod4$AICc; mod5$AICc
+mod1$AICc; mod2$AICc; mod3$AICc; mod4$AICc
+
+
+
 
 recon.mod1$rates.mat
 recon.mod2$rates.mat

### 2022-11-29 code/05_GeoHiSSE_plotting-output.R
Jenna's plotting code

@@ -1,0 +1,556 @@
+# This is an R script to plot output from log files of GeoHiSSE/MuSSE/MuHiSSE analyses
+# presented in McCullough et al. 2022 systematic biology. 
+
+# written by Jenna McCullough and Rosana Zenil-Ferguson 
+# last edited on 15 June 2022 by Jenna McCullough 
+
+# Modified by Putter Tiatragul
+# Last modified November 2022
+
+library(devtools)
+install_github("revbayes/RevGadgets")
+library(RevGadgets)
+library(ggplot2)
+
+#multiplot.R needs to be in your working directory for this to work. 
+source("code/utility/multiplot.R")
+
+#Remember, colours need to be in alphabetical order of your parameters!
+
+###########################################
+###########################################
+# Model 1 ANALYSIS: GEOHISSE ISLANDS
+###########################################
+###########################################
+islands_1.sse<-read.table("data/intermediate_data/geohisse/00_GeoHiSSE-blindsnake/output/comb_geohisse_blindsnake.log", header=TRUE)
+head(islands_1.sse)
+str(islands_1.sse)
+
+#anagenetic dispersal for model 1 
+anagenetic.disp_1<-data.frame(dens=c(islands_1.sse$anagenetic_dispersal_21, islands_1.sse$anagenetic_dispersal_31, islands_1.sse$anagenetic_dispersal_54, 
+                                     islands_1.sse$anagenetic_dispersal_64),Type=rep(c("Continent to widespread A","Island to widespread A","Continent to widespread B","Island to widespread B"), 
+                                                                                     each=length(islands_1.sse$anagenetic_dispersal_21)))
+anagenetic.colors=c("#336600","#330033","#66CC66","#996699") 
+
+#Continents A = #336600 
+#Islands A = #66CC66 
+#Continents B = #330033
+#Islands B = #996699 
+
+p1_1<-ggplot(anagenetic.disp_1, aes(x=dens, fill=Type))+labs(title="Anagenetic dispersal for Islands v. continents",x="Rate", y="Posterior Density")+geom_density(alpha=0.7)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+             panel.background = element_blank(), axis.line = element_line(colour = "black"))+scale_fill_manual( values =anagenetic.colors )
+p1_1
+
+#Extinction for model 1 
[…]

### 2022-11-30 manuscript.Rmd
minor updates

@@
 <!-- Consult Alex Skeels -->
-To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES].
+To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation].
 Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution.
@@
 
-## Biogeographic history and biome transitions
+## Ancestral biome reconstruction and biogeographic history
 
 <!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
-We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
+
+All BioGeoBEARS analyses inferred tropical grassland as the ancestral biome of *Anilios*.
 We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
@@
 
-Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous.
+We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
+
+Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, while we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), the ancestor leading to crown *Anilios* is ambiguous.
 We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
 We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
-Together the tropical grassland biome and the arid biome make up the majority (69.6%) of outward transitions.
+Together the tropical grassland biome and the arid biome make up the largest source of diversificttion accounting for 69.6% of outward transitions.
 
@@
 The largest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
+
+## Diversification rates between arid and non-arid biomes
 

### 2022-11-30 code/05_GeoHiSSE_plotting-output.R

@@ -19,49 +19,49 @@ source("code/utility/multiplot.R")
 
 ###########################################
 ###########################################
-# Model 1 ANALYSIS: GEOHISSE ISLANDS
+# Model 1 ANALYSIS: GEOHISSE non_arid
 ###########################################
 ###########################################
-islands_1.sse<-read.table("data/intermediate_data/geohisse/00_GeoHiSSE-blindsnake/output/comb_geohisse_blindsnake.log", header=TRUE)
-head(islands_1.sse)
-str(islands_1.sse)
+non_arid_1.sse<-read.table("data/intermediate_data/geohisse/00_GeoHiSSE-blindsnake/output/comb_geohisse_blindsnake.log", header=TRUE)
+head(non_arid_1.sse)
+str(non_arid_1.sse)
 
 #anagenetic dispersal for model 1 
-anagenetic.disp_1<-data.frame(dens=c(islands_1.sse$anagenetic_dispersal_21, islands_1.sse$anagenetic_dispersal_31, islands_1.sse$anagenetic_dispersal_54, 
-                                     islands_1.sse$anagenetic_dispersal_64),Type=rep(c("Continent to widespread A","Island to widespread A","Continent to widespread B","Island to widespread B"), 
-                                                                                     each=length(islands_1.sse$anagenetic_dispersal_21)))
+anagenetic.disp_1<-data.frame(dens=c(non_arid_1.sse$anagenetic_dispersal_21, non_arid_1.sse$anagenetic_dispersal_31, non_arid_1.sse$anagenetic_dispersal_54, 
+                                     non_arid_1.sse$anagenetic_dispersal_64),Type=rep(c("arid to widespread A","Island to widespread A","arid to widespread B","Island to widespread B"), 
+                                                                                     each=length(non_arid_1.sse$anagenetic_dispersal_21)))
 anagenetic.colors=c("#336600","#330033","#66CC66","#996699") 
 
-#Continents A = #336600 
-#Islands A = #66CC66 
-#Continents B = #330033
-#Islands B = #996699 
+#arids A = #336600 
+#non_arid A = #66CC66 
+#arids B = #330033
+#non_arid B = #996699 
 
[…]

### 2022-11-30 manuscript.Rmd
added some limitation statements and listed questions

@@
 
-In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
-
 <!--Why people should care-->
@@
 
-We 
-- Estimate the ancestral biome of Australian blindsnakes based on extant taxa.
-- Compare fit_env models between localised versus globalised temperature on the diversification of Australian blindsnake radiation, 
-- Assess model fit between paleo-aridity and paleo-climate on diversification rates, and 
-- Are there differences in the rates of speciation between arid and non-arid zones?
+In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
+Specifically, we sought to answer the following questions:
+(1) What is the ancestral biome of Australian _Anilios_? 
+() Which biomes are the largest source and sinks of biodiversity? 
+() What is the relationship between diversification rate and changes in paleo environment? 
+() Which paleo environment best explain current phylogeny?
+() Are there differences in diversification rates between lineages in arid and non-arid biome?
+
+Aims: 
+- Infer the ancestral biome of Australian blindsnakes based on near-complete phylogeny of extant taxa.
+- Investigate how paleo-environmental changes influence rates of diversification. 
+- Assess model fit between paleo-environment.  
+- Compare the rates of diversification between lineages in the arid versus non-arid biome.
+
+- Compare localised versus globalised temperature on the diversification of Australian blindsnake radiation,
 
@@
 
-## Biogeographic-dependent speciation
+## Biogeographic-dependent diversification analyses
 
@@
 
-## Reconstruction of Australian paleo climate
+## Reconstruction of Australian paleo environment
 <!-- Consult Alex Skeels -->
@@
 
-## Time ane environment-dependent diversification analyses
+## Environment-dependent diversification analyses
 
@@
 
-## Ancestral biome reconstruction and biogeographic history
+## Ancestral biome reconstruction and transitions
 
@@
 We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
+
 We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
 Together the tropical grassland biome and the arid biome make up the largest source of diversificttion accounting for 69.6% of outward transitions.
-
 On the other hand, the largest sink for all inward dispersal are the temperate grassland biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
-The largest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
+The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
 
@@
 The best-fitting geographic range independent model included a hidden unobserved state, indicating diversification rates is better explained by a hidden unobserved state.
-Compared to the three other models ,the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
+Compared to the three other models, the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
 The weighted average parameter estimates from all models indicate both diversification rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE).
@@
 Paleoclimate estimates for Australia differed between the three methods (\@ref(fig:fig-paleoclimate-compare))b.
-However, all three paleo temperature models indicates that Australia was hotter than the average global temperature.
-Paleoclimate in Australia were also relatively constant compared to global temperature that shows greater fluctuation then slight decrease towards the present.
+However, all three paleo temperature models indicate Australia was hotter than the average global temperature.
+Paleoclimate in Australia were also relatively constant (i.e.
+less fluctuation) compared to global temperature.
 
@@
 We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) with aridity index derived from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA).
-This model indicates that increase aridity is exponentially associate with decrease speciation rate (lambda = 1.32, alpha = -0.14) and with a constant background extinction (mu = 0.23).
+This best-fitting model indicates that increase aridity is exponentially associated with decrease speciation rate (lambda = 1.32, alpha = -0.14) and a constant background extinction (mu = 0.23).
 The second and third most competitive models (delta AICc < 2) also found increase aridity is exponentially associated with decrease speciation rate but without estimating extinction (Table-RPANDA-TOP10).
 Among exponential environment-dependent models, we found Australia-specific paleoclimate data provides a better fit to our phylogeny than global paleoclimate (see supp-TABLE_FULL_RPANDA_model_fit).
-
-## Rates of diversification
 
@@
 
-- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands (e.g. _A. longissimus_)
-
 - Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
@@
 Discuss species that are far apart...
+- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
 - Speciation could even be sympatric but no evidence to support this.
-
-- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation.
 
@@
 - Diversification rate decrease with aridification.
+But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
 - Our RPANDA model did not model environment-dependent extinction. 
 - Extinction is a big cavaet especially because we suspect there are more species
--
+
+## Limitations
+There are several limitations in our models. 
+- While taxonomic sampling the most representative of the genus, it is not comprehensive.
+We suspect there remains several undescribed species of blind snakes in Australia and New Guinea that may belong in the genus _Anilios_.
+Getting genetic data from _A.
+erycinus_ will allow us to place the species among other _Anilios_ and provide a more comprehensive picture of the evolution of this genus.
+- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands.
+For example *A.
+longissimus* has only been collected from 30m underground on Barrow Island located XX km the northwest coast of Western Australia.
+Placing that species with arid biome is not realistic but that area is considered the arid zone by our criteria.
+
+- Need to account for heterogeneity of the arid zone [@pepperLife2021].
 
 # Conclusion
-
-Account for heterogeneity of the arid zone [@pepperLife2021].
 

### 2022-12-01 manuscript.Rmd
added some discussions

@@
 **Keywords**
-scolecophidia, Diversification, paleoclimate, BioGeoBEARS
+scolecophidia, Diversification, paleoclimate, BioGeoBEARS, GeoHiSSE
 
@@
 
-- Compare localised versus globalised temperature on the diversification of Australian blindsnake radiation,
-
 #  Materials and methods
@@
 
-To investigate how the distribution of biome evolved in blind snakes we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
-BioGeoBEARS provides a flexible likelihood-based framework that models how the branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
+To investigate the biogeographic history of blind snakes we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
+BioGeoBEARS provides a flexible likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
 These shifts in biomes can occur within a single branch (anagenetic events including processes like dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant and founder-event speciation).
@@
 Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
+
 The phylogeny and distribution information were then applied to widely used historical biogeographical methods including likelihood implementations of Dispersal-Vicariance Analysis [@ronquistDispersalVicariance1997] (DIVALIKE), Dispersal-Extinction-Cladogenesis [@reeMaximum2008] (DECLIKE), and BayArea [@landisBayesian2013] (BAYAREALIKE) in `BioGeoBEARS` [@matzkeBioGeoBEARS2013].
@@
 
-## Biogeographic-dependent diversification analyses
-
-We further investigated whether diversification rates and biome transitions differ between arid and non-arid biomes.
-Using the R package `hisse` v.
-2.1.9 [@beaulieuDetecting2016], we fitted five biogeographic-dependent speciation-extinction (SSE) models under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework to the phylogeny and geographic range data.
-Since SSE models only allow two states, we restricted our geographic states to species that have distributions in arid areas, non-arid areas, and those that are widespread across both arid and non-arid areas.
-Consistent with our BioGeoBEARS anaylsis, we classified species that primarily occur in the arid zone as arid (n = 8), species that occur primarily outside of the arid zone as non-arid (n = 24), and those that occupy both the arid and non-arid zone as widespread (n = 5).
-We assumed uniform taxon sampling for 89% in arid areas, 71% in non-arid areas, and 83% in widespread areas based on number of species in the phylogeny and all currently described species of _Anilios_.
-The five SSE models vary by whether they allow diversification rates to be associated with geographic range and/or with unobserved 'hidden states.' Model 1 assumes equal diversification rates regardless of geographic range, model 2 allows geographic range to affect diversification rates (canonical GeoSSE model).
-Model 3 and 4 are derivations of models 1 and 2 but with unobserved hidden states (see Table GeoHiSSE table).
-The last model is a complementary model that removes cladogenetic effect entirely.
-We compared the AIC scores to select the best models.
-Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
-
-## Reconstruction of Australian paleo environment
+## Paleo environment reconstruction
 <!-- Consult Alex Skeels -->
 To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation].
-Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution.
+Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myrs resolution.
 We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
@@
 
---More information about summarising values and determining aridity index ---
+Based on the ____ value, we use XX as the threshold for arid. Thus aridity index is ...
 
-## Environment-dependent diversification analyses
+## Time- and environment-dependent diversification analyses
 
-To test whether and how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` function in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` functions in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleo environment as described above.
 We also fit the model to widely use mean global paleo temperature data set (CITE).
-We accounted for missing taxa using the fractioning function (f = XX) in RPANDA.
+To account for missing taxa, we used the fractioning function (f = XX) in `RPANDA`.
 We then assessed which diversification model best fit the data based on AICc scores.
 
-To visualise the dynamics of the radiation, we use ClaDS [@malietFast2022] implemented in `julia` to estimate speciation rates at the origin of each branch (hencforth branch-specific rates) and at the tip.
-ClaDS is designed to capture frequent shifts in speciation rates with small effects.
+## Geographic range and rates of diversification
+
+To test whether diversification rates differ between lineages in arid biomes and non-arid lienages, we used two approaches.
+First, we estimated speciation rates at the origin of each branch (hencforth branch-specific rates) and tip speciation rates using the Julia package `ClaDS` [@malietFast2022].
+`ClaDS` is designed to capture frequent shifts in speciation rates with small effects.
 At each speciation event, daughter lineages are assigned its own speciation rate sampled from a distribution parameterised on the parental rate [@malietFast2022].
 
-<!-- ## Niche evolution and spatial data -->
-<!-- We reconstructed ancestral niche from present-only data using NEMo package.  -->
-<!-- from the online database [Atlas of the Living Australia](doi.ala.org.au/doi/10.26197/ala.d92678b1-ad2d-437b-9457-9f52737ba003) (see Appendix-method-occurence).  -->
+In the second approach we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package `hisse` v.
+2.1.9 [@beaulieuDetecting2016].
+Since these SSE models only allow two states, we restricted our geographic states to arid and non-arid biomes.
+Based on the same categories we used in our BioGeoBEARS anaylses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily other biomes as non-arid (State 1; n = 24), and those that occupy both the arid and non-arid biomes as widespread (State 01; n = 5).
 
-## Data visualisation
-Base data visualisations were performed in run in R version 4.2 (CITE R) using baseR and `ggplot2` and post-processed in Adobe Illustrator.
-The Australian bioregion map visualised in `QGIS` v (CITE Australian Map).
+The four SSE models vary by whether they allow diversification rates to be associated with geographic states (range-dependent or range-independent) and whether they allow an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
+Model 1 assumes equal diversification rates regardless of geographic state. Model 2 allows geographic range to affect diversification rates (canonical GeoSSE model). 
+Models 3 and 4 are derivations of models 1 and 2 but also allow unmeasured trait ('hidden states') to contribute to diversification rates (see Table GeoHiSSE table).
+By incoporating hidden states, we are able to assess whether geographic range or an unmeasured trait contributes more to the observed diversification.
+For each model, we assumed uniform taxon sampling for 89% in arid lineages, 71% in non-arid lineages, and 83% in widespread lineages based on number of species in the phylogeny and all currently described species of _Anilios_.
+We then calculated AIC to identify the best-fitting models.
+Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
+
+## Data visualisation and software
+ClaDS was implemented in `Julia` v.1.8.1 [@bezansonJulia2012], all other analyses were performed in `R` v.4.2 (CITE R).
+Figures and graphs were prepared using a combination of baseR and `ggplot2` (CITE) to process raw output from R.
+The bioregion map was created using the Free and Open Source `QGIS` v.3.6 (CITE Australian Map).
+We arranged and finalised all figures using Adobe Illustrator v2022.
 
@@
 <!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
-
 All BioGeoBEARS analyses inferred tropical grassland as the ancestral biome of *Anilios*.
@@
 Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+We found that most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+We also observed that closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
 
-We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
+Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)).
+However, the ancestor leading to crown *Anilios* is ambiguous.
+We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny.
+The majority of these events were anagenetic dispersal (31.2 ± 1.3 events) and cladogenetic founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, while we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), the ancestor leading to crown *Anilios* is ambiguous.
-We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
-
-We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
-Together the tropical grassland biome and the arid biome make up the largest source of diversificttion accounting for 69.6% of outward transitions.
-On the other hand, the largest sink for all inward dispersal are the temperate grassland biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
+The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%), which makes up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
+Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
+On the other hand, the largest sink for all inward dispersals are the temperate grassland biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
 The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
-
-## Diversification rates between arid and non-arid biomes
-
-We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
-We found that the model where variation in diversification rates were not linked to geographic range fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
-The best-fitting geographic range independent model included a hidden unobserved state, indicating diversification rates is better explained by a hidden unobserved state.
-Compared to the three other models, the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
-The weighted average parameter estimates from all models indicate both diversification rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE).
-However, net diversification rates for the widespread lineages were slightly higher (see figure), but this may be due to low amount of lineages.
-
-Results from our ClaDS analysis show that variation in branch-specific diversification rates generally slowed down towards the present (fig-ClaDS), and did not differ between arid, non-arid, and widespread lineages (fig-ClaDS).
 
@@
 
+## Diversification rates between arid and non-arid biomes
+
+Results from our ClaDS analysis show that branch-specific diversification rates gradually decrease towards the present (fig-ClaDS).
+Mean branch-specific rates did did not differ between arid, non-arid, and widespread lineages (fig-ClaDS).
+
+We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
+We found that the range-independent model fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
+The best-fitting geographic range independent model included a hidden unobserved state, indicating diversification rates is better explained by a hidden unobserved state.
+Compared to the three other models, the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
+The weighted average parameter estimates from all models indicate both speciation rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE; see supplementary XX for net diversification rate and extinction rates).
+
 \newpage
@@
 
-## Localised climate model outperforms global models
+Our study
 
-## Ancestral biome reconstruction
-- Our result — that the ancestor of Anilios occupied tropical grasslands – is consistent with the result presented in @marinTracing2013.
+BIG SUMMARY
+
+SCIENTIFIC FINDING AS EXTENSION OF PREVIOUS WORK
+
+FINDINGS' SCIENTIFIC IMPLICATIONS
+
+HOW THIS WILL FACILITATE FUTURE RESEARCH
+
+FORWARD-LOOKING STATEMENTS ABOUT FIELD AS A WHOLE 
+- As more genetic data allow us to build more extensive phylogenies, there is a growing need to develop methods and ways that can estimate environmental conditions in the past.
+Having this will allow us to test further hypotheses about how different macroevolutionary processes influence diversification on a larger scale.
+
+IMPORTANCE OF THIS TECHNICAL ASPECT OF THIS WORK 
+- We show that reconstruction of localised paleo environment is possible and argue that it is more appropriate as a correlate to diversification rates. 
+- Building these paleoclimate models for regional areas will be of interested for studies where species are geographically restricted. 
+- Studies that aim to understand how aridification affect diversification rates in Australia should apply paleoclimate or use them to inform models and inferences. 
+- For example, models like NEMo allows bounding ancestral niche with estimates of paleo climate.
+
+## Ancestral biome of Australian *Anilios*
 
@@
 
-After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
+- We expected the biogeographic models with jump dispersal to be the best fitting model because _Anilios_ are believed to have non-Australian origins [@marinTracing2013;@vidalBlindsnake2010;@oliverPhylogenetic2017].
+
+- The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
+Events like floods are likely to have faciliated dispersals but in short distances.
+Discuss species that are far apart...
+
+- After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
 Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+
+- However, the rates of diversification slowed down as Australia started to become more arid.
+We found that transitions into the arid biome for many lineages actually pre-dated the intense aridification in the late Miocene, which indicates that blind snakes may be pre-adapted to arid environment.
+This pattern has been observed in XXX (Paul Oliver's paper about ____ reptile).
+Arid environemnts is heterogenous and some areas have acted as refugia during glaciation, which may cause isolation and subsequent speciation.
 
@@
 torresianus_, _A.
-leucoproctus_, ...) are found in Northern
-Australia Torres Strait Islands, and southern parts of New Guinea.
+leucoproctus_, ...) are found in Northern Australia.
   - Range expansion was important for diversification ...
@@
 
-- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
+## Rates of diversification between arid and non-arid lineages
+
+- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
+For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite).
+- It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.
+
+- The early burst of speciation is observed in lineages that colonise new areas, ecological release. For example. 
+- Interestingly blind snakes have diversified greatly compared to other fossorial vertebrates.
 
 Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
+
+- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
 
@@
 
-## Drivers of diversifcation 
-- We expected the biogeographic models with jump dispersal to be the best fitting model because _Anilios_ are believed to have non-Australian origins [@marinTracing2013;@vidalBlindsnake2010;@oliverPhylogenetic2017].
+- Diversification rate decrease with aridification.
+But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
+
+- Our RPANDA model did not model environment-dependent extinction. 
+- Extinction is a big cavaet especially because we suspect there are more species
+
 - Blind snakes are thought to not move very much and so formation of geological barriers during aridification may accelerate speciation between 20 – 10 MYA.
-- The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
-Events like floods are likely to have faciliated dispersals but in short distances.
-Discuss species that are far apart...
+
 - High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
@@
 
-## Environment-dependent diversification
-- Diversification rate decrease with aridification.
-But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
-- Our RPANDA model did not model environment-dependent extinction. 
-- Extinction is a big cavaet especially because we suspect there are more species
-
 ## Limitations
 There are several limitations in our models.
+
+- Without extensive sampling we can not confidently trace the pathways in which _Anilios_ colonised Australia.
+Due to the coarseness of our biomes, we also cannot infer exactly where the ancestral species landed in Australia.
+Current biogeographic models does not account for extinct species and cannot model complex dynamics of geological processes such as continents moving.
+
 - While taxonomic sampling the most representative of the genus, it is not comprehensive.
@@
 erycinus_ will allow us to place the species among other _Anilios_ and provide a more comprehensive picture of the evolution of this genus.
+
+- Need to account for heterogeneity of the arid zone [@pepperLife2021].
+
 - Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands.
@@
 Placing that species with arid biome is not realistic but that area is considered the arid zone by our criteria.
-
-- Need to account for heterogeneity of the arid zone [@pepperLife2021].
 

### 2022-12-01 manuscript.Rmd
update big summary of discussion

@@
 
-Our study
-
-BIG SUMMARY
+With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of scolecophidia in the world.
+In this study we inferred the the genus has a tropical grassland origin.
+Within a few million years the ancestral lineage diversified rapidly and the tropical grassland biome served as a major source of diversification as lineages transitioned into what we now consider the arid biome and subsequently into the mediterranean and temperate XXXXgrassland biomes.
+We found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree. 
+Although aridification has been shown to accelerate diversification in other groups of vertebrates, we found that *Anilios* blind snakes follow a different trajectory.
+Increased aridity from the late Miocene-early Pleistocene coincided with reduced diversification rates for genus.
 
@@
 - Studies that aim to understand how aridification affect diversification rates in Australia should apply paleoclimate or use them to inform models and inferences. 
-- For example, models like NEMo allows bounding ancestral niche with estimates of paleo climate.
+- For example, models like NEMo allows ancestral niche to be informed by paleo environment as a prior.
 

### 2022-12-06 code/04_BioGeoBEARS_results_plots.R
change output path

@@ -89,7 +89,7 @@ dev.off()
 
 
 
-# pdf('output/pie_ancestral.pdf')
+# pdf('output/bgb_pie_ancestral.pdf', width = 11.5, height = 8.5)
 # Pie chart
 BioGeoBEARS::plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j","x"), plotwhat="pie", 
                                       label.offset=0.45, tipcex=0.02, statecex=0.4, splitcex=0.4, titlecex=0.8, plotsplits=F, 

### 2022-12-06 manuscript.Rmd
rearranged discussion

--- 2022-12-01 manuscript.Rmd
@@
 
-Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [[@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches .
+For example, agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013] are reptiles that colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene.
 
+Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [[@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches.
 Other clades like geckos and blind snakes remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@marinHidden2013;@moritzCryptic2018;@zozayaAre2019].
-
 The species richness and distribution of squamate reptiles across different biomes makes them a good group to test hypotheses about the tempo and mode of diversification in Australia.
-
-For example, agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013] are reptiles that colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene.
 
@@
 eryncinus_, which is endemic to New Guinea.
-Blind snakes are morphologically conserved and often difficult to distinguis.
+Blind snakes are morphologically conserved and often difficult to distinguish.
 Extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
@@
 
-Our study
-
-BIG SUMMARY
-
-SCIENTIFIC FINDING AS EXTENSION OF PREVIOUS WORK
-
-FINDINGS' SCIENTIFIC IMPLICATIONS
-
-HOW THIS WILL FACILITATE FUTURE RESEARCH
-
-FORWARD-LOOKING STATEMENTS ABOUT FIELD AS A WHOLE 
-- As more genetic data allow us to build more extensive phylogenies, there is a growing need to develop methods and ways that can estimate environmental conditions in the past.
-Having this will allow us to test further hypotheses about how different macroevolutionary processes influence diversification on a larger scale.
-
-IMPORTANCE OF THIS TECHNICAL ASPECT OF THIS WORK 
-- We show that reconstruction of localised paleo environment is possible and argue that it is more appropriate as a correlate to diversification rates. 
-- Building these paleoclimate models for regional areas will be of interested for studies where species are geographically restricted. 
-- Studies that aim to understand how aridification affect diversification rates in Australia should apply paleoclimate or use them to inform models and inferences. 
-- For example, models like NEMo allows bounding ancestral niche with estimates of paleo climate.
+With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of scolecophidia in the world.
+In this study we inferred the extant species in Australia has a tropical grassland origin.
+Within a few million years, this ancestral lineage split with one clade staying in the tropical grassland biome while the other transitioned into the arid biome.
+Subsequent transitions from these two lineages resulted in the distribution of biomes we observed across the phylogeny today.
+We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree.
+Although aridification has been shown to accelerate diversification rates in other groups of vertebrates (CITE), we found no such difference for Australian *Anilios* blind snakes.
 
@@
 
-- Our result also supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the continent [@marinTracing2013], but exactly where they lived prior to this invasion is equivocal between the Bismarck Archipelago and New Guinea.
+Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent between 25.8 – 16 MYA [@marinTracing2013].
+Our analyses cannot conclude whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as previously proposed [@vidalBlindsnake2010].
+Instead, our results suggest the ancestor of Australian _Anilios_ occured either in the Bismarck Archipelago, New Guinea, or a combination of both.
 
@@
 
+After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
+Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid biome since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+For example elapid snakes were hypothesised to have colonised Australia from Southeast Asia (CITE) or New Guinea (cite) around the same time (cite Paul).
+
+However, we were not able to include the New Guinean endemic _Anilios erycinus_, which is the only recognised species of _Anilios_ not found in Australia.
+Therefore, we cannot make any conclusions about transitions between New Guinea and Australia even though at least two species (i.e.
+_A.
+torresianus_ and _A.
+leucoproctus_) are found between southern New Guinea and Far North Queensland
+
+Placement of _R. multilineatus_ in the phylogeny renders _Ramphotyphlops_ paraphyletic (as discussed in Tiatragul 2023) 
+The sister group to _Anilios_ is the genus _Ramphotyphlops_, which comprises XX species of blind snakes that occur in mesic areas across south and southeast Asia.
+
 - The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
 Events like floods are likely to have faciliated dispersals but in short distances.
-Discuss species that are far apart...
-
-- After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
-Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid zone since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
+- Range expansion was important for diversification ...
+- Topography dynamics [@brazModelling2021]
+- Solomon Island 40 MYA [@hallCenozoic2002]
+- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
 
@@
 We found that transitions into the arid biome for many lineages actually pre-dated the intense aridification in the late Miocene, which indicates that blind snakes may be pre-adapted to arid environment.
-This pattern has been observed in XXX (Paul Oliver's paper about ____ reptile).
+This pattern has been observed in XXX (Paul Oliver's paper about __ reptile).
 Arid environemnts is heterogenous and some areas have acted as refugia during glaciation, which may cause isolation and subsequent speciation.
 
-- We found transitions from XXX to XXX to be common as have been shown / as opposed to other Australian squamates (CITATIONS)
-- Arid zone was not arid until ~20 Million Years ago. Ancestors in deeper time probably occupied temperate or tropical areas that subsequently became the arid zone we know see today. 
-- The Australo-Papuan clade seems to have stayed in Australia for the most part but we were not able to include New Guinean endemic _Anilios erycinus_, the only recognised species of _Anilios_ not found in Australia.
-Therefore, we cannot make any conclusions about transitions between Australia and New Guinea even though at least three species (_A.
-torresianus_, _A.
-leucoproctus_, ...) are found in Northern Australia.
-  - Range expansion was important for diversification ...
-  - Topography dyanmics [@brazModelling2021]
-  - Solomon Island 40 MYA [@hallCenozoic2002]
+## Influence of paleoclimate on diversification rates
+
+Climate is one of the
+- Australian blind snakes follow a different trajectory, one in which increased aridity from the late Miocene to early Pleistocene coincided with reduced diversification rates for genus. 
+- In contrast to other vertebrates, our results suggest aridification reduces the rate of diversification in blind snakes. 
+[@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015]
+- Our study is the first to quantitatively assess the influence of aridity on the diversification rates of a group of vertebrate on a macroevolutionary scale.
+
+- Overall the average temperature from Australia specific reconstruction is consistently is warmer than global mean temperature.
+
+In the context of our birth-death models, our the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
+
+- Paleoclimate reconstructions based on DEM by Scotese and Straume were similar in profile.
+However, DEM base on Valdes differed from the other two.
+Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles.
+However, models that take in account thermal niche tolerance, such as NEMo, will benefit from paleoclimate reconstructions such as these to inform limits (boundaries). 
+- Aridity index is a good indicator, but aridity is geographically restricted to the arid zone. In our phylogeny, a lot of diversity is observed is in the mesic zones.
+- Warmer and dryer environment may not directly increase speciation rate because blind snakes are physically adapting to warmer temperatures, rather it is the fragmentation of the landscape. 
+- Thermal tolerance and preference for blind snakes remains unknown but fossorial species tend to live in pretty stable climate? 
+- Our RPANDA model did not model environment-dependent extinction. 
+- Studies that aim to understand how aridification affect diversification rates in Australia should apply paleoclimate or use them to inform models and inferences. 
+- For example, models like NEMo allows ancestral niche to be informed by paleo environment as a prior.
 
 ## Rates of diversification between arid and non-arid lineages
+- The early burst of speciation is observed in lineages that colonise new areas, ecological release. For example.
+
+- Diversification rate decrease with aridification.
+But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
+Turnover rate also ...
 
@@
 For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite).
-- It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.
 
-- The early burst of speciation is observed in lineages that colonise new areas, ecological release. For example. 
 - Interestingly blind snakes have diversified greatly compared to other fossorial vertebrates.
 
+  - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated. 
 Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017]. 
+  - Extinction is a big cavaet especially because we suspect there are more species.
+Arid adapted lineages may have experienced higher extinction OR not well sampled in remote Western Australia.
+Most species are sampled next to roads (Xia's paper or do supplemental analyses).
 
-- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
+  - Blind snakes are thought to not move very much and so formation of geological barriers during aridification may accelerate speciation between 20 – 10 MYA. 
+  - High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
+  - Speciation could even be sympatric but no evidence to support this.
 
@@
 Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have a hard time crossing the lines [@bryantLines2016]
-
-- Diversification rate decrease with aridification.
-But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
-
-- Our RPANDA model did not model environment-dependent extinction. 
-- Extinction is a big cavaet especially because we suspect there are more species
-
-- Blind snakes are thought to not move very much and so formation of geological barriers during aridification may accelerate speciation between 20 – 10 MYA.
-
-- High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
-- Speciation could even be sympatric but no evidence to support this.
 
@@
 
+As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleo environmental datasets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
+We show that tools like `gen3sis` can be used to reconstruct localised paleo environment that provides a more appropriate fit to our Australian blind snake phylogeny than generic paleo global temperature.
+We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.
+- Furthermore we provide the first quantitative measure of paleo aridity
+
 # Tables

### 2022-12-07 manuscript.Rmd
wrote more in discussion about biome transition and include ref to supplement showing dist of clades

@@
 Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent between 25.8 – 16 MYA [@marinTracing2013].
-Our analyses cannot conclude whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as previously proposed [@vidalBlindsnake2010].
-Instead, our results suggest the ancestor of Australian _Anilios_ occured either in the Bismarck Archipelago, New Guinea, or a combination of both.
+This is also consistent with other radiations of Australian squamates, where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
+Given what we know about the current distribution of species included in our phylogeny, our analyses support the idea that *Anilios* have non-Australian origins.
+However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as proposed by @vidalBlindsnake2010 is equivocal.
+Instead, our results suggest the ancestor of Australian _Anilios_ most likely occurred either in the Bismarck Archipelago, New Guinea, or a combination of both.
+Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and other closely related Typhlopidae snakes (e.g.
+*Ramphotyphlops*, *Acutotyphlops*) as outgroups will enable us to make inferences about the origin prior to colonising Australia.
 
-- We expected the biogeographic models with jump dispersal to be the best fitting model because _Anilios_ are believed to have non-Australian origins [@marinTracing2013;@vidalBlindsnake2010;@oliverPhylogenetic2017].
+## Biome transition and current distributions
 
-After arriving to Australia in the late Oligocene - early Miocene (~25 MYA), the ancestral _Anilios_ initially diversified rapidly within less than 1 million years.
-Our analysis is consistent with the conclusion that Australian blindsnakes, like several other Australian squamates, originated from tropical-mesic origins and expanded into the arid biome since the mid-Miocene [@chappleParallel2004;@melvilleEvolutionary2011].
-For example elapid snakes were hypothesised to have colonised Australia from Southeast Asia (CITE) or New Guinea (cite) around the same time (cite Paul).
+After their arrival to Australia, the ancestral _Anilios_ initially diversified rapidly (<1 million years) in the tropical grassland biome.
+The first transition out of the tropical grassland biome were to the Tropical Forest biome (for the branch leading to _A.
+silvia_) and the arid biome (\@ref(fig:fig-biogeobears))a.
 
-However, we were not able to include the New Guinean endemic _Anilios erycinus_, which is the only recognised species of _Anilios_ not found in Australia.
-Therefore, we cannot make any conclusions about transitions between New Guinea and Australia even though at least two species (i.e.
-_A.
-torresianus_ and _A.
-leucoproctus_) are found between southern New Guinea and Far North Queensland
+We found that the initial transition into the arid biome around 20 MYA pre-dated the continent wide aridification that intensified up in the late Miocene [@byrneBirth2008].
 
-Placement of _R. multilineatus_ in the phylogeny renders _Ramphotyphlops_ paraphyletic (as discussed in Tiatragul 2023) 
-The sister group to _Anilios_ is the genus _Ramphotyphlops_, which comprises XX species of blind snakes that occur in mesic areas across south and southeast Asia.
+The Pilbara region has both lineages that have deep coalescent histories and shallower divergences (_A. ammodytes_ and _A. ganei_)
+(Reference to phylogeography supplementary figure)
 
-- The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
+Arid environemnts is heterogenous and some areas have acted as refugia during glaciation, which may cause isolation and subsequent speciation.
+
+The western part of Australia dried up earlier than the east around 20 MYA (CITE Mitzy).
+
+Which indicates that blind snakes may be pre-adapted to arid environment. This pattern has been observed in XXX [@oliverPhylogenetic2017].
+
+- Range expansion was important for diversification ...
+The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
 Events like floods are likely to have faciliated dispersals but in short distances.
-- Range expansion was important for diversification ...
-- Topography dynamics [@brazModelling2021]
-- Solomon Island 40 MYA [@hallCenozoic2002]
-- Recent biogeographic reconstruction recovered India as the origin for superfamily Typhlopoidea (Gerrhopilidae, Xenotyphlopidae and Typhlopidae) [@kleinEvolution2021].
 
-- However, the rates of diversification slowed down as Australia started to become more arid.
-We found that transitions into the arid biome for many lineages actually pre-dated the intense aridification in the late Miocene, which indicates that blind snakes may be pre-adapted to arid environment.
-This pattern has been observed in XXX (Paul Oliver's paper about __ reptile).
-Arid environemnts is heterogenous and some areas have acted as refugia during glaciation, which may cause isolation and subsequent speciation.
+The rates of diversification slowed down as Australia started to become more arid
 
@@
 
+Therefore, we cannot make any conclusions about transitions between New Guinea and Australia even though at least two species (i.e.
+_A.
+torresianus_ and _A.
+leucoproctus_) are found between southern New Guinea and Far North Queensland
+
 # Conclusion

### 2022-12-08 manuscript.Rmd
discussion changes and add about pilbara and aridification

@@
 
-The four SSE models vary by whether they allow diversification rates to be associated with geographic states (range-dependent or range-independent) and whether they allow an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
-Model 1 assumes equal diversification rates regardless of geographic state. Model 2 allows geographic range to affect diversification rates (canonical GeoSSE model). 
-Models 3 and 4 are derivations of models 1 and 2 but also allow unmeasured trait ('hidden states') to contribute to diversification rates (see Table GeoHiSSE table).
-By incoporating hidden states, we are able to assess whether geographic range or an unmeasured trait contributes more to the observed diversification.
-For each model, we assumed uniform taxon sampling for 89% in arid lineages, 71% in non-arid lineages, and 83% in widespread lineages based on number of species in the phylogeny and all currently described species of _Anilios_.
+The four SSE models vary in whether they allow diversification rates to be associated with geographic states (range-dependent or range-independent) and whether they allow an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
+Model 1 assumes equal diversification rates regardless of geographic state. Model 2 allows geographic range to affect diversification rates (canonical geoSSE model).
+Models 3 and 4 are based on models 1 and 2, but they also allow unmeasured trait ('hidden states') to contribute to diversification rates (see Table GeoHiSSE table).
+By incorporating hidden states, we can assess whether geographic range or an unmeasured trait contributes more to the observed diversification.
+For each model, we assumed uniform taxon sampling for 89% in arid lineages, 71% in non-arid lineages, and 83% in widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_.
 We then calculated AIC to identify the best-fitting models.
@@
 <!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
-All BioGeoBEARS analyses inferred tropical grassland as the ancestral biome of *Anilios*.
-We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
-Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
-We found that most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-We also observed that closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).
+Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios* but the ancestor leading to the crown *Anilios* is ambiguous.
+We found that the DEC+_j_+_x_ model was the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
+A likelihood ratio test (LRT) showed that both founder events (+*j* parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)).
-However, the ancestor leading to crown *Anilios* is ambiguous.
-We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny.
-The majority of these events were anagenetic dispersal (31.2 ± 1.3 events) and cladogenetic founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of our best-fitting model.
+We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny.
+Most of these events were anagenetic dispersals (31.2 ± 1.3 events) and cladogenetic founder events (6.1 ± 1.5 events).
 
@@
 The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
+
+Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
+The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, but 12/21 lineages expanded or transitioned into other biomes.
+The common ancestor of *Group 2* transitioned into the arid biome, and 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.
 
@@
 We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) with aridity index derived from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA).
-This best-fitting model indicates that increase aridity is exponentially associated with decrease speciation rate (lambda = 1.32, alpha = -0.14) and a constant background extinction (mu = 0.23).
-The second and third most competitive models (delta AICc < 2) also found increase aridity is exponentially associated with decrease speciation rate but without estimating extinction (Table-RPANDA-TOP10).
-Among exponential environment-dependent models, we found Australia-specific paleoclimate data provides a better fit to our phylogeny than global paleoclimate (see supp-TABLE_FULL_RPANDA_model_fit).
+This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (lambda = 1.32, alpha = -0.14) and a constant background extinction rate (mu = 0.23).
+The second and third most competitive models (delta AICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10).
+Among the exponential environment-dependent models, we found that Australia-specific paleoclimate data provided a better fit to our phylogeny than global paleoclimate data (see supp-TABLE_FULL_RPANDA_model_fit).
 
@@
 
-Results from our ClaDS analysis show that branch-specific diversification rates gradually decrease towards the present (fig-ClaDS).
-Mean branch-specific rates did did not differ between arid, non-arid, and widespread lineages (fig-ClaDS).
+Results from our ClaDS analysis show that branch-specific diversification rates gradually decrease towards the present (see fig-ClaDS).
+The mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (see fig-ClaDS).
 
 We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
-We found that the range-independent model fit the data better than models where diversification rates were depdendent on geographic range (see supp-table-geohisse).
-The best-fitting geographic range independent model included a hidden unobserved state, indicating diversification rates is better explained by a hidden unobserved state.
-Compared to the three other models, the best-fitting model accounts for 0.95 of the AIC weight and had AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
-The weighted average parameter estimates from all models indicate both speciation rates and turnover rates did not differ between arid and non-arid ranges (REF FIGURE; see supplementary XX for net diversification rate and extinction rates).
+We found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see supp-table-geohisse).
+The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state.
+Compared to the three other models, the best-fitting model accounted for 0.95 of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see REF FIGURE and supplementary XX for net diversification rate and extinction rates).
 
@@
 In this study we inferred the extant species in Australia has a tropical grassland origin.
-Within a few million years, this ancestral lineage split with one clade staying in the tropical grassland biome while the other transitioned into the arid biome.
-Subsequent transitions from these two lineages resulted in the distribution of biomes we observed across the phylogeny today.
+Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioning into the arid biome.
+Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
 We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree.
@@
 
-After their arrival to Australia, the ancestral _Anilios_ initially diversified rapidly (<1 million years) in the tropical grassland biome.
-The first transition out of the tropical grassland biome were to the Tropical Forest biome (for the branch leading to _A.
-silvia_) and the arid biome (\@ref(fig:fig-biogeobears))a.
+Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by multiple transitions into the arid biome and subsequent expansion to the Mediterranean biome.
+However, our results show that the timing of these transition and diversification in the arid biome, for both species in groups 1 and 2, occured earlier (20 – 10 MYA) than previously proposed (10 MYA) by @marinTracing2013.
 
-We found that the initial transition into the arid biome around 20 MYA pre-dated the continent wide aridification that intensified up in the late Miocene [@byrneBirth2008].
+For Group 1 eastward transitions might have occured
 
-The Pilbara region has both lineages that have deep coalescent histories and shallower divergences (_A. ammodytes_ and _A. ganei_)
-(Reference to phylogeography supplementary figure)
+Pilbara in the west acted as a refugia and perhaps source of diversification 
+Arid environemnts is heterogenous and some areas have acted as refugia during glaciation (CITE), which may cause isolation and subsequent speciation.
+For example, the Pilbara region has both lineages that have deep coalescent histories and shallower divergences (_A.
+ammodytes_ and _A.
+ganei_).
+These species may have once been widely distributed but became extirpated as aridification intensified and are now found only in their place of refuge during glacial fluctuations.
 
-Arid environemnts is heterogenous and some areas have acted as refugia during glaciation, which may cause isolation and subsequent speciation.
+Alternatively, this suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy red centre of Australia [@byrneBirth2008].
+This pattern may indicate that blind snakes, like other fossorial animals, are pre-adapted to arid environments (CITE). Subterranean microhabitat is relatively stable compared to ground.
 
-The western part of Australia dried up earlier than the east around 20 MYA (CITE Mitzy).
-
-Which indicates that blind snakes may be pre-adapted to arid environment. This pattern has been observed in XXX [@oliverPhylogenetic2017].
-
-- Range expansion was important for diversification ...
+Our results suggest that range expansion was important for diversification.
 The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
 Events like floods are likely to have faciliated dispersals but in short distances.
-
-The rates of diversification slowed down as Australia started to become more arid
 

### 2022-12-09 manuscript.Rmd
update discussion

@@
 The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
-The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, but 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of *Group 2* transitioned into the arid biome, and 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.
+The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes.
+The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.
 
@@
 
-Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by multiple transitions into the arid biome and subsequent expansion to the Mediterranean biome.
-However, our results show that the timing of these transition and diversification in the arid biome, for both species in groups 1 and 2, occured earlier (20 – 10 MYA) than previously proposed (10 MYA) by @marinTracing2013.
+Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes.
+However, as opposedrecent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone [@marinHidden2013], our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA, during the time when the Australia was still relatively wet and humid [@byrneBirth2008].
+This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].
+
+One hypothesis to explain contemporary distribution in the arid biome is that *Anilios* may be pre-adapted to arid environments.
+Blind snakes have been found in xeric areas around the world, for example _Myriopholis macrorhyncha_ in the desert
+
+North American  and the Palearctic _Letheobia episcopus_) possibly because subterranean microhabitat has relatively stable climate compared to surface [@kinlawReview1999;@nevoAdaptive1979].
+
+Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.
+
+The fact that we observed multiple independent transitions into the arid biome.
 
@@
 These species may have once been widely distributed but became extirpated as aridification intensified and are now found only in their place of refuge during glacial fluctuations.
-
-Alternatively, this suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy red centre of Australia [@byrneBirth2008].
-This pattern may indicate that blind snakes, like other fossorial animals, are pre-adapted to arid environments (CITE). Subterranean microhabitat is relatively stable compared to ground.
 

### 2022-12-14 manuscript.Rmd
modify result

@@
 Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
-On the other hand, the largest sink for all inward dispersals are the temperate grassland biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
+On the other hand, the largest sink for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
 The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.

### 2022-12-19 manuscript.Rmd
update result secition biogeobears

@@
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of our best-fitting model.
-We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny.
-Most of these events were anagenetic dispersals (31.2 ± 1.3 events) and cladogenetic founder events (6.1 ± 1.5 events).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+_j_+_x_ model.
+We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see supplementary_table_BioGEOBEARS).
+Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
+Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2±2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
 
@@
 Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
+The model indicates two pulses of biogeographic transitions for each  of the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
 The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes.
@@
 We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree.
-Although aridification has been shown to accelerate diversification rates in other groups of vertebrates (CITE), we found no such difference for Australian *Anilios* blind snakes.
+Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found no such association for Australian *Anilios* blind snakes.
 
@@
 Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes.
-However, as opposedrecent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone [@marinHidden2013], our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA, during the time when the Australia was still relatively wet and humid [@byrneBirth2008].
+However, our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA and gradually diversified during the time when the Australia was still relatively wet and humid [@byrneBirth2008] as opposed more recent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone proposed by @marinHidden2013.
 This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].
@@
 
-Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.
+We found that closely related species share similar biome and distribution. 
+Our results indicate that cladogenetic processes (sympatry, vicariance, and founder event) made up the majority of range expansion events.
+Negative values for the distance parameter _X_ suggest that range expansion is expected to decrease as distance increase (CITE BioGeoBEARs intepetration?).
 
@@
 
-Our results suggest that range expansion was important for diversification.
-The fact that DEC+J+x model was preferred over DEC+J may be due to poor dispersal ability as closely related species occupy the same or nearby areas.
-Events like floods are likely to have faciliated dispersals but in short distances.
+While there's not study to show, blind snakes probably have small ranges and poor dispersal ability as closely related species occupy the same or nearby areas.
+Stochastic events like floods may have facilitated dispersals but in short distances.
+
+Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.
 

### 2022-12-20 manuscript.Rmd
paragraph about other desert blind snakes

@@
 
-One hypothesis to explain contemporary distribution in the arid biome is that *Anilios* may be pre-adapted to arid environments.
-Blind snakes have been found in xeric areas around the world, for example _Myriopholis macrorhyncha_ in the desert
-
-North American  and the Palearctic _Letheobia episcopus_) possibly because subterranean microhabitat has relatively stable climate compared to surface [@kinlawReview1999;@nevoAdaptive1979].
-
-We found that closely related species share similar biome and distribution. 
 Our results indicate that cladogenetic processes (sympatry, vicariance, and founder event) made up the majority of range expansion events.
@@
 Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.
+
+We also found that closely related species share similar biome and distribution. 
+One hypothesis to explain contemporary distribution in the arid biome is that *Anilios* (and other blind snakes in general) may be pre-adapted to living in arid environments. 
+Blind snakes have been found in xeric areas around the world, for example _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ semi-arid areas of from east Africa to the Middle East [@sindacoReptiles2013].
+To our knowledge, these species share similar dietary and morphological specialisations with their closest relatives, which is a pattern identified across all fossorial snakes [@cyriacDigging2018].
+Because subterranean microhabitat has relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised pheonotype may not be selected for.
+Future studies that investigate phenotypic variations across all blind snakes will further our understanding of how blind snakes evolved with seemingly conserved phenotypes.
 

### 2022-12-20 manuscript.Rmd
update manuscript

--- 2022-12-19 manuscript.Rmd
@@
 Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2±2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
+Our model also indicate that range expansion is expected to decrease as distance increase (free parameter +_X_ = -1.3).
 
@@
 The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.
+The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the arid biome.
 
@@
 
-Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes.
-However, our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA and gradually diversified during the time when the Australia was still relatively wet and humid [@byrneBirth2008] as opposed more recent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone proposed by @marinHidden2013.
-This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].
+Consistent with previous research by @marinTracing2013, we found that the initial diversification of Anilios occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes.
+However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013.
 
-One hypothesis to explain contemporary distribution in the arid biome is that *Anilios* may be pre-adapted to arid environments.
-Blind snakes have been found in xeric areas around the world, for example _Myriopholis macrorhyncha_ in the desert
+Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the Tropical Grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
+This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene, which has resulted in the dry and sandy landscape present today [@byrneBirth2008].
 
-North American  and the Palearctic _Letheobia episcopus_) possibly because subterranean microhabitat has relatively stable climate compared to surface [@kinlawReview1999;@nevoAdaptive1979].
-
-We found that closely related species share similar biome and distribution. 
-Our results indicate that cladogenetic processes (sympatry, vicariance, and founder event) made up the majority of range expansion events.
-Negative values for the distance parameter _X_ suggest that range expansion is expected to decrease as distance increase (CITE BioGeoBEARs intepetration?).
-
-The fact that we observed multiple independent transitions into the arid biome.
-
-For Group 1 eastward transitions might have occured
-
-Pilbara in the west acted as a refugia and perhaps source of diversification 
-Arid environemnts is heterogenous and some areas have acted as refugia during glaciation (CITE), which may cause isolation and subsequent speciation.
-For example, the Pilbara region has both lineages that have deep coalescent histories and shallower divergences (_A.
-ammodytes_ and _A.
-ganei_).
-These species may have once been widely distributed but became extirpated as aridification intensified and are now found only in their place of refuge during glacial fluctuations.
-
-While there's not study to show, blind snakes probably have small ranges and poor dispersal ability as closely related species occupy the same or nearby areas.
-Stochastic events like floods may have facilitated dispersals but in short distances.
-
-Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.
+A hypothesis to explain the contemporary distribution of _Anilios_ and other blind snakes in arid biomes is that they may be pre-adapted to living in such environments.
+Blind snakes have been found in xeric areas around the world, such as _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ in semi-arid areas from east Africa to the Middle East [@sindacoReptiles2013].
+To our knowledge, these species share similar dietary and morphological specialisations with their closest relatives, which is a pattern identified across all fossorial snakes [@cyriacDigging2018].
+Because the subterranean microhabitat has a relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for.
+Future studies that investigate phenotypic variations across all blind snakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
 
@@
 
-Climate is one of the
+Changes in paleoenvironmental changes have been linked with diversification rates via range contraction and expansion that influences niche availability. 
+We found that aridity played an important role in 
 - Australian blind snakes follow a different trajectory, one in which increased aridity from the late Miocene to early Pleistocene coincided with reduced diversification rates for genus.
-- In contrast to other vertebrates, our results suggest aridification reduces the rate of diversification in blind snakes. 
-[@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015]
-- Our study is the first to quantitatively assess the influence of aridity on the diversification rates of a group of vertebrate on a macroevolutionary scale.
+This is in contrast with other vertebrates, where diversification rates seem to increase with aridification [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015]
+- Warmer and dryer environment may not directly increase speciation rate because blind snakes are physically adapting to warmer temperatures.
+However, climatic fluctuations can contract and fragment available niche for blind snakes andthus influence diversification rates.
+- In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
+However, models that take in account thermal niche tolerance, such as NEMo [CITE], will benefit from paleoclimate reconstructions such as these to inform limits (boundaries).
+We suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these paleoclimate models and/or use them to inform inferences.
 
-- Overall the average temperature from Australia specific reconstruction is consistently is warmer than global mean temperature.
-
-In the context of our birth-death models, our the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
-
+ALEX SKEELS add to this discussion?
 - Paleoclimate reconstructions based on DEM by Scotese and Straume were similar in profile.
@@
 Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles.
-However, models that take in account thermal niche tolerance, such as NEMo, will benefit from paleoclimate reconstructions such as these to inform limits (boundaries). 
-- Aridity index is a good indicator, but aridity is geographically restricted to the arid zone. In our phylogeny, a lot of diversity is observed is in the mesic zones.
-- Warmer and dryer environment may not directly increase speciation rate because blind snakes are physically adapting to warmer temperatures, rather it is the fragmentation of the landscape. 
-- Thermal tolerance and preference for blind snakes remains unknown but fossorial species tend to live in pretty stable climate? 
-- Our RPANDA model did not model environment-dependent extinction. 
-- Studies that aim to understand how aridification affect diversification rates in Australia should apply paleoclimate or use them to inform models and inferences. 
-- For example, models like NEMo allows ancestral niche to be informed by paleo environment as a prior.
+to add discussion to this?
 
 ## Rates of diversification between arid and non-arid lineages
-- The early burst of speciation is observed in lineages that colonise new areas, ecological release. For example.
+
+- The early burst of speciation is observed in lineages that colonise new areas. For example.
 
@@
 - Interestingly blind snakes have diversified greatly compared to other fossorial vertebrates. 
+  - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.
 
-  - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated. 
 Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017]. 
@@
 Arid adapted lineages may have experienced higher extinction OR not well sampled in remote Western Australia.
-Most species are sampled next to roads (Xia's paper or do supplemental analyses).
+Most species are sampled next to roads [CITE Xia's paper or do supplemental analyses].
 
-  - Blind snakes are thought to not move very much and so formation of geological barriers during aridification may accelerate speciation between 20 – 10 MYA. 
-  - High rates of diversification observed in concert with changing climate may be explained by allopatric speciation. 
-  - Speciation could even be sympatric but no evidence to support this.
+- Vast arid biome is  the largest biome in Australia with lots of heterogeneity (cite Mitzy's paper). Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
 
-Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa.
-Two pulses of speciation early Miocene and then another early Pleistocene.
-Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have a hard time crossing the lines [@bryantLines2016]
+- Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa. 
+- Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have difficulty crossing the lines [@bryantLines2016]
 
@@
 
+- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands.
+For example *A.
+longissimus* has only been collected from 30m underground on Barrow Island located XX km the northwest coast of Western Australia.
+Placing that species with arid biome is not realistic but that area is considered the arid biome by the WWF Ecoregions map.
+
 - While taxonomic sampling the most representative of the genus, it is not comprehensive.
@@
 
-- Need to account for heterogeneity of the arid zone [@pepperLife2021].
-
-- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands.
-For example *A.
-longissimus* has only been collected from 30m underground on Barrow Island located XX km the northwest coast of Western Australia.
-Placing that species with arid biome is not realistic but that area is considered the arid zone by our criteria.
+- Our RPANDA model did not model environment-dependent extinction.
 
@@
 
+- Need to account for heterogeneity of the arid zone [@pepperLife2021].
+
 # Conclusion
@@
 As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleo environmental datasets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
-We show that tools like `gen3sis` can be used to reconstruct localised paleo environment that provides a more appropriate fit to our Australian blind snake phylogeny than generic paleo global temperature.
+We show that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
 We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.
-- Furthermore we provide the first quantitative measure of paleo aridity
 

### 2022-12-21 manuscript.Rmd
polished up some discussion for paleoclimate

@@
 
-Present-day distrubtion of extant species offers clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
-
-The aridification on continental Australia began around 20 million years ago with pockets of arid zone developing and subsequent extreme warming and cooling periods in last ~5 million years creating what is now known as the 'outback' [@byrneBirth2008].
-From there arid-adapted vegetation expanded and the mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
-This period of aridification is believed to involve rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
+The aridification of continental Australia began around 20 million years ago, with pockets of arid zone developing and subsequent extreme warming and cooling periods in the last ~5 million years creating what is now known as the 'outback' [@byrneBirth2008].
+During this period, arid-adapted vegetation expanded and mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
+This period of aridification is believed to have involved the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
 
@@
 Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified in the last 20 - 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
-
-For example, agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013] are reptiles that colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene.
+Examples of reptilest that have colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene include  agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013].
 
@@
 
-Roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021].
+Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021].
 All currently described species of _Anilios_ are found in Australia except for _A.
 eryncinus_, which is endemic to New Guinea.
-Blind snakes are morphologically conserved and often difficult to distinguish.
-Extensive phylogenetic studies and several efforts to revisit specimens revealed the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
-
-The diversification of _Anilios_ is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19-39 mya [@vidalBlindsnake2010;@marinTracing2013], before the onset of aridification over the past 20 million years [@byrneBirth2008].
-However, how and
-
-However, the study did not account for alternative biogeographic areas (e.g. WWF Ecoregions) and models, or uncertainty in phylogenetic relationships and divergence times.
-
-Despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
-The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in almost every biome.
+Blind snakes are morphologically conserved and often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
+The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19 - 39 mya [@vidalBlindsnake2010;@marinTracing2013].
+However, despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
+The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in every biome.
 
@@
 Recent development in geographic-based methods such as `gen3sis` modeling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
-In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia in the time span of interest, between the 45 MYA to present.
+In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia in the time span of interest, between the 30 MYA to present.
 These datasets will allow us to empirically test the influence of localised ecological processes on diversification rates over time.
 
-<!--Why people should care-->
-- Applying the cutting edge localised geographic-based paleoenvironment reconstruction,
+The aim of this study is to infer the ancestral biome of Australian blind snakes based on a near-complete phylogeny of extant taxa.
+Additionally, we aim to investigate how paleo-environmental changes have influenced the rates of diversification within this group.
+To do this, we assess the fit of models between paleo-environmental conditions and diversification rates, and compare the rates of diversification between lineages in the arid versus non-arid biome.
 
-In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
+<!--- In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
 Specifically, we sought to answer the following questions:
@@
 () Which paleo environment best explain current phylogeny?
-() Are there differences in diversification rates between lineages in arid and non-arid biome?
-
-Aims: 
-- Infer the ancestral biome of Australian blindsnakes based on near-complete phylogeny of extant taxa.
-- Investigate how paleo-environmental changes influence rates of diversification. 
-- Assess model fit between paleo-environment.  
-- Compare the rates of diversification between lineages in the arid versus non-arid biome.
+() Are there differences in diversification rates between lineages in arid and non-arid biome? --->
 
@@
 <!-- Based on our phylogenetic tree -->
-We employ the recent time-calibrated phylogenomic tree of Australian blindsnake with outgroups.
-This tree includes 38 operational taxonomic units (OTUs) representing  __taxa representing __ described species of _Anilios_.
-However, because our phylogenetic data suggest some cryptic species, we adjust our taxon sampling fraction to XX/XX species.
-The genetically distinct OTUs that are not fully recognised species include 3 '_Anilios grypus_' (2 the west and 1 in the east) and 2 '_Anilios ligatus_' (one from the east and one from the north).
-For ancestral range estimation analyses we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
+We employed a time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al.
+in review).
+This tree includes 38 operational taxonomic units (OTUs) representing 36 described species of _Anilios_.
+However, our phylogenetic data suggest the existence of some cryptic species, so we adjusted our taxon sampling fraction to 38/50 species.
+The genetically distinct OTUs that are not fully recognised as species include 3 '_Anilios grypus_' (2 in the west and 1 in the east) and 2 '_Anilios ligatus_' (one from the east and one from the north).
+For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.s
 
@@
 
-To investigate the biogeographic history of blind snakes we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
-BioGeoBEARS provides a flexible likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
-These shifts in biomes can occur within a single branch (anagenetic events including processes like dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant and founder-event speciation).
-The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range each species are found, and a historical biogeographic model.
+To investigate the biogeographic history of blind snakes, we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
+BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
+These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
+The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.
 We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records.
-Biome classification were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Austrlia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
+Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
 
-The phylogeny and distribution information were then applied to widely used historical biogeographical methods including likelihood implementations of Dispersal-Vicariance Analysis [@ronquistDispersalVicariance1997] (DIVALIKE), Dispersal-Extinction-Cladogenesis [@reeMaximum2008] (DECLIKE), and BayArea [@landisBayesian2013] (BAYAREALIKE) in `BioGeoBEARS` [@matzkeBioGeoBEARS2013].
-We included two additional free parameters that we thought is relevant.
-The first parameter *j* assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter _x_ modifies the probability of dispersal based on distance between biomes [@vandamEvaluating2016] (see Appendix-method-biogeobears).
-We used likelihood ratio tests to determine if the additional free parameter is a justified improvement over the simpler base models, and selected the model that fitted the data best based on corrected Akaike Information Criterion (AICc) scores.
-Based on the best model we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and trasition between biomes, and to estimate the frequency and directionality of dispersal events across biomes [@matzkeBioGeoBEARS2013].
+We applied the phylogeny and distribution information to widely used historical biogeographical methods, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in `BioGeoBEARS` [@matzkeBioGeoBEARS2013].
+We included two additional free parameters that we thought were relevant.
+The first parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (see Appendix-method-biogeobears).
+We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
+Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and directionality of dispersal events across biomes.[@matzkeBioGeoBEARS2013].
 
@@
 To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation].
-Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myrs resolution.
-We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021.
-From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day.
-Because reptiles are sensitive to extremely low and high temperature, we decided to to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature.
+Each DEM provides paleotopography at a resolution of 1&deg; x 1&deg; and 5 myr resolution.
+We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km, following the methods described in @hagenGen3sis2021.
+From the projected data, we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present day.
+Based on the ____ value, we use XX as the threshold for arid. Thus aridity index is ...
 
-Based on the ____ value, we use XX as the threshold for arid. Thus aridity index is ...
+<!--- Because reptiles are sensitive to extremely low and high temperatures, we decided to summarise the reconstructed paleo temperature data into three different datasets: minimum temperature, maximum temperature, and mean temperature.
+--->
 
@@
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleo environment as described above.
-We also fit the model to widely use mean global paleo temperature data set (CITE).
-To account for missing taxa, we used the fractioning function (f = XX) in `RPANDA`.
+We also fit the model to widely use mean global paleo temperature data set [@condamineMacroevolutionary2013].
+To account for missing taxa, we used the fractioning function (f = 38/50) in `RPANDA`.
 We then assessed which diversification model best fit the data based on AICc scores.
@@
 
-In the second approach we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package `hisse` v.
+In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package `hisse` v.
 2.1.9 [@beaulieuDetecting2016].
 Since these SSE models only allow two states, we restricted our geographic states to arid and non-arid biomes.
-Based on the same categories we used in our BioGeoBEARS anaylses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily other biomes as non-arid (State 1; n = 24), and those that occupy both the arid and non-arid biomes as widespread (State 01; n = 5).
+Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as non-arid (State 1; n = 24), and those that occupy both the arid and non-arid biomes as widespread (State 01; n = 5).
 
-The four SSE models vary in whether they allow diversification rates to be associated with geographic states (range-dependent or range-independent) and whether they allow an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
-Model 1 assumes equal diversification rates regardless of geographic state. Model 2 allows geographic range to affect diversification rates (canonical geoSSE model).
-Models 3 and 4 are based on models 1 and 2, but they also allow unmeasured trait ('hidden states') to contribute to diversification rates (see Table GeoHiSSE table).
-By incorporating hidden states, we can assess whether geographic range or an unmeasured trait contributes more to the observed diversification.
-For each model, we assumed uniform taxon sampling for 89% in arid lineages, 71% in non-arid lineages, and 83% in widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_.
+The four SSE models differed in whether they allowed diversification rates to be dependent on geographic states (range-dependent) or independent of them (range-independent), and whether they allowed an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
+Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates (the canonical geoSSE model).
+Models 3 and 4 were based on models 1 and 2, but they also allowed unmeasured traits ('hidden states') to contribute to diversification rates (see Table GeoHiSSE).
+By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification.
+For each model, we assumed uniform taxon sampling for 89% of arid lineages, 71% of non-arid lineages, and 83% of widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_.
 We then calculated AIC to identify the best-fitting models.
@@
 ## Data visualisation and software
-ClaDS was implemented in `Julia` v.1.8.1 [@bezansonJulia2012], all other analyses were performed in `R` v.4.2 (CITE R).
-Figures and graphs were prepared using a combination of baseR and `ggplot2` (CITE) to process raw output from R.
-The bioregion map was created using the Free and Open Source `QGIS` v.3.6 (CITE Australian Map).
+ClaDS was implemented in `Julia` v.1.8.1 [@bezansonJulia2012], all other analyses were performed in `R` v.4.2 [@rcoreteamLanguage2022].
+Figures and graphs were prepared using a combination of base R functions and `ggplot2` [@wickhamGgplot22016].
+All map graphics were created using the Free and Open Source `QGIS` v.3.6.
 We arranged and finalised all figures using Adobe Illustrator v2022.
@@
 We found that the DEC+_j_+_x_ model was the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
-A likelihood ratio test (LRT) showed that both founder events (+*j* parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+A likelihood ratio test (LRT) showed that both founder events (+*j* parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+_j_+_x_ model.
-We estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see supplementary_table_BioGEOBEARS).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+_j_+_x_ model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table-supp-BioGeoBEARS).
 Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
 Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2±2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
-Our model also indicate that range expansion is expected to decrease as distance increase (free parameter +_X_ = -1.3).
+Our model also indicate that range expansion is expected to decrease as distance increase (free parameter _x_ = -1.3).
 
@@
 
-Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-The model indicates two pulses of biogeographic transitions for each  of the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
-The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the arid biome.
+Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
+The common ancestor of *Group 1* remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
+The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
 
@@
 However, all three paleo temperature models indicate Australia was hotter than the average global temperature.
-Paleoclimate in Australia were also relatively constant (i.e.
-less fluctuation) compared to global temperature.
+Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene glacial cycles.
 
@@
 
-We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) with aridity index derived from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA).
+We found that the exponential aridity-dependent speciation and constant extinction (BEnvDCST) model based on DEM from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10).
 This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (lambda = 1.32, alpha = -0.14) and a constant background extinction rate (mu = 0.23).
@@
 
-Results from our ClaDS analysis show that branch-specific diversification rates gradually decrease towards the present (see fig-ClaDS).
-The mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (see fig-ClaDS).
+We applied ClaDS and SSE models to evaluate whether diversification rates rates differ between lineages in arid, non-arid, and widespread ranges.
+Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (\@ref(fig:fig-clads_sse).
+ClaDS also indicate branch-specific diversification rates gradually decrease towards the present.
 
-We applied GeoHiSSE models to evaluate whether speciation, extinction, and turnover rates differ between lineages in arid, non-arid, and widespread ranges.
-We found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see supp-table-geohisse).
+In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see supp-table-geohisse).
 The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state.
-Compared to the three other models, the best-fitting model accounted for 0.95 of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohiss).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see REF FIGURE and supplementary XX for net diversification rate and extinction rates).
-
-\newpage
+Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohisse).
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see \@ref(fig:fig-clads_sse) and supp-fig-clads-sse for net diversification rate and extinction rates).
 
@@
 In this study we inferred the extant species in Australia has a tropical grassland origin.
-Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioning into the arid biome.
+Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioned into the arid biome.
 Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
@@
 Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent between 25.8 – 16 MYA [@marinTracing2013].
-This is also consistent with other radiations of Australian squamates, where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
+This is also consistent with other radiations of Australian squamates, where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes as they emerge [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
 Given what we know about the current distribution of species included in our phylogeny, our analyses support the idea that *Anilios* have non-Australian origins.
@@
 Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and other closely related Typhlopidae snakes (e.g.
-*Ramphotyphlops*, *Acutotyphlops*) as outgroups will enable us to make inferences about the origin prior to colonising Australia.
+*Ramphotyphlops*, *Acutotyphlops*) as outgroups will enable us to make more informed inferences about the origin prior to colonising Australia.
 
@@
 However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013.
-
-Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the Tropical Grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
-This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene, which has resulted in the dry and sandy landscape present today [@byrneBirth2008].
+Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
+This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene [@byrneBirth2008].
 
@@
 
-## Influence of paleoclimate on diversification rates
+## Influence of paleoenvironment on diversification rates
 
-Changes in paleoenvironmental changes have been linked with diversification rates via range contraction and expansion that influences niche availability. 
-We found that aridity played an important role in 
-- Australian blind snakes follow a different trajectory, one in which increased aridity from the late Miocene to early Pleistocene coincided with reduced diversification rates for genus.
-This is in contrast with other vertebrates, where diversification rates seem to increase with aridification [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015]
-- Warmer and dryer environment may not directly increase speciation rate because blind snakes are physically adapting to warmer temperatures.
-However, climatic fluctuations can contract and fragment available niche for blind snakes andthus influence diversification rates.
-- In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
-However, models that take in account thermal niche tolerance, such as NEMo [CITE], will benefit from paleoclimate reconstructions such as these to inform limits (boundaries).
-We suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these paleoclimate models and/or use them to inform inferences.
+Our best-fitting model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
+Two of our reconstructed paleo aridity index show that aridity increased starting ~25 MYA ago with particular sharp increase from 10 MYA and intensification from 5 MYA onwards.
+Out of the 38 lineages we considered, 5/36 divergence event occured in the last 10 million years when the continent intensively aridified.
+In contrast, 31/36 divergence event occured between 25 - 10 MYA when the continent was getting warming but not entirely arid.
+Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 MYA.
+This pattern is consitent with other studies on vertebrates that have shown that the divergence time for some arid-adpated taxa pre-dated the intense aridifcation in the last 5 million years [@byrneBirth2008;@catulloAridification2014].
+
+While not the best fitting model, we observed that our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013].
+In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
+However, models that take in account thermal niche tolerance, such as NEMo [@huaAdapting2022], will benefit from paleoclimate reconstructions such as these to inform limits (boundaries).
+We suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these paleoclimate models to inform model inferences.
 
@@
 
-## Rates of diversification between arid and non-arid lineages
-
-- The early burst of speciation is observed in lineages that colonise new areas. For example.
+## Diversification in the arid zone
 
@@
 
-- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
-For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite).
+- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
+- Vast arid biome covers 70% of the continent, which the largest biome in Australia with lots of heterogeneity [@pepperLife2021]. 
+- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
+
+- Warmer and dryer environment may not directly decrease speciation rate because blind snakes are physically adapting to warmer temperatures.
+However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.
+
+- The early burst of speciation is observed in lineages that colonise new areas. For example ...
 
@@
 
-Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017]. 
+- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
+For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)
+
+- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
+
 - Extinction is a big cavaet especially because we suspect there are more species.
@@
 
-- Vast arid biome is  the largest biome in Australia with lots of heterogeneity (cite Mitzy's paper). Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
-
 - Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa. 
 - Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have difficulty crossing the lines [@bryantLines2016]
+
+@deepakAridification2018 - accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
+@catulloAridification2014 - multiple independent origing from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
+@fujitaDiversification2010 - independent origins of arid zone lineages in _Heteronotia_
+@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
+@oliverIndependent2014 - multiple independent invasion into the arid zone.
+@owenHow2017 - cicada increase speciation in arid zone compared to mesic zone (SSE)
+@toussaintUnveiling2015 - diving beetle boom and bust diversification. Invaded arid zone and changed ecologically
 
@@
 We show that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
+To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleo-aridity data.
 We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.

### 2022-12-22 manuscript.Rmd
upudate manuscript

@@
 
-Consistent with previous research by @marinTracing2013, we found that the initial diversification of Anilios occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes.
+Consistent with previous research by @marinTracing2013, we found that the initial diversification of _Anilios_ occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes.
 However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013.
@@
 Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles.
-to add discussion to this?
 
@@
 
-- Diversification rate decrease with aridification.
-But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes.
-Turnover rate also ...
+Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008].
+Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
+This pattern is consistent with several other groups vertebrates that started to diversify in the Miocene including goannas [@brennanPhylogenomicsMonitorLizards2021;@pavon-vazquezCompetition2022], frogs [@catulloAridification2014], and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
 
-- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
-- Vast arid biome covers 70% of the continent, which the largest biome in Australia with lots of heterogeneity [@pepperLife2021]. 
-- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
+Although the expansion of the arid zone and fragmentation of mesic environments throughout the Miocene has been linked with increase speciation [cite], studies have found that mountainous areas in the arid zone habour more diversity than the flat featureless desert [@pepperPalaeoclimate2011].
+
+-- new open dry habitats for different species of reptiles to fill [@sandersMolecular2008;@hugallPhylogeny2008], which
+
+our state speciation-extinction analyses on the blind snake phylogeny also indicate that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
 
@@
 However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.
+- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
+For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)
+- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
+
+- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
+- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
 
@@
   - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.
-
-- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
-For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)
-
-- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
 
@@
 
-@deepakAridification2018 - accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
-@catulloAridification2014 - multiple independent origing from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
+<!---
+@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
+@catulloAridification2014 - multiple independent origin from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
 @fujitaDiversification2010 - independent origins of arid zone lineages in _Heteronotia_
-@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
 @oliverIndependent2014 - multiple independent invasion into the arid zone.
@@
 @toussaintUnveiling2015 - diving beetle boom and bust diversification. Invaded arid zone and changed ecologically
+@deepakAridification2018 - India accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
+--->
 

### 2022-12-23 manuscript.Rmd
diversification in the arid zone discussion draft

@@
 To our knowledge, these species share similar dietary and morphological specialisations with their closest relatives, which is a pattern identified across all fossorial snakes [@cyriacDigging2018].
-Because the subterranean microhabitat has a relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for.
-Future studies that investigate phenotypic variations across all blind snakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
 
@@
 Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
-This pattern is consistent with several other groups vertebrates that started to diversify in the Miocene including goannas [@brennanPhylogenomicsMonitorLizards2021;@pavon-vazquezCompetition2022], frogs [@catulloAridification2014], and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
+This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including goannas [@brennanPhylogenomicsMonitorLizards2021;@pavon-vazquezCompetition2022], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020] and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
 
-Although the expansion of the arid zone and fragmentation of mesic environments throughout the Miocene has been linked with increase speciation [cite], studies have found that mountainous areas in the arid zone habour more diversity than the flat featureless desert [@pepperPalaeoclimate2011].
+For these terrestrial reptiles, increased rates of speciation for arid adapted taxa have been attributed to species filling niche in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
+However, results from our weighted-averaged state speciation-extinction analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
+One hypothesis is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much surface dwelling species do.
+For instance, because blind snakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
+Future studies that investigate phenotypic variations across all blind snakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
 
--- new open dry habitats for different species of reptiles to fill [@sandersMolecular2008;@hugallPhylogeny2008], which
+- Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa.
 
-our state speciation-extinction analyses on the blind snake phylogeny also indicate that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
+Warmer and dryer environment may not directly influence speciation rate because blind snakes are physically adapting to warmer temperatures. 
+However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.
 
-- Warmer and dryer environment may not directly decrease speciation rate because blind snakes are physically adapting to warmer temperatures.
-However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.
-- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species.
-For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)
 - Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
-
-- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
-- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.
-
-- The early burst of speciation is observed in lineages that colonise new areas. For example ...
 
@@
 
+##### Geographical things
+- Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have difficulty crossing the lines [@bryantLines2016]
+- Certain regions with complex topographic features tend to harbour more diveristy than flat featureless desert areas [@pepperPalaeoclimate2011]. 
+- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia.
+
+- The early burst of speciation is observed in lineages that colonise new areas. For example ... 
 - Extinction is a big cavaet especially because we suspect there are more species.
@@
 Most species are sampled next to roads [CITE Xia's paper or do supplemental analyses].
-
-- Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa. 
-- Geographic barrier for mesic adapted lineages in the temperate and tropical forests may have difficulty crossing the lines [@bryantLines2016]
 

### 2022-12-23 manuscript.Rmd
round out discussion and some parts of conclusion

@@
 
+<!---
 - Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa. 
-
 Warmer and dryer environment may not directly influence speciation rate because blind snakes are physically adapting to warmer temperatures. 
 However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates. 
-
 - Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017]. 
-
 - Interestingly blind snakes have diversified greatly compared to other fossorial vertebrates. 
@@
 Most species are sampled next to roads [CITE Xia's paper or do supplemental analyses].
-
+--->
 <!---
@@
 ## Limitations
-There are several limitations in our models.
-
-- Without extensive sampling we can not confidently trace the pathways in which _Anilios_ colonised Australia.
-Due to the coarseness of our biomes, we also cannot infer exactly where the ancestral species landed in Australia.
-Current biogeographic models does not account for extinct species and cannot model complex dynamics of geological processes such as continents moving.
-
-- Limitations is we can only coarsely assign species to one biome one some are range-limited endemics on islands.
-For example *A.
-longissimus* has only been collected from 30m underground on Barrow Island located XX km the northwest coast of Western Australia.
-Placing that species with arid biome is not realistic but that area is considered the arid biome by the WWF Ecoregions map.
-
-- While taxonomic sampling the most representative of the genus, it is not comprehensive.
+Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA) may be biased due to the small sample size of our _Anilios_ tree [@caetanoHidden2018;@malietModel2019;@sunEstimating2020].
+However, this near-complete phylogeny of _Anilios_ is is the most comprehensive data set available at this time.
 We suspect there remains several undescribed species of blind snakes in Australia and New Guinea that may belong in the genus _Anilios_.
@@
 
-- Our RPANDA model did not model environment-dependent extinction.
+Our BioGeoBEARS ancestral biome estimation is limited by the areas (e.g.
+in this case biomes) we can assign our extant species to.
+By following the WWF Bioregions maps, we had to assign species that occur in extremely different habitats in the same biome.
+For example, *A.
+longissimus*, a species that has only been collected from 30m underground on Barrow Island located 50 km the northwest off the Pilbara coast of Western Australia, shares the same biome with *A.
+centralis* from the arid interior of the continent.
+Current methods does not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat featureless desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011].
+However, modern software underdevelopment such as PhyBEARS may be able to account for more complex geological processes such as the movement of continent.
+
+Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today. Out of all our compartive analyses, only the SSE model explicit models extinction rates.
+
+<!---
+- Without extensive sampling we can not confidently trace the pathways in which _Anilios_ colonised Australia.
+Due to the coarseness of our biomes, we also cannot infer exactly where the ancestral species landed in Australia.
 
@@
 leucoproctus_) are found between southern New Guinea and Far North Queensland
+--->
 
-- Need to account for heterogeneity of the arid zone [@pepperLife2021].
+-
 
@@
 As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleo environmental datasets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
-We show that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
+We inferred that tropical grassland is the ancestral biome of Australian _Anilios_ and acted as a major source of biodiversity to other biomes.
+Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridify.
+
+We showed that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
 To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleo-aridity data.

### 2023-01-10 manuscript.Rmd
update references and include figure chunks

@@
     reference_docx: [C:/Users/ST/Documents/repo/AusARG_blindsnake/custom-reference-doc.docx]
-    number_sections: true
+    number_sections: false
     fig_caption: yes
@@
 **Keywords**
-scolecophidia, Diversification, paleoclimate, BioGeoBEARS, GeoHiSSE
+Scolecophidia, Diversification, paleoclimate, BioGeoBEARS, GeoHiSSE
 
@@
 The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
-The aridification of continental Australia began around 20 million years ago, with pockets of arid zone developing and subsequent extreme warming and cooling periods in the last ~5 million years creating what is now known as the 'outback' [@byrneBirth2008].
+The aridification of continental Australia began around 20 million years ago, with pockets of arid zone developing and subsequent extreme warming and cooling periods in the last ~5 million years creating what is now known as the "outback" [@byrneBirth2008].
 During this period, arid-adapted vegetation expanded and mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
@@
 Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified in the last 20 - 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
-Examples of reptilest that have colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene include agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013].
+Examples of reptiles that have colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene include agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013].
 
-Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [[@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches.
+Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches.
 Other clades like geckos and blind snakes remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@marinHidden2013;@moritzCryptic2018;@zozayaAre2019].
@@
 Blind snakes are morphologically conserved and often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
-The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19 - 39 mya [@vidalBlindsnake2010;@marinTracing2013].
+The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19 - 39 MYA [@vidalBlindsnake2010;@marinTracing2013].
 However, despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
@@
 Among others, studies on the influence of climate (e.g.
-temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@cantalapiedraDietary2014;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2021], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomicsMonitorLizards2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
-In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and other plants (reviewed in @crispHow2013).
+temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@cantalapiedraDietary2014;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@@brennanPhylogenomics2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
+In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants (reviewed in @crispHow2013).
 
@@
 The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019], these temperature data are reconstructed (interpolated) based on conversion of concentration of oxygen isotopes [@condamineMacroevolutionary2013].
-However, the relationship between diversification rates and aridity are attributed to some proxy such as the the timing of aridification or global temperature rather than quantifying the extent of aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
+However, the relationship between diversification rates and aridity are attributed to some proxy such as the timing of aridification or global temperature rather than quantifying the extent of local aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
 As such, no study has been able to quantitatively assess aridity as a continuous variable in environment-dependent models and compare it to other ecological processes.
-Recent development in geographic-based methods such as `gen3sis` modeling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
+Recent development in geographic-based methods such as `gen3sis` modelling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
 In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia in the time span of interest, between the 30 MYA to present.
-These datasets will allow us to empirically test the influence of localised ecological processes on diversification rates over time.
+These data sets allow us to empirically test the influence of localised ecological processes on diversification rates over time.
 
-The aim of this study is to infer the ancestral biome of Australian blind snakes based on a near-complete phylogeny of extant taxa.
-Additionally, we aim to investigate how paleo-environmental changes have influenced the rates of diversification within this group.
-To do this, we assess the fit of models between paleo-environmental conditions and diversification rates, and compare the rates of diversification between lineages in the arid versus non-arid biome.
+The aim of this study is to infer the ancestral biome of Australian blind snakes radiation and investigate how paleoenvironmental changes have influenced the rates of diversification within this group.
+To do this, we assess the fit of models between paleoenvironmental conditions and diversification rates, and compare the rates of diversification between lineages in the arid versus non-arid biome.
 
@@
 The genetically distinct OTUs that are not fully recognised as species include 3 '_Anilios grypus_' (2 in the west and 1 in the east) and 2 '_Anilios ligatus_' (one from the east and one from the north).
-For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.s
+For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
 
@@
 
-To investigate the biogeographic history of blind snakes, we used the R package `BioGeoBEARS` 1.1.1 [@matzkeBioGeoBEARS2013].
+To investigate the biogeographic history of blind snakes, we used the R package `BioGeoBEARS` v.1.1.1 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
@@
 We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records.
-Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World dataset [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
+Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
 
@@
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
-Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and directionality of dispersal events across biomes.[@matzkeBioGeoBEARS2013].
+Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and direction of dispersal events across biomes.[@matzkeBioGeoBEARS2013].
 
@@
 We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km, following the methods described in @hagenGen3sis2021.
-From the projected data, we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present day.
+From the projected data, we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 MYA to the present day.
 Based on the ____ value, we use XX as the threshold for arid. Thus aridity index is ...
 
-<!--- Because reptiles are sensitive to extremely low and high temperatures, we decided to summarise the reconstructed paleo temperature data into three different datasets: minimum temperature, maximum temperature, and mean temperature.
+<!--- Because reptiles are sensitive to extremely low and high temperatures, we decided to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature.
 --->
@@
 To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` functions in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
-Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleo environment as described above.
-We also fit the model to widely use mean global paleo temperature data set [@condamineMacroevolutionary2013].
+Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
+We also fit the model to widely use mean global paleotemperature data set [@condamineMacroevolutionary2013].
 To account for missing taxa, we used the fractioning function (f = 38/50) in `RPANDA`.
@@
 
-To test whether diversification rates differ between lineages in arid biomes and non-arid lienages, we used two approaches.
-First, we estimated speciation rates at the origin of each branch (hencforth branch-specific rates) and tip speciation rates using the Julia package `ClaDS` [@malietFast2022].
+To test whether diversification rates differ between lineages in arid biomes and non-arid lineages, we used two approaches.
+First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using the Julia package `ClaDS` [@malietFast2022].
 `ClaDS` is designed to capture frequent shifts in speciation rates with small effects.
@@
 Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates (the canonical geoSSE model).
-Models 3 and 4 were based on models 1 and 2, but they also allowed unmeasured traits ('hidden states') to contribute to diversification rates (see Table GeoHiSSE).
+Models 3 and 4 were based on models 1 and 2, but they also allowed unmeasured traits ('hidden states') to contribute to diversification rates.
 By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification.
@@
 Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios* but the ancestor leading to the crown *Anilios* is ambiguous.
-We found that the DEC+_j_+_x_ model was the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
-A likelihood ratio test (LRT) showed that both founder events (+*j* parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+We found that the DEC+j+x model was the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
+A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+_j_+_x_ model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table-supp-BioGeoBEARS).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table-supp-BioGeoBEARS).
 Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
-Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2±2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
-Our model also indicate that range expansion is expected to decrease as distance increase (free parameter _x_ = -1.3).
+Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2 ± 2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
+Our model also indicate that range expansion is expected to decrease as distance increase (free parameter x = -1.3).
 
@@
 
+```{r fig-biogeobears, fig.cap= 'Ancestral biome estimation and transitions of Australo-Papuan blind snakes in BioGeoBEARS, using the DEC+j+x model.
+(a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
+Inset map (left) shows the current distribution of biomes.
+Inset maps to the right of the tree shows distribution of species in Group 1 and Group 2, respectively.
+(b) Histogram showing frequency and composition of outward transitions from Australian biomes and combined non-Australian islands.
+(c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
+Abbreviations: TrG = Tropical grassland, TmF = Temperate forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate grassland.'}
+
+knitr::include_graphics(paste0(WD, '/assets/figures/biogeobears.jpg'))
+
+```
+
 ## Paleoclimate estimates
@@
 Paleoclimate estimates for Australia differed between the three methods (\@ref(fig:fig-paleoclimate-compare))b.
-However, all three paleo temperature models indicate Australia was hotter than the average global temperature.
+However, all three paleotemperature models indicate Australia was hotter than the average global temperature.
 Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene glacial cycles.
+
+```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based different data sets in gen3sis and RPANDA.
+(a) Estimated mean paleo temperature in Australia compared to RPANDA global paleotemperature data.
+(b) Estimated mean aridity index based on precipitation threshold.
+Abbreviations Pli = Paleocene and Ple = Pleistocene.'}
+
+knitr::include_graphics(paste0(WD, '/assets/figures/paleo_data_compare.jpg'))
+
+```
 
@@
 
+```{r Table-RPANDA-TOP10}
+
+# Location of table of top 10 RPANDA
+
+```
+
 ## Diversification rates between arid and non-arid biomes
 
-We applied ClaDS and SSE models to evaluate whether diversification rates rates differ between lineages in arid, non-arid, and widespread ranges.
-Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (\@ref(fig:fig-clads_sse).
+We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, non-arid, and widespread ranges.
+Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (\@ref(fig:fig-clads_sse)).
 ClaDS also indicate branch-specific diversification rates gradually decrease towards the present.
 
-In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see supp-table-geohisse).
+In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see Table-geohisse).
 The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state.
-Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see supp-table-geohisse).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see \@ref(fig:fig-clads_sse) and supp-fig-clads-sse for net diversification rate and extinction rates).
+Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see Table-geohisse).
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see (\@ref(fig:fig-clads_sse))) and supp-fig-clads-sse for net diversification rate and extinction rates).
+
+```{r fig-clads_sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models.
+(a) Branch-specific rates across the phylogeny.
+(b) Mean branch-specific rates over time.
+(c) Density plot of branch-specific rates by geographic range from ClaDS analysis.
+(d) Speciation rates from four geographic diversification rate models averaged by model weight.
+(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages.'}
+
+knitr::include_graphics(paste0(WD, '/assets/figures/fig-clads_sse.jpg'))
+
+```
+
+```{r Table-geohisse}
+
+# Location of table of geohisse results
+
+```
 
@@
 
-With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of scolecophidia in the world.
+With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
 In this study we inferred the extant species in Australia has a tropical grassland origin.
@@
 Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
-We also found that paleo aridity index provides the best fit to diversification dynamics observed in our blind snake tree.
+We also found that paleoaridity index provides the best fit to diversification dynamics observed in our blind snake tree.
 Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found no such association for Australian *Anilios* blind snakes.
@@
 Our best-fitting model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
-Two of our reconstructed paleo aridity index show that aridity increased starting ~25 MYA ago with particular sharp increase from 10 MYA and intensification from 5 MYA onwards.
+Two of our reconstructed paleoaridity index show that aridity increased starting ~25 MYA ago with particular sharp increase from 10 MYA and intensification from 5 MYA onwards.
 Out of the 38 lineages we considered, 5/36 divergence event occured in the last 10 million years when the continent intensively aridified.
-In contrast, 31/36 divergence event occured between 25 - 10 MYA when the continent was getting warming but not entirely arid.
+In contrast, 31/36 divergence event occurred between 25 - 10 MYA when the continent was getting warming but not entirely arid.
 Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 MYA.
-This pattern is consitent with other studies on vertebrates that have shown that the divergence time for some arid-adpated taxa pre-dated the intense aridifcation in the last 5 million years [@byrneBirth2008;@catulloAridification2014].
+This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 million years [@byrneBirth2008;@catulloAridification2014].
 
@@
 Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
-This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including goannas [@brennanPhylogenomicsMonitorLizards2021;@pavon-vazquezCompetition2022], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020] and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
+This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including goannas [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020] and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
 
@@
 However, modern software underdevelopment such as PhyBEARS may be able to account for more complex geological processes such as the movement of continent.
-
-Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today. Out of all our compartive analyses, only the SSE model explicit models extinction rates.
+Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today.
+Out of all our comparative analyses, only the SSE model explicit models extinction rates.
 
@@
 
--
-
 # Conclusion
 
-As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleo environmental datasets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
+As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleoenvironmental data sets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
 We inferred that tropical grassland is the ancestral biome of Australian _Anilios_ and acted as a major source of biodiversity to other biomes.
 Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridify.
-
 We showed that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
@@
 
-# Tables
-
-Table BioGeoBEARS. BioGeoBEARS model estimates and model AICc scores. Lower AICc scores indicate model is preferred. + Log likelihood
-Table RPANDA. Diversification model fit estimates and AICc score of model fit
+\newpage
 
@@
 
-Fig. Ancestral biome: Stochastic map of Ancestral biome from Australo-Papuan blind snakes
 Fig.
-Paleoclimate: Reconstructed paleoclimate values from different data sets.
-Mean temperature data.
-Aridity index based on precipitation threshold.
+Ancestral biome estimation and transitions of Australo-Papuan blind snakes in BioGeoBEARS, using the DEC+j+x model.
+(a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
+Inset map (left) shows the current distribution of biomes.
+Inset maps to the right of the tree shows distribution of species in Group 1 and Group 2, respectively.
+(b) Histogram showing frequency and composition of outward transitions from Australian biomes and combined non-Australian islands.
+(c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
+Abbreviations: TrG = Tropical grassland, TmF = Temperate forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate grassland.
+
+Fig.
+Paleoclimate estimation based different data sets in gen3sis and RPANDA.
+(a) Estimated mean paleo temperature in Australia compared to RPANDA global paleotemperature data.
+(b) Estimated mean aridity index based on precipitation threshold.
 Abbreviations Pli = Paleocene and Ple = Pleistocene.
-Fig. Speciation rates + phylogeny over time + lineages through time
+
+Fig.
+Diversification rates estimates from ClaDS and SSE models.
+(a) Branch-specific rates across the phylogeny.
+(b) Mean branch-specific rates over time.
+(c) Density plot of branch-specific rates by geographic range from ClaDS analysis.
+(d) Speciation rates from four geographic diversification rate models averaged by model weight.
+(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages
 

### 2023-01-10 code/01_env_data_plots.R
update env plotter

@@ -34,11 +34,21 @@ env_plottr <- function(.df){
   ggplot(data = ., aes(x = age, colour = source)) +
     geom_line(aes(y = vals), size = 1) +
     geom_point(aes(y = vals)) + 
-    deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
+    deeptime::coord_geo(xlim = c(26, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
     # scale_colour_manual(values = viridis(4), name = "Data source", labels = c("RPANDA Global", "Scotese Australia", "Straume Australia", "Valdes Australia")) +
-    scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Mean temperature (°C)") +
-    scale_x_reverse("Age (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() 
+    scale_y_continuous(limits = c(-5, 30), breaks = seq(-5, 30, 5)) + labs(y = "Mean temperature (°C)") +
+    scale_x_reverse("Age (Ma)", limits = c(25, 0), breaks = seq(25, 0, by = -5)) + theme_bw() 
 }
+# env_plottr <- function(.df){
+#   .df %>% 
+#   ggplot(data = ., aes(x = age, colour = source)) +
+#     geom_line(aes(y = vals), size = 1) +
+#     geom_point(aes(y = vals)) + 
+#     deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
+#     # scale_colour_manual(values = viridis(4), name = "Data source", labels = c("RPANDA Global", "Scotese Australia", "Straume Australia", "Valdes Australia")) +
+#     scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Mean temperature (°C)") +
+#     scale_x_reverse("Age (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() 
+# }
 
 # pdf(file = "output/paleo_env_data_plots.pdf", height = 10, width = 10)
 
@@ -48,14 +58,20 @@ env_df[which(env_df$source %in% c("mean_str", "max_str", "min_str")),] %>% env_p
 
[…]

### 2023-01-10 code/04_BioGeoBEARS_bsm_plots.R
change name of output

@@ -154,9 +154,9 @@ inwards_disp <- ggplot(data = mean_in_disp_area, aes(fill= from, y = count, x = 
   guides(fill = guide_legend(title = "From")) +
   theme(legend.position = "none")
 
-geom_text(data = df2,
-          aes(y = Freq + 1,
-              label = sprintf('%.2f%%', Prop)))# inwards dispersals geo state
+# geom_text(data = df2,
+#           aes(y = Freq + 1,
+#               label = sprintf('%.2f%%', Prop)))# inwards dispersals geo state
 inwards_disp_geo <- ggplot(data = mean_in_disp_area, aes(fill= state_from, y = count, x = state_to)) + 
   geom_bar(position="stack", stat="identity") +
   labs(x = "Inwards dispersal", y = "Mean inwards transition frequency") + 
@@ -181,7 +181,8 @@ inwards_disp_grp <- ggplot(data = mean_in_disp_area, aes(fill= subset_from, y = 
 ### stitch together with patchwork package
 
 # Save PDF to be included as inset in BioGeoBEARS plot
-pdf(file = 'output/biogeobears_dispersal_freq.pdf', width = 11.5, height = 5.7)
+# pdf(file = 'output/biogeobears_dispersal_freq.pdf', width = 11.5, height = 5.7)
+pdf(file = 'output/biogeobears_dispersal_mean.pdf', width = 11.5, height = 5.7)
 outwards_disp_grp +inwards_disp_grp
 dev.off()
 

### 2023-01-10 manuscript.Rmd
update fig references

@@
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
-Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and direction of dispersal events across biomes.[@matzkeBioGeoBEARS2013].
+Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and direction of dispersal events across biomes [@matzkeBioGeoBEARS2013].
 
@@
 
-The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%), which makes up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d.
+The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%), which makes up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (Fig.
+\@ref(fig:fig-biogeobears)c-d).
 Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
 On the other hand, the largest sink for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
-The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (\@ref(fig:fig-biogeobears)c-d.
+The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (Fig.
+\@ref(fig:fig-biogeobears)c-d).
 
 Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a.
+We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (Fig.
+\@ref(fig:fig-biogeobears)a).
 The common ancestor of *Group 1* remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
@@
 
-Paleoclimate reconstructions of aridity differed slightly between the two methods \@ref(fig:fig-paleoclimate-compare)a.
+Paleoclimate reconstructions of aridity differed slightly between the two methods (Fig.
+\@ref(fig:fig-paleoclimate-compare)a).
 Both aridity index shows increase in aridity post 10 MYA, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
-Paleoclimate estimates for Australia differed between the three methods (\@ref(fig:fig-paleoclimate-compare))b.
+Paleoclimate estimates for Australia differed between the three methods (Fig.
+\@ref(fig:fig-paleoclimate-compare)b).
 However, all three paleotemperature models indicate Australia was hotter than the average global temperature.
@@
 We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, non-arid, and widespread ranges.
-Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (\@ref(fig:fig-clads_sse)).
+Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (Fig.
+\@ref(fig:fig-clads-sse)).
 ClaDS also indicate branch-specific diversification rates gradually decrease towards the present.
@@
 Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see Table-geohisse).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see (\@ref(fig:fig-clads_sse))) and supp-fig-clads-sse for net diversification rate and extinction rates).
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see Fig.
+\@ref(fig:fig-clads-sse) and supp-fig-clads-sse for net diversification rate and extinction rates).
 
-```{r fig-clads_sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models.
+```{r fig-clads-sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models.
 (a) Branch-specific rates across the phylogeny.
@@
 
-knitr::include_graphics(paste0(WD, '/assets/figures/fig-clads_sse.jpg'))
+knitr::include_graphics(paste0(WD, '/assets/figures/fig-clads-sse.jpg'))
 
@@
 
+# Tables.
+
+Table-RPANDA-TOP10.
+Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
+The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (lambda = speciation rate at present, alpha = parameter controlling the dependency of speciation rate on time or paleoclimate, mu = extinction rate at present, beta = parameter controlling the dependency of extinction rate on time or paleoclimate).
+
+Table geo-hisse. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+
 \newpage

### 2023-01-16 manuscript.Rmd
modified paragraphs and cleaned up a few errors

@@
 During this period, arid-adapted vegetation expanded and mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
-This period of aridification is believed to have involved the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+This period of aridification is believed to have involved the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
 
-Today, the arid zone covers the majority of area in the Australian landscape (70%) and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
-Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which diversified in the last 20 - 6 myr [@brennanMass2017;@oliverPhylogenetic2017].
-Examples of reptiles that have colonised Australia from elsewhere and have diversified remarkably from the beginning of the Miocene include agamid dragons [@chenSexual2012], monitor lizards [@brennanPhylogenomics2021], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], _Gehyra_ geckos [@ashmanDiversification2018], elapid snakes [@leeDiversification2016], pythons [@esquerrePhylogenomics2020], and blind snakes [@marinTracing2013].
+Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which colonised the continent from elsewhere and have diversified in the last 20 – 6 MYRS [@brennanMass2017;@oliverPhylogenetic2017].
 
-Some clades, like Elapid snakes [@leeDiversification2016;@sandersMolecular2008] and scincid lizards [@raboskyExceptional2007] showed dramatic ecological diversity taking up different niches.
-Other clades like geckos and blind snakes remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@marinHidden2013;@moritzCryptic2018;@zozayaAre2019].
+Some clades, like elapid snakes and skinks, showed dramatic morphological and ecological differentitation [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007].
+Other clades, like some geckos and blind snakes, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
 The species richness and distribution of squamate reptiles across different biomes makes them a good group to test hypotheses about the tempo and mode of diversification in Australia.
 
-Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021].
+Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (**Tiatragul et al.
+2023 lumped on**).
 All currently described species of _Anilios_ are found in Australia except for _A.
@@
 Blind snakes are morphologically conserved and often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
-The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic adapted ancestor from Asia between 19 - 39 MYA [@vidalBlindsnake2010;@marinTracing2013].
+The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 MYA [@vidalBlindsnake2010;@marinTracing2013].
 However, despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
-The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted to the continent and can be found in every biome.
+The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted and can be found in every biome on mainland Australia.
 
@@
 Among others, studies on the influence of climate (e.g.
-temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@cantalapiedraDietary2014;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@@brennanPhylogenomics2021], CO<sub>2</sub> [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
+temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomics2021], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
 In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants (reviewed in @crispHow2013).
 
-There are several links between temperature and diversification rates of reptiles.
-First, warm temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
+In reptiles, variation in climate has been proposed to influence diversification rates.
+For instance, warmer temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
 Second, spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
@@
 Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
-Finally, aridification of Australia may offer new ecological niche for species to exploit, which have been shown to result in deep divergences in many arid-adapted [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013].
+Finally, aridification of Australia may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised Australia from elsewhere.
+ <!---which have been documented in several arid-adapted lineages with deep divergences [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013]. --->
 
-The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019], these temperature data are reconstructed (interpolated) based on conversion of concentration of oxygen isotopes [@condamineMacroevolutionary2013].
-However, the relationship between diversification rates and aridity are attributed to some proxy such as the timing of aridification or global temperature rather than quantifying the extent of local aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
-As such, no study has been able to quantitatively assess aridity as a continuous variable in environment-dependent models and compare it to other ecological processes.
-Recent development in geographic-based methods such as `gen3sis` modelling engine [@hagenGen3sis2021] enables us to simulate climatic, geological, and ecological factors that are spatially and temporally specific to certain areas.
-In practice, these outputs from `gen3sis` provide paleotemperature and aridity data specific to Australia in the time span of interest, between the 30 MYA to present.
+The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019].
+Traditionally, these temperature data are reconstructed (interpolated) based on conversion of global concentration of oxygen isotopes [@condamineMacroevolutionary2013].
+The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of local aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
+To our knowledge no study have assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity.
+However, recent development in geographic-based methods such as `gen3sis` modelling engine [@hagenGen3sis2021] enables us to simulate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
 These data sets allow us to empirically test the influence of localised ecological processes on diversification rates over time.
@@
 The aim of this study is to infer the ancestral biome of Australian blind snakes radiation and investigate how paleoenvironmental changes have influenced the rates of diversification within this group.
-To do this, we assess the fit of models between paleoenvironmental conditions and diversification rates, and compare the rates of diversification between lineages in the arid versus non-arid biome.
+To do this, we fit several birth-death models and estimate diversification rates under different conditions.
+We further compare diversification rates between arid-adapted lineages versus non arid-adapted lineages.
 
@@
 We employed a time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al.
-in review).
-This tree includes 38 operational taxonomic units (OTUs) representing 36 described species of _Anilios_.
-However, our phylogenetic data suggest the existence of some cryptic species, so we adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully recognised as species include 3 '_Anilios grypus_' (2 in the west and 1 in the east) and 2 '_Anilios ligatus_' (one from the east and one from the north).
+2023).
+This tree includes 38 operational taxonomic units (OTUs) representing 36/47 described species of _Anilios_.
+However, our phylogenetic data suggest the existence of at least three genetically distinct undescribed species, so we adjusted our taxon sampling fraction to 38/50 species.
+The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (see Tiatragul et al.
+2023).
 For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
@@
 
-To investigate the biogeographic history of blind snakes, we used the R package `BioGeoBEARS` v.1.1.1 [@matzkeBioGeoBEARS2013].
+To investigate the biogeographic history of blind snakes, we used the R package BioGeoBEARS v.1.1.1 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
 These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
-The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.
+<!--The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.  -->
 We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records.
@@
 
-We applied the phylogeny and distribution information to widely used historical biogeographical methods, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in `BioGeoBEARS` [@matzkeBioGeoBEARS2013].
+We applied the phylogeny and distribution information to widely used historical biogeographical methods, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
 We included two additional free parameters that we thought were relevant.
-The first parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (see Appendix-method-biogeobears).
+The first parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix-method-biogeobears).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation].
-Each DEM provides paleotopography at a resolution of 1&deg; x 1&deg; and 5 myr resolution.
+Each DEM provides paleotopography at a resolution of 1&deg; x 1&deg; and 5 million years intervals.
 We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km, following the methods described in @hagenGen3sis2021.
@@
 
-To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the `fit_bd` and `fit_env` functions in the R package `RPANDA` [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package RPANDA [@morlonRPANDA2016] to the _Anilios_ phylogeny.
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
 We also fit the model to widely use mean global paleotemperature data set [@condamineMacroevolutionary2013].
-To account for missing taxa, we used the fractioning function (f = 38/50) in `RPANDA`.
+To account for missing taxa, we used the fractioning function (f = 38/50) in RPANDA.
 We then assessed which diversification model best fit the data based on AICc scores.
@@
 To test whether diversification rates differ between lineages in arid biomes and non-arid lineages, we used two approaches.
-First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using the Julia package `ClaDS` [@malietFast2022].
-`ClaDS` is designed to capture frequent shifts in speciation rates with small effects.
+First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using the Julia package ClaDS [@malietFast2022].
+ClaDS is designed to capture frequent shifts in speciation rates with small effects.
 At each speciation event, daughter lineages are assigned its own speciation rate sampled from a distribution parameterised on the parental rate [@malietFast2022].
 
-In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package `hisse` v.
+In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package hisse v.
 2.1.9 [@beaulieuDetecting2016].
@@
 The four SSE models differed in whether they allowed diversification rates to be dependent on geographic states (range-dependent) or independent of them (range-independent), and whether they allowed an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
-Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates (the canonical geoSSE model).
+Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates.
 Models 3 and 4 were based on models 1 and 2, but they also allowed unmeasured traits ('hidden states') to contribute to diversification rates.
@@
 We then calculated AIC to identify the best-fitting models.
-Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters and ancestral geographic range estimate across all models according to their AIC weights.
+Following the approach in @caetanoHidden2018, we averaged the estimates of diversification parameters across all models according to their AIC weights.
 
 ## Data visualisation and software
-ClaDS was implemented in `Julia` v.1.8.1 [@bezansonJulia2012], all other analyses were performed in `R` v.4.2 [@rcoreteamLanguage2022].
-Figures and graphs were prepared using a combination of base R functions and `ggplot2` [@wickhamGgplot22016].
-All map graphics were created using the Free and Open Source `QGIS` v.3.6.
+ClaDS was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2 [@rcoreteamLanguage2022].
+Figures and graphs were prepared using a combination of base R functions and ggplot2 [@wickhamGgplot22016].
+All map graphics were created using the Free and Open Source QGIS v.3.6.
 We arranged and finalised all figures using Adobe Illustrator v2022.
@@
 Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios* but the ancestor leading to the crown *Anilios* is ambiguous.
-We found that the DEC+j+x model was the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS).
+We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, see Table-supp-BioGeoBEARS).
 A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
@@
 Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
-Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8±2.1 events) followed by subset sympatry (5.2 ± 2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
+Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8 ± 2.1 events) followed by subset sympatry (5.2 ± 2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
 Our model also indicate that range expansion is expected to decrease as distance increase (free parameter x = -1.3).
 
-The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%), which makes up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (Fig.
+The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events), which makes up  43.6% compared outward dispersals from the arid biome (9.7 ± 3.2 events) of 26% (Fig.
 \@ref(fig:fig-biogeobears)c-d).
@@
 Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
-We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (Fig.
+We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n = 21 lineages, not including _A.
+proximus_) and *Group 2* (n = 14 lineages, not including _A.
+silvia_) (Fig.
 \@ref(fig:fig-biogeobears)a).
@@
 knitr::include_graphics(paste0(WD, '/assets/figures/biogeobears.jpg'))
-
 ```
 
-## Paleoclimate estimates
+## Paleoenvironment estimates
 
-Paleoclimate reconstructions of aridity differed slightly between the two methods (Fig.
+Paleotemperature estimates from gen3sis for Australia differed between the three DEMs (Fig.
+\@ref(fig:fig-paleoclimate-compare)b).
+However, all three paleotemperature models indicate Australia was hotter than the average global temperature (Fig.
 \@ref(fig:fig-paleoclimate-compare)a).
+Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles.
+Reconstructions of paleoaridity differed slightly between the two methods (Fig.
+\@ref(fig:fig-paleoclimate-compare)b).
 Both aridity index shows increase in aridity post 10 MYA, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
-Paleoclimate estimates for Australia differed between the three methods (Fig.
-\@ref(fig:fig-paleoclimate-compare)b).
-However, all three paleotemperature models indicate Australia was hotter than the average global temperature.
-Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene glacial cycles.
 
 ```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based different data sets in gen3sis and RPANDA.
-(a) Estimated mean paleo temperature in Australia compared to RPANDA global paleotemperature data.
+(a) Estimated mean paleotemperature in Australia compared to RPANDA global paleotemperature data.
 (b) Estimated mean aridity index based on precipitation threshold.
@@
 
-We found that the exponential aridity-dependent speciation and constant extinction (BEnvDCST) model based on DEM from @scoteseAtlas2021 provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10).
-This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (lambda = 1.32, alpha = -0.14) and a constant background extinction rate (mu = 0.23).
-The second and third most competitive models (delta AICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10).
-Among the exponential environment-dependent models, we found that Australia-specific paleoclimate data provided a better fit to our phylogeny than global paleoclimate data (see supp-TABLE_FULL_RPANDA_model_fit).
+We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10).
+This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23).
+The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10).
+Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).
 
@@
 \@ref(fig:fig-clads-sse)).
-ClaDS also indicate branch-specific diversification rates gradually decrease towards the present.
+ClaDS also indicate branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present.
 
@@
 knitr::include_graphics(paste0(WD, '/assets/figures/fig-clads-sse.jpg'))
-
 ```
@@
 ```{r Table-geohisse}
-
 # Location of table of geohisse results
-
 ```
@@
 With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
-In this study we inferred the extant species in Australia has a tropical grassland origin.
+In this study we inferred the species in Australia occupied the tropical grassland biome.
 Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioned into the arid biome.
-Subsequent transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
-We also found that paleoaridity index provides the best fit to diversification dynamics observed in our blind snake tree.
-Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found no such association for Australian *Anilios* blind snakes.
+Subsequent biome transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
+We also found that paleoaridity index provides the best fit to the diversification dynamics observed in our time-calibrated phylogeny.
+Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found that aridification decreased diversification rates for Australian *Anilios* blind snakes.
 
@@
 However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as proposed by @vidalBlindsnake2010 is equivocal.
-Instead, our results suggest the ancestor of Australian _Anilios_ most likely occurred either in the Bismarck Archipelago, New Guinea, or a combination of both.
-Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and other closely related Typhlopidae snakes (e.g.
-*Ramphotyphlops*, *Acutotyphlops*) as outgroups will enable us to make more informed inferences about the origin prior to colonising Australia.
+Our results indicate the common ancestor of Australian _Anilios_, _Sundatyphlops polygrammicus_, _Ramphotyphlops multilineatus_ and _Acutotyphlops subocularis_ most likely occurred either in the Bismarck Archipelago, New Guinea, or a combination of both.
+Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin prior to colonising Australia.
 
@@
 
-Consistent with previous research by @marinTracing2013, we found that the initial diversification of _Anilios_ occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes.
+Consistent with previous research by @marinTracing2013, we found that the initial diversification of _Anilios_ occurred in the tropical grassland biome, followed by transitions and/or expansion into other biomes.
 However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013.
@@
 Out of the 38 lineages we considered, 5/36 divergence event occured in the last 10 million years when the continent intensively aridified.
-In contrast, 31/36 divergence event occurred between 25 - 10 MYA when the continent was getting warming but not entirely arid.
+In contrast, 31/36 divergence event occurred between 25 – 10 MYA when the continent was getting warming but not entirely arid.
 Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 MYA.
@@
 Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
-This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including goannas [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020] and several geckos [@fujitaDiversification2010;@oliverIndependent2014].
+This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
 
 For these terrestrial reptiles, increased rates of speciation for arid adapted taxa have been attributed to species filling niche in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
-However, results from our weighted-averaged state speciation-extinction analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
-One hypothesis is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much surface dwelling species do.
+However, results from our weighted-averaged SSE analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
+One hypothesis is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
 For instance, because blind snakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
@@
 ## Limitations
-Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA) may be biased due to the small sample size of our _Anilios_ tree [@caetanoHidden2018;@malietModel2019;@sunEstimating2020].
+Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA) may be biased due to the small sample size of our _Anilios_ phylogeny [@caetanoHidden2018;@malietModel2019;@sunEstimating2020].
 However, this near-complete phylogeny of _Anilios_ is is the most comprehensive data set available at this time.
@@
 Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridify.
-We showed that tools like `gen3sis` can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
+We showed that tools like gen3sis can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
 To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleo-aridity data.
@@
 Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
-The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (lambda = speciation rate at present, alpha = parameter controlling the dependency of speciation rate on time or paleoclimate, mu = extinction rate at present, beta = parameter controlling the dependency of extinction rate on time or paleoclimate).
+The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
 

### 2023-01-17 manuscript.Rmd
update abstract section and wrote some manuscript title

@@
 ---
-title: "Diversification of Australian blind snakes"
+title: ""
 date: "`r Sys.Date()`"
@@
 
+Possible titles:
+- Diversification rates of Australian blind snakes slows down as the continent became more dry up.
+- Reconstructed paleoaridity correlate with decreased diversification rates in Australian blind snakes.
+- Drying up speciation: aridification of Australia slowed down diversification of Australian blind snakes
+- Dull when dry: increased aridity correlate with decrease diversification in blind snakes
+
 **Abstract**
+
+**Aim**: Shifts in diversification rates of Australian flora and fauna has been associated with aridification, but the relationship between diversification rates and aridity ifself has never been quantified.
+We investigate how paleoenvironmental changes have influenced the diversification rates of the mostly endemic Australian blind snakes genus, Anilios.
+We explored how changes in temperature and aridity during the Neogene have influenced the diversification of the Australian blind snake and we further investigate whether diversification rates differ between arid-adapted lineages vs mesic-adapted lineages.
 
@@
 
-**Methods**:
+**Materials and Methods**: We infer the ancestral biome of Australian blind snakes using BioGeoBEARS.
+We reconstruct the paleoenvironments of Australia during the Neogene using the gen3sis engine.
+We fit several birth-death models and estimate diversification rates under paleoenvironmental conditions using RPANDA.
+We further compare diversification rates between arid-adapted lineages versus mesic-adapted lineages using ClaDS and GeoHiSSE.
 
-**Results**:
+**Results**: Results from our BioGeoBEARS analyses indicate that Australian blind snakes have tropical grassland origins.
+We found that local Australian paleoenvironments provide better fit than global paleoenvironments.
+Diversification rates of blind snakes decreased with increasing aridity.
+We found no difference in diversification rates between arid- and mesic-adapted lineages.
 
-**Main conclusions**:
+**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blind snakes diversified rapidly in mesic habitats during early Miocene.
+When the continent became more arid, diversification rates decreased.
+However, two clades emerged with one staying mostly in mesic habitats and the other exploiting the burgeoning arid biome.
+Both the arid and tropical biomes were important sources and sinks of subsequent diversification.
 
 **Keywords**
-Scolecophidia, Diversification, paleoclimate, BioGeoBEARS, GeoHiSSE
+Scolecophidia, Diversification, gen3sis, BioGeoBEARS, GeoHiSSE
 
@@
 We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10).
-This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23).
+This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23) (Fig.
+\@ref(fig:fig-RPANDA)).
 The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10).
 Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).
+
+```{r fig-RPANDA, fig.cap= 'Effect of aridity change on the diversification of Anilios estimated from the phylogeny.
+(a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
+(b) Speciation rates through time from the best diversification model.
+(c) Association between speciation rate and aridity index.
+We found that increased aridity is associated with decrease speciation rate.'}
+
+knitr::include_graphics(paste0(WD, '/assets/figures/RPANDA_fit_arid.jpg'))
+```
 
@@
 Fig.
+Effect of aridity change on the diversification of Anilios estimated from the phylogeny.
+(a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
+(b) Speciation rates through time from the best diversification model.
+(c) Association between speciation rate and aridity index.
+We found that increased aridity is associated with decrease speciation rate.
+
+Fig.
 Diversification rates estimates from ClaDS and SSE models.

### 2023-01-19 manuscript.Rmd
incorporated text from Alex Skeels and made changes in response to comments. Still need to explain why increase aridity decrease diversification

@@
 ---
-title: ""
+title: "Dessicated diversification: how aridification slowed down speciation of Australian blind snakes"
 date: "`r Sys.Date()`"
-author: 
+author: "Sarin Tiatragul^1^; Alexander Skeels^2,3^; J. Scott Keogh^1^"
 output: 
@@
 
-Possible titles:
+### Possible titles:
+
 - Diversification rates of Australian blind snakes slows down as the continent became more dry up.
+
 - Reconstructed paleoaridity correlate with decreased diversification rates in Australian blind snakes.
+
 - Drying up speciation: aridification of Australia slowed down diversification of Australian blind snakes
+
 - Dull when dry: increased aridity correlate with decrease diversification in blind snakes
 
-**Abstract**
+**Address**: Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
 
-**Aim**: Shifts in diversification rates of Australian flora and fauna has been associated with aridification, but the relationship between diversification rates and aridity ifself has never been quantified.
+1 Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
+
+2 Department of Environmental Systems Sciences, Landscape Ecology, Institute of Terrestrial Ecosystems, ETH Zürich, Zurich 8092, Switzerland;
+
+3 Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf 8903, Switzerland;
+
+**Corresponding author**: Sarin Tiatragul (sarin.tiatragul@anu.edu.au)
+
+\newpage
+
+### J. Biogeography format::
+
+**Aim**: Shifts in diversification rates of Australian flora and fauna has been associated with the process of aridification, but the relationship between diversification rates and aridity has never been quantified.
 We investigate how paleoenvironmental changes have influenced the diversification rates of the mostly endemic Australian blind snakes genus, Anilios.
@@
 
-**Materials and Methods**: We infer the ancestral biome of Australian blind snakes using BioGeoBEARS.
-We reconstruct the paleoenvironments of Australia during the Neogene using the gen3sis engine.
+**Materials and Methods**: We estimated the historical biogeography of blind snakes using BioGeoBEARS. 
+We synthesised multiple approaches to reconstruct paleotemperature and paleoaridity of Australia during the Neogene.
 We fit several birth-death models and estimate diversification rates under paleoenvironmental conditions using RPANDA.
@@
 
-**Results**: Results from our BioGeoBEARS analyses indicate that Australian blind snakes have tropical grassland origins.
-We found that local Australian paleoenvironments provide better fit than global paleoenvironments.
-Diversification rates of blind snakes decreased with increasing aridity.
+**Results**: Ancestral area estimation indicated Australian blind snakes have tropical grassland origins.
+We found that local Australian paleoenvironments provided a better explanation for diversification rate variation than global paleotemperature.
+Speciically, diversification rates of blind snakes decreased with increasing aridity.
 We found no difference in diversification rates between arid- and mesic-adapted lineages.
@@
 **Keywords**
-Scolecophidia, Diversification, gen3sis, BioGeoBEARS, GeoHiSSE
+Scolecophidia, Diversification, paleoenvironment, BioGeoBEARS, GeoHiSSE
 
@@
 The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
-The aridification of continental Australia began around 20 million years ago, with pockets of arid zone developing and subsequent extreme warming and cooling periods in the last ~5 million years creating what is now known as the "outback" [@byrneBirth2008].
+After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annums) with the development of pockets of aridity [@byrneDecline2011].
+Subsequent extreme warming and cooling periods in the last ~5 Ma creating what is now known as the "outback" [@byrneBirth2008;@pepperLife2021].
 During this period, arid-adapted vegetation expanded and mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
-This period of aridification is believed to have involved the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
+This period of aridification is believed to have promoted the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;
+@owenHow2017].
 
-Squamate reptiles (lizards and snakes) represent the largest vertebrate diversity in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which colonised the continent from elsewhere and have diversified in the last 20 – 6 MYRS [@brennanMass2017;@oliverPhylogenetic2017].
+<!-- Environmental correlates of diversification -->
+A variety of ecological correlates have been associated with diversification processes of a clade.
+Among others, studies on the influence of climate (e.g.
+temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomics2021], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
+In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
 
+<!-- paragraph 5 -->
+In reptiles, variation in climate has been proposed to influence diversification rates.
+For instance, warmer temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
+Spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
+Change in climate (e.g.
+glaciation and aridification) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
+Ectotherms are subjected to higher metabolic rates in warm climates (i.e.
+Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
+Finally, aridification of Australia may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised Australia from elsewhere.
+ <!---which have been documented in several arid-adapted lineages with deep divergences [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013]. --->
+
+<!-- paragraph 3 -->
+Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
+Squamate reptiles (lizards and snakes) are the largest terrestrial vertebrate group in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which colonised the continent from elsewhere and diversified in the last 20 – 6 Ma [@brennanMass2017;@oliverPhylogenetic2017;@brennanPhylogenomics2021;@esquerrePhylogenomics2020].
 Some clades, like elapid snakes and skinks, showed dramatic morphological and ecological differentitation [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007].
@@
 
+<!-- paragraph 4 -->
 Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (**Tiatragul et al.
-2023 lumped on**).
-All currently described species of _Anilios_ are found in Australia except for _A.
+under review**).
+All currently described species of _Anilios_ are fossorial and are found in Australia except for _A.
 eryncinus_, which is endemic to New Guinea.
-Blind snakes are morphologically conserved and often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
-The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 MYA [@vidalBlindsnake2010;@marinTracing2013].
-However, despite representing a substantial portion of snake biodiversity, there has been few investigations into the macroevolutionary dynamics of blind snakes.
+Blind snakes are morphologically conserved, species often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
+The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
+Despite representing a substantial portion of snake biodiversity, we know little about the macroevolutionary dynamics of blind snakes.
 The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted and can be found in every biome on mainland Australia.
 
-<!-- Environmental correlates of diversification -->
-<!-- agamid lizards [@harmonTempo2003], passerine birds [@ricklefsGlobal2006], amphibians [@rollandContribution2019], rosid plants [@sunRecent2020]).  -->
-A variety of ecological correlates have been associated with diversification processes of a clade.
-Among others, studies on the influence of climate (e.g.
-temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomics2021], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
-In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants (reviewed in @crispHow2013).
-
-In reptiles, variation in climate has been proposed to influence diversification rates.
-For instance, warmer temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
-Second, spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
-Third, change in climate (e.g.
-glaciation and aridification) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
-Fourth, ectotherms are subjected to higher metabolic rates in warm climates (i.e.
-Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
-Finally, aridification of Australia may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised Australia from elsewhere.
- <!---which have been documented in several arid-adapted lineages with deep divergences [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013]. --->
-
+<!-- paragraph 6 -->
 The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019].
@@
 The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of local aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
-To our knowledge no study have assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity.
-However, recent development in geographic-based methods such as `gen3sis` modelling engine [@hagenGen3sis2021] enables us to simulate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
-These data sets allow us to empirically test the influence of localised ecological processes on diversification rates over time.
 
-The aim of this study is to infer the ancestral biome of Australian blind snakes radiation and investigate how paleoenvironmental changes have influenced the rates of diversification within this group.
-To do this, we fit several birth-death models and estimate diversification rates under different conditions.
+<!-- Paragraph 7 -->
+To our knowledge no study has assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity.
+However, recent developments in spatially-explicit paleoenvironmental reconstructions from the lithological and fossil record as well as global circulation models (GCMs) now enables us to estimate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
+These data sets for the first time enables us to empirically test the effects of localised environmental changes on diversification rates over time.
+
+<!-- Paragraph 8 -->
+The aim of this study is to infer the historical biogeography of the Australian blind snake radiation and investigate how paleoenvironmental changes have influenced the rates of diversification within this group.
+To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimate diversification rates under these models.
 We further compare diversification rates between arid-adapted lineages versus non arid-adapted lineages.
@@
 <!-- Based on our phylogenetic tree -->
-We employed a time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al.
-2023).
+We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al.
+under review).
 This tree includes 38 operational taxonomic units (OTUs) representing 36/47 described species of _Anilios_.
+The remaining 11 species could not be included because no molecular data or tissues samples are available.
 However, our phylogenetic data suggest the existence of at least three genetically distinct undescribed species, so we adjusted our taxon sampling fraction to 38/50 species.
 The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (see Tiatragul et al.
-2023).
+under review).
 For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
@@
 BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
-These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal, extinction, or switching to adjacent biomes) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
+These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal and extinction) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
 <!--The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.  -->
@@
 
-We applied the phylogeny and distribution information to widely used historical biogeographical methods, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
-We included two additional free parameters that we thought were relevant.
-The first parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix-method-biogeobears).
+Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in the R package BioGeoBEARS [@matzkeBioGeoBEARS2013].
+We also fitted these three models, with two additional free parameters that we thought were relevant, both singly and together, for a total of 12 alternative models.
+The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix-method-biogeobears).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 <!-- Consult Alex Skeels -->
-To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation].
-Each DEM provides paleotopography at a resolution of 1&deg; x 1&deg; and 5 million years intervals.
-We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km, following the methods described in @hagenGen3sis2021.
-From the projected data, we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 MYA to the present day.
-Based on the ____ value, we use XX as the threshold for arid. Thus aridity index is ...
+
+To estimate paleoclimate in Australia from 40 Ma to the present-day, we used three two alternative approaches.
+The first approach used information derived from paleo-Köppen climate bands from @scoteseAtlas2021.
+Köppen bands are estimated from different fossil and lithological indicators and give a broad categorisation of spatial variation in major climate types (tropical, sub-tropical arid, warm temperate, cool temperate, and polar).
+We obtained estimated Köppen climate bands at 5 Ma intervals from 40 Ma to 0 Ma from and used the sub-tropical arid Köppen band as a binary aridity index.
+We removed sites found in the ocean by using digital elevation models (DEMs) to mask sites below sea-level.
+To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alterative DEMs [@straumeGlobSed2019;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity.
+To estimate paleotemperature from the Köppen bands, we followed the method of @scoteseAtlas2021 as implemented by @hagenEarth2021.
+Here, each Köppen climate band was assigned a temperature value based on present-day averages of those climate bands (tropical = 26, sub-tropical arid = 22, warm temperate = 16, cool temperate = 5, polar = -20).
+To account for global fluctuations in temperature, we then modified these values by a global temperature curve estimate from oxygen isotope data (66).
+Finally, to account for change of temperature with elevation, we used a lapse rate of 9.8°C/km.
+Following @hagenEarth2021, we model variation in the intensity of temperature oscillations during the Pleistocene.
+We also used the two alternative DEMs to give two alternative temperature curves using this approach [@straumeGlobSed2019;@scotesePALEOMAP2018].
+
+The second approach used a global circulation model (GCM) to obtain a model-based estimate of paleotemperature and paleoprecipitation @valdesDeep2021.
+The paleoclimatic data is based on a HadCM3L model which couples ocean, atmosphere and vegetation and simulated at 5 Ma intervals based on the paleogeography of @scoteseAtlas2021.
+To make it comparable with the binary aridity index derived from the Köppen band approach, we converted the paleoprecipitation into an aridity index by thresholding the precipitation data.
+We applied a threshold of 500mm/year, below which is considered arid, above which was considered non-arid.
+Each of the three global paleoaridity and paleotemperature maps, from the two approaches, were projected into a Behrmann equal-area projection using a spatial resolution of 110km x 110km.
+We then temporally interpolated the data from 5 Ma intervals to 100 ka using bilinear interpolation.
+Finally, we subset the data to include only the continent of Sahul from 40 Ma to the present-day.
+From these maps, we obtain multiple Sahulian curves for paleotemperature by taking the mean, minimum and maximum values within Sahul at each timestep.
+For paleoaridity, we took the mean value, as minimum and maximum showed little variation in the binary metric.
 
@@
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
-We also fit the model to widely use mean global paleotemperature data set [@condamineMacroevolutionary2013].
+We fitted a total of 12 fit_env models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each DEM.
+We also fit the phylogeny to the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironment.
 To account for missing taxa, we used the fractioning function (f = 38/50) in RPANDA.
@@
 Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates.
-Models 3 and 4 were based on models 1 and 2, but they also allowed unmeasured traits ('hidden states') to contribute to diversification rates.
+Models 3 and 4 were based on models 1 and 2, but they also allowed  unmeasured trait ('hidden state') to contribute to diversification rates.
 By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification.
@@
 Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
-On the other hand, the largest sink for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the mediterranean biome (7.0 ± 2.9 events; 18.8%).
-The greatest directional transition is from the arid biome to the mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (Fig.
+On the other hand, the largest sink for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%).
+The greatest directional transition is from the arid biome to the Mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (Fig.
 \@ref(fig:fig-biogeobears)c-d).
 
-Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10).
+Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
 We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n = 21 lineages, not including _A.
@@
 The common ancestor of *Group 1* remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
+The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
 
@@
 
-Paleotemperature estimates from gen3sis for Australia differed between the three DEMs (Fig.
+Paleotemperature estimates for Australia differed between the three different methods (Fig.
 \@ref(fig:fig-paleoclimate-compare)b).
@@
 Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles.
-Reconstructions of paleoaridity differed slightly between the two methods (Fig.
+Reconstructions of paleoaridity differed slightly between the two Köppen band methods, but these differed substantially to the aridity index derived from the GCM data (Fig.
 \@ref(fig:fig-paleoclimate-compare)b).
-Both aridity index shows increase in aridity post 10 MYA, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
+Both Köppen band-based aridity indices show increasing aridity from 10 Ma, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
 
@@
 
-We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10).
+Among all the birth-death, time-dependent, and environment-dependent models fitted in RPANDA, we found that environment-dependent models were preferred.
+Of the fitted environment-dependent models, we found the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table 1).
 This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23) (Fig.
 \@ref(fig:fig-RPANDA)).
-The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10).
+The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table 1).
 Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).
@@
 
-In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see Table-geohisse).
+In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see Table 2).
 The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state.
-Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see Table-geohisse).
+Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see Table 2).
 The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see Fig.
@@
 With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
-In this study we inferred the species in Australia occupied the tropical grassland biome.
+In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of invasion from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas.
 Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioned into the arid biome.
 Subsequent biome transitions in these two groups resulted in the distribution of biomes we observed across the phylogeny today.
-We also found that paleoaridity index provides the best fit to the diversification dynamics observed in our time-calibrated phylogeny.
-Although aridification has been linked elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], we found that aridification decreased diversification rates for Australian *Anilios* blind snakes.
+We also found that the paleoaridity index provides the best fit to the diversification dynamics observed in our time-calibrated phylogeny.
+Although aridification has been linked to elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], our results suggest that aridification decreased diversification rates for Australian *Anilios* blind snakes.
 
-## Ancestral biome of Australian *Anilios*
+## Historical biogeography of *Anilios*
 
-Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent between 25.8 – 16 MYA [@marinTracing2013].
-This is also consistent with other radiations of Australian squamates, where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes as they emerge [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
-Given what we know about the current distribution of species included in our phylogeny, our analyses support the idea that *Anilios* have non-Australian origins.
+Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent, which is consistent with previous studies [@marinTracing2013;@vidalBlindsnake2010].
+The observed pattern where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes as they emerge has also been documented in other radiations of Australian squamates [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
+Many Australian groups originated in Eurasia and crossed into Australia after the continental plates collided and caused the uplift of a number of 'stepping stones' including the Sunda islands, New Guinea, and the Soloman islands.
+Our historical biogeographic estimation of *Anilios* is consistent with the hypothesis that this group invaded Australia in the Miocene, first occupying tropical grasslands of Northern Australia before diversifying throughout the continent.
 However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as proposed by @vidalBlindsnake2010 is equivocal.
-Our results indicate the common ancestor of Australian _Anilios_, _Sundatyphlops polygrammicus_, _Ramphotyphlops multilineatus_ and _Acutotyphlops subocularis_ most likely occurred either in the Bismarck Archipelago, New Guinea, or a combination of both.
+Our results indicate the common ancestor of Australian _Anilios_, _Sundatyphlops polygrammicus_, _Ramphotyphlops multilineatus_ and _Acutotyphlops subocularis_ most likely occurred either in the Bismarck Archipelago or New Guinea.
 Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin prior to colonising Australia.
 
-## Biome transition and current distributions
+<!-- While we found that the initial diversification of _Anilios_ occurred in the tropical grassland biome, followed by transitions and/or expansion into other biomes,  -->
+Our analyses indicated that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by @marinTracing2013.
+Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
+This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of widespread aridification in the late Miocene [@byrneBirth2008].
 
-Consistent with previous research by @marinTracing2013, we found that the initial diversification of _Anilios_ occurred in the tropical grassland biome, followed by transitions and/or expansion into other biomes.
-However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013.
-Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
-This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene [@byrneBirth2008].
-
-A hypothesis to explain the contemporary distribution of _Anilios_ and other blind snakes in arid biomes is that they may be pre-adapted to living in such environments.
+We hypothesise that the contemporary distribution of _Anilios_ and other blind snakes in arid biomes is explained by pre-adaptation to living in such environments.
 Blind snakes have been found in xeric areas around the world, such as _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ in semi-arid areas from east Africa to the Middle East [@sindacoReptiles2013].
-To our knowledge, these species share similar dietary and morphological specialisations with their closest relatives, which is a pattern identified across all fossorial snakes [@cyriacDigging2018].
+These arid-adapted blind snake species share specialisations in diets and morphology with Australian _Anilios_, including a diet of pupae and larvae of social insects [@webbDietary1993;@cundallFeeding2000], smooth scales that protect them from biting adult ants/termites, relatively small body size compared to other snakes, and fossorial lifestyle.
+A combination of these traits allow blind snakes to avoid direct heat exposure and risk dessication in harsh arid environments.
 
@@
 Our best-fitting model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
-Two of our reconstructed paleoaridity index show that aridity increased starting ~25 MYA ago with particular sharp increase from 10 MYA and intensification from 5 MYA onwards.
-Out of the 38 lineages we considered, 5/36 divergence event occured in the last 10 million years when the continent intensively aridified.
-In contrast, 31/36 divergence event occurred between 25 – 10 MYA when the continent was getting warming but not entirely arid.
-Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 MYA.
+Two of our reconstructed paleoaridity index data show that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
+A particularly sharp increase from 10 Ma followed from the "Hill gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
+Out of the 38 lineages we considered, 5/36 divergence events occurred in the last 10 million years when the continent intensively aridified.
+In contrast, 31/36 divergence event occurred between 25 – 10 Ma when the continent was getting warming but not entirely arid.
+Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 Ma.
 This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 million years [@byrneBirth2008;@catulloAridification2014].
 
-While not the best fitting model, we observed that our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013].
+Our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013], although it was not the best fitting model.
 In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
 However, models that take in account thermal niche tolerance, such as NEMo [@huaAdapting2022], will benefit from paleoclimate reconstructions such as these to inform limits (boundaries).
-We suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these paleoclimate models to inform model inferences.
-
-ALEX SKEELS add to this discussion?
-- Paleoclimate reconstructions based on DEM by Scotese and Straume were similar in profile.
-However, DEM base on Valdes differed from the other two.
-Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles.
+As it stands, current estimates of climate at deep timescales are still developing and we see strong differences in climate estimates from different approaches.
+For example, while both GCM-based and Köppen band-based approaches showed a gradual increase in temperature from 25 – 5 Ma before a decrease from the Pliocene to the present-day, the GCM-based approach showed a much sharper increase in temperature.
+Similarly, aridity estimates from the GCM-based approach were starkly different to the Köppen band approaches.
+This may be because deep time GCM models are currently aimed at estimating general global fluctuations in climate, rather than regionally specific ones, which may be poorly approximated.
+Instead, the Köppen band approach is data driven and is based on known indicators of aridity from around the world [@boucotPhanerozoic2013].
+This could explain why the Köppen band approach for aridity provided the strongest explanation for the diversification rate data.
+Because climate estimates can vary widely, we suggest future studies that aim to understand how aridification affect diversification rates in Australia consider using these (and other) paleoclimate models to inform model inferences.
 
@@
 Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
-This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snake [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
+This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
 
-For these terrestrial reptiles, increased rates of speciation for arid adapted taxa have been attributed to species filling niche in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
+For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
 However, results from our weighted-averaged SSE analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
-One hypothesis is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
+A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
 For instance, because blind snakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
@@
 ## Limitations
-Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA) may be biased due to the small sample size of our _Anilios_ phylogeny [@caetanoHidden2018;@malietModel2019;@sunEstimating2020].
-However, this near-complete phylogeny of _Anilios_ is is the most comprehensive data set available at this time.
+Although we did not detect conflicting signal between our analyses, some of our results, particularly those that estimate diversification rates (GeoHiSSE, ClaDS, and RPANDA), may be biased due to the small sample size of our _Anilios_ phylogeny [@caetanoHidden2018;@malietModel2019;@sunEstimating2020].
+However, this near-complete phylogeny of _Anilios_ is the most comprehensive data set available at this time.
 We suspect there remains several undescribed species of blind snakes in Australia and New Guinea that may belong in the genus _Anilios_.
@@
 Our BioGeoBEARS ancestral biome estimation is limited by the areas (e.g.
-in this case biomes) we can assign our extant species to.
+in this case biomes) to which we can assign our extant species.
 By following the WWF Bioregions maps, we had to assign species that occur in extremely different habitats in the same biome.
 For example, *A.
-longissimus*, a species that has only been collected from 30m underground on Barrow Island located 50 km the northwest off the Pilbara coast of Western Australia, shares the same biome with *A.
+longissimus*, a species that has only been collected from 30 m underground on Barrow Island located 50 km the northwest off the Pilbara coast of Western Australia, shares the same biome with *A.
 centralis* from the arid interior of the continent.
-Current methods does not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat featureless desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011].
+Current methods do not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat seemingly featureless desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011].
 However, modern software underdevelopment such as PhyBEARS may be able to account for more complex geological processes such as the movement of continent.
-Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today.
+Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today [@oliverPhylogenetic2017].
 Out of all our comparative analyses, only the SSE model explicit models extinction rates.
@@
 
-As we continue generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods and complimentary paleoenvironmental data sets that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
+As we continue to generate more genetic data to build ever more extensive phylogenies, there is a growing need to develop methods, and complimentary paleoenvironmental data sets, that allow us to test hypotheses about how different macroevolutionary processes influence diversification.
 We inferred that tropical grassland is the ancestral biome of Australian _Anilios_ and acted as a major source of biodiversity to other biomes.
-Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridify.
-We showed that tools like gen3sis can be used to reconstruct localised paleoaridity data that provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature.
-To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleo-aridity data.
+Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified.
+We showed that localised paleoaridity data provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
+To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleoaridity data.
 We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.

### 2023-01-21 manuscript.Rmd
update needs more work on the aridification/diversification paragraph

@@
 The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
-After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annums) with the development of pockets of aridity [@byrneDecline2011].
+After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annums) with the development of pockets of aridity [@martinCenozoic2006;@byrneDecline2011].
 Subsequent extreme warming and cooling periods in the last ~5 Ma creating what is now known as the "outback" [@byrneBirth2008;@pepperLife2021].
@@
 Here, each Köppen climate band was assigned a temperature value based on present-day averages of those climate bands (tropical = 26, sub-tropical arid = 22, warm temperate = 16, cool temperate = 5, polar = -20).
-To account for global fluctuations in temperature, we then modified these values by a global temperature curve estimate from oxygen isotope data (66).
+To account for global fluctuations in temperature, we then modified these values by a global temperature curve estimate from oxygen isotope data [@songSeawater2019].
 Finally, to account for change of temperature with elevation, we used a lapse rate of 9.8°C/km.
@@
 A particularly sharp increase from 10 Ma followed from the "Hill gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
-Out of the 38 lineages we considered, 5/36 divergence events occurred in the last 10 million years when the continent intensively aridified.
-In contrast, 31/36 divergence event occurred between 25 – 10 Ma when the continent was getting warming but not entirely arid.
+Out of the 38 lineages we considered, 5/36 divergence events occurred in the last 10 Ma when the continent intensively aridified.
+In contrast, 31/36 divergence event occurred between 25 – 10 Ma when the continent was getting warmer but not entirely arid.
 Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 Ma.
-This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 million years [@byrneBirth2008;@catulloAridification2014].
+This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 Ma [@byrneBirth2008;@catulloAridification2014].
+
+The observed overall decrease in diversification rate with increase aridification may be associated with mass extinction.
+We hypothesise that mesic-adapted ancestral blind snakes may have once been widespread but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009], which possibly made even subterranean conditions uninhabitable for blind snakes and their social insect prey.
+The surviving population then became isolated in small patches of "mesic refugia" as the arid zone expanded, as has been proposed in Australian cicadas [@owenHow2017].
+Contemporary distribution of extant lineages supports this hypothesis.
+Occurrence data of the 13/38 arid lineages are usually found in pockets of climatically stable oasis rather than dry desolate sand dessert.
+
+However, extinction events are difficult to identify when fossil data is lacking [@loucaWhy2021;@raboskyExtinction2010].
+
+The surviving lineages in the arid zone are often associated with
+
+ due to their small size and home range,
+
+ as  when Australia
+
+Although many lineages occur in what is now the arid zone,  surviving lineages occur in the arid zone 
+For instance,
 
@@
 ## Diversification in the arid zone
-
 Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008].

### 2023-01-23 code/01_env_data_plots.R
change colours for arid and temperature plots

@@ -33,11 +33,11 @@ env_plottr <- function(.df){
   .df %>% 
   ggplot(data = ., aes(x = age, colour = source)) +
     geom_line(aes(y = vals), size = 1) +
-    geom_point(aes(y = vals)) + 
+    # geom_point(aes(y = vals)) + 
     deeptime::coord_geo(xlim = c(26, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
     # scale_colour_manual(values = viridis(4), name = "Data source", labels = c("RPANDA Global", "Scotese Australia", "Straume Australia", "Valdes Australia")) +
     scale_y_continuous(limits = c(-5, 30), breaks = seq(-5, 30, 5)) + labs(y = "Mean temperature (°C)") +
-    scale_x_reverse("Age (Ma)", limits = c(25, 0), breaks = seq(25, 0, by = -5)) + theme_bw() 
+    scale_x_reverse("Time (Ma)", limits = c(25, 0), breaks = seq(25, 0, by = -5)) + theme_bw() 
 }
 # env_plottr <- function(.df){
 #   .df %>% 
@@ -60,14 +60,33 @@ env_df[which(env_df$source %in% c("mean_val", "max_val", "min_val")),] %>% env_p
 
 dev.off()  
 
-temp_plots <- env_df[which(env_df$source %in% c("mean_sco", "mean_str", "mean_val", "global_temp")),] %>% env_plottr()
+# temp_plots <- env_df[which(env_df$source %in% c("mean_sco", "mean_str", "mean_val", "global_temp")),] %>% env_plottr()
 
-arid_plots <- env_df[which(env_df$source %in% c("arid_sco", "arid_str", "arid_val")),] %>% 
+
+temp_plots <- env_df[which(env_df$source %in% c("mean_sco", "mean_str", "mean_val", "global_temp")),] %>% 
   ggplot(data = ., aes(x = age, colour = source)) +
   geom_line(aes(y = vals), size = 1) +
-  scale_x_reverse("Age (Ma)", limits = c(25, 0), breaks = seq(25, 0, by = -5)) + theme_bw() +
+  # geom_point(aes(y = vals)) + 
   deeptime::coord_geo(xlim = c(26, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
-  labs(y = "Aridity index")
[…]

### 2023-01-23 code/04_BioGeoBEARS_results_plots.R
change dimensions

@@ -77,11 +77,11 @@ tr=tree.c
 dev.off()
 # States
 
-# pdf('output/bgb_letter_ancestral.pdf', width = 11.5, height = 8.5)
+pdf('output/bgb_letter_ancestral.pdf', width = 8.5, height = 11.33)
 resstates = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j","x"), plotwhat="text", label.offset=0.45,
                                      tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, 
                                      include_null_range=TRUE, tr=tr,
-                                     colors_list_for_states=colors_list_for_states,
+                                     # colors_list_for_states=colors_list_for_states,
                                      tipranges=tipranges)
 tiplabels(tr$tip.label, adj = c(0,0.5), bg = NULL, col = NULL, frame = "none")
 

### 2023-01-23 manuscript.Rmd
major update incorporating diversification-aridification discussion. Ref to figures and supplementary materials.

@@
 <!-- paragraph 4 -->
-Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (**Tiatragul et al.
-under review**).
+Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] [@tiatragulAustralia2022].
 All currently described species of _Anilios_ are fossorial and are found in Australia except for _A.
@@
 <!-- Based on our phylogenetic tree -->
-We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al.
-under review).
-This tree includes 38 operational taxonomic units (OTUs) representing 36/47 described species of _Anilios_.
-The remaining 11 species could not be included because no molecular data or tissues samples are available.
+We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups [@tiatragulAustralia2022].
+This tree includes 38 operational taxonomic units (OTUs) representing 35/47 described species of _Anilios_.
+The remaining 12 species could not be included because no molecular data or tissues samples are available.
 However, our phylogenetic data suggest the existence of at least three genetically distinct undescribed species, so we adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (see Tiatragul et al.
-under review).
+The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) [@tiatragulAustralia2022].
 For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
@@
 We also fitted these three models, with two additional free parameters that we thought were relevant, both singly and together, for a total of 12 alternative models.
-The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix-method-biogeobears).
+The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix 1).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 
-## Paleo environment reconstruction
+## Paleoenvironment reconstruction
 <!-- Consult Alex Skeels -->
@@
 
+<!-- Anilios_bituberculatus - distributed in the arid zone but also many occurrence in coastal areas in southern Australia. Classified as non-arid.
+Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
+Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
+Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
+Anilios_hamatus - species description says widespread through 'southern arid and semi arid regions'. Classified as Arid.   -->
+
 # Results
@@
 <!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
-Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios* but the ancestor leading to the crown *Anilios* is ambiguous.
-We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, see Table-supp-BioGeoBEARS).
+Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios*, but the ancestor leading to the crown *Anilios* is ambiguous (Fig.
+\@ref(fig:fig-biogeobears) and Fig.
+S1).
+We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, see Table S1).
 A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table-supp-BioGeoBEARS).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table S2).
 Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
@@
 proximus_) and *Group 2* (n = 14 lineages, not including _A.
-silvia_) (Fig.
+silvia_ and _A.
+leucoproctus_) (Fig.
 \@ref(fig:fig-biogeobears)a).
@@
 The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table 1).
-Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).
+Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (Table S3).
 
@@
 
-```{r Table-RPANDA-TOP10}
+<!-- {r Table-RPANDA-TOP10} -->
 
-# Location of table of top 10 RPANDA
+Table 1.
+Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
+The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
 
-```
+| TOP10       | Description |
+| ----------- | ----------- |
+| Header      | Title       |
+| Paragraph   | Text        |
 
@@
 
-In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (see Table 2).
+In our SSE analyses, we found that the range-independent model fit the data better than models where diversification rates were dependent on geographic range (Table 2).
 The best-fitting geographic range-independent model included a hidden unobserved state, indicating that diversification rates are better explained by this hidden state.
-Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (see Table 2).
+Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (Table 2).
 The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see Fig.
-\@ref(fig:fig-clads-sse) and supp-fig-clads-sse for net diversification rate and extinction rates).
+\@ref(fig:fig-clads-sse) and Fig.
+S2 for net diversification rate and extinction rates).
 
@@
 
-```{r Table-geohisse}
-# Location of table of geohisse results
-```
+<!-- Table-geohisse -->
+
+Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+
+| GeoHiSSE      | Description |
+| ----------- | ----------- |
+| Header      | Title       |
+| Paragraph   | Text        |
+
+<!--  -->
 
@@
 
-## Influence of paleoenvironment on diversification rates
-
-Our best-fitting model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
-Two of our reconstructed paleoaridity index data show that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
-A particularly sharp increase from 10 Ma followed from the "Hill gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
-Out of the 38 lineages we considered, 5/36 divergence events occurred in the last 10 Ma when the continent intensively aridified.
-In contrast, 31/36 divergence event occurred between 25 – 10 Ma when the continent was getting warmer but not entirely arid.
-Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 Ma.
-This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 Ma [@byrneBirth2008;@catulloAridification2014].
-
-The observed overall decrease in diversification rate with increase aridification may be associated with mass extinction.
-We hypothesise that mesic-adapted ancestral blind snakes may have once been widespread but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009], which possibly made even subterranean conditions uninhabitable for blind snakes and their social insect prey.
-The surviving population then became isolated in small patches of "mesic refugia" as the arid zone expanded, as has been proposed in Australian cicadas [@owenHow2017].
-Contemporary distribution of extant lineages supports this hypothesis.
-Occurrence data of the 13/38 arid lineages are usually found in pockets of climatically stable oasis rather than dry desolate sand dessert.
-
-However, extinction events are difficult to identify when fossil data is lacking [@loucaWhy2021;@raboskyExtinction2010].
-
-The surviving lineages in the arid zone are often associated with
-
- due to their small size and home range,
-
- as  when Australia
-
-Although many lineages occur in what is now the arid zone,  surviving lineages occur in the arid zone 
-For instance,
-
+## Local vs global paleoenvironment 
 Our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013], although it was not the best fitting model.
@@
 
-## Diversification in the arid zone
+## Aridification and diversification rates
+Our best-fitting RPANDA model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
+Two of our reconstructed paleoaridity index data show that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
+A particularly sharp increase from 10 Ma followed from the "Hill Gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
+Out of the 38 lineages we considered, 5/35 divergence events occurred in the last 10 Ma when the continent intensively aridified.
+In contrast, 31/35 divergence event occurred between 25 – 10 Ma when the continent was getting warmer but not entirely arid.
+Indeed 12/16 lineages that have current distributions in the arid biome diverged before 10 Ma.
+This pattern is consistent with other studies on vertebrates that have shown that the divergence time for some arid-adapted taxa pre-dated the intense aridification in the last 5 Ma [@byrneBirth2008;@catulloAridification2014].
+
+The observed decrease in diversification rate with increasing aridification may be linked to mass extinction.
+One hypothesis is that mesic-adapted ancestral blind snakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
+An extensive population genetics study of blind snake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
+Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events.
+However, the lack of *Anilios* fossils makes extinction events difficult to identify [@loucaWhy2021;@raboskyExtinction2010].
+An alternative explanation for decrease diversification is limited sampling of Australian blind snake, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
+The current phylogeny is missing 12 described species [@tiatragulAustralia2022], such as the arid-adapted _A.
+fossor_ and _A.
+minimus_, and the true diversity has been estimated to exceed 56 species [@marinHidden2013;@marinTracing2013;@ellisTyphlopid2019].
+It is possible that these lineages that are not represented in our phylogeny have diversified more recently when aridity is high.
+
+## Diversification rates in arid vs non-arid biomes
 Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008].
@@
 (d) Speciation rates from four geographic diversification rate models averaged by model weight.
-(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages
+(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages.
 
@@
 
-Table-RPANDA-TOP10.
+Table 1.
 Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
@@
 
-Table geo-hisse. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
 

### 2023-01-23 manuscript.Rmd
slight changes

@@
 <!--The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.  -->
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database as described in @tiatragulAustralia2022.
 Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
@@
 Paleotemperature estimates for Australia differed between the three different methods (Fig.
-\@ref(fig:fig-paleoclimate-compare)b).
+\@ref(fig:fig-paleoclimate-compare)a).
 However, all three paleotemperature models indicate Australia was hotter than the average global temperature (Fig.
@@
 Among all the birth-death, time-dependent, and environment-dependent models fitted in RPANDA, we found that environment-dependent models were preferred.
-Of the fitted environment-dependent models, we found the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table 1).
+Of the fitted environment-dependent models, we found the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity estimated from @scotesePALEOMAP2018 provided the best fit to the phylogeny (AICc = 238.43; see Table 1).
 This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23) (Fig.
@@
 Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (Table S3).
-
-```{r fig-RPANDA, fig.cap= 'Effect of aridity change on the diversification of Anilios estimated from the phylogeny.
-(a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
-(b) Speciation rates through time from the best diversification model.
-(c) Association between speciation rate and aridity index.
-We found that increased aridity is associated with decrease speciation rate.'}
-
-knitr::include_graphics(paste0(WD, '/assets/figures/RPANDA_fit_arid.jpg'))
-```
 
@@
 | Paragraph   | Text        |
+
+```{r fig-RPANDA, fig.cap= 'Effect of aridity change on the diversification of Anilios estimated from the phylogeny.
+(a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
+(b) Speciation rates through time from the best diversification model.
+(c) Association between speciation rate and aridity index.
+We found that increased aridity is associated with decrease speciation rate.'}
+
+knitr::include_graphics(paste0(WD, '/assets/figures/RPANDA_fit_arid.jpg'))
+```
 
@@
 
+<!-- Table-geohisse -->
+
+Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+
+| GeoHiSSE      | Description |
+| ----------- | ----------- |
+| Header      | Title       |
+| Paragraph   | Text        |
+
 ```{r fig-clads-sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models.
@@
 ```
-
-<!-- Table-geohisse -->
-
-Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
-
-| GeoHiSSE      | Description |
-| ----------- | ----------- |
-| Header      | Title       |
-| Paragraph   | Text        |
 

### 2023-01-24 code/01_env_data_plots.R
change label text

@@ -68,7 +68,7 @@ temp_plots <- env_df[which(env_df$source %in% c("mean_sco", "mean_str", "mean_va
   geom_line(aes(y = vals), size = 1) +
   # geom_point(aes(y = vals)) + 
   deeptime::coord_geo(xlim = c(26, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
-  scale_colour_manual(values = c("#F0027F", "#386CB0", "#FDC086", "#666666"), name = "Data source", labels = c("RPANDA Global", "Scotese and Wright 2018", "Straume et al. 2019", "Valdes et al. 2021")) +
+  scale_colour_manual(values = c("#F0027F", "#386CB0", "#FDC086", "#666666"), name = "Data source", labels = c("RPANDA Global", "Scotese and Wright 2018", "Straume et al., 2020", "Valdes et al., 2021")) +
   # scale_color_brewer(type='qual',palette='Set1', name = "Data source", labels = c("RPANDA Global", "Scotese and Wright 2018", "Straume et al. 2019", "Valdes et al. 2021")) +
   scale_y_continuous(limits = c(-5, 30), breaks = seq(-5, 30, 5)) + labs(y = "Mean temperature (°C)") +
   scale_x_reverse("Time (Ma)", limits = c(25, 0), breaks = seq(25, 0, by = -5)) + theme_bw() +
@@ -84,7 +84,7 @@ arid_plots <-env_df[which(env_df$source %in% c("arid_sco", "arid_str", "arid_val
   labs(y = "Aridity index") +
   # scale_color_brewer(type='qual',palette='Set1') +
   # scale_color_brewer(type='qual',palette='Set1', name = "Data source", labels = c("Scotese and Wright 2018", "Straume et al. 2019", "Valdes et al. 2021")) +
-  scale_colour_manual(values = c("#386CB0", "#FDC086", "#666666"), name = "Data source", labels = c("Scotese and Wright 2018", "Straume et al. 2019", "Valdes et al. 2021")) +
+  scale_colour_manual(values = c("#386CB0", "#FDC086", "#666666"), name = "Data source", labels = c("Scotese and Wright 2018", "Straume et al., 2020", "Valdes et al., 2021")) +
   theme(panel.border = element_blank(), axis.line.x = element_line(color = "black", size = 0.5), axis.line.y = element_line(color = "black", size = 0.5)) +
   theme(legend.position = "right")
 

### 2023-01-24 manuscript.Rmd
incorporated second rounds of suggestions from AS and SK

@@
 ---
-title: "Dessicated diversification: how aridification slowed down speciation of Australian blind snakes"
+title: "Reconstructing the dry: the influence of paleoaridity on the speciation of Australian blind snakes"
 date: "`r Sys.Date()`"
@@
 
-### Possible titles:
-
-- Diversification rates of Australian blind snakes slows down as the continent became more dry up.
-
-- Reconstructed paleoaridity correlate with decreased diversification rates in Australian blind snakes.
-
-- Drying up speciation: aridification of Australia slowed down diversification of Australian blind snakes
-
-- Dull when dry: increased aridity correlate with decrease diversification in blind snakes
-
 **Address**: Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
@@
 
-**Aim**: Shifts in diversification rates of Australian flora and fauna has been associated with the process of aridification, but the relationship between diversification rates and aridity has never been quantified.
-We investigate how paleoenvironmental changes have influenced the diversification rates of the mostly endemic Australian blind snakes genus, Anilios.
-We explored how changes in temperature and aridity during the Neogene have influenced the diversification of the Australian blind snake and we further investigate whether diversification rates differ between arid-adapted lineages vs mesic-adapted lineages.
+**Aim**: Shifts in diversification rates of Australian flora and fauna have been associated with aridification, but the relationship between diversification rates and aridity has never been quantified.
+We investigate how paleoenvironmental changes influenced the diversification rates of the Australian blind snakes genus, *Anilios*.
+We explored how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blind snakes and we further investigate whether diversification rates differ between arid-adapted lineages vs mesic-adapted lineages.
 
@@
 **Results**: Ancestral area estimation indicated Australian blind snakes have tropical grassland origins.
-We found that local Australian paleoenvironments provided a better explanation for diversification rate variation than global paleotemperature.
-Speciically, diversification rates of blind snakes decreased with increasing aridity.
+We found that Australia-specific regional paleoenvironments provided a better explanation for diversification rate variation than global paleotemperature.
+Specifically, diversification rates of blind snakes decreased with increasing aridity.
 We found no difference in diversification rates between arid- and mesic-adapted lineages.
 
-**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blind snakes diversified rapidly in mesic habitats during early Miocene.
-When the continent became more arid, diversification rates decreased.
-However, two clades emerged with one staying mostly in mesic habitats and the other exploiting the burgeoning arid biome.
-Both the arid and tropical biomes were important sources and sinks of subsequent diversification.
+**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blind snakes diversified rapidly in mesic habitats during the early Miocene.
+However, as the continent became increasingly arid, diversification rates decreased.
+To our knowledge, this is the first study that uses reconstructed paleoaridity data to empirically fit diversification models.
+We found that shifts in the environment led to the emergence of two major clades: one remaining in primarily mesic habitats and the other adapting to the expanding arid biome.
+Our results emphasise the importance of both arid and tropical biomes as sources and sinks of diversification for this group of snakes.
 
 **Keywords**
-Scolecophidia, Diversification, paleoenvironment, BioGeoBEARS, GeoHiSSE
+Scolecophidia, diversification, paleoenvironment, BioGeoBEARS, GeoHiSSE
 
@@
 
-<!-- General background  -->
+The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo of their diversification.
+In particular, one major aim of evolutionary biology is to understand the role that environmental factors have played in shaping rates of species diversification.
+A variety of environmental correlates have been associated with diversification processes across clades, including elevation [@esquerreHow2019;@boschmanMountain2022], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017], temperature [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014].
+However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regards to the regional scale at which many clades diversified.
+As such, it remains a key challenge to combine regional paleoenvironmental histories, with molecular phylogenies, to understand the drivers of diversification over deep-time scales.
 
-The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo and mode with which different groups diversified.
-After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annums) with the development of pockets of aridity [@martinCenozoic2006;@byrneDecline2011].
-Subsequent extreme warming and cooling periods in the last ~5 Ma creating what is now known as the "outback" [@byrneBirth2008;@pepperLife2021].
-During this period, arid-adapted vegetation expanded and mesic biomes became more fragmented [@crispHow2013;@martinCenozoic2006].
-This period of aridification is believed to have promoted the rapid speciation of many arid-adapted vertebrates [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] and the extinction of some mesic-adapted groups [@bryantLines2016;@cardilloPhylogeny2017;
-@owenHow2017].
-
-<!-- Environmental correlates of diversification -->
-A variety of ecological correlates have been associated with diversification processes of a clade.
-Among others, studies on the influence of climate (e.g.
-temperature, precipitation, and latitude) [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], elevation [@esquerreHow2019;@boschmanMountain2022], biological interactions [@ramirezAsynchronous2011;@druryEstimating2016;@brennanPhylogenomics2021], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017] and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014] have been documented.
-In Australia, shifts in diversification rates owing to changes in temperature and aridity has been reported in skinks [@raboskyExceptional2007], geckos [@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014], frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
-
-<!-- paragraph 5 -->
-In reptiles, variation in climate has been proposed to influence diversification rates.
-For instance, warmer temperature may increase resource availability, which in turn increase ecological niche in general [@clarkeClimate2006].
+Terrestrial ectotherms, such as squamate reptiles, are strongly influenced by the external environment and variation in climate may influence diversification rates through multiple mechanisms.
+For instance, warmer temperature may increase resource availability, which in turn increase ecological niche space [@clarkeClimate2006].
 Spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
-Change in climate (e.g.
-glaciation and aridification) may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
+Change in climate, such as glaciation or aridification, may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
 Ectotherms are subjected to higher metabolic rates in warm climates (i.e.
 Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
-Finally, aridification of Australia may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised Australia from elsewhere.
- <!---which have been documented in several arid-adapted lineages with deep divergences [@byrneBirth2008;@pepperGenetic2011;@marinTracing2013]. --->
+Finally, aridification may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised from elsewhere.
 
-<!-- paragraph 3 -->
+The Australian continent has one of the most diverse continental reptile faunas on the planet and this may have been driven by its rapid environmental change in the Neogene period.
+After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annum) with the development of pockets of aridity [@martinCenozoic2006;@byrneDecline2011].
+Subsequent extreme warming and cooling periods in the last ~5 Ma creating what is now known as the "outback" [@byrneBirth2008;@pepperLife2021].
+During this period, as aridification progressed, arid-adapted vegetation expanded while mesic biomes became increasingly fragmented [@crispHow2013;@martinCenozoic2006].
+This shift in the environment also led to a radiation of some reptile clades, such as elapid snakes and skinks, which diversified into morphologically and ecologically distinct lineages [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007].
+Conversely, other reptile clades, such as blind snakes and geckos, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
+Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
+Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification @raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+Whether environmental change in Australia has had generally positive or negative influence on diversification rates in squamate reptiles is still unknown.
+
 Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
-Squamate reptiles (lizards and snakes) are the largest terrestrial vertebrate group in Australia with over 1000 species (~800 lizards and ~200 snakes), many of which colonised the continent from elsewhere and diversified in the last 20 – 6 Ma [@brennanMass2017;@oliverPhylogenetic2017;@brennanPhylogenomics2021;@esquerrePhylogenomics2020].
-Some clades, like elapid snakes and skinks, showed dramatic morphological and ecological differentitation [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007].
-Other clades, like some geckos and blind snakes, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
-The species richness and distribution of squamate reptiles across different biomes makes them a good group to test hypotheses about the tempo and mode of diversification in Australia.
-
-<!-- paragraph 4 -->
-Roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] [@tiatragulAustralia2022].
-All currently described species of _Anilios_ are fossorial and are found in Australia except for _A.
+Of these, roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
+All currently described species of _Anilios_ are fossorial and are endemic to Australia except for _A.
 eryncinus_, which is endemic to New Guinea.
-Blind snakes are morphologically conserved, species often difficult to distinguish, and extensive phylogenetic studies and several efforts to revisit specimens have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
+Blind snakes are morphologically conserved, species can be difficult to distinguish, and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
 The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
 Despite representing a substantial portion of snake biodiversity, we know little about the macroevolutionary dynamics of blind snakes.
-The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic ecological changes because they are geographically restricted and can be found in every biome on mainland Australia.
+The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic environmental changes during the Cenozoic because they are geographically restricted and can be found in every biome on mainland Australia.
 
-<!-- paragraph 6 -->
+Advances in phylogenetic comparative methods and paleoenvironmental reconstructions allow us to better understand regional diversification dynamics.
 The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019].
 Traditionally, these temperature data are reconstructed (interpolated) based on conversion of global concentration of oxygen isotopes [@condamineMacroevolutionary2013].
-The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of local aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
+The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of regional aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
 
-<!-- Paragraph 7 -->
 To our knowledge no study has assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity.
 However, recent developments in spatially-explicit paleoenvironmental reconstructions from the lithological and fossil record as well as global circulation models (GCMs) now enables us to estimate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
-These data sets for the first time enables us to empirically test the effects of localised environmental changes on diversification rates over time.
+These data sets enable us to empirically test the effects of regionalised environmental changes on diversification rates over time for the first time.
 
-<!-- Paragraph 8 -->
-The aim of this study is to infer the historical biogeography of the Australian blind snake radiation and investigate how paleoenvironmental changes have influenced the rates of diversification within this group.
-To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimate diversification rates under these models.
-We further compare diversification rates between arid-adapted lineages versus non arid-adapted lineages.
-
-<!--- In this study, we use the Australian radiation of blindsnakes as a system to investigate the diversification of morphologically conserved squamates by testing hypotheses about the processes that promote speciation.
-Specifically, we sought to answer the following questions:
-(1) What is the ancestral biome of Australian _Anilios_? 
-() Which biomes are the largest source and sinks of biodiversity? 
-() What is the relationship between diversification rate and changes in paleo environment? 
-() Which paleo environment best explain current phylogeny?
-() Are there differences in diversification rates between lineages in arid and non-arid biome? --->
+In this study, we aimed to infer the historical biogeography of the Australian blind snake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on the diversification rates within this group.
+To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models.
+Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages inhabitating the arid biome have undergone faster or slower radiation.
 
@@
 <!-- Based on our phylogenetic tree -->
-We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups [@tiatragulAustralia2022].
+We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al., submitted MS).
 This tree includes 38 operational taxonomic units (OTUs) representing 35/47 described species of _Anilios_.
@@
 However, our phylogenetic data suggest the existence of at least three genetically distinct undescribed species, so we adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) [@tiatragulAustralia2022].
+The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., submitted MS).
 For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
@@
 These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal and extinction) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
-<!--The fundamental inputs for the software include a time-calibrated phylogeny, information about the biome/range in which each species is found, and a historical biogeographic model.  -->
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database as described in @tiatragulAustralia2022.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database as described in Tiatragul et al.
+(submitted).
 Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
@@
 Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in the R package BioGeoBEARS [@matzkeBioGeoBEARS2013].
-We also fitted these three models, with two additional free parameters that we thought were relevant, both singly and together, for a total of 12 alternative models.
+We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models.
 The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix 1).
-We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models, and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
+We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
 Based on the best model, we ran an additional 50 biogeographic stochastic maps to account for uncertainty in ancestral biome estimation and transition between biomes, and to estimate the frequency and direction of dispersal events across biomes [@matzkeBioGeoBEARS2013].
@@
 
-To estimate paleoclimate in Australia from 40 Ma to the present-day, we used three two alternative approaches.
+We used two alternative approaches to estimate paleoclimate in Australia from 40 Ma to the present-day.
 The first approach used information derived from paleo-Köppen climate bands from @scoteseAtlas2021.
@@
 We removed sites found in the ocean by using digital elevation models (DEMs) to mask sites below sea-level.
-To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alterative DEMs [@straumeGlobSed2019;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity.
+To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alterative DEMs [@straumeGlobal2020;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity.
 To estimate paleotemperature from the Köppen bands, we followed the method of @scoteseAtlas2021 as implemented by @hagenEarth2021.
@@
 Following @hagenEarth2021, we model variation in the intensity of temperature oscillations during the Pleistocene.
-We also used the two alternative DEMs to give two alternative temperature curves using this approach [@straumeGlobSed2019;@scotesePALEOMAP2018].
+We also used the two alternative DEMs to give two alternative temperature curves using this approach [@straumeGlobal2020;@scotesePALEOMAP2018].
 
@@
 
-To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environemnt-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package RPANDA [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environment-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package RPANDA [@morlonRPANDA2016] to the _Anilios_ phylogeny.
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
-We fitted a total of 12 fit_env models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each DEM.
-We also fit the phylogeny to the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironment.
+We fitted a total of 12 diversification models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each paleoenvironmental reconstruction method.
+We also fitted a environmental dependent diversification model with the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironmental data.
 To account for missing taxa, we used the fractioning function (f = 38/50) in RPANDA.
@@
 
-To test whether diversification rates differ between lineages in arid biomes and non-arid lineages, we used two approaches.
+To test whether diversification rates differ between lineages in arid biomes and mesic lineages, we used two approaches.
 First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using the Julia package ClaDS [@malietFast2022].
@@
 2.1.9 [@beaulieuDetecting2016].
-Since these SSE models only allow two states, we restricted our geographic states to arid and non-arid biomes.
-Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as non-arid (State 1; n = 24), and those that occupy both the arid and non-arid biomes as widespread (State 01; n = 5).
+Since these SSE models only allow two states, we restricted our geographic states to arid and mesic biomes.
+Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as mesic (State 1; n = 24), and those that occupy both the arid and mesic biomes as widespread (State 01; n = 5).
 
@@
 By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification.
-For each model, we assumed uniform taxon sampling for 89% of arid lineages, 71% of non-arid lineages, and 83% of widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_.
+For each model, we assumed uniform taxon sampling for 89% of arid lineages, 71% of mesic lineages, and 83% of widespread lineages, based on the number of species in the phylogeny and all currently described species of _Anilios_.
 We then calculated AIC to identify the best-fitting models.
@@
 
-<!-- Anilios_bituberculatus - distributed in the arid zone but also many occurrence in coastal areas in southern Australia. Classified as non-arid.
-Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as non-arid
-Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as non-arid.
+<!-- Anilios_bituberculatus - distributed in the arid zone but also many occurrence in coastal areas in southern Australia. Classified as mesic.
+Anilios_diversus - mostly in Top End with some distribution in the interior. Classified as mesic
+Anilios_grypusET - occur in eastern edge of the arid biome. Also occur in tropical savannah. Classified as mesic.
 Anilios_grypusNW - closest to the range of A. grypus in the tropical grassland but also occur in Pilbara. Classified as arid. 
@@
 The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
+Sources of paleoenvironment presented in the table: ^1^ @scotesePALEOMAP2018  and ^2^ @straumeGlobal2020
 
@@
 
-## Diversification rates between arid and non-arid biomes
+## Diversification rates between arid and mesic biomes
 
-We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, non-arid, and widespread ranges.
-Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, non-arid, and widespread ranges (Fig.
+We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges.
+Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, mesic, and widespread ranges (Fig.
 \@ref(fig:fig-clads-sse)).
@@
 Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (Table 2).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and non-arid ranges (see Fig.
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and mesic ranges (see Fig.
 \@ref(fig:fig-clads-sse) and Fig.
@@
 (d) Speciation rates from four geographic diversification rate models averaged by model weight.
-(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages.'}
+(e) Estimates of speciation rate and (d) turnover rates for arid, mesic, and widespread lineages.'}
 
@@
 
-<!--  -->
-
 # Discussion
 
-With 47 recognised species, the mostly Australian-endemic genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
+With 47 recognised species, the mostly Australian-endemic blind snake genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
 In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of invasion from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas.
@@
 Our historical biogeographic estimation of *Anilios* is consistent with the hypothesis that this group invaded Australia in the Miocene, first occupying tropical grasslands of Northern Australia before diversifying throughout the continent.
-However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia as proposed by @vidalBlindsnake2010 is equivocal.
+However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia, as proposed by @vidalBlindsnake2010, is equivocal.
 Our results indicate the common ancestor of Australian _Anilios_, _Sundatyphlops polygrammicus_, _Ramphotyphlops multilineatus_ and _Acutotyphlops subocularis_ most likely occurred either in the Bismarck Archipelago or New Guinea.
-Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin prior to colonising Australia.
+Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin of the group prior to colonising Australia.
 
-<!-- While we found that the initial diversification of _Anilios_ occurred in the tropical grassland biome, followed by transitions and/or expansion into other biomes,  -->
 Our analyses indicated that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by @marinTracing2013.
@@
 
-## Local vs global paleoenvironment 
+## Regional vs global paleoenvironment 
 Our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013], although it was not the best fitting model.
@@
 Our best-fitting RPANDA model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
-Two of our reconstructed paleoaridity index data show that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
+Our reconstructed paleoaridity index data indicate that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
 A particularly sharp increase from 10 Ma followed from the "Hill Gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
@@
 
-The observed decrease in diversification rate with increasing aridification may be linked to mass extinction.
+The observed decrease in diversification rate with increasing aridification may be linked to increasing speciation rates or decreasing extinction rates.
 One hypothesis is that mesic-adapted ancestral blind snakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
@@
 Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events.
-However, the lack of *Anilios* fossils makes extinction events difficult to identify [@loucaWhy2021;@raboskyExtinction2010].
-An alternative explanation for decrease diversification is limited sampling of Australian blind snake, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
-The current phylogeny is missing 12 described species [@tiatragulAustralia2022], such as the arid-adapted _A.
+Our best-fitting models of diversification support constant extinction rates with decreasing rates of speciation. 
+However, the lack of *Anilios* fossils makes extinction rates difficult to estimate [@loucaWhy2021;@raboskyExtinction2010]. An alternative explanation for 
+the negative relationship between aridification and diversification is decreasing speciation rates through time.
+This could be driven by the indirect effect of aridification on population size [@goldieEvolutionary2010].
+More arid sites have lower primary productivity and may support smaller population sizes.
+Under the integrated evolutionary speed hypothesis [@gillmanPatterns2013;@rohdeLatitudinal1992], smaller population sizes are predicted to have slower rates of molecular evolution and subsequently lower rates of speciation.
+Lower rates of molecular evolution have been found in Australian arid zone plants compared to mesic zone plants [@goldieEvolutionary2010].
+Yet, here we found no difference between arid zone and mesic zone speciation rates in blind snakes which is inconsistent with this hypothesis.
+Finally, another potential explanation for diversification patterns is the limited sampling of Australian blind snakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
+The current phylogeny is missing 12 described species (Tiatragul et al., submitted MS), such as the arid-adapted _A.
 fossor_ and _A.
@@
 
-## Diversification rates in arid vs non-arid biomes
+## Diversification rates in arid vs mesic biomes
 Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008].
-Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
+Our ancestral biome analyses show that several mesic ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
 This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
@@
 For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
-However, results from our weighted-averaged SSE analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.
+However, results from our weighted-averaged SSE analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes.
 A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
@@
 Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today [@oliverPhylogenetic2017].
-Out of all our comparative analyses, only the SSE model explicit models extinction rates.
+Out of all our comparative analyses, only the SSE and some RPANDA models estimated extinction rates.
 
@@
 Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified.
-We showed that localised paleoaridity data provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
+We showed that regionalised paleoaridity data provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
 To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleoaridity data.
@@
 (d) Speciation rates from four geographic diversification rate models averaged by model weight.
-(e) Estimates of speciation rate and (d) turnover rates for arid, non-arid, and widespread lineages.
+(e) Estimates of speciation rate and (d) turnover rates for arid, mesic, and widespread lineages.
 

### 2023-01-24 manuscript.Rmd
some spelling thing

@@
 However, as the continent became increasingly arid, diversification rates decreased.
-To our knowledge, this is the first study that uses reconstructed paleoaridity data to empirically fit diversification models.
 We found that shifts in the environment led to the emergence of two major clades: one remaining in primarily mesic habitats and the other adapting to the expanding arid biome.
@@
 A variety of environmental correlates have been associated with diversification processes across clades, including elevation [@esquerreHow2019;@boschmanMountain2022], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017], temperature [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014].
-However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regards to the regional scale at which many clades diversified.
+However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regard to the regional scale at which many clades diversified.
 As such, it remains a key challenge to combine regional paleoenvironmental histories, with molecular phylogenies, to understand the drivers of diversification over deep-time scales.
@@
 To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models.
-Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages inhabitating the arid biome have undergone faster or slower radiation.
+Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages that inhabit the arid biome have undergone faster or slower radiation.
 
@@
 We showed that regionalised paleoaridity data provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
-To our knowledge, this is the first study that empirically fit diversification models to reconstructed paleoaridity data.
+To our knowledge, this is the first study that empirically fit diversification models to Australia-specific reconstructed paleoaridity data.
 We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.

### 2023-01-25 manuscript.Rmd
reformat for JBG

@@
 date: "`r Sys.Date()`"
-author: "Sarin Tiatragul^1^; Alexander Skeels^2,3^; J. Scott Keogh^1^"
+author: ""
 output: 
@@
 
-**Address**: Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
+**Running title** Blind snake speciation slows down with aridification
 
-1 Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
+**Authors**: Sarin Tiatragul^1^; Alexander Skeels^2,3^; J. Scott Keogh^1^
 
-2 Department of Environmental Systems Sciences, Landscape Ecology, Institute of Terrestrial Ecosystems, ETH Zürich, Zurich 8092, Switzerland;
+^1^Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia
 
-3 Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf 8903, Switzerland;
+^2^Department of Environmental Systems Sciences, Landscape Ecology, Institute of Terrestrial Ecosystems, ETH Zürich, Zurich 8092, Switzerland
+
+^3^Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf 8903, Switzerland
 
@@
 
+# Acknowledgements
+ST would like to thank J Rolland and F.
+Condamine for helpful advice about fitting diversification models with RPANDA.
+Thanks to NJ Matzke and O Jiménez-Robles for help with BioGeoBEARS.
+Special thanks to IG Brennan, D Esquerré, M Pepper and CJ Pavón-Vázquez for helpful advice on the analysis and discussion.
+
 \newpage
 
-### J. Biogeography format::
+# Abstracts and keywords
 
 **Aim**: Shifts in diversification rates of Australian flora and fauna have been associated with aridification, but the relationship between diversification rates and aridity has never been quantified.
-We investigate how paleoenvironmental changes influenced the diversification rates of the Australian blind snakes genus, *Anilios*.
-We explored how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blind snakes and we further investigate whether diversification rates differ between arid-adapted lineages vs mesic-adapted lineages.
+We employed multiple approaches to accurately reconstruct paleoaridity of Australia for the first time.
+We  used this information and phylogenetic-based analyses to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blind snakes and we further test whether diversification rates differ between arid-adapted and mesic-adapted lineages.
 
@@
 
-**Location**: Australasia
+**Location**: Australia
 
@@
 **Results**: Ancestral area estimation indicated Australian blind snakes have tropical grassland origins.
-We found that Australia-specific regional paleoenvironments provided a better explanation for diversification rate variation than global paleotemperature.
-Specifically, diversification rates of blind snakes decreased with increasing aridity.
+We found that Australia-specific regional paleotemperature and paleoaridity provided a better explanation for diversification rate variation than global paleotemperature.
+Specifically, our best-fitting model indicate that speciation rates of blind snakes decreased with increasing aridity.
 We found no difference in diversification rates between arid- and mesic-adapted lineages.
@@
 We found that shifts in the environment led to the emergence of two major clades: one remaining in primarily mesic habitats and the other adapting to the expanding arid biome.
-Our results emphasise the importance of both arid and tropical biomes as sources and sinks of diversification for this group of snakes.
+Our results emphasise the importance of both arid and tropical biomes as sources and sinks of diversification.
 
 **Keywords**
-Scolecophidia, diversification, paleoenvironment, BioGeoBEARS, GeoHiSSE
+BioGeoBEARS, diversification, GeoHiSSE, paleoaridity, paleoenvironment reconstruction, Scolecophidia
 
@@
 A variety of environmental correlates have been associated with diversification processes across clades, including elevation [@esquerreHow2019;@boschmanMountain2022], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017], temperature [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014].
-However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regard to the regional scale at which many clades diversified.
+However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regard to the regional scale at which many clades radiate [@eyresClimatic2021].
 As such, it remains a key challenge to combine regional paleoenvironmental histories, with molecular phylogenies, to understand the drivers of diversification over deep-time scales.
@@
 Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
-Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification @raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
 Whether environmental change in Australia has had generally positive or negative influence on diversification rates in squamate reptiles is still unknown.
@@
 
-To investigate the biogeographic history of blind snakes, we used the R package BioGeoBEARS v.1.1.1 [@matzkeBioGeoBEARS2013].
+To investigate the biogeographic history of blind snakes, we used the R package 'BioGeoBEARS' v.1.1.1 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
@@
 
-Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in the R package BioGeoBEARS [@matzkeBioGeoBEARS2013].
+Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
 We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models.
-The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, _x_, multiplies the probability of dispersal by distance to the power of _x_ [@vandamEvaluating2016] (see Appendix 1).
+The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of *x* [@vandamEvaluating2016] (see Appendix 1).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 
-To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environment-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package RPANDA [@morlonRPANDA2016] to the _Anilios_ phylogeny.
+To evaluate how diversification rates correlated with climate and aridity through time, we fitted several time- and environment-dependent likelihood diversification models using the fit_bd and fit_env functions in the R package 'RPANDA' v.2.0  [@morlonRPANDA2016] to the _Anilios_ phylogeny.
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
@@
 To test whether diversification rates differ between lineages in arid biomes and mesic lineages, we used two approaches.
-First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using the Julia package ClaDS [@malietFast2022].
+First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using ClaDS function in the Julia package 'PANDA' v.1.6.1 [@malietFast2022].
 ClaDS is designed to capture frequent shifts in speciation rates with small effects.
@@
 
-In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package hisse v.
-2.1.9 [@beaulieuDetecting2016].
+In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package 'hisse' v.2.1.9 [@beaulieuDetecting2016].
 Since these SSE models only allow two states, we restricted our geographic states to arid and mesic biomes.
@@
 ClaDS was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2 [@rcoreteamLanguage2022].
-Figures and graphs were prepared using a combination of base R functions and ggplot2 [@wickhamGgplot22016].
+Figures were prepared using 'ggplot2' v.3.4.0 [@wickhamGgplot22016].
 All map graphics were created using the Free and Open Source QGIS v.3.6.
@@
 (c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
-Abbreviations: TrG = Tropical grassland, TmF = Temperate forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate grassland.'}
+Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.'}
 
@@
 
-Table 1.
+**Table 1**.
 Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
@@
 Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (Table 2).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and mesic ranges (see Fig.
-\@ref(fig:fig-clads-sse) and Fig.
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and mesic ranges (Fig.
+\@ref(fig:fig-clads-sse); Fig.
 S2 for net diversification rate and extinction rates).
@@
 
-Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+**Table 2**. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
 
@@
 However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia, as proposed by @vidalBlindsnake2010, is equivocal.
-Our results indicate the common ancestor of Australian _Anilios_, _Sundatyphlops polygrammicus_, _Ramphotyphlops multilineatus_ and _Acutotyphlops subocularis_ most likely occurred either in the Bismarck Archipelago or New Guinea.
+Our results indicate the common ancestor of Australian _Anilios_, _S.
+polygrammicus_, _R.
+multilineatus_ and _A.
+subocularis_ most likely occurred either in the Bismarck Archipelago or New Guinea.
 Increasing taxonomic sampling, particularly by including the New Guinean endemic _Anilios erycinus_ and more species from *Ramphotyphlops* and *Acutotyphlops* will enable us to make more informed inferences about the origin of the group prior to colonising Australia.
 
-Our analyses indicated that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by @marinTracing2013.
+Our analyses indicated that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
 Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
@@
 To our knowledge, this is the first study that empirically fit diversification models to Australia-specific reconstructed paleoaridity data.
-We believe these regional paleoclimate models will be of interested for studies where species are geographically restricted.
+We believe these regional paleoclimate models will be of interest for studies where species are geographically restricted.
 
 \newpage
+
+# Data availability statement
+
+Upon acceptance, data and analyses scripts will be available from the Dryad Repository (XXXXXXX)
 
@@
 (c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
-Abbreviations: TrG = Tropical grassland, TmF = Temperate forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate grassland.
+Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.
 
@@
 
-Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+Table 2. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework. AIC scores and AIC weights indicate which models best fit the data.
 
@@
 
-**References**
+# Biosketch
+
+Sarin Tiatragul is broadly interested in learning about how organisms came to be and their interaction with the environment.
+This work is part of his PhD work at the Australian National University on the macroevolution of Australian blind snakes.
+
+Alexander Skeels
+
+J. Scott Keogh is a professor of evolutionary biology with wide ranging interests in molecular phylogenetics and molecular ecology.
+
+# Author contributions: 
+S.T and J.S.K conceived the ideas; S.T and A.S collected the data; S.T analysed the data; S.T led the writing; A.S and J.S.K edited the manuscript.
+
+# References
 <!-- # References -->

### 2023-01-26 code/07_NEMo.R
delete

@@ -1,2 +1,0 @@
-# 07_NEMo.R
-

### 2023-01-26 manuscript.Rmd
change title, update Alex info

@@
 ---
-title: "Reconstructing the dry: the influence of paleoaridity on the speciation of Australian blind snakes"
+title: "Paleoenvironmental models for Australia and the impact of aridification on blind snake diversification"
 date: "`r Sys.Date()`"
@@
 Thanks to NJ Matzke and O Jiménez-Robles for help with BioGeoBEARS.
-Special thanks to IG Brennan, D Esquerré, M Pepper and CJ Pavón-Vázquez for helpful advice on the analysis and discussion.
+Special thanks to IG Brennan, D Esquerré, M Pepper, CJ Pavón-Vázquez, and O Hagen for helpful advice on the analysis and discussion.
 
@@
 We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models.
-The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of *x* [@vandamEvaluating2016] (see Appendix 1).
+The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of *x* [@vandamEvaluating2016] (see Table S1 in Supporting Information).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 ## Paleoenvironment reconstruction
-<!-- Consult Alex Skeels -->
 
@@
 Specifically, we fitted different combinations of pure-birth and birth-death models with speciation and/or extinction rates being constant or otherwise varying linearly or exponentially with paleoenvironment as described above.
-We fitted a total of 12 diversification models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each paleoenvironmental reconstruction method.
+We fitted 12 diversification models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each paleoenvironmental reconstruction method.
 We also fitted a environmental dependent diversification model with the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironmental data.
@@
 
-Alexander Skeels
+Alexander Skeels researches the evolution of biodiversity and how deep-time processes such as plate tectonics and climate change shape diversification patterns in plants and animals.
 

### 2023-01-27 code/01_env_data_plots.R
plot together 3 paleotemp

@@ -94,3 +94,44 @@ pdf(file = "output/paleo_data_compare_plots.pdf", height = 11.33, width = 8.5)
 temp_plots /
   arid_plots
 dev.off()
+
+
+
+p_sco <- env_df[which(env_df$source %in% c("mean_sco", "max_sco", "min_sco")),] %>% 
+  ggplot(data = ., aes(x = age, colour = source)) +
+  geom_line(aes(y = vals), size = 1) +
+  geom_point(aes(y = vals)) +
+  deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
+  scale_colour_manual(values = viridis(3), name = "Scotese & Wright, 2018", labels = c("Max. temp", "Meann. temp", "Min. temp")) +
+  scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Temperature (°C)") +
+  scale_x_reverse("Time (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() +
+  theme(legend.position = "top")
+
+
+p_str <- env_df[which(env_df$source %in% c("mean_str", "max_str", "min_str")),] %>% 
+  ggplot(data = ., aes(x = age, colour = source)) +
+  geom_line(aes(y = vals), size = 1) +
+  geom_point(aes(y = vals)) +
+  deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
+  scale_colour_manual(values = viridis(3), name = "Straume et al., 2020", labels = c("Max. temp", "Meann. temp", "Min. temp")) +
+  scale_y_continuous(limits = c(-5, 40), breaks = seq(-5, 40, 5)) + labs(y = "Temperature (°C)") +
+  scale_x_reverse("Time (Ma)", limits = c(45, 0), breaks = seq(45, 0, by = -5)) + theme_bw() +
+  theme(legend.position = "top")
+
+p_val <- env_df[which(env_df$source %in% c("mean_val", "max_val", "min_val")),] %>% 
+  ggplot(data = ., aes(x = age, colour = source)) +
+  geom_line(aes(y = vals), size = 1) +
+  geom_point(aes(y = vals)) +
+  deeptime::coord_geo(xlim = c(46, 0), dat = c("epochs"), pos = 'bottom', expand = TRUE) +
[…]

### 2023-01-27 code/05_GeoHiSSE_fit.R
added statistical tests

@@ -354,11 +354,57 @@ mod1$AIC; mod2$AIC; mod3$AIC; mod4$AIC
 
 mod1$AICc; mod2$AICc; mod3$AICc; mod4$AICc
 
-
-
-
 recon.mod1$rates.mat
 recon.mod2$rates.mat
 recon.mod3$rates.mat
 recon.mod4$rates.mat
 
+
+
+# Summary_statistics ------------------------------------------------------
+
+±
+
+geohisse_summary <- rates_df %>% 
+  dplyr::group_by(range) %>% 
+  dplyr::summarize(mean_speciation = mean(speciation), 
+                                     sd_speciation = sd(speciation),
+                                     mean_net.div = mean(net.div), 
+                                     sd_net.div = sd(net.div),
+                                     mean_turnover = mean(turnover),
+                                     sd_turnover = sd(turnover),
+                                     mean_extinction = mean(extinction),
+                                     sd_extinction = sd(extinction)
+                   )
+
+geohisse_summary_transposed <- t(geohisse_summary)
+
+write.table(geohisse_summary_transposed, file = "output/geohisse_summary.txt", sep = "\t", row.names = TRUE, quote = FALSE)
+
+
+
+rates_df$speciation ~ rates_df$range
+
+fit_1 <- lm(rates_df$speciation ~ rates_df$range)
+fit_2 <- lm(rates_df$net.div ~ rates_df$range)
+fit_3 <- lm(rates_df$turnover ~ rates_df$range)
+fit_4 <- lm(rates_df$extinction ~ rates_df$range)
+
+
+library(stats)
+
+
+
+anova(fit_1)
+TukeyHSD(aov(rates_df$speciation ~ rates_df$range))
+anova(fit_2)
+TukeyHSD(aov(rates_df$net.div ~ rates_df$range))
+anova(fit_3)
+TukeyHSD(aov(rates_df$turnover ~ rates_df$range))
+anova(fit_4)
+TukeyHSD(aov(rates_df$extinction ~ rates_df$range))
+
+summary(fit_1)
+summary(fit_2)
+summary(fit_3)
+summary(fit_4)

### 2023-01-27 manuscript.Rmd
finalise manuscript ready for submission to J Biogeography

@@
 ---
-title: "Paleoenvironmental models for Australia and the impact of aridification on blind snake diversification"
+title: "Paleoenvironmental models for Australia and the impact of aridification on blindsnake diversification"
 date: "`r Sys.Date()`"
@@
 
-**Running title** Blind snake speciation slows down with aridification
+**Running title** blindsnake speciation slows down with aridification
 
@@
 We employed multiple approaches to accurately reconstruct paleoaridity of Australia for the first time.
-We  used this information and phylogenetic-based analyses to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blind snakes and we further test whether diversification rates differ between arid-adapted and mesic-adapted lineages.
+We  used this information and phylogenetic-based analyses to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blindsnakes and we further test whether diversification rates differ between arid-adapted and mesic-adapted lineages.
 
-**Taxon**: Blind snakes
+**Taxon**: blindsnakes
 
@@
 
-**Materials and Methods**: We estimated the historical biogeography of blind snakes using BioGeoBEARS.
+**Materials and Methods**: We estimated the historical biogeography of blindsnakes using BioGeoBEARS.
 We synthesised multiple approaches to reconstruct paleotemperature and paleoaridity of Australia during the Neogene.
@@
 
-**Results**: Ancestral area estimation indicated Australian blind snakes have tropical grassland origins.
+**Results**: Ancestral area estimation indicated Australian blindsnakes have tropical grassland origins.
 We found that Australia-specific regional paleotemperature and paleoaridity provided a better explanation for diversification rate variation than global paleotemperature.
-Specifically, our best-fitting model indicate that speciation rates of blind snakes decreased with increasing aridity.
+Specifically, our best-fitting model indicate that speciation rates of blindsnakes decreased with increasing aridity.
 We found no difference in diversification rates between arid- and mesic-adapted lineages.
 
-**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blind snakes diversified rapidly in mesic habitats during the early Miocene.
+**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blindsnakes diversified rapidly in mesic habitats during the early Miocene.
 However, as the continent became increasingly arid, diversification rates decreased.
@@
 
-Terrestrial ectotherms, such as squamate reptiles, are strongly influenced by the external environment and variation in climate may influence diversification rates through multiple mechanisms.
+Terrestrial ectotherms, such as squamate reptiles, are highly sensitive to changes in external environment and climate.
+This can influence their diversification rates through multiple mechanisms.
 For instance, warmer temperature may increase resource availability, which in turn increase ecological niche space [@clarkeClimate2006].
-Spatial and temporal heterogeneity in temperature may lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
-Change in climate, such as glaciation or aridification, may cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
+Spatial and temporal heterogeneity in temperature can lead to climatic niche divergence, such as local adaptations in different geographic areas (Pepper et al., 2013), and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
+Change in climate, such as glaciation or aridification, has been shown to cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
 Ectotherms are subjected to higher metabolic rates in warm climates (i.e.
-Kleiber's law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
+Kleiber’s law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
 Finally, aridification may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised from elsewhere.
 
-The Australian continent has one of the most diverse continental reptile faunas on the planet and this may have been driven by its rapid environmental change in the Neogene period.
+The Australian continent has one of the most diverse reptile faunas on the planet and this may have been driven by its rapid environmental change in the Neogene period.
 After a long period of relatively stable climate and mesic habitats, the aridification of continental Australia began around 20 Ma (mega annum) with the development of pockets of aridity [@martinCenozoic2006;@byrneDecline2011].
@@
 This shift in the environment also led to a radiation of some reptile clades, such as elapid snakes and skinks, which diversified into morphologically and ecologically distinct lineages [@leeDiversification2016;@sandersMolecular2008;@raboskyExceptional2007].
-Conversely, other reptile clades, such as blind snakes and geckos, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
+Conversely, other reptile clades, such as blindsnakes and geckos, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
 Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
@@
 Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
-Of these, roughly a quarter of the snake species diversity in Australia is represented by blind snakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
+Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
 All currently described species of _Anilios_ are fossorial and are endemic to Australia except for _A.
 eryncinus_, which is endemic to New Guinea.
-Blind snakes are morphologically conserved, species can be difficult to distinguish, and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
+blindsnakes are morphologically conserved, species can be difficult to distinguish, and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
 The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
-Despite representing a substantial portion of snake biodiversity, we know little about the macroevolutionary dynamics of blind snakes.
+Despite representing a substantial portion of snake biodiversity, we know little about the macroevolutionary dynamics of blindsnakes.
 The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic environmental changes during the Cenozoic because they are geographically restricted and can be found in every biome on mainland Australia.
@@
 However, recent developments in spatially-explicit paleoenvironmental reconstructions from the lithological and fossil record as well as global circulation models (GCMs) now enables us to estimate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
-These data sets enable us to empirically test the effects of regionalised environmental changes on diversification rates over time for the first time.
+The use of these data sets allows us to conduct empirically analyses of the impacts of regionalised environmental changes on diversification rates over time for the first time.
 
-In this study, we aimed to infer the historical biogeography of the Australian blind snake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on the diversification rates within this group.
+In this study, we aimed to infer the historical biogeography of the Australian blindsnake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on the diversification rates within this group.
 To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models.
@@
 <!-- Based on our phylogenetic tree -->
-We employed the most comprehensive time-calibrated phylogenomic tree of Australian blind snakes with outgroups (Tiatragul et al., submitted MS).
+We employed the most comprehensive time-calibrated phylogeny of Australian blindsnakes with outgroups (Tiatragul et al., submitted MS).
 This tree includes 38 operational taxonomic units (OTUs) representing 35/47 described species of _Anilios_.
@@
 
-To investigate the biogeographic history of blind snakes, we used the R package 'BioGeoBEARS' v.1.1.1 [@matzkeBioGeoBEARS2013].
+To investigate the biogeographic history of blindsnakes, we used the R package 'BioGeoBEARS' v.1.1.2 [@matzkeBioGeoBEARS2013].
 BioGeoBEARS provides a flexible, likelihood-based framework that models how branches in the phylogeny (representing extant taxa and their ancestors) have shifted among a set of discrete biomes over time.
@@
 
-Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis(DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
+Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis (DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
 We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models.
-The first additional parameter, *j*, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, *x*, multiplies the probability of dispersal by distance to the power of *x* [@vandamEvaluating2016] (see Table S1 in Supporting Information).
+The first additional parameter, j, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, x, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (see Table S1 in Supporting Information).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 
-The second approach used a global circulation model (GCM) to obtain a model-based estimate of paleotemperature and paleoprecipitation @valdesDeep2021.
+The second approach used a global circulation model (GCM) to obtain a model-based estimate of paleotemperature and paleoprecipitation [@valdesDeep2021].
 The paleoclimatic data is based on a HadCM3L model which couples ocean, atmosphere and vegetation and simulated at 5 Ma intervals based on the paleogeography of @scoteseAtlas2021.
@@
 We applied a threshold of 500mm/year, below which is considered arid, above which was considered non-arid.
-Each of the three global paleoaridity and paleotemperature maps, from the two approaches, were projected into a Behrmann equal-area projection using a spatial resolution of 110km x 110km.
+Each of the three global paleoaridity and paleotemperature maps, from the two approaches, were projected into a Behrmann equal-area projection using a spatial resolution of 110 km x 110 km.
 We then temporally interpolated the data from 5 Ma intervals to 100 ka using bilinear interpolation.
@@
 We fitted 12 diversification models based on reconstructed paleoenvironmental data to the phylogeny including three paleotemperature (mean, minimum, and maximum) and the paleoaridity index from each paleoenvironmental reconstruction method.
-We also fitted a environmental dependent diversification model with the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironmental data.
+We also fitted an additional environmental dependent diversification model with the widely use mean global paleotemperature data set [@condamineMacroevolutionary2013] to provide comparison to our reconstructed paleoenvironmental data.
 To account for missing taxa, we used the fractioning function (f = 38/50) in RPANDA.
@@
 
-To test whether diversification rates differ between lineages in arid biomes and mesic lineages, we used two approaches.
-First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) and tip speciation rates using ClaDS function in the Julia package 'PANDA' v.1.6.1 [@malietFast2022].
+To test whether diversification rates differ between arid-adapted and mesic-adapted lineages, we used two approaches.
+First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) using ClaDS function in the Julia package 'PANDA' v.1.6.1 [@malietFast2022].
 ClaDS is designed to capture frequent shifts in speciation rates with small effects.
@@
 
-In the second approach, we fit four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package 'hisse' v.2.1.9 [@beaulieuDetecting2016].
-Since these SSE models only allow two states, we restricted our geographic states to arid and mesic biomes.
-Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as mesic (State 1; n = 24), and those that occupy both the arid and mesic biomes as widespread (State 01; n = 5).
+In the second approach, we fitted four state-dependent speciation and extinction (SSE) models to the phylogeny and geographic range data under the GeoSSE [@goldbergPhylogenetic2011] and GeoHiSSE [@caetanoHidden2018] framework using the R package 'hisse' v.2.1.9 [@beaulieuDetecting2016].
+Since these SSE models only allow two states, we restricted our geographic states to arid and mesic.
+Based on the same categories we used in our BioGeoBEARS analyses, we classified species that primarily occur in the arid biome as arid (State 0; n = 8), species that occur primarily in other biomes as mesic (State 1; n = 25), and those that occupy both the arid and mesic biomes as widespread (State 01; n = 5).
 
-The four SSE models differed in whether they allowed diversification rates to be dependent on geographic states (range-dependent) or independent of them (range-independent), and whether they allowed an unmeasured trait ('hidden states') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
+The four SSE models differed in whether they allowed diversification rates to be dependent on geographic states (range-dependent) or independent of them (range-independent), and whether they allowed an unmeasured trait ('hidden state') to influence diversification rates [@beaulieuDetecting2016;@caetanoHidden2018].
 Model 1 assumed equal diversification rates regardless of geographic state. Model 2 allowed geographic range to affect diversification rates.
-Models 3 and 4 were based on models 1 and 2, but they also allowed  unmeasured trait ('hidden state') to contribute to diversification rates.
+Models 3 and 4 were based on models 1 and 2, but they also allowed  a hidden state to contribute to diversification rates.
 By incorporating hidden states, we could assess whether geographic range or an unmeasured trait contributed more to the observed diversification.
@@
 ## Data visualisation and software
-ClaDS was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2 [@rcoreteamLanguage2022].
+ClaDS was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2.1 [@rcoreteamLanguage2022].
 Figures were prepared using 'ggplot2' v.3.4.0 [@wickhamGgplot22016].
@@
 Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios*, but the ancestor leading to the crown *Anilios* is ambiguous (Fig.
-\@ref(fig:fig-biogeobears) and Fig.
+\@ref(fig:fig-biogeobears)a and Fig.
 S1).
-We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, see Table S1).
+We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, Table S2).
 A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table S2).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table S3).
 Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
@@
 The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events), which makes up  43.6% compared outward dispersals from the arid biome (9.7 ± 3.2 events) of 26% (Fig.
-\@ref(fig:fig-biogeobears)c-d).
+\@ref(fig:fig-biogeobears)b).
 Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
-On the other hand, the largest sink for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%).
+On the other hand, the largest sinks for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%).
 The greatest directional transition is from the arid biome to the Mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (Fig.
-\@ref(fig:fig-biogeobears)c-d).
+\@ref(fig:fig-biogeobears)c).
 
 Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
-We identified divergent biogeographic transitions between two major groups of species, which we will refer to as *Group 1* (n = 21 lineages, not including _A.
-proximus_) and *Group 2* (n = 14 lineages, not including _A.
+We identified divergent biogeographic transitions between two major groups of species, which we will refer to as Group 1 (n = 21 lineages, not including _A.
+proximus_) and Group 2 (n = 14 lineages, not including _A.
 silvia_ and _A.
@@
 \@ref(fig:fig-biogeobears)a).
-The common ancestor of *Group 1* remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
+The common ancestor of Group 1 remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
+The common ancestor of Group 2 first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
 
-```{r fig-biogeobears, fig.cap= 'Ancestral biome estimation and transitions of Australo-Papuan blind snakes in BioGeoBEARS, using the DEC+j+x model.
+```{r fig-biogeobears, fig.cap= 'Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model.
 (a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
 Inset map (left) shows the current distribution of biomes.
-Inset maps to the right of the tree shows distribution of species in Group 1 and Group 2, respectively.
-(b) Histogram showing frequency and composition of outward transitions from Australian biomes and combined non-Australian islands.
-(c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
+Inset maps to the right of the phylogeny shows distribution of species in Group 1 and Group 2.
+(b) Bar plot showing frequency and composition of outward transitions and (c) inward transitions from Australian biomes and combined non-Australian islands.
 Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.'}
@@
 However, all three paleotemperature models indicate Australia was hotter than the average global temperature (Fig.
-\@ref(fig:fig-paleoclimate-compare)a).
-Paleoclimate in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles.
+\@ref(fig:fig-paleoclimate-compare)a; Fig.
+S2).
+Paleotemperature in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles
 Reconstructions of paleoaridity differed slightly between the two Köppen band methods, but these differed substantially to the aridity index derived from the GCM data (Fig.
@@
 
-```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based different data sets in gen3sis and RPANDA.
-(a) Estimated mean paleotemperature in Australia compared to RPANDA global paleotemperature data.
-(b) Estimated mean aridity index based on precipitation threshold.
-Abbreviations Pli = Paleocene and Ple = Pleistocene.'}
+```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based on three different methods and RPANDA.
+(a) Estimated mean temperature over time in Australia compared to RPANDA global paleotemperature data.
+(b) Estimated mean aridity index over time based on precipitation threshold.
+Abbreviations Pli = Pliocene and Ple = Pleistocene.'}
 
@@
 The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table 1).
-Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (Table S3).
+Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (Table S4).
 
@@
 **Table 1**.
-Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
+Results of the 10 most supported RPANDA diversification models on the Australian blindsnake radiation.
 The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
@@
 
-We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges.
-Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between lineages in arid, mesic, and widespread ranges (Fig.
+We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges (Fig.
 \@ref(fig:fig-clads-sse)).
-ClaDS also indicate branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present.
+Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between the geographic ranges (Fig.
+\@ref(fig:fig-clads-sse)c).
+ClaDS also indicate branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present (Fig.
+\@ref(fig:fig-clads-sse)b).
 
@@
 Compared to the three other models, the best-fitting model accounted for 95% of the AIC weight and had an AIC value 7.02 units lower than that of the second best-fitting model (Table 2).
-The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ significantly between arid and mesic ranges (Fig.
+The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ between arid and mesic ranges (Fig.
 \@ref(fig:fig-clads-sse); Fig.
-S2 for net diversification rate and extinction rates).
+S3; Table S5).
 
@@
 
-**Table 2**. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework.
+**Table 2**. Diversification rate models investigated in the GeoSSE and GeoHiSSE framework. AIC scores and AIC weights indicate which models best fit the data.
 
@@
 
-With 47 recognised species, the mostly Australian-endemic blind snake genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
+With 47 recognised species, the mostly Australian-endemic blindsnake genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
 In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of invasion from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas.
@@
 We also found that the paleoaridity index provides the best fit to the diversification dynamics observed in our time-calibrated phylogeny.
-Although aridification has been linked to elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], our results suggest that aridification decreased diversification rates for Australian *Anilios* blind snakes.
+Although aridification has been linked to elevated diversification rates in other plants and animals [@brennanMass2017;@raboskyAnalysis2014;@hernandez-hernandezAridification2014;@jhaPhylogeny2021], our results suggest that aridification decreased diversification rates for Australian *Anilios* blindsnakes.
 
@@
 
-We hypothesise that the contemporary distribution of _Anilios_ and other blind snakes in arid biomes is explained by pre-adaptation to living in such environments.
-Blind snakes have been found in xeric areas around the world, such as _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ in semi-arid areas from east Africa to the Middle East [@sindacoReptiles2013].
-These arid-adapted blind snake species share specialisations in diets and morphology with Australian _Anilios_, including a diet of pupae and larvae of social insects [@webbDietary1993;@cundallFeeding2000], smooth scales that protect them from biting adult ants/termites, relatively small body size compared to other snakes, and fossorial lifestyle.
-A combination of these traits allow blind snakes to avoid direct heat exposure and risk dessication in harsh arid environments.
+We hypothesise that the contemporary distribution of _Anilios_ and other blindsnakes in arid biomes is explained by pre-adaptation to living in such environments.
+blindsnakes have been found in xeric areas around the world, such as _Rena humilis_ in the desert of North America and _Myriopholis macrorhyncha_ in semi-arid areas from east Africa to the Middle East [@sindacoReptiles2013].
+These arid-adapted blindsnake species share specialisations in diets and morphology with Australian _Anilios_, including a diet of pupae and larvae of social insects [@webbDietary1993;@cundallFeeding2000], smooth scales that protect them from biting adult ants/termites, relatively small body size compared to other snakes, and fossorial lifestyle.
+A combination of these traits allows blindsnakes to avoid direct heat exposure and risk desiccation  in harsh arid environments.
 
@@
 The observed decrease in diversification rate with increasing aridification may be linked to increasing speciation rates or decreasing extinction rates.
-One hypothesis is that mesic-adapted ancestral blind snakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
-An extensive population genetics study of blind snake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
+One hypothesis is that mesic-adapted ancestral blindsnakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
+An extensive population genetics study of blindsnake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
 Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events.
@@
 Lower rates of molecular evolution have been found in Australian arid zone plants compared to mesic zone plants [@goldieEvolutionary2010].
-Yet, here we found no difference between arid zone and mesic zone speciation rates in blind snakes which is inconsistent with this hypothesis.
-Finally, another potential explanation for diversification patterns is the limited sampling of Australian blind snakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
+Yet, here we found no difference between arid zone and mesic zone speciation rates in blindsnakes which is inconsistent with this hypothesis.
+Finally, another potential explanation for diversification patterns is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
 The current phylogeny is missing 12 described species (Tiatragul et al., submitted MS), such as the arid-adapted _A.
@@
 Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008].
-Our ancestral biome analyses show that several mesic ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
+Our ancestral biome analyses show that several mesic ancestors of _Anilios_ blindsnakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny.
 This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
@@
 For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
-However, results from our weighted-averaged SSE analyses on the Australian blind snake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes.
+However, results from our weight-averaged SSE analyses on the Australian blindsnake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes.
 A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
-For instance, because blind snakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
-Future studies that investigate phenotypic variations across all blind snakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
+For instance, because blindsnakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
+Future studies that investigate phenotypic variations across all blindsnakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
 
@@
 - Future phylogeographic study should consider densely sampling species to see whether there are more expansion into or out of arid zone for widespread taxa.
-Warmer and dryer environment may not directly influence speciation rate because blind snakes are physically adapting to warmer temperatures.
-However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.
+Warmer and dryer environment may not directly influence speciation rate because blindsnakes are physically adapting to warmer temperatures.
+However, climatic fluctuations can contract and fragment available niche for blindsnakes and thus influence diversification rates.
 - Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].
-- Interestingly blind snakes have diversified greatly compared to other fossorial vertebrates.
+- Interestingly blindsnakes have diversified greatly compared to other fossorial vertebrates.
   - It is also possible that fossorial species are more cryptic and thus the full extent of their diversity is under appreciated.
@@
 However, this near-complete phylogeny of _Anilios_ is the most comprehensive data set available at this time.
-We suspect there remains several undescribed species of blind snakes in Australia and New Guinea that may belong in the genus _Anilios_.
+We suspect there remains several undescribed species of blindsnakes in Australia and New Guinea that may belong in the genus _Anilios_.
 Getting genetic data from _A.
@@
 Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified.
-We showed that regionalised paleoaridity data provides a better fit to our Australian blind snake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
+We showed that regionalised paleoaridity data provides a better fit to our Australian blindsnake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
 To our knowledge, this is the first study that empirically fit diversification models to Australia-specific reconstructed paleoaridity data.
@@
 Fig.
-Ancestral biome estimation and transitions of Australo-Papuan blind snakes in BioGeoBEARS, using the DEC+j+x model.
+Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model.
 (a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
@@
 Table 1.
-Results of the 10 most supported RPANDA diversification models on the Australian blind snake radiation.
+Results of the 10 most supported RPANDA diversification models on the Australian blindsnake radiation.
 The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
@@
 Sarin Tiatragul is broadly interested in learning about how organisms came to be and their interaction with the environment.
-This work is part of his PhD work at the Australian National University on the macroevolution of Australian blind snakes.
+This work is part of his PhD work at the Australian National University on the macroevolution of Australian blindsnakes.
 

### 2023-01-27 code/05_GeoHiSSE_fit.R
clean up tukey hsd

@@ -379,8 +379,7 @@ geohisse_summary <- rates_df %>%
 
 geohisse_summary_transposed <- t(geohisse_summary)
 
-write.table(geohisse_summary_transposed, file = "output/geohisse_summary.txt", sep = "\t", row.names = TRUE, quote = FALSE)
-
+# write.table(geohisse_summary_transposed, file = "output/geohisse_summary.txt", sep = "\t", row.names = TRUE, quote = FALSE)
 
 
 rates_df$speciation ~ rates_df$range
@@ -391,10 +390,7 @@ fit_3 <- lm(rates_df$turnover ~ rates_df$range)
 fit_4 <- lm(rates_df$extinction ~ rates_df$range)
 
 
-library(stats)
-
-
-
+# Anova and TukeyHSD
 anova(fit_1)
 TukeyHSD(aov(rates_df$speciation ~ rates_df$range))
 anova(fit_2)
@@ -403,6 +399,11 @@ anova(fit_3)
 TukeyHSD(aov(rates_df$turnover ~ rates_df$range))
 anova(fit_4)
 TukeyHSD(aov(rates_df$extinction ~ rates_df$range))
+
+# Number of samples
+rates_df %>% 
+  group_by(range) %>% 
+  count()
 
 summary(fit_1)
 summary(fit_2)

### 2023-04-17 manuscript.Rmd
changes made in response to suggestions from reviewers round 1

@@
 
-**Running title** blindsnake speciation slows down with aridification
+**Running title** Decrease in blindsnake speciation correlate with aridification
 
@@
 Special thanks to IG Brennan, D Esquerré, M Pepper, CJ Pavón-Vázquez, and O Hagen for helpful advice on the analysis and discussion.
+JSK thanks the Australian Research Council for ongoing support.
 
@@
 **Aim**: Shifts in diversification rates of Australian flora and fauna have been associated with aridification, but the relationship between diversification rates and aridity has never been quantified.
-We employed multiple approaches to accurately reconstruct paleoaridity of Australia for the first time.
-We  used this information and phylogenetic-based analyses to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blindsnakes and we further test whether diversification rates differ between arid-adapted and mesic-adapted lineages.
+We employed multiple approaches to accurately reconstruct paleoenvironments of Australia for the first time.
+We used this information, and phylogenetic-based analyses, to explore how changes in temperature and increasing aridity during the Neogene influenced the diversification of the Australian blindsnakes.
+We tested whether diversification rates differ between arid-adapted and mesic-adapted lineages.
 
-**Taxon**: blindsnakes
+**Taxon**: Typhlopidae, _Anilios_ blindsnakes
 
@@
 We found that Australia-specific regional paleotemperature and paleoaridity provided a better explanation for diversification rate variation than global paleotemperature.
-Specifically, our best-fitting model indicate that speciation rates of blindsnakes decreased with increasing aridity.
+Specifically, our best-fitting model indicated that speciation rates of blindsnakes decreased with increasing aridity.
 We found no difference in diversification rates between arid- and mesic-adapted lineages.
 
-**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blindsnakes diversified rapidly in mesic habitats during the early Miocene.
+**Main conclusions**: Soon after dispersing to Australia, the common ancestors of Australian blindsnakes diversified rapidly in mesic habitats during the early Miocene.
 However, as the continent became increasingly arid, diversification rates decreased.
@@
 The present-day distribution of extant species can provide clues about their biogeographic history and insight into the tempo of their diversification.
-In particular, one major aim of evolutionary biology is to understand the role that environmental factors have played in shaping rates of species diversification.
+In particular, one major aim of evolutionary biology is to understand the role that environmental factors have played in shaping rates of species diversification [@simpsonTempo1944].
 A variety of environmental correlates have been associated with diversification processes across clades, including elevation [@esquerreHow2019;@boschmanMountain2022], CO<sub>2</sub> concentration [@kergoatOpposite2018], microhabitats [@bars-closelDiversification2017], temperature [@ricklefsGlobal2006;@mayhewBiodiversity2012;@condamineMacroevolutionary2013;@condamineClimate2019;@rollandContribution2019;@sunRecent2020], and aridification [@raboskyExceptional2007;@pepperGenetic2011;@deepakAridification2018;@catulloAridification2014].
@@
 
-Terrestrial ectotherms, such as squamate reptiles, are highly sensitive to changes in external environment and climate.
+<!-- Terrestrial ectotherms, such as squamate reptiles, are highly sensitive to changes in external environment and climate.
 This can influence their diversification rates through multiple mechanisms.
-For instance, warmer temperature may increase resource availability, which in turn increase ecological niche space [@clarkeClimate2006].
-Spatial and temporal heterogeneity in temperature can lead to climatic niche divergence, such as local adaptations in different geographic areas (Pepper et al., 2013), and change in activity time of prey/predators [@santosTemperature2011;@lawsonLatitudinal2014].
-Change in climate, such as glaciation or aridification, has been shown to cause fragmentation of continuous habitats leading to isolated populations, which may lead to divergence [@oliverRocks2019].
-Ectotherms are subjected to higher metabolic rates in warm climates (i.e.
+For instance, warmer temperature may increase resource availability, which in turn increases ecological niche space [@clarkeClimate2006].
+Spatial and temporal heterogeneity in temperature can lead to climatic niche divergence, such as local adaptations in different geographic areas [@pepperGeodiversity2013], and changes in activity times of prey and predators [@santosTemperature2011;@lawsonLatitudinal2014].
+Change in climate, leading to glaciation or aridification, has been shown to cause fragmentation of continuous habitats leading to isolated populations, which may lead to species-level divergence [@oliverRocks2019].
+Ectotherms have higher metabolic rates in warm climates (i.e.
 Kleiber’s law), which may increase mutation rates and ultimately greater speciation rates [@allenKinetic2006;@makarievaMean2008].
 Finally, aridification may offer new ecological niche for species to exploit, particularly mesic-adapted lineages that have colonised from elsewhere.
+Despite many possible pathways, little is known about whether environmental change is associated with variable rates of diversification.
+-->
 
@@
 Conversely, other reptile clades, such as blindsnakes and geckos, remain morphologically conserved and were thought to be species poor until recent molecular data suggested otherwise [@moritzCryptic2018;@zozayaAre2019;@marinHidden2013].
-Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], saltbush [@kadereitMolecular2010], *Acacia* [@millerComparative2013], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013].
-Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013;@kadereitMolecular2010;@millerComparative2013].
+Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have had suppressed diversification or been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
 Whether environmental change in Australia has had generally positive or negative influence on diversification rates in squamate reptiles is still unknown.
 
-Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia's terrestrial reptiles, termites, and ants [@byrneBirth2008].
-Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus _Anilios_, which comprises **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
-All currently described species of _Anilios_ are fossorial and are endemic to Australia except for _A.
-eryncinus_, which is endemic to New Guinea.
-blindsnakes are morphologically conserved, species can be difficult to distinguish, and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
-The diversification of _Anilios_ in Australia is reported to have originated from a single colonisation event by a mesic-adapted ancestor from Asia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
-Despite representing a substantial portion of snake biodiversity, we know little about the macroevolutionary dynamics of blindsnakes.
+Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species  **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
+All but three species of currenty described _Anilios_ are endemic to Australia, _A.
+torresianus_ and _A.
+leucoproctus_ also occur in southern parts of New Guinea, while _A.
+eryncinus_ has only been recorded in New Guinea and the Maluku Islands.
+Blindsnakes are morphologically conserved and previous phylogenetic and morphological studies have revealed that the true species diversity is underestimated [@marinHidden2013;@pyronSystematics2014;@hedgesTaxonomic2014;@ellisTyphlopid2019].
+The diversification of _Anilios_ in Australia is reported to have originated from a single dispersal event by a mesic-adapted ancestor from Southeast Asia or Indonesia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
+However, phylogenetic analyses from Tiatragul et al.
+(2023) indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin
+Despite representing a substantial portion of snake diversity, we know little about the macroevolutionary dynamics of blindsnakes.
 The monophyly of the genus makes _Anilios_ well suited to test hypotheses about how lineages diversified with dramatic environmental changes during the Cenozoic because they are geographically restricted and can be found in every biome on mainland Australia.
@@
 However, recent developments in spatially-explicit paleoenvironmental reconstructions from the lithological and fossil record as well as global circulation models (GCMs) now enables us to estimate paleotemperature and paleoaridity data specific to Australia in the time span of interest.
-The use of these data sets allows us to conduct empirically analyses of the impacts of regionalised environmental changes on diversification rates over time for the first time.
+The use of these data sets allows us to conduct empirical analyses of the impacts of regionalised environmental changes on diversification rates over time for the first time.
 
-In this study, we aimed to infer the historical biogeography of the Australian blindsnake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on the diversification rates within this group.
+In this study, we aimed to infer the historical biogeography of the Australian blindsnake radiation and investigate the impact of paleoenvironmental changes, specifically changes in temperature and aridity associated with Neogene aridification, on diversification rates within this group.
 To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models.
@@
 ## Phylogenetic hypothesis and taxon sampling
-<!-- Based on our phylogenetic tree -->
-We employed the most comprehensive time-calibrated phylogeny of Australian blindsnakes with outgroups (Tiatragul et al., submitted MS).
-This tree includes 38 operational taxonomic units (OTUs) representing 35/47 described species of _Anilios_.
-The remaining 12 species could not be included because no molecular data or tissues samples are available.
-However, our phylogenetic data suggest the existence of at least three genetically distinct undescribed species, so we adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and 2 '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., submitted MS).
-For ancestral biome estimation analyses, we also included _Ramphotyphlops multilineatus_, _Sundatyphlops polygrammicus_, and _Acutotyphlops subocularis_ as outgroup taxa.
+
+We used the most updated phylogeny for *Anilios* and closely related outgroups from Tiatragul et al., submitted MS.
+The species tree was built with the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] using maximum likelihood gene trees inferred in IQTREE from 4930 loci (322 Anchored Hybrid Enrichment, 4574 ultraconserved elements, and 34 legacy genes) for 35/47 recognised *Anilios* species.
+Tiatragul et al.
+(submitted) time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007] (see Tiatragul et al., submitted MS for details).
+Phylogenetic data in Tiatragul et al.
+(submitted) suggest the existence of at least three genetically distinct undescribed species, so we followed their designation and adjusted our taxon sampling fraction to 38/50 species.
+The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., submitted MS).
+For ancestral biome estimation analyses, we included three outgroup blindsnakes species from Southeast Asia, Indonesia, and Melanesia, where the ancestor of *Anilios* has been hypothesised to have originated [@vidalBlindsnake2010].
+The three species include *Ramphotyphlops multilineatus* (Western New Guinea, Kai Island, and Salawati), *Sundatyphlops polygrammicus* (Moyo Island, Indonesia), and *Acutotyphlops subocularis* (Bismarck Archipelago).
+We were not able to include a *Ramphotyphlops* from continental Southeast Asia in the time-calibrated phylogeny because insufficient data was available for divergence time estimation.
 
@@
 These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal and extinction) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database as described in Tiatragul et al.
-(submitted).
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database.
 Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
 
-Using the phylogeny and distribution data, we fitted three widely used historical biogeographical methods, which vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis (DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
+Using the phylogeny and distribution data, we fitted models that vary in the kinds of cladogenetic and anagenetic events they model, including likelihood implementations of Dispersal-Vicariance Analysis (DIVALIKE) [@ronquistDispersalVicariance1997], Dispersal-Extinction-Cladogenesis (DEC) [@reeMaximum2008], and BayArea (BAYAREALIKE) [@landisBayesian2013] in BioGeoBEARS [@matzkeBioGeoBEARS2013].
 We also fitted these three models, with two  additional relevant free parameters, both singly and together, for a total of 12 alternative models.
-The first additional parameter, j, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, x, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (see Table S1 in Supporting Information).
+The first additional parameter, j, assigns per-event weight to founder-events [@matzkeModel2014], and the second parameter, x, multiplies the probability of dispersal by distance to the power of x [@vandamEvaluating2016] (Table S1).
 We used likelihood ratio tests to determine if the additional free parameter was a justified improvement over the simpler base models and selected the model that best fitted the data based on corrected Akaike Information Criterion (AICc) scores.
@@
 Köppen bands are estimated from different fossil and lithological indicators and give a broad categorisation of spatial variation in major climate types (tropical, sub-tropical arid, warm temperate, cool temperate, and polar).
-We obtained estimated Köppen climate bands at 5 Ma intervals from 40 Ma to 0 Ma from and used the sub-tropical arid Köppen band as a binary aridity index.
+We obtained estimated Köppen climate bands at 5 Ma intervals from 40 Ma to 0 Ma and used the sub-tropical arid Köppen band as a binary aridity index.
 We removed sites found in the ocean by using digital elevation models (DEMs) to mask sites below sea-level.
-To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alterative DEMs [@straumeGlobal2020;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity.
+To account for uncertainty in paleogeographic reconstructions of continental movements and sea-levels, we used two alternative DEMs [@straumeGlobal2020;@scotesePALEOMAP2018] to mask the ocean sites and provide two estimates of paleoaridity.
 To estimate paleotemperature from the Köppen bands, we followed the method of @scoteseAtlas2021 as implemented by @hagenEarth2021.
@@
 To make it comparable with the binary aridity index derived from the Köppen band approach, we converted the paleoprecipitation into an aridity index by thresholding the precipitation data.
-We applied a threshold of 500mm/year, below which is considered arid, above which was considered non-arid.
+We applied a threshold of 500mm/year, below which is considered arid, above which was considered non-arid [@holzapfelDeserts2008;@waltonArid1969].
 Each of the three global paleoaridity and paleotemperature maps, from the two approaches, were projected into a Behrmann equal-area projection using a spatial resolution of 110 km x 110 km.
@@
 
-## Data visualisation and software
-ClaDS was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2.1 [@rcoreteamLanguage2022].
+PANDA was implemented in Julia v.1.8.1 [@bezansonJulia2012], all other analyses were performed in R v.4.2.1 [@rcoreteamLanguage2022].
 Figures were prepared using 'ggplot2' v.3.4.0 [@wickhamGgplot22016].
@@
 <!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
-Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios*, but the ancestor leading to the crown *Anilios* is ambiguous (Fig.
+Our BioGeoBEARS analyses all inferred tropical grassland as the ancestral biome of *Anilios*, but the ancestral biome for the common ancestor between *Ramphotyphlops multilineatus* and *Anilios* is ambiguous (Fig.
 \@ref(fig:fig-biogeobears)a and Fig.
 S1).
-We found that the DEC+j+x model was the most preferred biogeographical history model (Δ AICc = -4.47, Table S2).
-A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide a statistically significant improvement over the simpler DEC model, DEC+x, and DEC+j (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
+We found that the DEC+j+x model was the best-fitting biogeographical history model (Δ AICc = -4.47, Table S2).
+A likelihood ratio test (LRT) showed that both founder events (+j parameter) and distance between biomes (+x parameter) together provide statistically significant improvements over the simpler DEC, DEC+x, and DEC+j models (χ^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).
 
-Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occurred throughout the phylogeny (see Table S3).
-Of these 40 ± 0 were cladogenetic events compared to 31.2 ± 1.3 anagenetic dispersal events.
-Out of the 40 cladogenetic dispersal events, most were narrow sympatry range expansion events (23.8 ± 2.1 events) followed by subset sympatry (5.2 ± 2.3 events), vicariance (4.9 ± 1.3), and founder events (6.1 ± 1.5).
+Based on 50 BioGeoBEARS biogeographic stochastic mappings (BSM) of the DEC+j+x model, we estimated that an average of 71.2 ± 1.3 (mean ± SD) total biome shifts occurred throughout the phylogeny (see Table S3).
+Of these 40 ± 0 were cladogenetic (branching) events compared to 31.2 ± 1.3 anagenetic (within a single branch) dispersal events.
+Out of the 40 branching events, in 23.8 ± 2.1 events sister species maintained their ancestral biome (“narrow sympatry”) and in 5.2 ± 2.3 events one sister species maintain their ancestral biome while the other sister occupies only a subset of the ancestral range (“subset sympatry”).
+The remaining branching events were made up by 6.1 ± 1.5 founder events where one of the sister species occupies a new range, and 4.9 ± 1.3 vicariance events where the sister species split their ancestral range.
 Our model also indicate that range expansion is expected to decrease as distance increase (free parameter x = -1.3).
 
-The largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events), which makes up  43.6% compared outward dispersals from the arid biome (9.7 ± 3.2 events) of 26% (Fig.
-\@ref(fig:fig-biogeobears)b).
-Together the tropical grassland biome and the arid biome contributes the largest source of diversification accounting for 69.6% of outward transitions.
+Based on current distributions from ALA, most Anilios species are found entirely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
+Apart from A.
+silvia, A.
+leucoproctus, and A.
+proximus that appeared to form independent divergent lineages (Fig.
+\@ref(fig:fig-biogeobears)a), we identified divergent biogeographic transitions between two major groups of species, which we refer to as Group 1 (n = 21 lineages) and Group 2 (n = 14 lineages).
+The common ancestor of Group 1 remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
+The common ancestor of Group 2 first transitioned into the arid biome, four lineages remained in the arid biome while ten lineages transitioned or expanded into other biomes.
+
+The tropical grassland biome is the largest source of outward dispersal (16.2 ± 4.9 events; 46.6%) followed by the arid biome (9.7 ± 3.2 events; 26%)(Fig.
+1b).
 On the other hand, the largest sinks for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%).
-The greatest directional transition is from the arid biome to the Mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland into the temperate forest biome (5.1 ± 1.3 events) (Fig.
+The greatest directional transition is from the arid biome to the Mediterranean biome (5.3 ± 1.5 events) and from the tropical grassland biome into the temperate forest biome (5.1 ± 1.3 events) (Fig.
 \@ref(fig:fig-biogeobears)c).
-
-Based on current distributions from ALA, most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
-We identified divergent biogeographic transitions between two major groups of species, which we will refer to as Group 1 (n = 21 lineages, not including _A.
-proximus_) and Group 2 (n = 14 lineages, not including _A.
-silvia_ and _A.
-leucoproctus_) (Fig.
-\@ref(fig:fig-biogeobears)a).
-The common ancestor of Group 1 remained mostly in the tropical grassland biome, then 12/21 lineages expanded or transitioned into other biomes.
-The common ancestor of Group 2 first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and temperate forest biomes, and/or transitioned back into the arid biome.
 
@@
 S2).
-Paleotemperature in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles
+Paleotemperature in Australia were also relatively constant compared to global temperature until the Plio-Pleistocene, a period characterised by fluctuations corresponding to the glacial cycles.
 Reconstructions of paleoaridity differed slightly between the two Köppen band methods, but these differed substantially to the aridity index derived from the GCM data (Fig.
 \@ref(fig:fig-paleoclimate-compare)b).
-Both Köppen band-based aridity indices show increasing aridity from 10 Ma, consistent with the onset of aridification of the continent [@byrneBirth2008;@pepperLife2021].
+Both Köppen band-based aridity indices show increasing aridity from 10 Ma.
 
@@
 Results of the 10 most supported RPANDA diversification models on the Australian blindsnake radiation.
-The table includes the model, number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
+The table includes the models (BEnvVar = Speciation rate varying with environment; DCST = Constant extinction rate; DEnvVar = Extinction rate varying with environment; (EXPO) = rates varying exponentially with environment; (LIN) = rates varying linearly with environment) number of parameters, the estimated log-likelihood (logL), the corrected Akaike information criterion (AICc), and corresponding parameter estimates (λ = speciation rate at present, α = parameter controlling the dependency of speciation rate on time or paleoclimate, µ = extinction rate at present, ß = parameter controlling the dependency of extinction rate on time or paleoclimate).
 Sources of paleoenvironment presented in the table: ^1^ @scotesePALEOMAP2018  and ^2^ @straumeGlobal2020
@@
 (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
-(b) Speciation rates through time from the best diversification model.
-(c) Association between speciation rate and aridity index.
-We found that increased aridity is associated with decrease speciation rate.'}
+(b) Fitted diversification rates through time reconstructed from parameter estimates from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”).
+(c) Fitted diversification rates at different aridity index values calculated from parameter estimates of the previous model.
+Increased aridity is associated with decrease net diversification rates in this model..'}
 
@@
 
-We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges (Fig.
-\@ref(fig:fig-clads-sse)).
-Results from our ClaDS analysis show mean branch-specific rates did not differ significantly between the geographic ranges (Fig.
+We applied ClaDS and SSE models to evaluate whether diversification rates differ between lineages in arid, mesic, and widespread ranges (Fig. \@ref(fig:fig-clads-sse)). 
+Results from our ClaDS analysis show branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present (Fig.
+\@ref(fig:fig-clads-sse)b).
+ClaDS also indicate mean branch-specific rates did not differ significantly between the geographic ranges (Fig.
 \@ref(fig:fig-clads-sse)c).
-ClaDS also indicate branch-specific diversification rates was high when the genus first diversified and gradually decreased towards the present (Fig.
-\@ref(fig:fig-clads-sse)b).
 
@@
 The weighted average parameter estimates from all models indicate that both speciation rates and turnover rates did not differ between arid and mesic ranges (Fig.
-\@ref(fig:fig-clads-sse); Fig.
-S3; Table S5).
+S3 and S4; Table S5).
 
@@
 ```{r fig-clads-sse, fig.cap= 'Diversification rates estimates from ClaDS and SSE models.
-(a) Branch-specific rates across the phylogeny.
+(a) Branch-specific rates (log(λi)) across the phylogeny.
 (b) Mean branch-specific rates over time.
-(c) Density plot of branch-specific rates by geographic range from ClaDS analysis.
-(d) Speciation rates from four geographic diversification rate models averaged by model weight.
-(e) Estimates of speciation rate and (d) turnover rates for arid, mesic, and widespread lineages.'}
+Thin grey lines: individual draws from the MCMC.
+Blue lines: 95% credibility interval in the MCMC.
+Dotted green line: inferred mean rate through time estimated from the MCMC iterations.
+(c) Density plot of branch-specific rates by geographic range from ClaDS analysis.'}
 
@@
 
-Our result supports the hypothesis that the ancestor of the Australian clade first invaded the mesic part of the Australian continent, which is consistent with previous studies [@marinTracing2013;@vidalBlindsnake2010].
+Our result supports the hypothesis that the ancestor of the Australian clade first dispersed to the mesic part of the Australian continent, which is consistent with previous studies [@marinTracing2013;@vidalBlindsnake2010].
 The observed pattern where the ancestral lineage first inhabited tropical-mesic regions before later expanding into other biomes as they emerge has also been documented in other radiations of Australian squamates [@chappleParallel2004;@melvilleEvolutionary2011;@oliverPhylogenetic2017].
 Many Australian groups originated in Eurasia and crossed into Australia after the continental plates collided and caused the uplift of a number of 'stepping stones' including the Sunda islands, New Guinea, and the Soloman islands.
-Our historical biogeographic estimation of *Anilios* is consistent with the hypothesis that this group invaded Australia in the Miocene, first occupying tropical grasslands of Northern Australia before diversifying throughout the continent.
+Our historical biogeographic estimation of *Anilios* is consistent with the hypothesis that this group arrived in Australia during the Miocene, first occupying tropical grasslands of Northern Australia before diversifying throughout the continent.
 However, whether the Australian radiation was a result of an oceanic dispersal from Southeast Asia or Indonesia, as proposed by @vidalBlindsnake2010, is equivocal.
@@
 
-Our analyses indicated that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
+Tiatragul et al., (submitted) proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
 Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
@@
 ## Regional vs global paleoenvironment 
-Our reconstructed Australia-specific paleotemperature data provided a better fit than the global temperature data set [@condamineMacroevolutionary2013], although it was not the best fitting model.
+Our reconstructed Australia-specific paleotemperature data provided a better fit to the environment-dependent diversification models we considered in RPANDA than the global temperature data set [@condamineMacroevolutionary2013], although models that used paleoaridity provided better fit (Table 1; Table S4).
 In the context of our birth-death models, the absolute value of temperature might not have made much of a difference because our environment-dependent models fit the spline rather than absolute temperature.
@@
 As it stands, current estimates of climate at deep timescales are still developing and we see strong differences in climate estimates from different approaches.
-For example, while both GCM-based and Köppen band-based approaches showed a gradual increase in temperature from 25 – 5 Ma before a decrease from the Pliocene to the present-day, the GCM-based approach showed a much sharper increase in temperature.
-Similarly, aridity estimates from the GCM-based approach were starkly different to the Köppen band approaches.
+For example, while both GCM-based and Köppen band-based approaches showed a gradual increase in temperature from 25 – 5 Ma before a decrease from the Pliocene to the present-day, the GCM-based approach showed a much sharper increase in temperature (Fig.
+\@ref(fig:fig-paleoclimate-compare)a).
+Similarly, aridity estimates from the GCM-based approach were starkly different to the Köppen band approaches (Fig.
+\@ref(fig:fig-paleoclimate-compare)b).
 This may be because deep time GCM models are currently aimed at estimating general global fluctuations in climate, rather than regionally specific ones, which may be poorly approximated.
@@
 ## Aridification and diversification rates
-Our best-fitting RPANDA model show that aridity is negatively associated with speciation, where speciation decreased as aridity increased.
+
+Our best-fitting diversification model show that speciation is negatively correlated with aridity (Fig.
+\@ref(fig:fig-RPANDA)c).
 Our reconstructed paleoaridity index data indicate that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
@@
 
-The observed decrease in diversification rate with increasing aridification may be linked to increasing speciation rates or decreasing extinction rates.
+The observed decrease in diversification rate with increasing aridification may be linked to decreasing speciation rates or increasing extinction rates.
 One hypothesis is that mesic-adapted ancestral blindsnakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
 An extensive population genetics study of blindsnake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
-Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events.
+Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events or intraspecific diversity not yet recognised.
 Our best-fitting models of diversification support constant extinction rates with decreasing rates of speciation. 
@@
 Yet, here we found no difference between arid zone and mesic zone speciation rates in blindsnakes which is inconsistent with this hypothesis. 
-Finally, another potential explanation for diversification patterns is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
+Finally, another potential explanation for the negative correlation between diversification and aridity is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
 The current phylogeny is missing 12 described species (Tiatragul et al., submitted MS), such as the arid-adapted _A.
@@
 minimus_, and the true diversity has been estimated to exceed 56 species [@marinHidden2013;@marinTracing2013;@ellisTyphlopid2019].
-It is possible that these lineages that are not represented in our phylogeny have diversified more recently when aridity is high.
+It is possible that unrecognised species may be hidden under nominal species and that these missing lineages have diversified more recently when aridity is high.
+If all suspected cryptic OTUs were treated as recognised species, we could see a different relationship with aridification.
 
@@
 This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
-
 For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
 However, results from our weight-averaged SSE analyses on the Australian blindsnake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes.
-A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
+This lack of difference between biomes could be due in part to unrecognised species from the often under sampled arid zone [@pepperLife2021]
+
+<!-- A potential explanation is that fossorial species may not be subjected to the same selective pressures present in the arid zone, such as extreme temperatures [@pepperLife2021], as much as surface dwelling species do.
 For instance, because blindsnakes live in subterranean microhabitat with relatively stable climate compared to the surface [@kinlawReview1999;@nevoAdaptive1979], modification from the already specialised phenotype may not be selected for [@cyriacDigging2018].
 Future studies that investigate phenotypic variations across all blindsnakes will further our understanding of how these snakes evolved with seemingly conserved phenotypes.
+-->
 
@@
 centralis* from the arid interior of the continent.
-Current methods do not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat seemingly featureless desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011].
-However, modern software underdevelopment such as PhyBEARS may be able to account for more complex geological processes such as the movement of continent.
+Current methods do not allow us to account for heterogeneity of the biome even though we know the arid biome consists of everything from flat, seemingly featureless, desert in the interior to topologically complex mountain ranges on the fringes [@pepperLife2021;@pepperGenetic2011].
+However, modern software under development such as PhyBEARS may be able to account for more complex geological processes such as the movement of continents.
 Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today [@oliverPhylogenetic2017].
@@
 We inferred that tropical grassland is the ancestral biome of Australian _Anilios_ and acted as a major source of biodiversity to other biomes.
-Compared to other Australian reptile radiation in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified.
-We showed that regionalised paleoaridity data provides a better fit to our Australian blindsnake phylogeny than generic global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
+Compared to other Australian reptile radiations in the Miocene, we observe a burst of diversification when the lineage first arrived on the continent and gradual decrease in diversification as the continent aridified.
+We showed that regionalised paleoaridity data provides a better fit to our Australian blindsnake phylogeny than global paleotemperature, highlighting the need to include and improve estimates of continental scale paleoenvironmental data in diversification analyses.
 To our knowledge, this is the first study that empirically fit diversification models to Australia-specific reconstructed paleoaridity data.
@@
 
-Upon acceptance, data and analyses scripts will be available from the Dryad Repository (XXXXXXX)
+Data and analyses scripts are available from the Dryad Repository (https://doi.org/10.5061/dryad.t4b8gtj5x)
 
@@
 Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model.
-(a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
-Inset map (left) shows the current distribution of biomes.
-Inset maps to the right of the tree shows distribution of species in Group 1 and Group 2, respectively.
-(b) Histogram showing frequency and composition of outward transitions from Australian biomes and combined non-Australian islands.
-(c) Histogram showing frequency and composition of inward transitions from Australian biomes and combined non-Australian islands.
+(a) Inset map (left) shows biome classifications of the regions considered in this study.
+Coloured boxes at the tips correspond to biomes on the map.
+Pie charts represent ancestral biome reconstruction based on 50 biogeographic stochastic maps (BSM) for each given nodes.
+Colours present on the pie charts but not in the biome map indicate mixture of more than one biome.
+Coloured boxes below pie charts indicate the most probable biome(s) at visually ambiguous nodes.
+Inset maps (right of the phylogeny) shows distribution of species in Group 1 and Group 2.
+(b) Barplots showing frequency and composition of outward transitions and (c) inward transitions from Australian biomes and combined non-Australian islands.
 Abbreviations: TrG = Tropical Grassland, TmF = Temperate Forest, AR = Arid, MD = Mediterranean, TrF = Tropical Forest, and TmG = Temperate Grassland.
@@
 Fig.
-Effect of aridity change on the diversification of Anilios estimated from the phylogeny.
+Effect of aridity change on the diversification of *Anilios* estimated from the phylogeny.
 (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
-(b) Speciation rates through time from the best diversification model.
-(c) Association between speciation rate and aridity index.
-We found that increased aridity is associated with decrease speciation rate.
+(b) Fitted diversification rates through time reconstructed from parameter estimates from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”).
+(c) Fitted diversification rates at different aridity index values calculated from parameter estimates of the previous model.
+Increased aridity is associated with decrease net diversification rates in this model.
 
@@
 Diversification rates estimates from ClaDS and SSE models.
-(a) Branch-specific rates across the phylogeny.
+(a) Branch-specific rates (log(λi)) across the phylogeny.
 (b) Mean branch-specific rates over time.
+Thin grey lines: individual draws from the MCMC.
+Blue lines: 95% credibility interval in the MCMC.
+Dotted green line: inferred mean rate through time estimated from the MCMC iterations.
 (c) Density plot of branch-specific rates by geographic range from ClaDS analysis.
-(d) Speciation rates from four geographic diversification rate models averaged by model weight.
-(e) Estimates of speciation rate and (d) turnover rates for arid, mesic, and widespread lineages.
 
@@
 
-Alexander Skeels researches the evolution of biodiversity and how deep-time processes such as plate tectonics and climate change shape diversification patterns in plants and animals.
+Alexander Skeels is a postdoctoral researcher who works on the evolution of biodiversity and how deep-time processes such as plate tectonics and climate change shape diversification patterns in plants and animals.
 

### 2023-05-16 manuscript.Rmd
update to match word manuscript after revision

@@
 
-**Running title** Decrease in blindsnake speciation correlate with aridification
+**Running title** Blindsnake speciation slows down with aridification
 
-**Authors**: Sarin Tiatragul^1^; Alexander Skeels^2,3^; J. Scott Keogh^1^
+**Authors**: Sarin Tiatragul^1^; Alexander Skeels1^,^2,3^; J. Scott Keogh^1^
 
@@
 ST would like to thank J Rolland and F.
-Condamine for helpful advice about fitting diversification models with RPANDA.
+Condamine for their helpful advice about fitting diversification models with RPANDA.
 Thanks to NJ Matzke and O Jiménez-Robles for help with BioGeoBEARS.
-Special thanks to IG Brennan, D Esquerré, M Pepper, CJ Pavón-Vázquez, and O Hagen for helpful advice on the analysis and discussion.
+Special thanks to IG Brennan, D Esquerré, M Pepper, CJ Pavón-Vázquez, and O Hagen for helpful advice on the analyses and discussion.
 JSK thanks the Australian Research Council for ongoing support.
@@
 
-Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species  **47** described species [@uetzQuarter2021] (Tiatragul et al., submitted MS).
+Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species **47** described species [@uetzQuarter2021] (Tiatragul et al., accepted).
 All but three species of currenty described _Anilios_ are endemic to Australia, _A.
@@
 However, phylogenetic analyses from Tiatragul et al.
-(2023) indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin
+(accepted) indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin.
 Despite representing a substantial portion of snake diversity, we know little about the macroevolutionary dynamics of blindsnakes.
@@
 Tiatragul et al.
-(submitted) time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007] (see Tiatragul et al., submitted MS for details).
+(accepted) time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007].
 Phylogenetic data in Tiatragul et al.
 (submitted) suggest the existence of at least three genetically distinct undescribed species, so we followed their designation and adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., submitted MS).
-For ancestral biome estimation analyses, we included three outgroup blindsnakes species from Southeast Asia, Indonesia, and Melanesia, where the ancestor of *Anilios* has been hypothesised to have originated [@vidalBlindsnake2010].
+The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., accepted).
+For ancestral biome estimation analyses, we included three outgroup blindsnakes species from Southeast Asia, Indonesia, and Melanesia, where the ancestor of *Anilios* has been hypothesised to have originated [@vidalBlindsnake2010; Tiatragul et al., accepted].
 The three species include *Ramphotyphlops multilineatus* (Western New Guinea, Kai Island, and Salawati), *Sundatyphlops polygrammicus* (Moyo Island, Indonesia), and *Acutotyphlops subocularis* (Bismarck Archipelago).
@@
 These shifts in biomes can occur within a single branch (anagenetic events, including processes such as dispersal and extinction) or at the time of branching (cladogenetic events involving sympatric, vicariant, and founder-event speciation).
-We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database.
+We divided Australo-Papua into nine discrete biomes and assigned each Australian species to one or more biomes based on their occurrence records from the Atlas of Living Australia database (ALA).
 Biome classifications were based on the World Wildlife Fund's (WWF) Terrestrial Ecoregions of the World data set [@olsonTerrestrial2001] and followed the nomenclature of published Australia-specific studies [@cardilloPhylogeny2017;@pavon-vazquezComprehensive2021].
@@
 To test whether diversification rates differ between arid-adapted and mesic-adapted lineages, we used two approaches.
-First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) using ClaDS function in the Julia package 'PANDA' v.1.6.1 [@malietFast2022].
+First, we estimated speciation rates at the origin of each branch (henceforth branch-specific rates) using ClaDS function in the Julia package PANDA v.1.6.1 [@malietFast2022].
 ClaDS is designed to capture frequent shifts in speciation rates with small effects.
@@
 
-Based on current distributions from ALA, most Anilios species are found entirely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
-Apart from A.
-silvia, A.
-leucoproctus, and A.
-proximus that appeared to form independent divergent lineages (Fig.
+Based on current distributions from ALA, most *Anilios* species are found entirely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and Mediterranean biome (n = 10).
+Apart from *A.
+silvia*, *A.
+leucoproctus*, and *A.
+proximus* that appeared to form independent divergent lineages (Fig.
 \@ref(fig:fig-biogeobears)a), we identified divergent biogeographic transitions between two major groups of species, which we refer to as Group 1 (n = 21 lineages) and Group 2 (n = 14 lineages).
@@
 The tropical grassland biome is the largest source of outward dispersal (16.2 ± 4.9 events; 46.6%) followed by the arid biome (9.7 ± 3.2 events; 26%)(Fig.
-1b).
+\@ref(fig:fig-biogeobears)b).
 On the other hand, the largest sinks for all inward dispersals are the tropical forest biome (7.9 ± 4.2 events; 21.2%) and the Mediterranean biome (7.0 ± 2.9 events; 18.8%).
@@
 ```{r fig-biogeobears, fig.cap= 'Ancestral biome estimation and transitions of Australo-Papuan blindsnakes in BioGeoBEARS, using the DEC+j+x model.
-(a) Pie charts based on 50 biogeographic stochastic map (BSM) representing probabilities of each biome at speciation event.
-Inset map (left) shows the current distribution of biomes.
-Inset maps to the right of the phylogeny shows distribution of species in Group 1 and Group 2.
+(a) Inset map (left) shows biome classifications of the regions considered in this study.
+Coloured boxes at the tips correspond to biomes on the map.
+Pie charts represent ancestral biome reconstruction based on 50 biogeographic stochastic maps (BSM) for each node.
+Colours present on the pie charts but not in the biome map indicate mixture of more than one biome.
+Coloured boxes below pie charts indicate the most probable biome(s) at visually ambiguous nodes.
+Inset maps (right of the phylogeny) shows distribution of species in Group 1 and Group 2.
 (b) Bar plot showing frequency and composition of outward transitions and (c) inward transitions from Australian biomes and combined non-Australian islands.
@@
 
-```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation based on three different methods and RPANDA.
+```{r fig-paleoclimate-compare, fig.cap= 'Paleoclimate estimation derived from RPANDA (pink),  a global circulation model (GCM) approach (black = Valdes et al., 2021), and two paleo-Köppen climate bands approach (orange = Straume et al., 2020; blue = Scotese & Wright, 2018).
 (a) Estimated mean temperature over time in Australia compared to RPANDA global paleotemperature data.
-(b) Estimated mean aridity index over time based on precipitation threshold.
+(b) Estimated mean aridity index over time based on precipitation threshold (arid < 500mm/year).
 Abbreviations Pli = Pliocene and Ple = Pleistocene.'}
@@
 (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line).
-(b) Fitted diversification rates through time reconstructed from parameter estimates from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”).
+(b) Fitted diversification rates through time reconstructed from parameter estimates (λ = 1.32, α = -0.14, µ = 0.23) from the best diversification model (speciation rates vary exponentially with paleoaridity and constant extinction rate; “BEnvVar + DCST (EXPO)”).
 (c) Fitted diversification rates at different aridity index values calculated from parameter estimates of the previous model.
-Increased aridity is associated with decrease net diversification rates in this model..'}
+Increased aridity is associated with decrease net diversification rates in this model.'}
 
@@
 
-Tiatragul et al., (submitted) proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
+Tiatragul et al., (accepted) proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
 Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
@@
 We suspect there remains several undescribed species of blindsnakes in Australia and New Guinea that may belong in the genus _Anilios_.
-Getting genetic data from _A.
-erycinus_ will allow us to place the species among other _Anilios_ and provide a more comprehensive picture of the evolution of this genus.
+Future studies should endeavour to include more samples per species to increase the chance of detecting cryptic OTUs.
+We also wish to highlight that obtaining genetic data from _A.
+erycinus_ should be a priority as this will allow us to place the species among other _Anilios_ and provide a more comprehensive picture of the evolution of this genus.
 

### 2023-05-29 manuscript.Rmd
update manuscript

@@
 The influence of temperature on diversification rates has been empirically tested with birth-death models [@condamineMacroevolutionary2013;@kergoatOpposite2018;@condamineClimate2019;@rollandContribution2019].
-Traditionally, these temperature data are reconstructed (interpolated) based on conversion of global concentration of oxygen isotopes [@condamineMacroevolutionary2013].
+Global paleotemperature data for these analyses was first presented in @condamineMacroevolutionary2013, whereby relative proportions of different oxygen isotopes in samples of benthic foraminifer shells [@zachosEarly2008] over geological time was converted to absolute temperature using a formula from @epsteinRevised1953.
 The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of regional aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].

### 2023-07-06 manuscript.Rmd
change citations

@@
 However, despite advancements in our understanding of the evolutionary relationships and divergence times of many clades, our knowledge of the impact of paleoenvironmental change on these clades remains limited, particularly in regard to the regional scale at which many clades radiate [@eyresClimatic2021].
-As such, it remains a key challenge to combine regional paleoenvironmental histories, with molecular phylogenies, to understand the drivers of diversification over deep-time scales.
+As such, it remains a key challenge to combine regional paleoenvironmental histories with molecular phylogenies to understand the drivers of diversification over deep-time scales.
 
@@
 Furthermore, shifts in diversification rates due to changes in Australian temperature and aridity have been reported in some squamate reptile clades including skinks [@raboskyExceptional2007] and geckos [@fujitaDiversification2010;@oliverContrasting2014;@oliverIndependent2014;@pepperGenetic2011;@pepperPalaeoclimate2011] as well in frogs [@catulloAridification2014;@vidal-garciaConvergent2015], insects [@owenHow2017;@toussaintUnveiling2015], and plants [@crispHow2013;@kadereitMolecular2010;@millerComparative2013].
-Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have had suppressed diversification or been driven extinct [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
+Yet, the role of aridification and temperature changes on diversification is highly variable, with some arid-adapted vertebrates radiating following aridification [@raboskyExceptional2007;@pepperPalaeoclimate2011;@brennanMass2017] while other mesic-adapted groups have had suppressed diversification or been driven to extinction [@bryantLines2016;@cardilloPhylogeny2017;@owenHow2017].
 Whether environmental change in Australia has had generally positive or negative influence on diversification rates in squamate reptiles is still unknown.
 
-Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species **47** described species [@uetzQuarter2021] (Tiatragul et al., accepted).
+Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species **47** described species [@uetzQuarter2021;@tiatragulAustralia2023].
 All but three species of currenty described _Anilios_ are endemic to Australia, _A.
@@
 The diversification of _Anilios_ in Australia is reported to have originated from a single dispersal event by a mesic-adapted ancestor from Southeast Asia or Indonesia between 39 – 19 Ma [@vidalBlindsnake2010;@marinTracing2013].
-However, phylogenetic analyses from Tiatragul et al.
-(accepted) indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin.
+However, phylogenetic analyses from @tiatragulAustralia2023 indicated that the closest relative to all *Anilios* is *Ramphotyphlops multilineatus*, a species that occur in Western New Guinea, Kai Islands, and Salawati, suggesting the Australian radiation may have a Melanesian origin.
 Despite representing a substantial portion of snake diversity, we know little about the macroevolutionary dynamics of blindsnakes.
@@
 To do this, we fitted several birth-death models with reconstructed paleotemperature and aridity, and estimated diversification rates under these models.
-Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages that inhabit the arid biome have undergone faster or slower radiation.
+Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages that inhabit the arid biome have undergone faster or slower diversification.
 
@@
 The species tree was built with the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] using maximum likelihood gene trees inferred in IQTREE from 4930 loci (322 Anchored Hybrid Enrichment, 4574 ultraconserved elements, and 34 legacy genes) for 35/47 recognised *Anilios* species.
-Tiatragul et al.
-(accepted) time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007].
-Phylogenetic data in Tiatragul et al.
-(submitted) suggest the existence of at least three genetically distinct undescribed species, so we followed their designation and adjusted our taxon sampling fraction to 38/50 species.
-The genetically distinct OTUs that are not fully described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) (Tiatragul et al., accepted).
+@tiatragulAustralia2023 time-calibrated the phylogeny using 27 outgroup squamate fossils as primary calibrations using MCMCTree from the PAML package [@yangPAML2007].
+Phylogenetic data in @tiatragulAustralia2023 suggest the existence of at least three genetically distinct undescribed species, so we followed their designation and adjusted our taxon sampling fraction to 38/50 species.
+The genetically distinct OTUs that are not formally described species include three '_Anilios grypus_' (two in Western Australia and one in Queensland) and two '_Anilios ligatus_' (one from the eastern part of the range and one from the northern part) [@tiatragulAustralia2023].
 For ancestral biome estimation analyses, we included three outgroup blindsnakes species from Southeast Asia, Indonesia, and Melanesia, where the ancestor of *Anilios* has been hypothesised to have originated [@vidalBlindsnake2010; Tiatragul et al., accepted].
@@
 
-Tiatragul et al., (accepted) proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
+@tiatragulAustralia2023 proposed that the earliest divergence of *Anilios* occurred around 25 Ma (28 – 22 Ma; 95% HPD), which is slightly older than 21.9 Ma (30.2 – 15.1 Ma; 95% Bayesian credibility interval) proposed by previous studies [@marinTracing2013;@vidalBlindsnake2010].
 Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008].
@@
 Finally, another potential explanation for the negative correlation between diversification and aridity is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
-The current phylogeny is missing 12 described species (Tiatragul et al., submitted MS), such as the arid-adapted _A.
+The current phylogeny is missing 12 described species [@tiatragulAustralia2023], such as the arid-adapted _A.
 fossor_ and _A.

### 2023-07-10 manuscript.Rmd
second revisions submitted to JBI

@@
 
-**Running title** Blindsnake speciation slows down with aridification
+**Running title** Blindsnake speciation slows down with increased aridification
 
@@
 Today, the arid zone covers roughly 70% of the Australian landscape and provides habitat to 43% of all of Australia’s terrestrial reptiles [@byrneBirth2008].Of these, roughly a quarter of the snake species diversity in Australia is represented by blindsnakes of the genus *Anilios*, which currently comprises 47 described species **47** described species [@uetzQuarter2021;@tiatragulAustralia2023].
-All but three species of currenty described _Anilios_ are endemic to Australia, _A.
+All but three species of formally described _Anilios_ are endemic to Australia, _A.
 torresianus_ and _A.
@@
 The relationship between diversification rates and aridity, however, have been attributed to some proxies such as the timing of aridification or global temperature rather than quantifying the extent of regional aridity itself [@dornStrange2014;@deepakAridification2018;@catulloAridification2014;@fujitaDiversification2010;@pepperGenetic2011;@pepperPalaeoclimate2011;@oliverIndependent2014;@oliverContrasting2014;@owenHow2017;@crispHow2013;@toussaintUnveiling2015].
-
 To our knowledge no study has assessed aridity as a continuous variable in an environment-dependent birth-death model framework, probably due to the difficulty of quantifying paleoaridity.
@@
 Additionally, we compared diversification rates between arid-adapted lineages versus mesic-adapted lineages to investigate if lineages that inhabit the arid biome have undergone faster or slower diversification.
+We predicted that arid-adapted lineages will have higher diversification rates compared to mesic-adapted lineages.
+This expectation is based on the hypothesis that aridification can cause  fragmentation of continuous habitats, leading to isolated populations in distinct microhabitats and refugia in arid areas, which may lead to greater species level divergence [@brennanMiocene2018].
+Additionally, the ecological opportunities or selective pressures resulting from aridification may drive the evolution of specialised traits in mesic-adapted ancestral species, enabling them to exploit previously unoccupied niches and promote speciation.
 
@@
 With 47 recognised species, the mostly Australian-endemic blindsnake genus *Anilios* is currently the most diverse genus of Scolecophidia in the world.
-In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of invasion from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas.
+In this study we estimated that the ancestral species in Australia occupied the tropical grassland biome, most likely as a result of dispersal from other tropical mesic areas including New Guinea, Bismarck Archipelago, and the Lesser Sundas.
 Within a few million years, this ancestral lineage split with one group staying in the tropical grassland biome while the other group transitioned into the arid biome.
@@
 \@ref(fig:fig-RPANDA)c).
-Our reconstructed paleoaridity index data indicate that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antartic circumpolar current [@byrneBirth2008].
+Our reconstructed paleoaridity index data indicate that aridity increased from ~25 Ma following the opening of the Southern Ocean and the onset of the Antarctic circumpolar current [@byrneBirth2008].
 A particularly sharp increase from 10 Ma followed from the "Hill Gap" period of cooling [@hillHistory1994], and intensification from 5 Ma onwards following the Pliocene transition to a contemporary arid Australian climate [@byrneBirth2008].
@@
 One hypothesis is that mesic-adapted ancestral blindsnakes may have been widespread in the past, but populations later became extinct due to extreme aridity in the last 5 Ma [@fujiokaAustralian2009;@byrneDecline2011].
-An extensive population genetics study of blindsnake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the truly arid environments.
+An extensive population genetics study of blindsnake by @marinHidden2013 proposed that during widespread aridification, mesic-adapted lineages went extinct or were confined to mesic refugia, such as the Pilbara and Kimberley, where they became pre-adapted to seasonal aridity, which enable subsequent invasion into the more arid environments.
 Additionally, long naked branches in our current phylogeny, particularly in Group 1, suggests past extinction events or intraspecific diversity not yet recognised.
@@
 Lower rates of molecular evolution have been found in Australian arid zone plants compared to mesic zone plants [@goldieEvolutionary2010].
-Yet, here we found no difference between arid zone and mesic zone speciation rates in blindsnakes which is inconsistent with this hypothesis. 
+Yet, here we found no difference between arid biome and mesic biome speciation rates in blindsnakes which is inconsistent with this hypothesis. 
 Finally, another potential explanation for the negative correlation between diversification and aridity is the limited sampling of Australian blindsnakes, particularly in major sand deserts [@pepperLife2021] or even in areas of high cryptic diversity such as the Kimberley [@ellisTyphlopid2019].
@@
 This pattern is consistent with several other groups of reptiles that started to diversify in the Miocene including monitor lizards [@brennanPhylogenomics2021;@pavon-vazquezCompetition2022], _Sphenomorphus_ and _Eugongylus_ skinks [@skinnerLygosomine2011], elapid snakes [@sandersMolecular2008], pythons [@esquerrePhylogenomics2020;@esquerreEvolutionary2022] and several geckos [@ashmanDiversification2018;@fujitaDiversification2010;@oliverIndependent2014].
+
 For these terrestrial reptiles, increased rates of speciation for arid-adapted taxa have been attributed to species filling niches in the saxicolous and xeric habitats that emerged with the contraction of mesic forest throughout the Miocene [@sandersMolecular2008;@hugallPhylogeny2008].
-However, results from our weight-averaged SSE analyses on the Australian blindsnake phylogeny indicated that speciation, extinction, and turnover rates did not differ between lineages in arid and mesic biomes.
-This lack of difference between biomes could be due in part to unrecognised species from the often under sampled arid zone [@pepperLife2021]
+Contrary to our predictions, results from our analyses on the Australian blindsnake phylogeny revealed no differences in speciation, extinction, and turnover rates between arid and mesic lineages.
+This suggests that the influence of aridity on diversification rates is more complex and does not follow the pattern we expected if species are filling new niches.
+Several factors may explain the lack of difference in diversification rates between biomes.
+One possible biological reason is that ancestral mesic-adapted blindsnakes already possessed specialised traits suitable for both arid and mesic subterranean environments, which reduced the potential for divergent diversification rates.
+A possible lack of differentiation in prey and availability of these prey items, particularly ant and termite larvae and pupae, in both biomes could also contribute to comparable diversification rates observed.
+Additionally, all species of blindsnakes exclusively inhabit subterranean environments during the day, which shield them against surface climatic variations such as extreme heat in the arid zone [@ehmannFamily1993].
+While other reptiles may seek shade cover, many species must also physiologically adapt to extreme temperatures.
+Finally, incomplete sampling may obscure the true diversification patterns, particularly because there may be a larger proportion of unrecognised species from the arid zone[@pepperLife2021].
 
@@
 However, modern software under development such as PhyBEARS may be able to account for more complex geological processes such as the movement of continents.
+For our SSE analyses, it is also noteworthy that the lineages we designated as “mesic-adapted” in reality may occupy a variety of biomes with varying degree of aridity from tropical rainforests to tropical savannah.
 Finally, (mass) extinction likely played important roles in the pattern of diversity we observed today [@oliverPhylogenetic2017].
